### 2015-01-22 msc-publication.tex
barebones MSc publication document

\documentclass[a4paper]{article}
\usepackage{amsmath}
\usepackage{mathtools}

\title{Comparison of Terrain Following and Cut Cell Grids with a Non-Hydrostatic Model}
\author{James Shaw}
\date{TODO}

\begin{document}
\maketitle

\section{Conclusions}

\end{document}

### 2015-01-23 msc-publication.tex
start drafting conclusions and figures for msc publication; add matrix of tests to thesis

@@
 \documentclass[a4paper]{article}
+\usepackage{xcolor}
 \usepackage{amsmath}
 \usepackage{mathtools}
+\usepackage{siunitx}
 
@@
 \begin{document}
+\newcommand{\TODO}[1]{\textcolor{purple}{TODO: \emph{#1}}}
+\newcommand{\textcite}[1]{\textbf{#1}}
+
 \maketitle
 
+\section{Results}
+
+\begin{itemize}
+	\item Horizontal advection
+	\item \TODO{Maybe wobblyTracerAdvection}
+	\item Resting atmosphere
+	\item Gravity waves
+\end{itemize}
+
+\begin{figure}
+	btf-cubicUpwindCPCFit-divFree \\
+	\input{../advectTracer-horizontal-btf-cubicUpwindCPCFit-divFree/tracer-contour}
+	\input{../advectTracer-horizontal-btf-cubicUpwindCPCFit-divFree/error-contour}
+
+	schaer-4thorder \\
+	\includegraphics[height=1.5in]{schaer-btf-4thorder.png}
+	\includegraphics[height=1.5in]{schaer-btf-4thorder-error.png}
+%
+	\caption{Comparing our BTF cubicUpwindCPCFit result with Schaer's 4th order result}
+	\label{fig:advection}
+\end{figure}
+
+\begin{figure}
+	\includegraphics[height=2in]{wobblyTracerAdvection.png}
+%
+	\caption{tracer advection in terrain following velocity field}
+	\label{fig:wobblyTracerAdvection}
+\end{figure}
+
+\begin{figure}
+	\includegraphics[height=3in]{resting-atmosphere-w.png}
+%
+	\caption{Comparing our results with \textcite{klemp2011}'s}
+	\label{fig:resting}
+\end{figure}
+
+\begin{figure}
+	\includegraphics[height=2in]{gw-w-btf.png}
+	\includegraphics[height=2in]{melvin-7a.png}
+%
+	\caption{Gravity wave vertical velocities and thermal anomalies from initial thermal profile}
+	\label{fig:gw-w}
+\end{figure}
+
+\begin{figure}
+	\includegraphics[height=3in]{gw-theta.png}
+	\includegraphics[height=3in]{gw-theta-zoom.png}
+%
+	\caption{Gravity wave thermal anomalies from initial thermal profile}
+	\label{fig:gw-theta}
+\end{figure}
+
+\begin{figure}
+	\includegraphics[height=3in]{gw-exner-theta.png}
+%
+	\caption{Vertical sample line of Exner function of pressure and theta to demonstrate Lorenz computational mode}
+	\label{fig:gw-exner-theta}
+\end{figure}
+
 \section{Conclusions}
+\begin{enumerate}
+	\item BTF grid isn't as bad as people say it is.  We found that:
+	\begin{itemize}
+		\item Upwind-biased cubic advection scheme accurately advects a tracer in non-divergent flows (Figure~\ref{fig:advection})
+		\item Advection in a TF velocity field, if we include it, gives better results on BTF than SnapCol (Figure~\ref{fig:wobblyTracerAdvection})
+		\item spurious vertical velocities are small in resting atmosphere (Figure~\ref{fig:resting})  Could also compare with \textcite{zaengl2012}, \textcite{good2014}
+		\item gravity waves results visually as good as reference solution from \textcite{melvin2010} (figure~\ref{fig:gw-w})
+	\end{itemize}
+
+	\item Cut cell grids can be worse than TF grids:
+	\begin{itemize}
+		\item Lorenz computational mode found on SnapCol grid only (figure~\ref{fig:gw-theta}, \ref{fig:gw-exner-theta})
+	\end{itemize}
+\end{enumerate}
+
+Miscellany:
+\begin{itemize}
+	\item Advection accuracy depends on alignment of the flow with grid layers
+	\item Little/no evidence of small cell problem in GW test
+	\item Lack of energy conservation (evidenced in resting atmosphere test, Fig 4.12d in MSc dissertation)
+\end{itemize}
+
+\section{Further work}
+\begin{itemize}
+	\item Lorenz computational mode motivates formulation of C-P staggering on unstructured grids
+	\item Find out what excites the Lorenz computational mode
+	\item Fix lack of energy conservation
+\end{itemize}
 

### 2015-01-27 msc-publication.tex
add table of stats for advection

@@
 \usepackage{siunitx}
+\usepackage{booktabs}
 
@@
 \end{figure}
+
+\begin{table}
+\centering
+\begin{tabular}{ l l l l l l }
+\toprule
+& \multicolumn{3}{c}{Cubic upwind-biased} & \multicolumn{2}{c}{Sch\"ar 4th order} \\
+& $\ell^2$ error & min & max & min & max \\
+\midrule
+Analytic  & 0 & 0 & 1 & \multicolumn{1}{c}{---} & \multicolumn{1}{c}{---} \\
+BTF 	  & TODO & TODO & TODO & \num{-0.058} & \num{1.001} \\
+SnapCol   & TODO & TODO & TODO & \multicolumn{1}{c}{---} & \multicolumn{1}{c}{---} \\
+noOrography & TODO & TODO & TODO & \num{-0.002} & \num{0.984} \\
+\bottomrule
+\end{tabular}
+%
+\caption{Min, max and error norms compared with \textcite{schaer2002}}
+\label{tab:advection-stats}
+\end{table}
 
@@
 	\begin{itemize}
-		\item Upwind-biased cubic advection scheme accurately advects a tracer in non-divergent flows (Figure~\ref{fig:advection})
+		\item Upwind-biased cubic advection scheme accurately advects a tracer in non-divergent flows (Figure~\ref{fig:advection}, Table~\ref{tab:advection-stats})
 		\item Advection in a TF velocity field, if we include it, gives better results on BTF than SnapCol (Figure~\ref{fig:wobblyTracerAdvection})
@@
 	\item Little/no evidence of small cell problem in GW test
-	\item Lack of energy conservation (evidenced in resting atmosphere test, Fig 4.12d in MSc dissertation)
+	\item Formulation does not conserve energy (evidenced in resting atmosphere test, Fig 4.12d in MSc dissertation)
 \end{itemize}
@@
 \begin{itemize}
-	\item Lorenz computational mode motivates formulation of C-P staggering on unstructured grids
+	\item Lorenz computational mode motivates formulation of C-P staggering for cut cell grids
 	\item Find out what excites the Lorenz computational mode
-	\item Fix lack of energy conservation
 \end{itemize}

### 2015-01-27 msc-publication.tex
beef up notes for each test case

@@
 
-\section{Results}
+Builds on the work of \textcite{schaer2002} (MWR), \textcite{klemp2011} (MWR), \textcite{weller-shahrokhi2014} (MWR)
+
+Results also compared with \textcite{melvin2010} (QRJ), maybe \textcite{zaengl2012} (MWR), \textcite{good2014} (Atmos Sci Letters)
+
+\section{Introduction}
 
 \begin{itemize}
-	\item Horizontal advection
-	\item \TODO{Maybe wobblyTracerAdvection}
-	\item Resting atmosphere
-	\item Gravity waves
+	\item Small cell problem
+\end{itemize}
+
+\section{Results}
+\begin{itemize}
+	\item \TODO{How much do I need to say about the model?  Just cite \textcite{weller-shahrokhi2014}?}
+	\item \TODO{Include BTF equations from \textcite{galchen-somerville1975}?}
+	\item \TODO{Explain how SnapCol grid is constructed?}
+\end{itemize}
+
+\subsection{Advection}
+\begin{itemize}
+	\item Motivation: tests advection scheme accuracy, especially with grid distortions
+	\item Specify domain, terrain, discretisation in time and space, tracer
+	\item Specify wind field, projection method for non-divergence
+	\item Advection equation in flux form
+	\item Upwind-biased cubic advection scheme, non-monotonic, not flux corrected
+	\item d/dt scheme (RK2)
+\end{itemize}
+
+We compare the following grids:
+\begin{itemize}
+	\item No orography (which tells us how good the model can get)
+	\item BTF
+	\item SLEVE (only briefly since it's not adding so much to our argument)
+	\item SnapCol (briefly because it's the same as noOrography)
+\end{itemize}
+
+Analysis:
+\begin{itemize}
+	\item Upwind-biased cubic is qualitatively better than Sch\"ar's 4th order leapfrog on BTF grid (figure~\ref{fig:advection})
+	\item SnapCol error, min and max is almost identical to noOrography (unsurprisingly) (table~\ref{tab:advection})  This agrees with \textcite{good2014}
+	\item SLEVE error, min and max are part-way between BTF and noOrography (again, unsurprisingly)
+	\item noOrography, BTF and SLEVE min and max comparable to \textcite{schaer2002}
+	\item Conclusion: advection more accurate on more regular grids, but nevertheless satisfactory on all grids
 \end{itemize}
@@
 \caption{Min, max and error norms compared with \textcite{schaer2002}}
-\label{tab:advection-stats}
+\label{tab:advection}
 \end{table}
 
-\begin{figure}
-	\includegraphics[height=2in]{wobblyTracerAdvection.png}
-%
-	\caption{tracer advection in terrain following velocity field}
-	\label{fig:wobblyTracerAdvection}
-\end{figure}
+\subsection{Resting atmosphere}
+
+\TODO{do we want to discuss energy measures?
+1.
+we can't compare our results with anybody else, 2.
+we'll have to explain how we calculate the energy measures 3.
+it could open a can of worms w.r.t.
+lack of conservation}
+
+\begin{itemize}
+	\item Motivation: challenge accuracy of H operator (\TODO{if we mention H op here we'll have to explain it beforehand})
+	\item Specify domain, thermodynamics, discretisation in time and space
+\end{itemize}
+
+Analysis:
+\begin{itemize}
+	\item Spurious $w$ significantly less than those found by \textcite{klemp2011} for BTF: \SI{0.35}{\meter\per\second} vs \SI{10}{\meter\per\second}
+	\item SnapCol $w$ (\SI{1e-3}{\meter\per\second}) significantly lower and energy conservation better than BTF or SLEVE
+	\item Formulation is not energy conserving
+	\item \TODO{Could also compare with \textcite{zaengl2012}, \textcite{good2014}?}
+	\item Conclusion: non-orthogonality is a significant cause of numerical error in this test but, again, satisfactory on all grids
+\end{itemize}
 
@@
 \end{figure}
+
+\subsection{Gravity waves}
+\begin{itemize}
+	\item \TODO{Motivation: it models a real dynamic process?}
+	\item Specify domain, thermodynamics, prescribed inlet wind
+	\item Specify sponge layers, BCs
+	\item Compare BTF and SnapCol results (\TODO{no need to include SLEVE here as the results are visually identical to BTF?})
+\end{itemize}
+
+\begin{itemize}
+	\item Little/no evidence of small cell problem
+\end{itemize}
 
@@
 	\begin{itemize}
-		\item Upwind-biased cubic advection scheme accurately advects a tracer in non-divergent flows (Figure~\ref{fig:advection}, Table~\ref{tab:advection-stats})
-		\item Advection in a TF velocity field, if we include it, gives better results on BTF than SnapCol (Figure~\ref{fig:wobblyTracerAdvection})
-		\item spurious vertical velocities are small in resting atmosphere (Figure~\ref{fig:resting})  Could also compare with \textcite{zaengl2012}, \textcite{good2014}
+		\item Upwind-biased cubic advection scheme accurately advects a tracer in non-divergent flows (Figure~\ref{fig:advection}, Table~\ref{tab:advection})
+		\item spurious vertical velocities are small in resting atmosphere (Figure~\ref{fig:resting})
 		\item gravity waves results visually as good as reference solution from \textcite{melvin2010} (figure~\ref{fig:gw-w})
@@
 	\end{itemize}
+
+	\item Cut cell grids can also be better than TF grids in more artificial test cases:
+	\begin{itemize}
+		\item SnapCol $w$ two orders of magnitude smaller in resting atmosphere test
+		\item SnapCol advection test as good as noOrography
+	\end{itemize}
 \end{enumerate}
+
+\textbf{With the exception of the Lorenz computational mode on the SnapCol grid in the gravity waves test, results were satisfactory across BTF, SLEVE and SnapCol grids in all three test cases.}
 
@@
 \begin{itemize}
-	\item Advection accuracy depends on alignment of the flow with grid layers
-	\item Little/no evidence of small cell problem in GW test
+	\item Advection accuracy depends on alignment of the flow with grid layers (\TODO{we kinda need wobblyTracerAdvection to reach this conclusion})
 	\item Formulation does not conserve energy (evidenced in resting atmosphere test, Fig 4.12d in MSc dissertation)

### 2015-01-28 msc-publication.tex
more bullet points on gravity waves

@@
 \begin{itemize}
+	\item Comparing results BTF, SLEVE and cut cell style grid for three standard two-dimensional test cases
 	\item Small cell problem
@@
 	\item \TODO{How much do I need to say about the model?  Just cite \textcite{weller-shahrokhi2014}?}
-	\item \TODO{Include BTF equations from \textcite{galchen-somerville1975}?}
+	\item \TODO{Include BTF equations from \textcite{galchen-somerville1975} and SLEVE equations from \textcite{schaer2002}?}
 	\item \TODO{Explain how SnapCol grid is constructed?}
@@
 	\item Specify sponge layers, BCs
-	\item Compare BTF and SnapCol results (\TODO{no need to include SLEVE here as the results are visually identical to BTF?})
+	\item Compare BTF, SLEVE and SnapCol results (SLEVE only briefly as visually identical to BTF)
 \end{itemize}
 
+Analysis:
 \begin{itemize}
-	\item Little/no evidence of small cell problem
+	\item $w$ contours visually similar on all grids, agree with \textcite{melvin2010} (Figure~\ref{fig:gw-w})
+	\item \TODO{don't know if divergence is relevant -- it's greater on the SnapCol grid (MSc dissertation Figure 4.15e)}
+	\item $\theta$ anomalies similar on all grids EXCEPT...
+	\item ... on SnapCol grid in lee of mountain near the ground. (Figure~\ref{fig:gw-theta})
+	\item \TODO{worth mentioning implications of this?
+1.
+reduced stability, although not enough to create vertical motion in this instance 2.
+model has no viscosity so thermal mixing should not occur in theory}
+	\item Lorenz computation mode has been excited because Exner sample line at $x = \SI{50}{\kilo\meter}$ is in hydrostatic balance (Figure~\ref{fig:gw-exner-theta})
+	\item \TODO{We can speculate on what excites the computational mode, but perhaps better not to?}
+	\item Little/no evidence of small cell problem -- \TODO{don't know that we can say much here without another vertical momentum test; our hypothesis about the quasi-horizontal flow isn't backed up with a test case}
+	\item Conclusion: results similar on all grids, agree with literature, except for Lorenz computational mode on SnapCol grid
 \end{itemize}

### 2015-01-28 msc-publication.tex
more introduction
add ametsoc style and a bibtex reference

@@
-\documentclass[a4paper]{article}
+\documentclass{ametsoc}
 \usepackage{xcolor}
@@
 \usepackage{siunitx}
-\usepackage{booktabs}
+
+\bibpunct{(}{)}{;}{a}{}{,}
 
 \title{Comparison of Terrain Following and Cut Cell Grids with a Non-Hydrostatic Model}
-\author{James Shaw}
-\date{TODO}
+\authors{James Shaw\correspondingauthor{Dept., Institution, Address, City, State/Country.}}
+\affiliation{}
+\email{j.shaw@pgr.reading.ac.uk}
+
+\abstract{Enter the text of your abstract here.}
 
@@
 
-Builds on the work of \textcite{schaer2002} (MWR), \textcite{klemp2011} (MWR), \textcite{weller-shahrokhi2014} (MWR)
+Builds on the work of \citet{schaer2002} (MWR), \textcite{klemp2011} (MWR), \textcite{weller-shahrokhi2014} (MWR)
 
@@
 	\item Comparing results BTF, SLEVE and cut cell style grid for three standard two-dimensional test cases
-	\item Small cell problem
+	\item TF grids often shown to have lesser accuracy than cut cells in some test cases, often significantly so.
+Our results demonstrate that accuracy need not be much reduced and can even be better than cut cells.
+	\item Small cell problem common in cut cell grids (\TODO{citations})
 \end{itemize}
@@
 \end{figure}
-
-\begin{table}
-\centering
-\begin{tabular}{ l l l l l l }
-\toprule
-& \multicolumn{3}{c}{Cubic upwind-biased} & \multicolumn{2}{c}{Sch\"ar 4th order} \\
-& $\ell^2$ error & min & max & min & max \\
-\midrule
-Analytic  & 0 & 0 & 1 & \multicolumn{1}{c}{---} & \multicolumn{1}{c}{---} \\
-BTF 	  & TODO & TODO & TODO & \num{-0.058} & \num{1.001} \\
-SnapCol   & TODO & TODO & TODO & \multicolumn{1}{c}{---} & \multicolumn{1}{c}{---} \\
-noOrography & TODO & TODO & TODO & \num{-0.002} & \num{0.984} \\
-\bottomrule
-\end{tabular}
-%
-\caption{Min, max and error norms compared with \textcite{schaer2002}}
-\label{tab:advection}
-\end{table}
 
@@
 	\item Find out what excites the Lorenz computational mode
+	\item \TODO{anything else?}
 \end{itemize}
 
+\acknowledgments
+Start acknowledgments here.
+
+% REFERENCES
+
+\bibliographystyle{ametsoc2014}
+\bibliography{references}
+
+% TABLES
+
+\begin{table}[t]
+\caption{Min, max and error norms compared with \textcite{schaer2002}}
+\label{tab:advection}
+%
+\begin{center}
+\begin{tabular}{ l l l l l l }
+\hline\hline
+& \multicolumn{3}{c}{Cubic upwind-biased} & \multicolumn{2}{c}{Sch\"ar 4th order} \\
+& $\ell^2$ error & min & max & min & max \\
+\hline
+Analytic  & 0 & 0 & 1 & \multicolumn{1}{c}{---} & \multicolumn{1}{c}{---} \\
+BTF 	  & TODO & TODO & TODO & \num{-0.058} & \num{1.001} \\
+SnapCol   & TODO & TODO & TODO & \multicolumn{1}{c}{---} & \multicolumn{1}{c}{---} \\
+noOrography & TODO & TODO & TODO & \num{-0.002} & \num{0.984} \\
+\hline
+\end{tabular}
+\end{center}
+\end{table}
+
+% FIGURES
+
 \end{document}

### 2015-01-29 msc-publication.tex
add more references in AMS bibtex format

@@
-\documentclass{ametsoc}
+\documentclass[twocol]{ametsoc}
+\journal{mwr}
+
 \usepackage{xcolor}
@@
 
-Builds on the work of \citet{schaer2002} (MWR), \textcite{klemp2011} (MWR), \textcite{weller-shahrokhi2014} (MWR)
+Builds on the work of \citet{schaer2002}, \citet{klemp2011}, \citet{weller-shahrokhi2014} (all MWR)
 
-Results also compared with \textcite{melvin2010} (QRJ), maybe \textcite{zaengl2012} (MWR), \textcite{good2014} (Atmos Sci Letters)
+Results also compared with \citet{melvin2010} (QRJ), maybe \textcite{zaengl2012} (MWR), \citet{good2014} (Atmos Sci Letters)
 
@@
 \begin{itemize}
-	\item \TODO{How much do I need to say about the model?  Just cite \textcite{weller-shahrokhi2014}?}
-	\item \TODO{Include BTF equations from \textcite{galchen-somerville1975} and SLEVE equations from \textcite{schaer2002}?}
+	\item \TODO{How much do I need to say about the model?  Just cite \citet{weller-shahrokhi2014}?}
+	\item \TODO{Include BTF equations from \textcite{galchen-somerville1975} and SLEVE equations from \citet{schaer2002}?}
 	\item \TODO{Explain how SnapCol grid is constructed?}
@@
 	\item Upwind-biased cubic is qualitatively better than Sch\"ar's 4th order leapfrog on BTF grid (figure~\ref{fig:advection})
-	\item SnapCol error, min and max is almost identical to noOrography (unsurprisingly) (table~\ref{tab:advection})  This agrees with \textcite{good2014}
+	\item SnapCol error, min and max is almost identical to noOrography (unsurprisingly) (table~\ref{tab:advection})  This agrees with \citet{good2014}
 	\item SLEVE error, min and max are part-way between BTF and noOrography (again, unsurprisingly)
-	\item noOrography, BTF and SLEVE min and max comparable to \textcite{schaer2002}
+	\item noOrography, BTF and SLEVE min and max comparable to \citet{schaer2002}
 	\item Conclusion: advection more accurate on more regular grids, but nevertheless satisfactory on all grids
@@
 \begin{itemize}
-	\item Spurious $w$ significantly less than those found by \textcite{klemp2011} for BTF: \SI{0.35}{\meter\per\second} vs \SI{10}{\meter\per\second}
+	\item Spurious $w$ significantly less than those found by \citet{klemp2011} for BTF: \SI{0.35}{\meter\per\second} vs \SI{10}{\meter\per\second}
 	\item SnapCol $w$ (\SI{1e-3}{\meter\per\second}) significantly lower and energy conservation better than BTF or SLEVE
 	\item Formulation is not energy conserving
-	\item \TODO{Could also compare with \textcite{zaengl2012}, \textcite{good2014}?}
+	\item \TODO{Could also compare with \textcite{zaengl2012}, \citet{good2014}?}
 	\item Conclusion: non-orthogonality is a significant cause of numerical error in this test but, again, satisfactory on all grids
@@
 \begin{figure}
-	\includegraphics[height=3in]{resting-atmosphere-w.png}
+	\includegraphics[height=2in]{resting-atmosphere-w.png}
 %
-	\caption{Comparing our results with \textcite{klemp2011}'s}
+	\caption{Comparing our results with \citet{klemp2011}'s}
 	\label{fig:resting}
@@
 \begin{itemize}
-	\item $w$ contours visually similar on all grids, agree with \textcite{melvin2010} (Figure~\ref{fig:gw-w})
+	\item $w$ contours visually similar on all grids, agree with \citet{melvin2010} (Figure~\ref{fig:gw-w})
 	\item \TODO{don't know if divergence is relevant -- it's greater on the SnapCol grid (MSc dissertation Figure 4.15e)}
@@
 	\includegraphics[height=3in]{gw-theta.png}
-	\includegraphics[height=3in]{gw-theta-zoom.png}
+	\includegraphics[height=2in]{gw-theta-zoom.png}
 %
@@
 \begin{figure}
-	\includegraphics[height=3in]{gw-exner-theta.png}
+	\includegraphics[height=2in]{gw-exner-theta.png}
 %
@@
 		\item spurious vertical velocities are small in resting atmosphere (Figure~\ref{fig:resting})
-		\item gravity waves results visually as good as reference solution from \textcite{melvin2010} (figure~\ref{fig:gw-w})
+		\item gravity waves results visually as good as reference solution from \citet{melvin2010} (figure~\ref{fig:gw-w})
 	\end{itemize}
@@
 \begin{table}[t]
-\caption{Min, max and error norms compared with \textcite{schaer2002}}
+\caption{Min, max and error norms compared with \citet{schaer2002}}
 \label{tab:advection}

### 2015-01-30 msc-publication.tex
don't discuss energy measures for resting atmosphere

@@
 \begin{itemize}
+	\item A sentence or two about why orography is important to represent in NWP
 	\item Comparing results BTF, SLEVE and cut cell style grid for three standard two-dimensional test cases
@@
 \begin{itemize}
-	\item \TODO{How much do I need to say about the model?  Just cite \citet{weller-shahrokhi2014}?}
+	\item \TODO{How much do I need to say about the model?  -- Give a summary: what equation set? finite volume with curl-free pressure grad; lorenz grid; advection scheme}
 	\item \TODO{Include BTF equations from \textcite{galchen-somerville1975} and SLEVE equations from \citet{schaer2002}?}
@@
 
+\TODO{should we just compare cubicUpwindCPCFit with Sch\"ar's 4th order?  We could also compare our skewCorrected linear with Sch\"ar's linear but I'm not sure what value it would have...}
+
 Analysis:
@@
 	\includegraphics[height=1.5in]{schaer-btf-4thorder-error.png}
+
+	\TODO{also include SnapCol results}
+	\TODO{check error contours are correct (0.01?)}
 %
@@
 \subsection{Resting atmosphere}
-
-\TODO{do we want to discuss energy measures?
-1.
-we can't compare our results with anybody else, 2.
-we'll have to explain how we calculate the energy measures 3.
-it could open a can of worms w.r.t.
-lack of conservation}
 
@@
 	\item Spurious $w$ significantly less than those found by \citet{klemp2011} for BTF: \SI{0.35}{\meter\per\second} vs \SI{10}{\meter\per\second}
-	\item SnapCol $w$ (\SI{1e-3}{\meter\per\second}) significantly lower and energy conservation better than BTF or SLEVE
-	\item Formulation is not energy conserving
+	\item SnapCol $w$ (\SI{1e-3}{\meter\per\second}) significantly than BTF or SLEVE
 	\item \TODO{Could also compare with \textcite{zaengl2012}, \citet{good2014}?}
@@
 	\item Advection accuracy depends on alignment of the flow with grid layers (\TODO{we kinda need wobblyTracerAdvection to reach this conclusion})
-	\item Formulation does not conserve energy (evidenced in resting atmosphere test, Fig 4.12d in MSc dissertation)
 \end{itemize}

### 2015-02-02 fig-advection-tracer.tex
put advection contours into their own PDF, include in main PDF

\documentclass{article}
\usepackage[paperwidth=3in,paperheight=4.6in,margin=0in]{geometry}
\usepackage{mathptmx}%{times}
\usepackage{graphicx}
\usepackage{siunitx}

\setlength{\parindent}{0pt}
\begin{document}
\centering
\hspace{0.05in}\includegraphics[width=2.75in]{img/schaer-btf-4thorder.png}

\footnotesize
\input{../advectTracer-horizontal-btf-cubicUpwindCPCFit-divFree/tracer-contour}

\input{../advectTracer-horizontal-snapCol-cubicUpwindCPCFit-divFree/tracer-contour}
\end{document}

### 2015-02-02 msc-publication.tex
put advection contours into their own PDF, include in main PDF

@@
 \begin{itemize}
-	\item Upwind-biased cubic is qualitatively better than Sch\"ar's 4th order leapfrog on BTF grid (figure~\ref{fig:advection})
+	\item Upwind-biased cubic is qualitatively better than Sch\"ar's 4th order leapfrog on BTF grid (figure~\ref{fig:advection-tracer})
 	\item SnapCol error, min and max is almost identical to noOrography (unsurprisingly) (table~\ref{tab:advection})  This agrees with \citet{good2014}
@@
 \begin{figure}
-	btf-cubicUpwindCPCFit-divFree \\
-	\input{../advectTracer-horizontal-btf-cubicUpwindCPCFit-divFree/tracer-contour}
-	\input{../advectTracer-horizontal-btf-cubicUpwindCPCFit-divFree/error-contour}
+	\centering
+	\includegraphics{../msc-publication-fig-advection-tracer/fig-advection-tracer.pdf}
 
-	schaer-4thorder \\
-	\includegraphics[height=1.5in]{schaer-btf-4thorder.png}
-	\includegraphics[height=1.5in]{schaer-btf-4thorder-error.png}
+%	btf-cubicUpwindCPCFit-divFree \\
+%	\input{../advectTracer-horizontal-btf-cubicUpwindCPCFit-divFree/error-contour}
 
-	\TODO{also include SnapCol results}
-	\TODO{check error contours are correct (0.01?)}
+%	schaer-4thorder \\
+%	\includegraphics[height=1.5in]{schaer-btf-4thorder-error.png}
 %
-	\caption{Comparing our BTF cubicUpwindCPCFit result with Schaer's 4th order result}
-	\label{fig:advection}
+	\caption{Horizontally advected tracer contours at $t = \SI{0}{\second}$, \SI{5000}{\second} and \SI{10000}{\second}.  \TODO{identify subfigures}  Contour intervals are every 0.1.}
+	\label{fig:advection-tracer}
+\end{figure}
+
+\begin{figure}
+	\centering
+	\caption{Errors in horizontal tracer advection at $t = \SI{10000}{\second}$.  \TODO{identify subfigures}  Contour intervals are every 0.01 with negative contours denoted by dashed lines.}
+	\label{fig:advection-error}
 \end{figure}
@@
 	\begin{itemize}
-		\item Upwind-biased cubic advection scheme accurately advects a tracer in non-divergent flows (Figure~\ref{fig:advection}, Table~\ref{tab:advection})
+		\item Upwind-biased cubic advection scheme accurately advects a tracer in non-divergent flows (Figure~\ref{fig:advection-tracer}, \ref{fig:advection-error}, Table~\ref{tab:advection})
 		\item spurious vertical velocities are small in resting atmosphere (Figure~\ref{fig:resting})

### 2015-02-02 fig-advection-error.tex
get advection plots aligned for msc publication

--- 2015-02-02 fig-advection-tracer.tex
@@
 \documentclass{article}
-\usepackage[paperwidth=3in,paperheight=4.6in,margin=0in]{geometry}
+\usepackage[paperwidth=3in,paperheight=3.0in,margin=0in]{geometry}
 \usepackage{mathptmx}%{times}
@@
 \begin{document}
-\centering
-\hspace{0.05in}\includegraphics[width=2.75in]{img/schaer-btf-4thorder.png}
+\hspace{0.15in}\includegraphics[width=2.8in,height=1.32in]{img/schaer-btf-4thorder-error.png}
 
 \footnotesize
-\input{../advectTracer-horizontal-btf-cubicUpwindCPCFit-divFree/tracer-contour}
-
-\input{../advectTracer-horizontal-snapCol-cubicUpwindCPCFit-divFree/tracer-contour}
+\input{../advectTracer-horizontal-btf-cubicUpwindCPCFit-divFree/error-contour-x-only}
 \end{document}

### 2015-02-02 fig-advection-tracer.tex
get advection plots aligned for msc publication

@@
 \begin{document}
-\centering
-\hspace{0.05in}\includegraphics[width=2.75in]{img/schaer-btf-4thorder.png}
+\hspace{0.15in}\includegraphics[width=2.75in]{img/schaer-btf-4thorder.png}
 
 \footnotesize
-\input{../advectTracer-horizontal-btf-cubicUpwindCPCFit-divFree/tracer-contour}
+\input{../advectTracer-horizontal-btf-cubicUpwindCPCFit-divFree/tracer-contour-y-only}
 

### 2015-02-04 msc-publication.tex
get advection plots aligned for msc publication
get resting atmosphere working for snapCol, but only by hacking at the controlDict (from binary to ascii)
start writing intro

@@
 \newcommand{\TODO}[1]{\textcolor{purple}{TODO: \emph{#1}}}
-\newcommand{\textcite}[1]{\textbf{#1}}
 
@@
 
-Builds on the work of \citet{schaer2002}, \citet{klemp2011}, \citet{weller-shahrokhi2014} (all MWR)
-
-Results also compared with \citet{melvin2010} (QRJ), maybe \textcite{zaengl2012} (MWR), \citet{good2014} (Atmos Sci Letters)
+\TODO{Builds on the work of \citet{schaer2002}, \citet{klemp2011}, \citet{weller-shahrokhi2014} (all MWR) \\
+Results also compared with \citet{melvin2010} (QRJ), maybe zaengl2012 (MWR), \citet{good2014} (Atmos Sci Letters)}
 
 \section{Introduction}
+Representing orography in numerical weather prediction systems is necessary to model downslope winds and local precipitation patterns.
+Terrain also affects global circulation, giving rise to planetary and gravity waves.
+
+There are two main approaches to represent orography on a grid.
+With terrain following (TF) layers the terrain's influence decays with height: the bottommost layers follow the underlying surface closely while the uppermost layers are flat.
+TF layers are usually implemented on a rectangular computational grid using transformed coordinates.
+
+It is well-known that TF coordinates perform badly in the presence of steep orography: as model resolution increases, steeper gradients lead to larger truncation errors in calculating the horizontal pressure gradientwhich result in spurious winds \citep{dempsey-davis1998} \TODO{or steppeler2002}.
+Much work has been done to reduce error associated with TF coordinates: firstly by smoothing the effects of terrain with height \citep{schaer2002,klemp2011} and, secondly, by improving the accuracy in calculating the horizontal pressure gradient itself \citep{klemp2011} \TODO{and zaengl2012?}.
+
+Despite their associated numerical errors, TF coordinates are attractive because their rectangular structure is simple to process by computer, boundary layer resolution can be increased with variable spacing of vertical layers \citep{schaer2002}, and cell sizes remain almost constant \citep{jebens2011}.
+
+Cut cells is an alternative method in which cells that lie entirely below the terrain are removed, and those that intersect the surface are modified in shape so that they more closely fit the terrain.
+This modification means that some cells become very small, which can reduce computational efficiency \TODO{(klein2009)}, and several approaches have been tried to alleviate the problem \TODO{(steppeler2002 thin wall approx,yamazaki-satomura2010 cell combining, jebens2011 implicit method in cut cells)}.
+
+Several studies have found that cut cells produce more accurate results when compared to TF coordinates.
+Spurious winds seen in TF coordinates are not present and errors do not increase with steeper terrain \citep{good2014}.
+A comparison of TF and cut cells using real initial data by \TODO{steppeler2006} found that precipitation patterns, temperature and wind fields were forecast more accurately in the cut cell model.
+
+\TODO{that's the high-level intro, now onto BTF/SLEVE grids...}  A variety of vertical decay functions have been formulated that make TF layers flatter with height.
+\citet{galchen-somerville1975} proposed a basic terrain following (BTF) coordinate defined as \(z = \left( H - h \right) \left( z^\star / H \right) + h\) where, in two dimensions, \(z(x, z^\star)\) is the height of the coordinate surface at level \(z^\star\), \(H\) is the height of the domain, and \(h(x)\) is the height of the terrain surface.
+
+The smooth level vertical (SLEVE) coordinate proposed by \citet{schaer2002} achieves a more regular TF grid in the middle and top of the domain.
+The terrain height is split into large-scale and small-scale components, \(h_1\) and \(h_2\), such that \(h = h_1 + h_2\), with each component having a different exponential decay.
+The transformation is defined as \(z = z^\star + h_1 b_1 + h_2 b_2\), where the vertical decay functions are given by
+\begin{displaymath}
+	b_i = \frac{\sinh \left( \left( H - z^\star \right) / s_i \right)}{\sinh \left( H / s_i \right)}
+\end{displaymath}
+with \(s_1\) and \(s_2\) are the scale heights of large-scale and small-scale terrain respectively.
+
+\TODO{say what we're going to do in this paper}
+
+\TODO{final para of intro: in section 2, section 3 etc etc}
 
 \begin{itemize}
-	\item A sentence or two about why orography is important to represent in NWP
 	\item Comparing results BTF, SLEVE and cut cell style grid for three standard two-dimensional test cases
@@
 Our results demonstrate that accuracy need not be much reduced and can even be better than cut cells.
-	\item Small cell problem common in cut cell grids (\TODO{citations})
+\TODO{this is too confrontational!}
 \end{itemize}
@@
 	\item \TODO{How much do I need to say about the model?  -- Give a summary: what equation set? finite volume with curl-free pressure grad; lorenz grid; advection scheme}
-	\item \TODO{Include BTF equations from \textcite{galchen-somerville1975} and SLEVE equations from \citet{schaer2002}?}
 	\item \TODO{Explain how SnapCol grid is constructed?}
@@
 
+\begin{figure}
+	\centering
+	\includegraphics{../msc-publication-fig-advection-tracer/fig-advection-tracer.pdf}
+%
+	\caption{Horizontally advected tracer contours at $t = \SI{0}{\second}$, \SI{5000}{\second} and \SI{10000}{\second}.  \TODO{identify subfigures}  Contour intervals are every 0.1.}
+	\label{fig:advection-tracer}
+\end{figure}
+
 Analysis:
@@
 	\centering
-	\includegraphics{../msc-publication-fig-advection-tracer/fig-advection-tracer.pdf}
-
-%	btf-cubicUpwindCPCFit-divFree \\
-%	\input{../advectTracer-horizontal-btf-cubicUpwindCPCFit-divFree/error-contour}
-
-%	schaer-4thorder \\
-%	\includegraphics[height=1.5in]{schaer-btf-4thorder-error.png}
+	\includegraphics{../msc-publication-fig-advection-error/fig-advection-error.pdf}
 	%
-	\caption{Horizontally advected tracer contours at $t = \SI{0}{\second}$, \SI{5000}{\second} and \SI{10000}{\second}.  \TODO{identify subfigures}  Contour intervals are every 0.1.}
-	\label{fig:advection-tracer}
-\end{figure}
-
-\begin{figure}
-	\centering
 	\caption{Errors in horizontal tracer advection at $t = \SI{10000}{\second}$.  \TODO{identify subfigures}  Contour intervals are every 0.01 with negative contours denoted by dashed lines.}
@@
 	\item SnapCol $w$ (\SI{1e-3}{\meter\per\second}) significantly than BTF or SLEVE
-	\item \TODO{Could also compare with \textcite{zaengl2012}, \citet{good2014}?}
+	\item \TODO{Could also compare with zaengl2012, \citet{good2014}?}
 	\item Conclusion: non-orthogonality is a significant cause of numerical error in this test but, again, satisfactory on all grids
@@
 
-\begin{figure}
-	\includegraphics[height=2in]{resting-atmosphere-w.png}
+\begin{figure*}
+	\centering
+	\includegraphics{../msc-publication-fig-resting/fig-resting.pdf}
 %
-	\caption{Comparing our results with \citet{klemp2011}'s}
+	\caption{Maximum spurious vertical velocity $w$ in the resting atmosphere test \TODO{identify subfigures}.  Note that vertical scales differ.}
 	\label{fig:resting}
-\end{figure}
+\end{figure*}
 

### 2015-02-04 msc-publication.tex
reorder intro paras, say what this study is trying to do

@@
 
+\citet{galchen-somerville1975} proposed a basic terrain following (BTF) coordinate defined as 
+\begin{equation}
+	z = \left( H - h \right) \left( z^\star / H \right) + h
+\end{equation}
+where, in two dimensions, \(z(x, z^\star)\) is the height of the coordinate surface at level \(z^\star\), \(H\) is the height of the domain, and \(h(x)\) is the height of the terrain surface.
+Using this coordinate, the terrain's influence decays linearly with height but disappears only at the top of the domain.
+
 It is well-known that TF coordinates perform badly in the presence of steep orography: as model resolution increases, steeper gradients lead to larger truncation errors in calculating the horizontal pressure gradientwhich result in spurious winds \citep{dempsey-davis1998} \TODO{or steppeler2002}.
+
 Much work has been done to reduce error associated with TF coordinates: firstly by smoothing the effects of terrain with height \citep{schaer2002,klemp2011} and, secondly, by improving the accuracy in calculating the horizontal pressure gradient itself \citep{klemp2011} \TODO{and zaengl2012?}.
+
+The smooth level vertical (SLEVE) coordinate proposed by \citet{schaer2002} achieves a more regular TF grid in the middle and top of the domain than the BTF coordinate.
+The terrain height is split into large-scale and small-scale components, \(h_1\) and \(h_2\), such that \(h = h_1 + h_2\), with each component having a different exponential decay.
+The transformation is defined as
+\begin{align}
+	z &= z^\star + h_1 b_1 + h_2 b_2
+\intertext{where the vertical decay functions are given by}
+	b_i &= \frac{\sinh \left( \left( H - z^\star \right) / s_i \right)}{\sinh \left( H / s_i \right)}
+\end{align}
+with \(s_1\) and \(s_2\) are the scale heights of large-scale and small-scale terrain respectively.
 
@@
 
-\TODO{that's the high-level intro, now onto BTF/SLEVE grids...}  A variety of vertical decay functions have been formulated that make TF layers flatter with height.
-\citet{galchen-somerville1975} proposed a basic terrain following (BTF) coordinate defined as \(z = \left( H - h \right) \left( z^\star / H \right) + h\) where, in two dimensions, \(z(x, z^\star)\) is the height of the coordinate surface at level \(z^\star\), \(H\) is the height of the domain, and \(h(x)\) is the height of the terrain surface.
-
-The smooth level vertical (SLEVE) coordinate proposed by \citet{schaer2002} achieves a more regular TF grid in the middle and top of the domain.
-The terrain height is split into large-scale and small-scale components, \(h_1\) and \(h_2\), such that \(h = h_1 + h_2\), with each component having a different exponential decay.
-The transformation is defined as \(z = z^\star + h_1 b_1 + h_2 b_2\), where the vertical decay functions are given by
-\begin{displaymath}
-	b_i = \frac{\sinh \left( \left( H - z^\star \right) / s_i \right)}{\sinh \left( H / s_i \right)}
-\end{displaymath}
-with \(s_1\) and \(s_2\) are the scale heights of large-scale and small-scale terrain respectively.
-
-\TODO{say what we're going to do in this paper}
-
-\TODO{final para of intro: in section 2, section 3 etc etc}
+This study uses the nonhydrostatic model from \citet{weller-shahrokhi2014} to enable a like-for-like comparison between BTF, SLEVE and cut cell grids for idealised, two-dimensional test cases from the literature.
+\TODO{final bit of intro: in section 2, section 3 etc etc}
 
 \begin{itemize}
-	\item Comparing results BTF, SLEVE and cut cell style grid for three standard two-dimensional test cases
 	\item TF grids often shown to have lesser accuracy than cut cells in some test cases, often significantly so.
@@
 \begin{itemize}
-	\item \TODO{How much do I need to say about the model?  -- Give a summary: what equation set? finite volume with curl-free pressure grad; lorenz grid; advection scheme}
+	\item \TODO{Give a summary of model: fully-compressible, finite volume with curl-free pressure grad; lorenz grid; advection scheme}
 	\item \TODO{Explain how SnapCol grid is constructed?}
@@
 \begin{itemize}
-	\item Motivation: challenge accuracy of H operator (\TODO{if we mention H op here we'll have to explain it beforehand})
-	\item Specify domain, thermodynamics, discretisation in time and space
+	\item Motivation: challenge accuracy of H operator (\TODO{if we mention H op here we'll have to explain it beforehand}), curl-free pressure gradient formulation
+	\item Cite weller-shahrokhi2014 for domain, thermodynamics, discretisation in time and space
 \end{itemize}
@@
 	\item $w$ contours visually similar on all grids, agree with \citet{melvin2010} (Figure~\ref{fig:gw-w})
-	\item \TODO{don't know if divergence is relevant -- it's greater on the SnapCol grid (MSc dissertation Figure 4.15e)}
+	\item \TODO{divergence might help explain computational mode -- it's greater on the SnapCol grid (MSc dissertation Figure 4.15e)}
 	\item $\theta$ anomalies similar on all grids EXCEPT...
@@
 reduced stability, although not enough to create vertical motion in this instance 2.
+would disrupt clouds 2.
 model has no viscosity so thermal mixing should not occur in theory}
@@
 	\item \TODO{We can speculate on what excites the computational mode, but perhaps better not to?}
-	\item Little/no evidence of small cell problem -- \TODO{don't know that we can say much here without another vertical momentum test; our hypothesis about the quasi-horizontal flow isn't backed up with a test case}
+	\item No evidence of small cell problem -- \TODO{don't know that we can say much here without another vertical momentum test; our hypothesis about the quasi-horizontal flow isn't backed up with a test case}
 	\item Conclusion: results similar on all grids, agree with literature, except for Lorenz computational mode on SnapCol grid

### 2015-02-05 msc-publication.tex
tweak intro to identify TF and cut cells earlier on

@@
 
-There are two main approaches to represent orography on a grid.
-With terrain following (TF) layers the terrain's influence decays with height: the bottommost layers follow the underlying surface closely while the uppermost layers are flat.
+There are two main approaches to represent orography on a grid: terrain following layers, and cut cells.
+With terrain following (TF) layers the terrain's influence decays with height so that the bottommost layers follow the underlying surface closely while the uppermost layers are flat.
 TF layers are usually implemented on a rectangular computational grid using transformed coordinates.

### 2015-02-09 msc-publication.tex
add w contour plot for gravityWaves-btf
update to use gnuplot 5, start adding exner/theta sample line plot
tweak alignment of figures

@@
 
-\begin{figure}
-	\includegraphics[height=2in]{gw-w-btf.png}
-	\includegraphics[height=2in]{melvin-7a.png}
+\begin{figure*}
+	\centering
+	\includegraphics{../msc-publication-fig-gravityWaves/fig-gravityWaves.pdf}
 %
@@
 	\label{fig:gw-w}
-\end{figure}
+\end{figure*}
 
@@
 \begin{figure}
-	\includegraphics[height=2in]{gw-exner-theta.png}
+	\centering
+	\includegraphics{../msc-publication-fig-gravityWaves-sampleLine/fig-gravityWaves-sampleLine.pdf}
 %
-	\caption{Vertical sample line of Exner function of pressure and theta to demonstrate Lorenz computational mode}
+	\caption{Vertical profiles of the Exner function of pressure, $\Pi$, and potential temperature, $\theta$, in the gravity waves test.
+The computational mode is manifested as a zig-zag in potential temperature on the SnapCol grid.
+Results on the SLEVE grid (not shown) are qualitatively the same as those on the BTF grid.}
 	\label{fig:gw-exner-theta}

### 2015-02-09 fig-gravityWaves-theta.tex
plot thetaDiff for gravityWaves

\documentclass{article}
\usepackage[paperwidth=5.4in,paperheight=4in,margin=0in]{geometry}
\usepackage{mathptmx}%{times}
\usepackage{graphicx}
\usepackage{siunitx}

\setlength{\parindent}{0pt}
\begin{document}
\includegraphics[height=2in,angle=270]{../gravityWaves-btf/18000/thetaDiff.eps}

\includegraphics[height=2in,angle=270]{../gravityWaves-snapCol/18000/thetaDiff.eps}
\end{document}

### 2015-02-10 fig-gravityWaves-theta.tex
add thetaDiff[Zoom] plots

@@
 \documentclass{article}
-\usepackage[paperwidth=5.4in,paperheight=4in,margin=0in]{geometry}
+\usepackage[paperwidth=6.4in,paperheight=4.1in,margin=0in]{geometry}
 \usepackage{mathptmx}%{times}
@@
 \begin{document}
-\includegraphics[height=2in,angle=270]{../gravityWaves-btf/18000/thetaDiff.eps}
+\includegraphics[height=2.2in,angle=270]{../gravityWaves-btf/18000/thetaDiff.eps}
+\hspace{0.1in}
+\includegraphics[height=4in,angle=270]{../gravityWaves-btf/18000/thetaDiffZoom.eps}
+\vspace*{0.1in}
 
-\includegraphics[height=2in,angle=270]{../gravityWaves-snapCol/18000/thetaDiff.eps}
+\includegraphics[height=2.2in,angle=270]{../gravityWaves-snapCol/18000/thetaDiff.eps}
+\hspace{0.1in}
+\includegraphics[height=4in,angle=270]{../gravityWaves-snapCol/18000/thetaDiffZoom.eps}
+\vspace*{0.1in}
+
+\includegraphics[height=2.2in,angle=270]{../gravityWaves-diagnostics/thetaDiffLegend.eps}
+\hspace{0.2in}
+\includegraphics[height=3.9in,angle=270]{../gravityWaves-diagnostics/thetaDiffZoomLegend.eps}
 \end{document}

### 2015-02-10 msc-publication.tex
plot thetaDiff for gravityWaves
add thetaDiff[Zoom] plots

@@
 \section{Introduction}
-Representing orography in numerical weather prediction systems is necessary to model downslope winds and local precipitation patterns.
-Terrain also affects global circulation, giving rise to planetary and gravity waves.
+Representing orography accurately in numerical weather prediction systems is necessary to model downslope winds and local precipitation patterns.
+There are two main approaches to represent orography on a grid: terrain following layers and cut cells.
+Both methods align cells in vertical columns.
 
-There are two main approaches to represent orography on a grid: terrain following layers, and cut cells.
 With terrain following (TF) layers the terrain's influence decays with height so that the bottommost layers follow the underlying surface closely while the uppermost layers are flat.
@@
 
+It is well-known that TF coordinates perform badly in the presence of steep orography: as model resolution increases, steeper gradients lead to larger truncation errors in calculating the horizontal pressure gradientwhich result in spurious winds \citep{dempsey-davis1998,steppeler2002}.
+
+Much work has been done to reduce error associated with TF coordinates: firstly by smoothing the effects of terrain with height \citep{schaer2002,klemp2011} and, secondly, by improving the accuracy in calculating the horizontal pressure gradient itself \citep{klemp2011} \TODO{and zaengl2012?}.
+
+Despite their associated numerical errors, TF coordinates are attractive because their rectangular structure is simple to process by computer, boundary layer resolution can be increased with variable spacing of vertical layers \citep{schaer2002}, and cell sizes remain almost constant \citep{jebens2011}.
+
+Cut cells is an alternative method in which cells that lie entirely below the terrain are removed, and those that intersect the surface are modified in shape so that they more closely fit the terrain.
+This modification means that some cells become very small, which can reduce computational efficiency \citep{klein2009}, and several approaches have been tried to alleviate the problem \citep{steppeler2002,yamazaki-satomura2010,jebens2011}.
+
+Several studies have found that cut cells produce more accurate results when compared to TF coordinates.
+Spurious winds seen in TF coordinates are not present and errors do not increase with steeper terrain \citep{good2014}.
+A comparison of TF and cut cells using real initial data by \citet{steppeler2006} found that precipitation patterns, temperature and wind fields were forecast more accurately in the cut cell model.
+
+Many comparisons of TF and cut cell grids have made between different models.
+On the contrary, this study uses the nonhydrostatic model from \citet{weller-shahrokhi2014} to enable a like-for-like comparison between BTF, SLEVE and cut cell grids for idealised, two-dimensional test cases from the literature.
+\TODO{final bit of intro: in section 2, section 3 etc etc}
+
+\TODO{TF grids often shown to have lesser accuracy than cut cells in some test cases, often significantly so.
+Our results demonstrate that accuracy need not be much reduced and can even be better than cut cells.}
+
+\section{Grids}
 \citet{galchen-somerville1975} proposed a basic terrain following (BTF) coordinate defined as 
@@
 Using this coordinate, the terrain's influence decays linearly with height but disappears only at the top of the domain.
-
-It is well-known that TF coordinates perform badly in the presence of steep orography: as model resolution increases, steeper gradients lead to larger truncation errors in calculating the horizontal pressure gradientwhich result in spurious winds \citep{dempsey-davis1998} \TODO{or steppeler2002}.
-
-Much work has been done to reduce error associated with TF coordinates: firstly by smoothing the effects of terrain with height \citep{schaer2002,klemp2011} and, secondly, by improving the accuracy in calculating the horizontal pressure gradient itself \citep{klemp2011} \TODO{and zaengl2012?}.
 
@@
 
-Despite their associated numerical errors, TF coordinates are attractive because their rectangular structure is simple to process by computer, boundary layer resolution can be increased with variable spacing of vertical layers \citep{schaer2002}, and cell sizes remain almost constant \citep{jebens2011}.
-
-Cut cells is an alternative method in which cells that lie entirely below the terrain are removed, and those that intersect the surface are modified in shape so that they more closely fit the terrain.
-This modification means that some cells become very small, which can reduce computational efficiency \TODO{(klein2009)}, and several approaches have been tried to alleviate the problem \TODO{(steppeler2002 thin wall approx,yamazaki-satomura2010 cell combining, jebens2011 implicit method in cut cells)}.
-
-Several studies have found that cut cells produce more accurate results when compared to TF coordinates.
-Spurious winds seen in TF coordinates are not present and errors do not increase with steeper terrain \citep{good2014}.
-A comparison of TF and cut cells using real initial data by \TODO{steppeler2006} found that precipitation patterns, temperature and wind fields were forecast more accurately in the cut cell model.
-
-This study uses the nonhydrostatic model from \citet{weller-shahrokhi2014} to enable a like-for-like comparison between BTF, SLEVE and cut cell grids for idealised, two-dimensional test cases from the literature.
-\TODO{final bit of intro: in section 2, section 3 etc etc}
-
-\begin{itemize}
-	\item TF grids often shown to have lesser accuracy than cut cells in some test cases, often significantly so.
-Our results demonstrate that accuracy need not be much reduced and can even be better than cut cells.
-\TODO{this is too confrontational!}
-\end{itemize}
+\section{Numerical method}
 
@@
 \end{itemize}
-
-\TODO{should we just compare cubicUpwindCPCFit with Sch\"ar's 4th order?  We could also compare our skewCorrected linear with Sch\"ar's linear but I'm not sure what value it would have...}
 
@@
 \begin{itemize}
-	\item Motivation: challenge accuracy of H operator (\TODO{if we mention H op here we'll have to explain it beforehand}), curl-free pressure gradient formulation
+	\item Motivation: challenge curl-free pressure gradient formulation?
 	\item Cite weller-shahrokhi2014 for domain, thermodynamics, discretisation in time and space
@@
 	\item \TODO{Could also compare with zaengl2012, \citet{good2014}?}
-	\item Conclusion: non-orthogonality is a significant cause of numerical error in this test but, again, satisfactory on all grids
+	\item Conclusion: non-orthogonality is a significant cause of numerical error in this test but, again, \TODO{satisfactory?} on all grids
 \end{itemize}
@@
 
-\begin{figure*}
+\begin{figure}
 	\centering
@@
 %
-	\caption{Gravity wave vertical velocities and thermal anomalies from initial thermal profile}
+	\caption{Vertical cross section of vertical velocity contours in the gravity waves test after 5 hours on (a) the BTF grid compared with (b) the mass-conserving semi-implicit semi-Lagrangian solution from \citet{melvin2010}.
+Contours are every \SI{5e-2}{\meter\per\second} with solid lines denoting ascent and dashed lines descent.}
 	\label{fig:gw-w}
+\end{figure}
+
+\begin{figure*}
+	\centering
+	\includegraphics{../msc-publication-fig-gravityWaves-theta/fig-gravityWaves-theta.pdf}
+%
+	\caption{Anomalies in potential temperature in the gravity waves test after 5 hours.  \TODO{identify subfigures}}
+	\label{fig:gw-theta}
 \end{figure*}
-
-\begin{figure}
-	\includegraphics[height=3in]{gw-theta.png}
-	\includegraphics[height=2in]{gw-theta-zoom.png}
-%
-	\caption{Gravity wave thermal anomalies from initial thermal profile}
-	\label{fig:gw-theta}
-\end{figure}
 
@@
 BTF 	  & TODO & TODO & TODO & \num{-0.058} & \num{1.001} \\
+SLEVE 	  & TODO & TODO & TODO & TODO & TODO \\
 SnapCol   & TODO & TODO & TODO & \multicolumn{1}{c}{---} & \multicolumn{1}{c}{---} \\

### 2015-02-10 fig-advection-error.tex
add subfigure labels, improve captioning

@@
 \usepackage{siunitx}
+\usepackage[absolute,overlay]{textpos}
 
@@
 \begin{document}
-\hspace{0.15in}\includegraphics[width=2.8in,height=1.32in]{img/schaer-btf-4thorder-error.png}
+\begin{textblock}{1}(2.5,0.4)
+\normalsize
+(a)
+\end{textblock}
+\begin{textblock}{1}(2.5,8.5)
+\normalsize
+(b)
+\end{textblock}
+\hspace{0.35in}\includegraphics[width=2.6in,height=1.32in]{img/schaer-btf-4thorder-error.png}
 
 \footnotesize
-\input{../advectTracer-horizontal-btf-cubicUpwindCPCFit-divFree/error-contour-x-only}
+\input{../advectTracer-horizontal-btf-cubicUpwindCPCFit-divFree/error-contour}
 \end{document}

### 2015-02-10 msc-publication.tex
add subfigure labels, improve captioning

@@
 
-It is well-known that TF coordinates perform badly in the presence of steep orography: as model resolution increases, steeper gradients lead to larger truncation errors in calculating the horizontal pressure gradientwhich result in spurious winds \citep{dempsey-davis1998,steppeler2002}.
+It is well-known that TF coordinates perform badly in the presence of steep orography: as model resolution increases, steeper gradients lead to larger truncation errors in calculating the horizontal pressure gradient which result in spurious winds \citep{dempsey-davis1998,steppeler2002}.
 
-Much work has been done to reduce error associated with TF coordinates: firstly by smoothing the effects of terrain with height \citep{schaer2002,klemp2011} and, secondly, by improving the accuracy in calculating the horizontal pressure gradient itself \citep{klemp2011} \TODO{and zaengl2012?}.
+Much work has been done to reduce error associated with TF coordinates: firstly by smoothing the effects of terrain with height \citep{schaer2002,klemp2011} and, secondly, by improving the accuracy in calculating the horizontal pressure gradient itself \citep{klemp2011,zaengl2012}.
 
@@
 
+\TODO{cut cell method}
+
 \section{Numerical method}
+We use a finite volume discretisation of the fully-compressible Euler equations described by \citet{weller-shahrokhi2014}.
+The model has a curl-free pressure gradient formulation, an upwind-biased cubic advection scheme, and uses a Lorenz staggering of thermodynamic variables.
 
@@
 %
-	\caption{Horizontally advected tracer contours at $t = \SI{0}{\second}$, \SI{5000}{\second} and \SI{10000}{\second}.  \TODO{identify subfigures}  Contour intervals are every 0.1.}
+	\caption{Horizontally advected tracer contours at $t = \SI{0}{\second}$, \SI{5000}{\second} and \SI{10000}{\second} using (a) the fourth-order centred difference scheme with sigma coordinates from \citet{schaer2002}, and the upwind-biased cubic scheme on (b) BTF grid, (c) cut cell grid.
+Contour intervals are every 0.1.}
 	\label{fig:advection-tracer}
@@
 	%
-	\caption{Errors in horizontal tracer advection at $t = \SI{10000}{\second}$.  \TODO{identify subfigures}  Contour intervals are every 0.01 with negative contours denoted by dashed lines.}
+	\caption{Errors in horizontal tracer advection at $t = \SI{10000}{\second}$ using (a) the fourth-order centred difference scheme with sigma coordinates from \citet{schaer2002}, (b) the upwind-biased cubic scheme on a BTF grid.
+Contour intervals are every 0.01 with negative contours denoted by dashed lines.}
 	\label{fig:advection-error}
@@
 %
-	\caption{Maximum spurious vertical velocity $w$ in the resting atmosphere test \TODO{identify subfigures}.  Note that vertical scales differ.}
+	\caption{Maximum spurious vertical velocity $w$ in the resting atmosphere test with results on (a) BTF, SLEVE, Hybrid Terrain Following (HTF) and Smoothed Terrain Following (STF) coordinates from \citet{klemp2011}, (b) BTF, SLEVE, cut cell and regular grids using the model from \citet{weller-shahrokhi2014}.
+Note that vertical scales differ.}
 	\label{fig:resting}
@@
 %
-	\caption{Anomalies in potential temperature in the gravity waves test after 5 hours.  \TODO{identify subfigures}}
+	\caption{Anomalies in potential temperature in the gravity waves test after 5 hours.
+The central domain in the lowest \SI{12}{\kilo\meter} is shown on (a) the BTF grid, and (c) the cut cell grid.
+The four lowest layers of each grid are shown for (b) BTF, and (d) cut cell grids, using a narrower potential temperature scale.}
 	\label{fig:gw-theta}
@@
 	\caption{Vertical profiles of the Exner function of pressure, $\Pi$, and potential temperature, $\theta$, in the gravity waves test.
-The computational mode is manifested as a zig-zag in potential temperature on the SnapCol grid.
+The computational mode is manifested as a zig-zag in potential temperature on the cut cell grid.
 Results on the SLEVE grid (not shown) are qualitatively the same as those on the BTF grid.}

### 2015-02-11 msc-publication.tex
start describing advection test, leuenberger SLEVE optimisation

@@
 
-Much work has been done to reduce error associated with TF coordinates: firstly by smoothing the effects of terrain with height \citep{schaer2002,klemp2011} and, secondly, by improving the accuracy in calculating the horizontal pressure gradient itself \citep{klemp2011,zaengl2012}.
+Much work has been done to reduce error associated with TF coordinates: firstly by smoothing the effects of terrain with height \citep{schaer2002,leuenberger-2010,klemp2011} and, secondly, by improving the accuracy in calculating the horizontal pressure gradient itself \citep{klemp2011,zaengl2012}.
 
@@
 \section{Grids}
+\TODO{introduce this section}
 \citet{galchen-somerville1975} proposed a basic terrain following (BTF) coordinate defined as 
@@
 \intertext{where the vertical decay functions are given by}
-	b_i &= \frac{\sinh \left( \left( H - z^\star \right) / s_i \right)}{\sinh \left( H / s_i \right)}
+	b_i &= \frac{\sinh \left( \left( H / s_i \right)^n - \left( z^\star / s_i \right)^n \right)}{\sinh \left( H / s_i \right)^n}
 \end{align}
 with \(s_1\) and \(s_2\) are the scale heights of large-scale and small-scale terrain respectively.
+The exponent \(n\) was introduced by \citet{leuenberger2010} in order to increase cell thickness in the layers nearest the ground, allowing longer timesteps and permitting more accurate calculation of parameterised low-level heat and momentum fluxes.
+\citet{leuenberger2010} found the exponent has an optimal value of \(n = 1.35\).
+Choosing \(n = 1\) gives the decay functions used by \citet{schaer2002}.
 
-\TODO{cut cell method}
+\TODO{Explain how SnapCol grid is constructed?}
 
 \section{Numerical method}
+\TODO{Hilary suggested adding this as a section but, if I'm not sure what to include.  For now, I've put a summary of the model in the opening para of Results}
+
+\section{Results}
+A series of standard, two-dimensional tests were performed over idealised orography.  For each test, results on the BTF, SLEVE and cut cell grid are compared.
 We use a finite volume discretisation of the fully-compressible Euler equations described by \citet{weller-shahrokhi2014}.
@@
 
-\section{Results}
-\begin{itemize}
-	\item \TODO{Give a summary of model: fully-compressible, finite volume with curl-free pressure grad; lorenz grid; advection scheme}
-	\item \TODO{Explain how SnapCol grid is constructed?}
-\end{itemize}
+\subsection{Advection}
 
-\subsection{Advection}
-\begin{itemize}
-	\item Motivation: tests advection scheme accuracy, especially with grid distortions
-	\item Specify domain, terrain, discretisation in time and space, tracer
-	\item Specify wind field, projection method for non-divergence
-	\item Advection equation in flux form
-	\item Upwind-biased cubic advection scheme, non-monotonic, not flux corrected
-	\item d/dt scheme (RK2)
-\end{itemize}
+Following \citet{schaer2002}, a tracer is transported above wave-shaped terrain by solving the advection equation for a prescribed horizontal wind.
+This challenges the accuracy of the advection scheme in the presence of grid distortions.
+The spatial domain, terrain profile, wind field, tracer, SLEVE scale heights are given by \citet{schaer2002}.
+The optimisation of SLEVE by \citet{leuenberger2010} is not used, so the exponent $n = 1$.
+
+\TODO{define the continuous wind profile, say that wind is prescribed at cell faces, and explain how we make it discretely non-divergent}
+
+The flux form of the advection equation, \(\partial \phi / \partial t + \nabla \cdot \left( \bm{u} \phi \right) = 0\), is solved using an upwind-biased cubic advection scheme which is non-monotonic and not flux corrected.
+The time derivative is solved using a second order Runge-Kutta scheme.
+\TODO{I ought to understand this method myself!}
+
+%The tracer is positioned such that it is centred above the mountain at \(t = \SI{5000}{\second}\).
+
+Unlike \citet{schaer2002} who use periodic lateral boundaries, we use a fixed value of 0 at the inlet boundary and zero gradient boundaries elsewhere.
+Tests are integrated forward in time for \SI{10000}{\second} with a timestep of \(\Delta t = \SI{25}{\second}\).
 
@@
 %
-	\caption{Horizontally advected tracer contours at $t = \SI{0}{\second}$, \SI{5000}{\second} and \SI{10000}{\second} using (a) the fourth-order centred difference scheme with sigma coordinates from \citet{schaer2002}, and the upwind-biased cubic scheme on (b) BTF grid, (c) cut cell grid.
+	\caption{Horizontally advected tracer contours at \(t = \SI{0}{\second}\), \SI{5000}{\second} and \SI{10000}{\second} using (a) the fourth-order centred difference scheme with sigma coordinates from \citet{schaer2002}, and the upwind-biased cubic scheme on (b) BTF grid, (c) cut cell grid.
 Contour intervals are every 0.1.}
@@
 	%
-	\caption{Errors in horizontal tracer advection at $t = \SI{10000}{\second}$ using (a) the fourth-order centred difference scheme with sigma coordinates from \citet{schaer2002}, (b) the upwind-biased cubic scheme on a BTF grid.
+	\caption{Errors in horizontal tracer advection at \(t = \SI{10000}{\second}\) using (a) the fourth-order centred difference scheme with sigma coordinates from \citet{schaer2002}, (b) the upwind-biased cubic scheme on a BTF grid.
 Contour intervals are every 0.01 with negative contours denoted by dashed lines.}
@@
 \begin{itemize}
-	\item Spurious $w$ significantly less than those found by \citet{klemp2011} for BTF: \SI{0.35}{\meter\per\second} vs \SI{10}{\meter\per\second}
-	\item SnapCol $w$ (\SI{1e-3}{\meter\per\second}) significantly than BTF or SLEVE
+	\item Spurious \(w\) significantly less than those found by \citet{klemp2011} for BTF: \SI{0.35}{\meter\per\second} vs \SI{10}{\meter\per\second}
+	\item SnapCol \(w\) (\SI{1e-3}{\meter\per\second}) significantly than BTF or SLEVE
 	\item \TODO{Could also compare with zaengl2012, \citet{good2014}?}
@@
 %
-	\caption{Maximum spurious vertical velocity $w$ in the resting atmosphere test with results on (a) BTF, SLEVE, Hybrid Terrain Following (HTF) and Smoothed Terrain Following (STF) coordinates from \citet{klemp2011}, (b) BTF, SLEVE, cut cell and regular grids using the model from \citet{weller-shahrokhi2014}.
+	\caption{Maximum spurious vertical velocity \(w\) in the resting atmosphere test with results on (a) BTF, SLEVE, Hybrid Terrain Following (HTF) and Smoothed Terrain Following (STF) coordinates from \citet{klemp2011}, (b) BTF, SLEVE, cut cell and regular grids using the model from \citet{weller-shahrokhi2014}.
 Note that vertical scales differ.}
@@
 \begin{itemize}
-	\item $w$ contours visually similar on all grids, agree with \citet{melvin2010} (Figure~\ref{fig:gw-w})
+	\item \(w\) contours visually similar on all grids, agree with \citet{melvin2010} (Figure~\ref{fig:gw-w})
 	\item \TODO{divergence might help explain computational mode -- it's greater on the SnapCol grid (MSc dissertation Figure 4.15e)}
-	\item $\theta$ anomalies similar on all grids EXCEPT...
+	\item \(\theta\) anomalies similar on all grids EXCEPT...
 	\item ... on SnapCol grid in lee of mountain near the ground. (Figure~\ref{fig:gw-theta})
@@
 %
-	\caption{Vertical profiles of the Exner function of pressure, $\Pi$, and potential temperature, $\theta$, in the gravity waves test.
+	\caption{Vertical profiles of the Exner function of pressure, \(\Pi\), and potential temperature, \(\theta\), in the gravity waves test.
 The computational mode is manifested as a zig-zag in potential temperature on the cut cell grid.
@@
 & \multicolumn{3}{c}{Cubic upwind-biased} & \multicolumn{2}{c}{Sch\"ar 4th order} \\
-& $\ell^2$ error & min & max & min & max \\
+& \(\ell^2\) error & min & max & min & max \\
 \hline

### 2015-02-11 msc-publication.tex
start introducing advection results; say how sigma is equivalent to BTF

@@
 
+The sigma coordinate transform of \citet{phillips1957} is equivalent to the BTF coordinate transform since they both decay linearly.
+However, since they decay with pressure rather than height, sigma coordinates also change with horizontal variations in pressure.
+
 The smooth level vertical (SLEVE) coordinate proposed by \citet{schaer2002} achieves a more regular TF grid in the middle and top of the domain than the BTF coordinate.
@@
 The optimisation of SLEVE by \citet{leuenberger2010} is not used, so the exponent $n = 1$.
+A tracer is placed above and to the left of the mountain with a magnitude of 1 at its centre and 0 at the perimeter.
 
@@
 Tests are integrated forward in time for \SI{10000}{\second} with a timestep of \(\Delta t = \SI{25}{\second}\).
+
+The test was executed on the BTF, SLEVE and cut cell grids, and on a regular grid with flat terrain.
+Tracer contours at \(t = \SI{0}{\second}, \SI{5000}{\second}\) and \(\SI{10000}{\second}\) are shown in Figure~\ref{fig:advection-tracer}.
+The result from \citet{schaer2002} using a fourth-order centred scheme with sigma coordinates (\ref{fig:advection-tracer}a) is compared with the upwind-biased cubic scheme on (\ref{fig:advection-tracer}b) the BTF grid, and (\ref{fig:advection-tracer}c) cut cell grid.
 

### 2015-02-16 msc-publication.tex
add double height mountain to sample line

@@
 	\caption{Vertical profiles of the Exner function of pressure, \(\Pi\), and potential temperature, \(\theta\), in the gravity waves test.
-The computational mode is manifested as a zig-zag in potential temperature on the cut cell grid.
-Results on the SLEVE grid (not shown) are qualitatively the same as those on the BTF grid.}
+Exner profile is visually identical on all grids for both mountain heights; for clarity, the Exner profile is only plotted for the BTF grid.
+The computational mode is manifested as a zig-zag in potential temperature on the cut cell grid which.
+Doubling the mountain height from \SI{250}{\meter} to \SI{500}{\meter} increases the severity of the computational mode but has negligible effect on the BTF grid.
+Results on the SLEVE grid (not shown) are qualitatively the same as those on the BTF grid for both mountain heights.}
 	\label{fig:gw-exner-theta}

### 2015-02-18 msc-publication.tex
describe projection method for divergence free velocity field

@@
 
-\TODO{define the continuous wind profile, say that wind is prescribed at cell faces, and explain how we make it discretely non-divergent}
+While the continuous wind field specified by \citet{schaer2002} is non-divergent, this is not true of the discrete wind field on non-orthogonal grids.
+A potential, $p$, is used to correct the wind field, $\bm{u}$, so that is non-divergent such that
+\begin{align}
+	\nabla \cdot \left( \bm{u} + \nabla p \right) = 0,
+\intertext{which can be rearranged to give a Poisson equation}
+	\nabla^2 p = - \nabla \cdot \bm{u}.
+\end{align}
 
@@
 
-%The tracer is positioned such that it is centred above the mountain at \(t = \SI{5000}{\second}\).
-
-Unlike \citet{schaer2002} who use periodic lateral boundaries, we use a fixed value of 0 at the inlet boundary and zero gradient boundaries elsewhere.
+Unlike \citet{schaer2002} who use periodic lateral boundaries, a fixed value of 0 is used at the inlet boundary and all other boundaries have zero gradient.
 Tests are integrated forward in time for \SI{10000}{\second} with a timestep of \(\Delta t = \SI{25}{\second}\).
@@
 \end{figure}
+
+\begin{table}[t]
+\caption{Min, max and error norms compared with \citet{schaer2002}}
+\label{tab:advection}
+%
+\centering
+\footnotesize
+\begin{tabular}{ l l l l l l }
+\hline\hline
+& \multicolumn{3}{c}{Cubic upwind-biased} & \multicolumn{2}{c}{Sch\"ar 4th order} \\
+& \(\ell^2\) error & min & max & min & max \\
+\hline
+Analytic  & 0 & 0 & 1 & 0 & 1 \\
+BTF 	  & \num{0.00758} & \num{-0.034} & \num{0.928} & \num{-0.058} & \num{1.001} \\
+SLEVE 	  & \num{0.00110} & \num{-0.011} & \num{0.981} & \num{-0.002} & \num{0.984} \\
+SnapCol   & \num{0.00058} & \num{-0.007} & \num{0.982} & \multicolumn{1}{c}{---} & \multicolumn{1}{c}{---} \\
+noOrography & \num{0.00058} & \num{-0.007} & \num{0.982} & \num{-0.002} & \num{0.984} \\
+\hline
+\end{tabular}
+\end{table}
 
@@
 
-\begin{table}[t]
-\caption{Min, max and error norms compared with \citet{schaer2002}}
-\label{tab:advection}
-%
-\begin{center}
-\begin{tabular}{ l l l l l l }
-\hline\hline
-& \multicolumn{3}{c}{Cubic upwind-biased} & \multicolumn{2}{c}{Sch\"ar 4th order} \\
-& \(\ell^2\) error & min & max & min & max \\
-\hline
-Analytic  & 0 & 0 & 1 & \multicolumn{1}{c}{---} & \multicolumn{1}{c}{---} \\
-BTF 	  & TODO & TODO & TODO & \num{-0.058} & \num{1.001} \\
-SLEVE 	  & TODO & TODO & TODO & TODO & TODO \\
-SnapCol   & TODO & TODO & TODO & \multicolumn{1}{c}{---} & \multicolumn{1}{c}{---} \\
-noOrography & TODO & TODO & TODO & \num{-0.002} & \num{0.984} \\
-\hline
-\end{tabular}
-\end{center}
-\end{table}
-
 % FIGURES

### 2015-02-18 msc-publication.tex
finish off advection test

@@
 
-\section{Numerical method}
-\TODO{Hilary suggested adding this as a section but, if I'm not sure what to include.  For now, I've put a summary of the model in the opening para of Results}
-
 \section{Results}
@@
 Tracer contours at \(t = \SI{0}{\second}, \SI{5000}{\second}\) and \(\SI{10000}{\second}\) are shown in Figure~\ref{fig:advection-tracer}.
-The result from \citet{schaer2002} using a fourth-order centred scheme with sigma coordinates (\ref{fig:advection-tracer}a) is compared with the upwind-biased cubic scheme on (\ref{fig:advection-tracer}b) the BTF grid, and (\ref{fig:advection-tracer}c) cut cell grid.
+The result from \citet{schaer2002} using a fourth-order centred scheme and sigma coordinates (\ref{fig:advection-tracer}a) is compared with the upwind-biased cubic scheme on (\ref{fig:advection-tracer}b) the BTF grid, and (\ref{fig:advection-tracer}c) cut cell grid.
 
-We compare the following grids:
-\begin{itemize}
-	\item No orography (which tells us how good the model can get)
-	\item BTF
-	\item SLEVE (only briefly since it's not adding so much to our argument)
-	\item SnapCol (briefly because it's the same as noOrography)
-\end{itemize}
+Tracer magnitude and shape are well-preserved on all grids, both above the mountain at \(t = \SI{5000}{\second}\) and past the mountain at \(t = \SI{10000}{\second}\).
+Advection is most accurate on the cut cell grid (figure~\ref{fig:advection-tracer}c) and regular grid (not shown).
+As found by \citet{good2014}, the result is the same on both grids.
+This is to be expected since the wind is zero in the region of the ground and both grids are orthogonal elsewhere.
+On the BTF grid, the tracer is less distorted by the cubic upwind-biased scheme (figure~\ref{fig:advection-tracer}b) compared to the fourth-order centred scheme from \citet{schaer2002} (figure~\ref{fig:advection-tracer}a).
+
+Minimum and maximum tracer values and \(\ell^2\) error norms on the BTF, SLEVE, cut cell and regular grids are summarised in table~\ref{tab:advection}.
+The results of the cubic upwind-biased scheme on TF and regular grids are comparable with those for the fourth-order centred scheme from \citet{schaer2002}.
+Error is largest on the BTF grid with \(\ell^2 = \num{0.00758}\) but significant reduced on the SLEVE grid with \(\ell^2 = \num{0.00110}\).
+The error is approximately halved by changing from the SLEVE grid to the cut cell grid.
+
+The results of the tracer advection test show that, given an advection scheme of a sufficiently high order, distortions in the grid do not significantly distort the tracer.
 
@@
 
+\begin{figure}
+	\centering
+	\includegraphics{../msc-publication-fig-advection-error/fig-advection-error.pdf}
+	%
+	\caption{Errors in horizontal tracer advection at \(t = \SI{10000}{\second}\) using (a) the fourth-order centred difference scheme with sigma coordinates from \citet{schaer2002}, (b) the upwind-biased cubic scheme on a BTF grid.
+Contour intervals are every 0.01 with negative contours denoted by dashed lines.}
+	\label{fig:advection-error}
+\end{figure}
+
 \begin{table}[t]
-\caption{Min, max and error norms compared with \citet{schaer2002}}
+	\caption{Minimum and maximum tracer magnitudes and \(\ell^2\) error norms at \(t = \SI{10000}{\second}\) in the tracer advection test.
+Results of the cubic upwind-biased scheme are compared with the fourth-order centred scheme from \citet{schaer2002}.}
 \label{tab:advection}
@@
 SLEVE 	  & \num{0.00110} & \num{-0.011} & \num{0.981} & \num{-0.002} & \num{0.984} \\
-SnapCol   & \num{0.00058} & \num{-0.007} & \num{0.982} & \multicolumn{1}{c}{---} & \multicolumn{1}{c}{---} \\
-noOrography & \num{0.00058} & \num{-0.007} & \num{0.982} & \num{-0.002} & \num{0.984} \\
+Cut cell  & \num{0.00058} & \num{-0.007} & \num{0.982} & \multicolumn{1}{c}{---} & \multicolumn{1}{c}{---} \\
+No orography & \num{0.00058} & \num{-0.007} & \num{0.982} & \num{-0.002} & \num{0.984} \\
 \hline
@@
 \end{table}
-
-Analysis:
-\begin{itemize}
-	\item Upwind-biased cubic is qualitatively better than Sch\"ar's 4th order leapfrog on BTF grid (figure~\ref{fig:advection-tracer})
-	\item SnapCol error, min and max is almost identical to noOrography (unsurprisingly) (table~\ref{tab:advection})  This agrees with \citet{good2014}
-	\item SLEVE error, min and max are part-way between BTF and noOrography (again, unsurprisingly)
-	\item noOrography, BTF and SLEVE min and max comparable to \citet{schaer2002}
-	\item Conclusion: advection more accurate on more regular grids, but nevertheless satisfactory on all grids
-\end{itemize}
-
-\begin{figure}
-	\centering
-	\includegraphics{../msc-publication-fig-advection-error/fig-advection-error.pdf}
-	%
-	\caption{Errors in horizontal tracer advection at \(t = \SI{10000}{\second}\) using (a) the fourth-order centred difference scheme with sigma coordinates from \citet{schaer2002}, (b) the upwind-biased cubic scheme on a BTF grid.
-Contour intervals are every 0.01 with negative contours denoted by dashed lines.}
-	\label{fig:advection-error}
-\end{figure}
 

### 2015-02-24 msc-publication.tex
describe resting atmosphere test

@@
 The model has a curl-free pressure gradient formulation, an upwind-biased cubic advection scheme, and uses a Lorenz staggering of thermodynamic variables.
+No explicit diffusion is used in any of the tests.
 
@@
 \subsection{Resting atmosphere}
+An idealised terrain profile is defined along with a stratified atmosphere in hydrostatic balance.
+The analytic solution is time-invariant, but numerical errors in calculating the horizontal pressure gradient can give rise to spurious velocities which become more severe over steeper terrain \citet{klemp2011}.
+The test setup follows the specification by \citet{klemp2011} but uses the smaller domain and rigid boundaries from \cite{weller-shahrokhi2014}.
 
-\begin{itemize}
-	\item Motivation: challenge curl-free pressure gradient formulation?
-	\item Cite weller-shahrokhi2014 for domain, thermodynamics, discretisation in time and space
-\end{itemize}
+The test was integrated forward by 5 hours on the BTF, SLEVE and cut cell grids, and a regular grid with flat terrain.
+Maximum vertical velocities are compared with the results from \citet{klemp2011} in figure~\ref{fig:resting} (note different vertical scales).
+In agreement with \citet{klemp2011}, vertical velocities are larger on more distorted grids.
+However, magnitudes are smaller comparing results on the terrain following grids with those from \citet{klemp2011}, with $w$ reaching a maximum of \(\sim \SI{0.35}{\meter\per\second}\) on the BTF grid in our test compared with a maximum of \(\sim \SI{10}{\meter\per\second}\) found by \citet{klemp2011}.
 
-Analysis:
-\begin{itemize}
-	\item Spurious \(w\) significantly less than those found by \citet{klemp2011} for BTF: \SI{0.35}{\meter\per\second} vs \SI{10}{\meter\per\second}
-	\item SnapCol \(w\) (\SI{1e-3}{\meter\per\second}) significantly than BTF or SLEVE
-	\item \TODO{Could also compare with zaengl2012, \citet{good2014}?}
-	\item Conclusion: non-orthogonality is a significant cause of numerical error in this test but, again, \TODO{satisfactory?} on all grids
-\end{itemize}
+Unlike the result from \citet{klemp2011}, the SLEVE grid does not significantly reduce vertical velocities compared to the BTF grid.
+However, errors are three orders of magnitude smaller on the cut cell grid with vertical velocities of \(\sim \SI{1e-3}{\meter\per\second}\).
+The smallest error of \(\sim \SI{1e-10}{\meter\per\second}\) is found on the regular grid.
+
+\citet{good2014} found the maximum vertical velocity in their cut cell model was \SI{1e-12}{\meter\per\second}, which is better than any result obtained using the model by \citet{weller-shahrokhi2014}.
+It is worth noting that \citet{good2014} used a timestep of \SI{1.01}{\second} instead of the \SI{100}{\second} timestep specified by \citet{klemp2011} that was used to obtain the results presented here.
+In addition, in the model used by \citet{good2014}, cell centres are in the centre of the uncut cell, resulting in the centre of some cut cells being below the ground (S.-J.
+Lock 2014, personal communication).
+This means that the grid is effectively regular when calculating horizontal and vertical gradients.
+These two modelling decisions may account for the very small velocities found by \citet{good2014}.
+
+In summary, spurious velocities in the resting atmosphere test were similar on both terrain following grids, with much lower errors compared to those from \citet{klemp2011}.
+The maximum vertical velocity was significantly decreased on the cut cell grid, so we conclude that non-orthogonality is a significant cause of numerical error in this test.
 
@@
 %
-	\caption{Maximum spurious vertical velocity \(w\) in the resting atmosphere test with results on (a) BTF, SLEVE, Hybrid Terrain Following (HTF) and Smoothed Terrain Following (STF) coordinates from \citet{klemp2011}, (b) BTF, SLEVE, cut cell and regular grids using the model from \citet{weller-shahrokhi2014}.
+	\caption{Maximum spurious vertical velocity, \(w\) (\si{\meter\per\second}), in the resting atmosphere test with results on (a) BTF, SLEVE, Hybrid Terrain Following (HTF) and Smoothed Terrain Following (STF) coordinates from \citet{klemp2011}, (b) BTF, SLEVE, cut cell and regular grids using the model from \citet{weller-shahrokhi2014}.
 Note that vertical scales differ.}

### 2015-03-02 msc-publication.tex
remove 250m from gravity wave sampleLine plot, include test case setups and cut cell/snappyHexMesh description in paper

@@
 
-Much work has been done to reduce error associated with TF coordinates: firstly by smoothing the effects of terrain with height \citep{schaer2002,leuenberger-2010,klemp2011} and, secondly, by improving the accuracy in calculating the horizontal pressure gradient itself \citep{klemp2011,zaengl2012}.
+Much work has been done to reduce error associated with TF coordinates: firstly by smoothing the effects of terrain with height \citep{schaer2002,leuenberger2010,klemp2011} and, secondly, by improving the accuracy in calculating the horizontal pressure gradient itself \citep{klemp2011,zaengl2012}.
 
@@
 \section{Grids}
+\label{sec:grid}
+
+\begin{figure*}
+	\centering
+	TODO
+	%
+	\caption{Examples of (a) BTF, (b) SLEVE, and (c) a cut cell grid, all constructed with OpenFOAM.
+The two dimensional grids are \SI{20}{\kilo\meter} wide and \SI{2}{\kilo\meter} high.
+The cut cell grid was created by intersecting the terrain surface with a regular grid as described in section~\ref{sec:grid}.}
+	\label{fig:grid}
+\end{figure*}
+
 \TODO{introduce this section}
@@
 
-\TODO{Explain how SnapCol grid is constructed?}
+Most implementations of terrain following layers use a coordinate system that makes the domain rectangular, but introduces metric terms into the equations of motion.
+Instead, the model employed in this study uses Cartesian coordinates and unstructured grids.
+By doing so, results from the same model can be compared between terrain following and cut cell grids without modifying the equation set or discretisation.
+
+The OpenFOAM utility \texttt{snappyHexMesh} was used to create a grid that approximates the cut cell method.
+First, a custom utility was used to move points beneath the surface up to the surface creating small cells near mountain peaks.
+Second, the surface faces were taken from the BTF grid and \texttt{snappyHexMesh} was used to intersect the surface with the grid.
+The tool removes cells whose centres are below the surface.
+An example of the resulting grid is shown in figure~\ref{fig:grid}c.
+
+There are two details of grid construction which mean that resulting cut cell grids can differ slightly from a typical cut cell grid created using a shaving method, as described by \citet{adcroft1997}.
+First, when \texttt{snappyHexMesh} moves points along the surface according to its heuristics, some points are moved horizontally.
+Second, the utility does not create new points necessary for pentagonal cells.
 
@@
 This challenges the accuracy of the advection scheme in the presence of grid distortions.
-The spatial domain, terrain profile, wind field, tracer, SLEVE scale heights are given by \citet{schaer2002}.
+
+The domain is \SI{300}{\kilo\meter} wide and \SI{25}{\kilo\meter} high which is discretized onto a grid with \(\Delta x = \SI{1}{\kilo\meter}\) and \(\Delta z^\star = \SI{500}{\meter}\).
+The terrain is wave-shaped, specified by the surface height, \(h\), such that
+\begin{subequations}
+\begin{align}
+	h(x) &= \cos^2 \left( \frac{\pi x}{\lambda} \right) h^\star
+%
+	\intertext{where}
+%
+	h^\star(x) &= \left\{ \begin{array}{l l}
+		h_0 \cos^2 \left( \frac{\pi x}{2a} \right) & \quad \text{if $| x | < a$} \\
+		0 & \quad \text{otherwise}
+	\end{array} \right.
+\end{align}
+\end{subequations}
+where $a = \SI{25}{\kilo\meter}$ is the mountain half-width, $h_0 = \SI{3}{\kilo\meter}$ is the maximum mountain height, and $\lambda = \SI{8}{\kilo\meter}$ is the wavelength.
+On the SLEVE grid, the large-scale component $h_1$ is given by \(h_1(x) = h^\star(x) / 2\)
+and $s_1 = \SI{15}{\kilo\meter}$ is the large scale height, and $s_2 = \SI{2.5}{\kilo\meter}$ is the small scale height.
 The optimisation of SLEVE by \citet{leuenberger2010} is not used, so the exponent $n = 1$.
-A tracer is placed above and to the left of the mountain with a magnitude of 1 at its centre and 0 at the perimeter.
 
-While the continuous wind field specified by \citet{schaer2002} is non-divergent, this is not true of the discrete wind field on non-orthogonal grids.
+The wind is entirely horizontal and is prescribed as
+\begin{align}
+	u(z) = u_0 \left\{ \begin{array}{l l}
+		1 & \quad \text{if $z \geq z_2$} \\
+		\sin^2 \left( \frac{\pi}{2} \frac{z - z_1}{z_2 - z_1} \right) & \quad \text{if $z_1 < z < z_2$} \\
+		0 & \quad \text{otherwise}
+	\end{array} \right.	
+\end{align}
+where $u_0 = \SI{10}{\meter\per\second}$, $z_1 = \SI{4}{\kilo\meter}$ and $z_2 = \SI{5}{\kilo\meter}$.
+This results in a constant wind aloft, and zero flow at \SI{4}{\kilo\meter} and below.
+
+While the continuous wind field is non-divergent, this is not true of the discrete wind field on non-orthogonal grids.
 A potential, $p$, is used to correct the wind field, $\bm{u}$, so that is non-divergent such that
@@
 \end{align}
+
+A tracer $\varphi$ is positioned upstream above the height of the terrain.  It has the shape
+\begin{align}
+	\varphi(x, z) &= \varphi_0 \left\{ \begin{array}{l l}
+		\cos^2 \left( \frac{\pi r}{2} \right) & \quad \text{if $r \leq 1$} \\
+		0 & \quad \text{otherwise}
+	\end{array} \right.
+%
+\intertext{having radius $r$ given by}
+%
+	r &= \sqrt{
+		\left( \frac{x - x_0}{A_x} \right)^2 + 
+		\left( \frac{z - z_0}{A_z} \right)^2
+	}
+\end{align}
+where $A_x = \SI{25}{\kilo\meter}$, $A_z = \SI{3}{\kilo\meter}$ are the horizontal and vertical half-widths respectively, and $\varphi_0 = 1$ is the maximum magnitude of the anomaly.
+At $t = \SI{0}{\second}$, the anomaly is centred at $(x_0, z_0) = (\SI{-50}{\kilo\meter}, \SI{9}{\kilo\meter})$ so that the anomaly is upwind of the mountain and well above the maximum terrain height of \SI{3}{\kilo\meter}.
+Analytic solutions can be found by adjusting the anomaly centre such that $x_0 = ut$.
 
@@
 	\caption{Horizontally advected tracer contours at \(t = \SI{0}{\second}\), \SI{5000}{\second} and \SI{10000}{\second} using (a) the fourth-order centred difference scheme with sigma coordinates from \citet{schaer2002}, and the upwind-biased cubic scheme on (b) BTF grid, (c) cut cell grid.
-Contour intervals are every 0.1.}
+Contour intervals are every 0.1.
+The terrain profile is also shown immediately above the ground.}
 	\label{fig:advection-tracer}
@@
 Analytic  & 0 & 0 & 1 & 0 & 1 \\
-BTF 	  & \num{0.00758} & \num{-0.034} & \num{0.928} & \num{-0.058} & \num{1.001} \\
-SLEVE 	  & \num{0.00110} & \num{-0.011} & \num{0.981} & \num{-0.002} & \num{0.984} \\
-Cut cell  & \num{0.00058} & \num{-0.007} & \num{0.982} & \multicolumn{1}{c}{---} & \multicolumn{1}{c}{---} \\
-No orography & \num{0.00058} & \num{-0.007} & \num{0.982} & \num{-0.002} & \num{0.984} \\
+BTF 	  & \num{-1} & \num{-1} & \num{-1} & \num{-0.058} & \num{1.001} \\
+SLEVE 	  & \num{-1} & \num{-1} & \num{-1} & \num{-0.002} & \num{0.984} \\
+Cut cell  & \num{-1} & \num{-1} & \num{-1} & \multicolumn{1}{c}{---} & \multicolumn{1}{c}{---} \\
+No orography & \num{-1} & \num{-1} & \num{-1} & \num{-0.002} & \num{0.984} \\
 \hline
@@
 \subsection{Resting atmosphere}
-An idealised terrain profile is defined along with a stratified atmosphere in hydrostatic balance.
+An idealised terrain profile is defined along with a stratified atmosphere at rest in hydrostatic balance.
 The analytic solution is time-invariant, but numerical errors in calculating the horizontal pressure gradient can give rise to spurious velocities which become more severe over steeper terrain \citet{klemp2011}.
-The test setup follows the specification by \citet{klemp2011} but uses the smaller domain and rigid boundaries from \cite{weller-shahrokhi2014}.
+
+The test setup follows the specification by \cite{klemp2011}, but has a narrower domain that is \SI{20}{\kilo\meter} wide and \SI{20}{\kilo\meter} high in order to reduce simulation time, as used by \citet{weller-shahrokhi2014}.
+The grid resolution is \(\Delta x = \Delta z^\star = \SI{500}{\meter}\) as originally specified by \citet{klemp2011}.
+All boundary conditions are no normal flow.
+
+The wave-shaped mountain profile has a surface height, $h$, given by
+\begin{align}
+	h(x) = h_0 \exp \left( - \left( \frac{x}{a} \right)^2 \right) \cos^2 \left( \frac{\pi x}{\lambda} \right) \label{eqn:resting:mountain}
+\end{align}
+where $a = \SI{5}{\kilo\meter}$ is the mountain half-width, $h_0 = \SI{1}{\kilo\meter}$ is the maximum mountain height and $\lambda = \SI{4}{\kilo\meter}$ is the wavelength.
+For the optimised SLEVE grid, the large-scale component $h_1$ is specified as
+\begin{align}
+h_1(x) = \frac{1}{2} h_0 \exp \left( - \left( \frac{x}{a} \right)^2 \right)
+\end{align}
+and, following \cite{leuenberger2010}, $s_1 = \SI{4}{\kilo\meter}$ is the large scale height, $s_2 = \SI{1}{\kilo\meter}$ is the small scale height, and the optimal exponent value of $n = 1.35$ is used.
+
+The initial thermodynamic conditions are in discrete hydrostatic balance, having a surface temperature of $\theta_0 = \SI{288}{\kelvin}$ and constant stability with Brunt-V\"ais\"al\"a frequency $N = \SI{0.01}{\per\second}$ everywhere, except for a more stable layer of $N = \SI{0.02}{\per\second}$ between $\SI{2}{\kilo\meter} \leq z \leq \SI{3}{\kilo\meter}$.
+Unlike \citet{klemp2011}, there is no eddy diffusion in the equation set.
 
@@
 \subsection{Gravity waves}
-\begin{itemize}
-	\item \TODO{Motivation: it models a real dynamic process?}
-	\item Specify domain, thermodynamics, prescribed inlet wind
-	\item Specify sponge layers, BCs
-	\item Compare BTF, SLEVE and SnapCol results (SLEVE only briefly as visually identical to BTF)
-\end{itemize}
+\TODO{introduce and motivate the test... it models a real dynamic process}
+
+Following \citet{melvin2010}, the domain is \SI{300}{\kilo\meter} wide and \SI{30}{\kilo\meter} high.
+The mountain profile has the same form as equation~\ref{eqn:resting:mountain}.
+Tests are performed with mountain heights of $h_0 = \SI{250}{\meter}$ and \(h_0 = \SI{500}{\meter}\).
+As in the resting atmosphere test, $a = \SI{5}{\kilo\meter}$ is the mountain half-width and $\lambda = \SI{4}{\kilo\meter}$ is the wavelength.
+On the optimised SLEVE grid, $s_1 = \SI{5}{\kilo\meter}$ is the large scale height, $s_2 = \SI{2}{\kilo\meter}$ is the small scale height and the optimal exponent value $n = 1.35$ is used.
+
+The initial thermodynamic conditions have a surface temperature of $\theta_0 = \SI{288}{\kelvin}$ and constant stability with $N = \SI{0.01}{\per\second}$ everywhere.
+A constant horizontal wind $u = \SI{10}{\meter\per\second}$ is prescribed at the inlet boundary.
+
+Sponge layers are added to the upper \SI{10}{\kilo\meter} and leftmost \SI{10}{\kilo\meter} at the inlet boundary to damp the reflection of waves.
+The term $\mu \rho \bf{u}$ is subtracted from the momentum equation \TODO{include the equation set in this paper, or cite weller-shahrokhi2014 here?} where the damping function $\mu$ is adapted from \citet{melvin2010} such that
+\begin{align}
+	\mu(x, z) &= \mu_\mathrm{upper} + \mu_\mathrm{inlet} \\
+	\mu_\mathrm{upper}(z) &= \begin{cases}
+		\overline{\mu} \sin^2 \left( \frac{\pi}{2} \frac{z - z_B}{H - z_B} \right) & \text{if } z \geq z_B \\
+		0 & \text{otherwise} \\
+	\end{cases} \\
+	\mu_\mathrm{inlet}(x) &= \begin{cases}
+		\overline{\mu} \sin^2 \left( \frac{\pi}{2} \frac{x_I - x}{x_I - x_0} \right) & \text{if } x < x_I \\
+		0 & \text{otherwise}
+	\end{cases}
+\end{align}
+where $\overline{\mu} = 1.2$ is the damping coefficient, $z_B = \SI{20}{\kilo\meter}$ is the bottom of the sponge layer, $H = \SI{30}{\kilo\meter}$ is the top of the domain, $x_0 = \SI{-150}{\kilo\meter}$ is the leftmost limit of the domain and $x_I = \SI{-140}{\kilo\meter}$ is the rightmost extent of the inlet sponge layer.
+The sponge layer is only active on faces whose normal is vertical so that it damps vertical momentum only.
+
+Note that, while the domain itself is \SI{30}{\kilo\meter} in height, for the purposes of generating of BTF and SLEVE grids, the domain height is set to \SI{20}{\kilo\meter} because the sponge layer occupies the uppermost \SI{10}{\kilo\meter}.
+
+No normal flow is imposed at the top and bottom boundaries and the outlet is zero gradient.
+For Exner, hydrostatic balance is prescribed on all boundaries.
+The simulation is integrated forward by 5 hours with a timestep $\Delta t = \SI{8}{\second}$.
 
 Analysis:
+\TODO{We compare BTF, SLEVE and SnapCol results (SLEVE only briefly as visually identical to BTF)}
 \begin{itemize}
@@
 %
-	\caption{Vertical profiles of the Exner function of pressure, \(\Pi\), and potential temperature, \(\theta\), in the gravity waves test.
+	\caption{Vertical profiles of the Exner function of pressure, \(\Pi\), and potential temperature, \(\theta\), in the gravity waves test with a mountain height of \(h_0 = \SI{500}{\meter}\).
 Exner profile is visually identical on all grids for both mountain heights; for clarity, the Exner profile is only plotted for the BTF grid.
 The computational mode is manifested as a zig-zag in potential temperature on the cut cell grid which.
-Doubling the mountain height from \SI{250}{\meter} to \SI{500}{\meter} increases the severity of the computational mode but has negligible effect on the BTF grid.
-Results on the SLEVE grid (not shown) are qualitatively the same as those on the BTF grid for both mountain heights.}
+Results on the SLEVE grid (not shown) are qualitatively the same as those on the BTF grid.}
 	\label{fig:gw-exner-theta}

### 2015-03-04 msc-publication.tex
improve introduction, say more about the horizontal pressure gradient errors and how TF and cut cells proposed to reduce this error

@@
 Both methods align cells in vertical columns.
+Existing studies of cut cells have compared results with tests on terrain following grids from previous studies that used different models.
+On the contrary, this study uses a single model to enable a like-for-like comparison between terrain following and cut cell grids.
+
+With increasing model resolution, terrain can become steeper, making accurate calculation of the horizontal pressure gradient more difficult \citep{steppeler2002}.
+Numerical errors in this calculation result in spurious winds and can cause numerical instability \citep{fast2003,webster2003}.
+Terrain following and cut cell methods seek to reduce the error that is associated with steep orography.
 
 With terrain following (TF) layers the terrain's influence decays with height so that the bottommost layers follow the underlying surface closely while the uppermost layers are flat.
-TF layers are usually implemented on a rectangular computational grid using transformed coordinates.
+There are two main approaches to minimizing errors associated with TF layers.
+First, by smoothing the effects of terrain with height, the influence of the terrain is reduced, hence errors in the calculated horizontal pressure gradient are also reduced aloft \citep{schaer2002,leuenberger2010,klemp2011}.
+However, the error is not reduced at the ground where steep terrain remains unmodified.
 
-It is well-known that TF coordinates perform badly in the presence of steep orography: as model resolution increases, steeper gradients lead to larger truncation errors in calculating the horizontal pressure gradient which result in spurious winds \citep{dempsey-davis1998,steppeler2002}.
+Second, numerical errors can also be reduced by improving the accuracy in calculating the horizontal pressure gradient itself.
+TF layers are usually implemented using a coordinate transformation onto a rectangular computational domain, which introduces metric terms into the equations of motion.
+The techniques proposed by \citet{klemp2011} and \cite{zaengl2012} both involve the calculation of the horizontal pressure gradient in the physical domain.
+They found that, by avoiding the calculation of metric terms associated with the computational domain, numerical errors were reduced.
 
-Much work has been done to reduce error associated with TF coordinates: firstly by smoothing the effects of terrain with height \citep{schaer2002,leuenberger2010,klemp2011} and, secondly, by improving the accuracy in calculating the horizontal pressure gradient itself \citep{klemp2011,zaengl2012}.
-
-Despite their associated numerical errors, TF coordinates are attractive because their rectangular structure is simple to process by computer, boundary layer resolution can be increased with variable spacing of vertical layers \citep{schaer2002}, and cell sizes remain almost constant \citep{jebens2011}.
+Despite their associated numerical errors, TF layers are in widespread operational use \citep{steppeler2003}.
+They are attractive because their rectangular structure is simple to process by computer, boundary layer resolution can be increased with variable spacing of vertical layers \citep{schaer2002}, and cell sizes remain almost constant \citep{jebens2011}.
 
 Cut cells is an alternative method in which cells that lie entirely below the terrain are removed, and those that intersect the surface are modified in shape so that they more closely fit the terrain.
-This modification means that some cells become very small, which can reduce computational efficiency \citep{klein2009}, and several approaches have been tried to alleviate the problem \citep{steppeler2002,yamazaki-satomura2010,jebens2011}.
+The resulting grid is orthogonal everywhere except where cells have been cut.
+Hence, errors are still introduced when calculating the horizontal pressure gradient between cut and uncut cells.
+
+The cut cell method can create some very small cells which reduce computational efficiency \citep{klein2009}, and several approaches have been tried to alleviate the problem \citep{steppeler2002,yamazaki-satomura2010,jebens2011}.
 
@@
 Spurious winds seen in TF coordinates are not present and errors do not increase with steeper terrain \citep{good2014}.
-A comparison of TF and cut cells using real initial data by \citet{steppeler2006} found that precipitation patterns, temperature and wind fields were forecast more accurately in the cut cell model.
+\TODO{anything else?}
 
-Many comparisons of TF and cut cell grids have made between different models.
-On the contrary, this study uses the nonhydrostatic model from \citet{weller-shahrokhi2014} to enable a like-for-like comparison between BTF, SLEVE and cut cell grids for idealised, two-dimensional test cases from the literature.
+This study uses the nonhydrostatic model from \citet{weller-shahrokhi2014} to enable a like-for-like comparison between BTF, SLEVE and cut cell grids for idealised, two-dimensional test cases from the literature.
 \TODO{final bit of intro: in section 2, section 3 etc etc}
@@
 As found by \citet{good2014}, the result is the same on both grids.
-This is to be expected since the wind is zero in the region of the ground and both grids are orthogonal elsewhere.
+This is to be expected since the wind is zero in the region of the ground and flow aloft is aligned with the grids.
 On the BTF grid, the tracer is less distorted by the cubic upwind-biased scheme (figure~\ref{fig:advection-tracer}b) compared to the fourth-order centred scheme from \citet{schaer2002} (figure~\ref{fig:advection-tracer}a).

### 2015-03-05 fig-gravityWaves-theta.tex
add subfigure labels, improve captioning
switch now that plotPatchData spits out PDFs for rawValue plots

@@
 \usepackage{siunitx}
+\usepackage[absolute,overlay]{textpos}
+\usepackage{color}
 
@@
 \begin{document}
-\includegraphics[height=2.2in,angle=270]{../gravityWaves-btf/18000/thetaDiff.eps}
+\TPMargin{3pt}
+\textblockcolour{white}
+\begin{textblock}{0.75}(1.2,1.2)
+\normalsize
+\centering
+(a)
+\end{textblock}
+\begin{textblock}{0.75}(7,1.2)
+\normalsize
+\centering
+(b)
+\end{textblock}
+\begin{textblock}{0.75}(1.2,8)
+\normalsize
+\centering
+(c)
+\end{textblock}
+\begin{textblock}{0.75}(7,8)
+\normalsize
+\centering
+(d)
+\end{textblock}
+\includegraphics[width=2.2in]{../gravityWaves-btf-250m/18000/thetaDiff.pdf}
 \hspace{0.1in}
-\includegraphics[height=4in,angle=270]{../gravityWaves-btf/18000/thetaDiffZoom.eps}
+\includegraphics[width=4in]{../gravityWaves-btf-250m/18000/thetaDiffZoom.pdf}
 \vspace*{0.1in}
 
-\includegraphics[height=2.2in,angle=270]{../gravityWaves-snapCol/18000/thetaDiff.eps}
+\includegraphics[width=2.2in]{../gravityWaves-snapCol-250m/18000/thetaDiff.pdf}
 \hspace{0.1in}
-\includegraphics[height=4in,angle=270]{../gravityWaves-snapCol/18000/thetaDiffZoom.eps}
+\includegraphics[width=4in]{../gravityWaves-snapCol-250m/18000/thetaDiffZoom.pdf}
 \vspace*{0.1in}

### 2015-03-09 msc-publication.tex
start discussing gravity waves results, thermal anomaly on the cut cell grid

@@
 \maketitle
-
-\TODO{Builds on the work of \citet{schaer2002}, \citet{klemp2011}, \citet{weller-shahrokhi2014} (all MWR) \\
-Results also compared with \citet{melvin2010} (QRJ), maybe zaengl2012 (MWR), \citet{good2014} (Atmos Sci Letters)}
 
@@
 \TODO{final bit of intro: in section 2, section 3 etc etc}
-
-\TODO{TF grids often shown to have lesser accuracy than cut cells in some test cases, often significantly so.
-Our results demonstrate that accuracy need not be much reduced and can even be better than cut cells.}
 
@@
 
+Test results are compared between the BTF, SLEVE and cut cell grids.
+Vertical velocities on the BTF grid are shown in figure~\ref{fig:gw-w}a, which have few visible differences from results on the SLEVE and cut cell grids (not shown).
+Vertical velocities on all grids are in agreement with the high resolution solution from \citet{melvin2010}, shown in figure~\ref{fig:gw-w}b.
+
+The initial thermal profile is subtracted from the potential temperature field at the end of the integration to reveal the structure of thermal anomalies.
+Once again, the results are similar on all three grids, and results are shown on the BTF and cut cell grids in figures~\ref{fig:gw-theta}a and \ref{fig:gw-theta}c respectively.
+However, examining more closely the anomalies in the lee of the mountain, figure~\ref{fig:gw-theta}d shows that the bottommost layer is anomalously warm and the layer above it is anomalously cold.
+This feature is not present on the BTF grid (figure~\ref{fig:gw-theta}b) or the SLEVE grid (not shown).
+
+In a further test, the mountain height is doubled from \SI{250}{\meter} to \SI{500}{\meter} with all other parameter values unchanged.
+The same spurious anomaly in potential temperature is present on the cut cell grid but its amplitude increases.
+Figure~\ref{fig:gw-exner-theta} shows vertical profiles of the Exner function of pressure and potential temperature in the lowest \SI{1}{\kilo\meter} in the lee of the mountain at \(x = \SI{50}{\kilo\meter}\).
+
 Analysis:
-\TODO{We compare BTF, SLEVE and SnapCol results (SLEVE only briefly as visually identical to BTF)}
 \begin{itemize}
-	\item \(w\) contours visually similar on all grids, agree with \citet{melvin2010} (Figure~\ref{fig:gw-w})
 	\item \TODO{divergence might help explain computational mode -- it's greater on the SnapCol grid (MSc dissertation Figure 4.15e)}
-	\item \(\theta\) anomalies similar on all grids EXCEPT...
-	\item ... on SnapCol grid in lee of mountain near the ground. (Figure~\ref{fig:gw-theta})
 	\item \TODO{worth mentioning implications of this?
@@
 %
-	\caption{Anomalies in potential temperature in the gravity waves test after 5 hours.
+	\caption{Anomalies in potential temperature in the gravity waves test after 5 hours with a mountain height, \(h_0 = \SI{250}{\meter}\).
 The central domain in the lowest \SI{12}{\kilo\meter} is shown on (a) the BTF grid, and (c) the cut cell grid.
-The four lowest layers of each grid are shown for (b) BTF, and (d) cut cell grids, using a narrower potential temperature scale.}
+The four lowest layers of each grid are shown for (b) BTF, and (d) cut cell grids, using a narrower potential temperature scale.
+The results on the SLEVE grid (not shown) are qualitatively identical to results on the BTF grid.}
 	\label{fig:gw-theta}
@@
 The computational mode is manifested as a zig-zag in potential temperature on the cut cell grid which.
-Results on the SLEVE grid (not shown) are qualitatively the same as those on the BTF grid.}
+Results on the SLEVE grid (not shown) are qualitatively the same as those on the BTF grid.
+The thermal profile with a lesser mountain height of \(h_0 = \SI{250}{\meter}\) (not shown) exhibits a computational mode with smaller amplitude.}
 	\label{fig:gw-exner-theta}

### 2015-03-09 msc-publication.tex
introduce gravity waves test a bit

@@
 \subsection{Gravity waves}
-\TODO{introduce and motivate the test... it models a real dynamic process}
+The test originally specified by \citet{schaer2002} prescribes flow over terrain with small-scale and large-scale undulations which induces propagating and evanescent gravity waves.
 
@@
 In a further test, the mountain height is doubled from \SI{250}{\meter} to \SI{500}{\meter} with all other parameter values unchanged.
-The same spurious anomaly in potential temperature is present on the cut cell grid but its amplitude increases.
+The same spurious anomaly in potential temperature is again present on the cut cell grid but its amplitude increases.
 Figure~\ref{fig:gw-exner-theta} shows vertical profiles of the Exner function of pressure and potential temperature in the lowest \SI{1}{\kilo\meter} in the lee of the mountain at \(x = \SI{50}{\kilo\meter}\).

### 2015-03-11 msc-publication.tex
add test setup and motivate wobblyTracerAdvection

@@
 
-With increasing model resolution, terrain can become steeper, making accurate calculation of the horizontal pressure gradient more difficult \citep{steppeler2002}.
+With increasing model resolution, terrain can become steeper, making accurate calculation of the horizontal pressure gradient more difficult \citep{gary1973,steppeler2002}.
 Numerical errors in this calculation result in spurious winds and can cause numerical instability \citep{fast2003,webster2003}.
@@
 \begin{equation}
-	z = \left( H - h \right) \left( z^\star / H \right) + h
+	z = \left( H - h \right) \left( z^\star / H \right) + h \label{eqn:btf}
 \end{equation}
@@
 
-\subsection{Advection}
+\subsection{Horizontal advection}
 
@@
 \begin{align}
-	h(x) &= \cos^2 \left( \frac{\pi x}{\lambda} \right) h^\star
+	h(x) &= h^\star \cos^2 ( \alpha x )
 %
@@
 	h^\star(x) &= \left\{ \begin{array}{l l}
-		h_0 \cos^2 \left( \frac{\pi x}{2a} \right) & \quad \text{if $| x | < a$} \\
+	h_0 \cos^2 ( \beta x ) & \quad \text{if $| x | < a$} \\
 		0 & \quad \text{otherwise}
@@
 \end{subequations}
-where $a = \SI{25}{\kilo\meter}$ is the mountain half-width, $h_0 = \SI{3}{\kilo\meter}$ is the maximum mountain height, and $\lambda = \SI{8}{\kilo\meter}$ is the wavelength.
+where $a = \SI{25}{\kilo\meter}$ is the mountain half-width, $h_0 = \SI{3}{\kilo\meter}$ is the maximum mountain height, $\lambda = \SI{8}{\kilo\meter}$ is the wavelength, \(\alpha = \pi / \lambda\) and \(\beta = \pi / 2a\).
 On the SLEVE grid, the large-scale component $h_1$ is given by \(h_1(x) = h^\star(x) / 2\)
@@
 \begin{table}[t]
-	\caption{Minimum and maximum tracer magnitudes and \(\ell^2\) error norms at \(t = \SI{10000}{\second}\) in the tracer advection test.
+	\caption{\TODO{add terrain-following results to this table} Minimum and maximum tracer magnitudes and \(\ell^2\) error norms at \(t = \SI{10000}{\second}\) in the tracer advection test.
 Results of the cubic upwind-biased scheme are compared with the fourth-order centred scheme from \citet{schaer2002}.}
@@
 \end{table}
+
+\subsection{Terrain following advection}
+In the horizontal advection test, results were more accurate where the flow was aligned with the grid layers, and distortions in the BTF grid led to increased errors.
+We formulate a new tracer advection test in which the velocity field is everywhere tangential to the basic terrain following coordinate surfaces.
+In this new test, flow is aligned with the BTF grid layers.
+Misalignment with the SLEVE and cut cell grids is designed to challenge the advection scheme.
+
+The spatial domain, mountain profile, initial tracer profile and discretisation are the same as those in the horizontal tracer advection test.
+The velocity field is defined using a streamfunction, $\Psi$, so that the continuous velocity field is non-divergent and follows the BTF coordinate surfaces given by equation~\ref{eqn:btf} such that
+\begin{equation}
+	\Psi(x,z) = H \frac{z - h}{H - h}
+\end{equation}
+The horizontal and vertical components of velocity, $u$ and $w$ are then given by
+\begin{align}
+	u &= \frac{\partial \Psi}{\partial z} = \frac{H}{H - h}, \quad w = -\frac{\partial \Psi}{\partial x} = H \frac{\partial h}{\partial x} \frac{H - z}{\left( H - h \right)^2}, \nonumber \\
+	\frac{\partial h}{\partial x} &= - h_0 \left[ 
+		\beta \cos^2 \left( \alpha x \right) \sin \left( 2 \beta x \right) +
+		\alpha \cos^2 \left( \beta x \right) \sin \left( 2 \alpha x \right)
+	\right]
+\end{align}
+Unlike the horizontal advection test, flow extends from the top of the domain all the way to the ground.
+
+At $t = \SI{10000}{\second}$ the tracer has passed over the mountain.  The horizontal position of the tracer centre can be calculated by integrating \(\mathrm{d}t = \mathrm{d}x / u\) to find
+\begin{multline}
+t = x_0 + \frac{x}{u_0} - \frac{h_0}{16 u_0 H} \left[ 4x + \frac{\sin 2 (\alpha + \beta) x}{\alpha + \beta} \right.+ \\
+\left. \frac{\sin 2(\alpha - \beta) x}{\alpha - \beta} + 2 \left( \frac{\sin 2\alpha x}{\alpha} + \frac{\sin 2\beta x}{\beta} \right) \right]
+\end{multline}
+Solving numerically we find that \(x(t=\SI{10000}{\second}) = \SI{51.5}{\kilo\meter}\).
+Because the velocity field is non-divergent, the flow accelerates over mountain ridges and the tracer travels \SI{1.5}{\kilo\meter} further compared to advection in the horizontal velocity field.
+Tracer height is unchanged because advection is along the terrain following coordinate surface.
+
+\begin{itemize}
+	\item \TODO{should I run this test on the SLEVE grid as well for completeness?  I would expect results to be half-way between BTF and cut cell grids.}
+	\item Analysis
+	\item Conclusions/summary for this test
+\end{itemize}
 

### 2015-03-12 msc-publication.tex
plot meshes for MWR paper
introduce the tests we're performing

@@
 	\centering
-	TODO
+	\includegraphics{../msc-publication-fig-meshes/fig-meshes.pdf}
 	%
@@
 \section{Results}
-A series of standard, two-dimensional tests were performed over idealised orography.  For each test, results on the BTF, SLEVE and cut cell grid are compared.
+A series of standard, two-dimensional tests are performed over idealised orography.
+For each test, results on the BTF, SLEVE and cut cell grid are compared.
+The first test from \citet{schaer2002} advects a tracer in a horizontal velocity field.
+Second, a new tracer advection test is formulated employing a terrain following velocity field to challenge the advection scheme on more orthogonal grids.
+The third test solves the Euler equations for a stably stratified atmosphere at rest, following \citet{klemp2011}.
+Finally, as specified by \citet{schaer2002}, a test of orographically-induced gravity waves is performed.
+
 We use a finite volume discretisation of the fully-compressible Euler equations described by \citet{weller-shahrokhi2014}.
@@
 \begin{table}[t]
-	\caption{\TODO{add terrain-following results to this table} Minimum and maximum tracer magnitudes and \(\ell^2\) error norms at \(t = \SI{10000}{\second}\) in the tracer advection test.
+	\caption{\TODO{add terrain-following results to this table or put in separate table?} Minimum and maximum tracer magnitudes and \(\ell^2\) error norms at \(t = \SI{10000}{\second}\) in the tracer advection test.
 Results of the cubic upwind-biased scheme are compared with the fourth-order centred scheme from \citet{schaer2002}.}

### 2015-03-12 msc-publication.tex
a few punctuation tweaks and TODOs for MWR paper

@@
 \begin{table}[t]
-	\caption{\TODO{add terrain-following results to this table or put in separate table?} Minimum and maximum tracer magnitudes and \(\ell^2\) error norms at \(t = \SI{10000}{\second}\) in the tracer advection test.
+	\caption{\TODO{add terrain-following results to this table or put in separate table?
+will add another three columns} Minimum and maximum tracer magnitudes and \(\ell^2\) error norms at \(t = \SI{10000}{\second}\) in the tracer advection test.
 Results of the cubic upwind-biased scheme are compared with the fourth-order centred scheme from \citet{schaer2002}.}
@@
 In this new test, flow is aligned with the BTF grid layers.
-Misalignment with the SLEVE and cut cell grids is designed to challenge the advection scheme.
+Misalignment on the SLEVE and cut cell grids is designed to challenge the advection scheme.
 
@@
 \end{equation}
-The horizontal and vertical components of velocity, $u$ and $w$ are then given by
+The horizontal and vertical components of velocity, $u$ and $w$, are then given by
 \begin{align}
@@
 \begin{itemize}
-	\item \TODO{should I run this test on the SLEVE grid as well for completeness?  I would expect results to be half-way between BTF and cut cell grids.}
-	\item Analysis
+	\item l2 error is $\sim 10 \times$ worse on SLEVE compared to BTF
+	\item \TODO{find out what results are for cut cell grid}
 	\item Conclusions/summary for this test
@@
 \begin{enumerate}
-	\item BTF grid isn't as bad as people say it is.  We found that:
+	\item Results can be accurate on a highly-distorted grid such as BTF.  We found that:
 	\begin{itemize}
 		\item Upwind-biased cubic advection scheme accurately advects a tracer in non-divergent flows (Figure~\ref{fig:advection-tracer}, \ref{fig:advection-error}, Table~\ref{tab:advection})
+		\item Advection accuracy depends on alignment of the flow with grid layers
 		\item spurious vertical velocities are small in resting atmosphere (Figure~\ref{fig:resting})
@@
 
-\textbf{With the exception of the Lorenz computational mode on the SnapCol grid in the gravity waves test, results were satisfactory across BTF, SLEVE and SnapCol grids in all three test cases.}
-
-Miscellany:
-\begin{itemize}
-	\item Advection accuracy depends on alignment of the flow with grid layers (\TODO{we kinda need wobblyTracerAdvection to reach this conclusion})
-\end{itemize}
+\textbf{With the exception of the Lorenz computational mode on the SnapCol grid in the gravity waves test, results were ``satisfactory'' across BTF, SLEVE and SnapCol grids in all three test cases.}
 
@@
 \acknowledgments
-Start acknowledgments here.
+Start acknowledgments here.  \TODO{John, Terry, S-J Lock.  Assuming only Hilary is co-author...?}
 

### 2015-03-16 msc-publication.tex
add more conclusions to terrain-following advection

@@
 \begin{itemize}
-	\item l2 error is $\sim 10 \times$ worse on SLEVE compared to BTF
-	\item \TODO{find out what results are for cut cell grid}
-	\item Conclusions/summary for this test
+	\item l2 error is $\sim 10 \times$ worse on SLEVE compared to BTF, slightly worse again on cut cell grid
+	\item Conclusions/summary: Accuracy depends upon alignment of flow with grid.
+Accuracy not reduced by grid distortions.
+Accuracy on BTF grid comparable with accuracy on SLEVE grid in horizontal advection test.
 \end{itemize}

### 2015-04-10 msc-publication.tex
add stats for horizontal and TF advection, write some analysis of TF advection

--- 2015-03-12 msc-publication.tex
@@
 This results in a constant wind aloft, and zero flow at \SI{4}{\kilo\meter} and below.
-
-While the continuous wind field is non-divergent, this is not true of the discrete wind field on non-orthogonal grids.
-A potential, $p$, is used to correct the wind field, $\bm{u}$, so that is non-divergent such that
-\begin{align}
-	\nabla \cdot \left( \bm{u} + \nabla p \right) = 0,
-\intertext{which can be rearranged to give a Poisson equation}
-	\nabla^2 p = - \nabla \cdot \bm{u}.
-\end{align}
+To ensure the discrete velocity field is non-divergent, velocities are prescribed at cell faces by summing the streamfunction at the centre of all face edges.
 
@@
 The time derivative is solved using a second order Runge-Kutta scheme.
-\TODO{I ought to understand this method myself!}
 
@@
 
-\begin{table}[t]
-	\caption{\TODO{add terrain-following results to this table or put in separate table?
-will add another three columns} Minimum and maximum tracer magnitudes and \(\ell^2\) error norms at \(t = \SI{10000}{\second}\) in the tracer advection test.
-Results of the cubic upwind-biased scheme are compared with the fourth-order centred scheme from \citet{schaer2002}.}
-\label{tab:advection}
-%
-\centering
-\footnotesize
-\begin{tabular}{ l l l l l l }
-\hline\hline
-& \multicolumn{3}{c}{Cubic upwind-biased} & \multicolumn{2}{c}{Sch\"ar 4th order} \\
-& \(\ell^2\) error & min & max & min & max \\
-\hline
-Analytic  & 0 & 0 & 1 & 0 & 1 \\
-BTF 	  & \num{-1} & \num{-1} & \num{-1} & \num{-0.058} & \num{1.001} \\
-SLEVE 	  & \num{-1} & \num{-1} & \num{-1} & \num{-0.002} & \num{0.984} \\
-Cut cell  & \num{-1} & \num{-1} & \num{-1} & \multicolumn{1}{c}{---} & \multicolumn{1}{c}{---} \\
-No orography & \num{-1} & \num{-1} & \num{-1} & \num{-0.002} & \num{0.984} \\
-\hline
-\end{tabular}
-\end{table}
-
 \subsection{Terrain following advection}
@@
 \begin{equation}
-	\Psi(x,z) = H \frac{z - h}{H - h}
+	\Psi(x,z) = u_0 H \frac{z - h}{H - h}
 \end{equation}
+where $u_0 = \SI{10}{\meter\per\second}$, which is the horizontal wind speed where $h(x) = 0$.
 The horizontal and vertical components of velocity, $u$ and $w$, are then given by
 \begin{align}
-	u &= \frac{\partial \Psi}{\partial z} = \frac{H}{H - h}, \quad w = -\frac{\partial \Psi}{\partial x} = H \frac{\partial h}{\partial x} \frac{H - z}{\left( H - h \right)^2}, \nonumber \\
+	u &= \frac{\partial \Psi}{\partial z} = u_0 \frac{H}{H - h}, \quad w = -\frac{\partial \Psi}{\partial x} = u_0 H \frac{\mathrm{d} h}{\mathrm{d} x} \frac{H - z}{\left( H - h \right)^2}, \nonumber \\
 	\frac{\partial h}{\partial x} &= - h_0 \left[ 
@@
 Unlike the horizontal advection test, flow extends from the top of the domain all the way to the ground.
+The discrete velocity field is calculated using the streamfunction in the same way as the horizontal advection test.
 
-At $t = \SI{10000}{\second}$ the tracer has passed over the mountain.  The horizontal position of the tracer centre can be calculated by integrating \(\mathrm{d}t = \mathrm{d}x / u\) to find
+At $t = \SI{10000}{\second}$ the tracer has passed over the mountain.
+The horizontal position of the tracer centre can be calculated by integrating along the trajectory to find $t$, the time taken to pass from one side of the mountain to the other:
 \begin{multline}
-t = x_0 + \frac{x}{u_0} - \frac{h_0}{16 u_0 H} \left[ 4x + \frac{\sin 2 (\alpha + \beta) x}{\alpha + \beta} \right.+ \\
+t = \frac{x}{u_0} - \frac{h_0}{16 u_0 H} \left[ 4x + \frac{\sin 2 (\alpha + \beta) x}{\alpha + \beta} \right.+ \\
 \left. \frac{\sin 2(\alpha - \beta) x}{\alpha - \beta} + 2 \left( \frac{\sin 2\alpha x}{\alpha} + \frac{\sin 2\beta x}{\beta} \right) \right]
 \end{multline}
-Solving numerically we find that \(x(t=\SI{10000}{\second}) = \SI{51.5}{\kilo\meter}\).
-Because the velocity field is non-divergent, the flow accelerates over mountain ridges and the tracer travels \SI{1.5}{\kilo\meter} further compared to advection in the horizontal velocity field.
+Hence, we find that \(x(t=\SI{10000}{\second}) = \SI{51577.4}{\meter}\).
+Because the velocity field is non-divergent, the flow accelerates over mountain ridges and the tracer travels \SI{1577.4}{\meter} further compared to advection in the horizontal velocity field.
 Tracer height is unchanged because advection is along the terrain following coordinate surface.
 
-\begin{itemize}
-	\item l2 error is $\sim 10 \times$ worse on SLEVE compared to BTF
-	\item \TODO{find out what results are for cut cell grid}
-	\item Conclusions/summary for this test
-\end{itemize}
+\begin{table*}[t]
+	\caption{Minimum and maximum tracer magnitudes and \(\ell^2\) error norms at \(t = \SI{10000}{\second}\) in the horizontal and terrain following tracer advection tests.
+In the horizontal advection test, results of the cubic upwind-biased scheme are compared with the fourth-order centred scheme from \citet{schaer2002}.}
+\label{tab:advection}
+%
+\centering
+\footnotesize
+\begin{tabular}{ l l l l l l l l l l }
+\hline\hline
+& \multicolumn{5}{c}{Horizontal advection} & \multicolumn{3}{c}{Terrain following advection} \\
+& \multicolumn{2}{c}{Sch\"ar 4th order} & \multicolumn{3}{c}{Cubic upwind-biased} & \multicolumn{3}{c}{Cubic upwind-biased} \\
+& min & max & \(\ell^2\) error & min & max & \(\ell^2\) error & min & max \\
+\hline
+Analytic  & 0 & 0 & 1 & 0 & 1 & 0 & 0 & 1 \\
+BTF 	  & \num{-0.058} & \num{1.001} & \num{0.00767} & \num{-0.0433} & \num{0.928} & \num{0.00154} & \num{-0.0105} & \num{0.983} \\
+SLEVE 	  & \num{-0.002} & \num{0.984} & \num{0.00108} & \num{-0.0106} & \num{0.981} & \num{0.0122} & \num{-0.0273} & \num{0.864} \\
+Cut cell  & \multicolumn{1}{c}{---} & \multicolumn{1}{c}{---} & \num{0.000579} & \num{-0.00670} & \num{0.982} & \num{0.0136} & \num{-0.0297} & \num{0.850} \\
+No orography & \num{-0.002} & \num{0.984} & \num{0.000577} & \num{-0.00670} & \num{0.982} & \multicolumn{1}{c}{---} & \multicolumn{1}{c}{---} & \multicolumn{1}{c}{---} \\
+\hline
+\end{tabular}
+\end{table*}
+
+$\ell^2$ errors and tracer extrema for this test are compared with the horizontal advection results in table~\ref{tab:advection}.
+In the terrain following velocity field, tracer accuracy is greatest on the BTF grid.
+Errors are about ten times larger on the SLEVE and cut cell grids compared to the BTF grid.
+
+We conclude from this test that accuracy depends upon alignment of the flow with the grid, and accuracy is not significantly reduced by grid distortions.
+Accuracy on the BTF grid is comparable with accuracy on the SLEVE grid in the horizontal advection test.
 
@@
 An idealised terrain profile is defined along with a stratified atmosphere at rest in hydrostatic balance.
-The analytic solution is time-invariant, but numerical errors in calculating the horizontal pressure gradient can give rise to spurious velocities which become more severe over steeper terrain \citet{klemp2011}.
+The analytic solution is time-invariant, but numerical errors in calculating the horizontal pressure gradient can give rise to spurious velocities which become more severe over steeper terrain \citep{klemp2011}.
 
@@
 \citet{good2014} found the maximum vertical velocity in their cut cell model was \SI{1e-12}{\meter\per\second}, which is better than any result obtained using the model by \citet{weller-shahrokhi2014}.
-It is worth noting that \citet{good2014} used a timestep of \SI{1.01}{\second} instead of the \SI{100}{\second} timestep specified by \citet{klemp2011} that was used to obtain the results presented here.
-In addition, in the model used by \citet{good2014}, cell centres are in the centre of the uncut cell, resulting in the centre of some cut cells being below the ground (S.-J.
+It is worth noting that, in the model used by \citet{good2014}, cell centres are in the centre of the uncut cell, resulting in the centre of some cut cells being below the ground (S.-J.
 Lock 2014, personal communication).
 This means that the grid is effectively regular when calculating horizontal and vertical gradients.
-These two modelling decisions may account for the very small velocities found by \citet{good2014}.
+This may account for the very small velocities found by \citet{good2014}.
 
@@
 	\item Lorenz computation mode has been excited because Exner sample line at $x = \SI{50}{\kilo\meter}$ is in hydrostatic balance (Figure~\ref{fig:gw-exner-theta})
-	\item \TODO{We can speculate on what excites the computational mode, but perhaps better not to?}
-	\item No evidence of small cell problem -- \TODO{don't know that we can say much here without another vertical momentum test; our hypothesis about the quasi-horizontal flow isn't backed up with a test case}
 	\item Conclusion: results similar on all grids, agree with literature, except for Lorenz computational mode on SnapCol grid

### 2015-04-10 msc-publication.tex
more discussion of Lorenz computational mode, add introduction to sections, introduce grids section

@@
 
-This study uses the nonhydrostatic model from \citet{weller-shahrokhi2014} to enable a like-for-like comparison between BTF, SLEVE and cut cell grids for idealised, two-dimensional test cases from the literature.
-\TODO{final bit of intro: in section 2, section 3 etc etc}
+This study uses the nonhydrostatic model from \citet{weller-shahrokhi2014} to enable a like-for-like comparison between terrain following and cut cell grids for idealised, two-dimensional test cases from the literature.
+Section 2 presents the formulation of the terrain following and cut cell grids used in the experiments that follow.
+Section 3 analyses the results from two tracer advection tests, a test of a stratified atmosphere at rest, and orographically induced gravity waves.
+Concluding remarks are made in section 4.
 
@@
 
-\TODO{introduce this section}
+In this section we describe the formulation of the terrain following grids and the method of cut cell grid construction.
+The techniques presented here are used to define the grids for the experiments in the subsequent section.
+
 \citet{galchen-somerville1975} proposed a basic terrain following (BTF) coordinate defined as 
@@
 Figure~\ref{fig:gw-exner-theta} shows vertical profiles of the Exner function of pressure and potential temperature in the lowest \SI{1}{\kilo\meter} in the lee of the mountain at \(x = \SI{50}{\kilo\meter}\).
+We find that the Exner function of pressure decreases linearly with height on all grids.
+The potential temperature increases linearly with height on the BTF and SLEVE grids.
+On the cut cell grid, the potential temperature anomalies seen in figure~\ref{fig:gw-theta}d appear as a zig--zag in figure~\ref{fig:gw-exner-theta}.
 
-Analysis:
+The model by \citet{weller-shahrokhi2014} has a Lorenz staggering of variables so that $\theta$ is stored at cell centres and velocity at cell faces.
+The inaccuracies in potential temperature become invisible when calculating momentum because $\theta$ is interpolated onto cell faces, so discrete hydrostatic balance is maintained.
+This is a manifestation of the Lorenz computational mode, which is a disadvantage of the Lorenz grid that has been widely discussed in the literature \citep{arakawa-konor1996,holdaway2013b}.
+
 \begin{itemize}
-	\item \TODO{divergence might help explain computational mode -- it's greater on the SnapCol grid (MSc dissertation Figure 4.15e)}
 	\item \TODO{worth mentioning implications of this?
@@
 model has no viscosity so thermal mixing should not occur in theory}
-	\item Lorenz computation mode has been excited because Exner sample line at $x = \SI{50}{\kilo\meter}$ is in hydrostatic balance (Figure~\ref{fig:gw-exner-theta})
 	\item Conclusion: results similar on all grids, agree with literature, except for Lorenz computational mode on SnapCol grid

### 2015-04-12 msc-publication.tex
start writing conclusion

@@
 
-\section{Grids}
-\label{sec:grid}
-
 \begin{figure*}
@@
 
-In this section we describe the formulation of the terrain following grids and the method of cut cell grid construction.
+\section{Grids}
+\label{sec:grid}
+
+Here we describe the formulation of the terrain following grids and the method of cut cell grid construction.
 The techniques presented here are used to define the grids for the experiments in the subsequent section.
@@
 The terrain is wave-shaped, specified by the surface height, \(h\), such that
-\begin{subequations}
 \begin{align}
-	h(x) &= h^\star \cos^2 ( \alpha x )
-%
-	\intertext{where}
-%
-	h^\star(x) &= \left\{ \begin{array}{l l}
-	h_0 \cos^2 ( \beta x ) & \quad \text{if $| x | < a$} \\
+	h(x) &= \left\{ \begin{array}{l l}
+	h_0 \cos^2 ( \alpha x ) \cos^2 ( \beta x ) & \quad \text{if $| x | < a$} \\
 		0 & \quad \text{otherwise}
@@
 \end{align}
-\end{subequations}
 where $a = \SI{25}{\kilo\meter}$ is the mountain half-width, $h_0 = \SI{3}{\kilo\meter}$ is the maximum mountain height, $\lambda = \SI{8}{\kilo\meter}$ is the wavelength, \(\alpha = \pi / \lambda\) and \(\beta = \pi / 2a\).
@@
 
-The model by \citet{weller-shahrokhi2014} has a Lorenz staggering of variables so that $\theta$ is stored at cell centres and velocity at cell faces.
+The model by \citet{weller-shahrokhi2014} has a Lorenz staggering of variables such that $\theta$ is stored at cell centres and velocity at cell faces.
 The inaccuracies in potential temperature become invisible when calculating momentum because $\theta$ is interpolated onto cell faces, so discrete hydrostatic balance is maintained.
@@
 
-\begin{itemize}
-	\item \TODO{worth mentioning implications of this?
+\TODO{what might be the physical implications?
 1.
 reduced stability, although not enough to create vertical motion in this instance 2.
-would disrupt clouds 2.
+would disrupt clouds 3.
 model has no viscosity so thermal mixing should not occur in theory}
-	\item Conclusion: results similar on all grids, agree with literature, except for Lorenz computational mode on SnapCol grid
-\end{itemize}
+
+To summarize, results of the gravity waves test on all grids are in good agreement with the reference solution from \citet{melvin2010}.
+The most prominent errors are found only on the cut cell grid, where the potential temperature errors near the ground are maintained by the Lorenz computational mode.
 
@@
 \section{Conclusions}
+We have presented a like-for-like comparison between terrain following and cut cell grids using a single model.
+Accuracy on the BTF, SLEVE and cut cell grids was evaluated in a series of two-dimensional tests.
+
+Across all tests, a high degree of accuracy was achieved for all grids.
+Even on the highly-distorted BTF grid, which is often found to give poor results \citet{schaer2002,klemp2011,good2014}, errors were often small in the tests presented here.
+In the first two tests, tracers were advected in horizontal and terrain following velocity fields.
+We found that the accuracy of the upwind-biased cubic advection scheme depended upon alignment of the flow with the grid.
+Spurious vertical velocities were small in the resting atmosphere test, reaching a maximum of $\sim \SI{0.35}{\meter\per\second}$ on the BTF grid, compared to a maximum of $\sim \SI{10}{\meter\per\second}$ found by \citet{klemp2011}.
+In the gravity waves test, vertical velocities were in good agreement with the reference solution from \citet{melvin2010} across all grids.
+
 \begin{enumerate}
-	\item Results can be accurate on a highly-distorted grid such as BTF.  We found that:
-	\begin{itemize}
-		\item Upwind-biased cubic advection scheme accurately advects a tracer in non-divergent flows (Figure~\ref{fig:advection-tracer}, \ref{fig:advection-error}, Table~\ref{tab:advection})
-		\item Advection accuracy depends on alignment of the flow with grid layers
-		\item spurious vertical velocities are small in resting atmosphere (Figure~\ref{fig:resting})
-		\item gravity waves results visually as good as reference solution from \citet{melvin2010} (figure~\ref{fig:gw-w})
-	\end{itemize}
-
 	\item Cut cell grids can be worse than TF grids:
@@
 
-\section{Further work}
-\begin{itemize}
-	\item Lorenz computational mode motivates formulation of C-P staggering for cut cell grids
-	\item Find out what excites the Lorenz computational mode
-	\item \TODO{anything else?}
-\end{itemize}
+\TODO{Further work: Lorenz computational mode motivates formulation of C-P staggering for cut cell grids}
 

### 2015-04-12 msc-publication.tex
add abstract

@@
 
-\abstract{Enter the text of your abstract here.}
+%but are known to suffer from horizontal pressure gradient errors near steep terrain.
+\abstract{Terrain following coordinates are widely used in operational models.
+More recently, the cut cell method has been proposed as an alternative that can more accurately represent orography.
+In order to compare the accuracy of the different grids, it is typically necessary to use different models.
+Instead, we use a single model to enable a like-for-like comparison of terrain following and cut cell grids.
+A series of standard two-dimensional tests using idealised terrain are performed: tracer advection in a prescribed horizontal velocity field, a test of a stably stratified atmosphere at rest, and orographically induced gravity waves.
+In addition, we formulate a new tracer advection test having a velocity field that is tangential to the basic terrain following coordinate surfaces.
+The results of the tracer advection tests demonstrate that tracer accuracy depends upon alignment of the flow with the grid.
+In the gravity waves test, results on all grids are in good agreement with existing results from the literature.
+However, we find that the Lorenz computational mode is manifested as a vertical zig--zag in potential temperature only on the cut cell grid.
+Although errors do vary between grids, overall, a high degree of accuracy is obtained for all tests on all grids.}
 
@@
 \footnotesize
-\begin{tabular}{ l l l l l l l l l l }
+\begin{tabular}{ r @{\hspace{2em}} l l @{\hspace{1.5em}} l l l @{\hspace{2em}} l l l }
 \hline\hline
@@
 This is a manifestation of the Lorenz computational mode, which is a disadvantage of the Lorenz grid that has been widely discussed in the literature \citep{arakawa-konor1996,holdaway2013b}.
-
-\TODO{what might be the physical implications?
-1.
-reduced stability, although not enough to create vertical motion in this instance 2.
-would disrupt clouds 3.
-model has no viscosity so thermal mixing should not occur in theory}
+In models that include moist processes, the Lorenz computational mode can disrupt clouds and generate spurious precipitation \citep{hollingsworth1995}.
 
@@
 
-\begin{enumerate}
-	\item Cut cell grids can be worse than TF grids:
-	\begin{itemize}
-		\item Lorenz computational mode found on SnapCol grid only (figure~\ref{fig:gw-theta}, \ref{fig:gw-exner-theta})
-	\end{itemize}
+Cut cell grids reduced errors in two of the four tests.
+First, in the horizontal advection test, tracer accuracy on the cut cell grid was almost as good as accuracy on a regular grid with no mountain.
+Second, in the resting atmosphere test, spurious vertical velocities were two orders of magnitude smaller on the cut cell grid compared with the terrain folllowing grids.
 
-	\item Cut cell grids can also be better than TF grids in more artificial test cases:
-	\begin{itemize}
-		\item SnapCol $w$ two orders of magnitude smaller in resting atmosphere test
-		\item SnapCol advection test as good as noOrography
-	\end{itemize}
-\end{enumerate}
-
-\textbf{With the exception of the Lorenz computational mode on the SnapCol grid in the gravity waves test, results were ``satisfactory'' across BTF, SLEVE and SnapCol grids in all three test cases.}
-
-\TODO{Further work: Lorenz computational mode motivates formulation of C-P staggering for cut cell grids}
+Conversely, in the terrain following advection test, errors were large on the SLEVE and cut cell grids where velocities were misaligned with the grids.
+In the gravity waves test, the Lorenz computational mode was manifested as a zig--zag in potential temperature in the lowest layers in the lee of the mountain.
+This spurious error was excited only on the cut cell grid.
+This test motivates further work to formulate a Charney--Phillips staggering of variables on cut cell grids.
 

### 2015-04-13 msc-publication.tex
revert the mountain profile equation to include h^\star, fix typos

@@
 \title{Comparison of Terrain Following and Cut Cell Grids with a Non-Hydrostatic Model}
-\authors{James Shaw\correspondingauthor{Dept., Institution, Address, City, State/Country.}}
+\authors{James Shaw\correspondingauthor{Department of Meteorology, University of Reading, Earley Gate, PO Box 243, Reading, RG6 6BB, UK.}}
 \affiliation{}
@@
 %but are known to suffer from horizontal pressure gradient errors near steep terrain.
-\abstract{Terrain following coordinates are widely used in operational models.
-More recently, the cut cell method has been proposed as an alternative that can more accurately represent orography.
-In order to compare the accuracy of the different grids, it is typically necessary to use different models.
-Instead, we use a single model to enable a like-for-like comparison of terrain following and cut cell grids.
+\abstract{Terrain following coordinates are widely used in operational models but, more recently, the cut cell method has been proposed as an alternative that can more accurately represent the orography and associated atmospheric dynamics.
+Because the type of grid is usually chosen during model implementation, it is typically necessary to use different models in order to compare the accuracy of different grids.
+On the contrary, we use a single model to enable a like-for-like comparison of terrain following and cut cell grids.
 A series of standard two-dimensional tests using idealised terrain are performed: tracer advection in a prescribed horizontal velocity field, a test of a stably stratified atmosphere at rest, and orographically induced gravity waves.
 In addition, we formulate a new tracer advection test having a velocity field that is tangential to the basic terrain following coordinate surfaces.
-The results of the tracer advection tests demonstrate that tracer accuracy depends upon alignment of the flow with the grid.
+The results of the advection tests demonstrate that tracer accuracy depends upon alignment of the flow with the grid.
 In the gravity waves test, results on all grids are in good agreement with existing results from the literature.
@@
 Both methods align cells in vertical columns.
-Existing studies of cut cells have compared results with tests on terrain following grids from previous studies that used different models.
-On the contrary, this study uses a single model to enable a like-for-like comparison between terrain following and cut cell grids.
+Because most models are designed for a particular type of grid, existing studies of cut cells have compared results with terrain following grids using different models.
+Instead, this study uses a single model to enable a like-for-like comparison between terrain following and cut cell grids.
 
@@
 Cut cells is an alternative method in which cells that lie entirely below the terrain are removed, and those that intersect the surface are modified in shape so that they more closely fit the terrain.
-The resulting grid is orthogonal everywhere except where cells have been cut.
+The resulting grid is orthogonal everywhere except near cells that have been cut.
 Hence, errors are still introduced when calculating the horizontal pressure gradient between cut and uncut cells.
@@
 Spurious winds seen in TF coordinates are not present and errors do not increase with steeper terrain \citep{good2014}.
-\TODO{anything else?}
+\TODO{any other studies?}
 
@@
 Section 2 presents the formulation of the terrain following and cut cell grids used in the experiments that follow.
-Section 3 analyses the results from two tracer advection tests, a test of a stratified atmosphere at rest, and orographically induced gravity waves.
+Section 3 analyses the results from two tracer advection tests, a test of a stably stratified atmosphere at rest, and orographically induced gravity waves.
 Concluding remarks are made in section 4.
@@
 The two dimensional grids are \SI{20}{\kilo\meter} wide and \SI{2}{\kilo\meter} high.
+SLEVE parameters are specified in the resting atmosphere test in section~\ref{sec:results}\ref{sec:resting}.
 The cut cell grid was created by intersecting the terrain surface with a regular grid as described in section~\ref{sec:grid}.}
@@
 Here we describe the formulation of the terrain following grids and the method of cut cell grid construction.
-The techniques presented here are used to define the grids for the experiments in the subsequent section.
+The techniques presented are used to define the grids for the experiments in the subsequent section.
 
@@
 Using this coordinate, the terrain's influence decays linearly with height but disappears only at the top of the domain.
+An example is shown in figure~\ref{fig:grid}a.
 
@@
 Choosing \(n = 1\) gives the decay functions used by \citet{schaer2002}.
+An example of the SLEVE grid can be seen in figure~\ref{fig:grid}b.
 
@@
 \section{Results}
+\label{sec:results}
+
 A series of standard, two-dimensional tests are performed over idealised orography.
@@
 Following \citet{schaer2002}, a tracer is transported above wave-shaped terrain by solving the advection equation for a prescribed horizontal wind.
-This challenges the accuracy of the advection scheme in the presence of grid distortions.
+This test challenges the accuracy of the advection scheme in the presence of grid distortions.
 
-The domain is \SI{300}{\kilo\meter} wide and \SI{25}{\kilo\meter} high which is discretized onto a grid with \(\Delta x = \SI{1}{\kilo\meter}\) and \(\Delta z^\star = \SI{500}{\meter}\).
+The domain is \SI{300}{\kilo\meter} wide and \SI{25}{\kilo\meter} high, discretized onto a grid with \(\Delta x = \SI{1}{\kilo\meter}\) and \(\Delta z^\star = \SI{500}{\meter}\).
 The terrain is wave-shaped, specified by the surface height, \(h\), such that
+\begin{subequations}
 \begin{align}
-	h(x) &= \left\{ \begin{array}{l l}
-	h_0 \cos^2 ( \alpha x ) \cos^2 ( \beta x ) & \quad \text{if $| x | < a$} \\
+   h(x) &= h^\star \cos^2 ( \alpha x )
+%
+\intertext{where}
+%
+   h^\star(x) &= \left\{ \begin{array}{l l}
+       h_0 \cos^2 ( \beta x ) & \quad \text{if $| x | < a$} \\
 	0 & \quad \text{otherwise}
@@
 \end{align}
+\end{subequations}
 where $a = \SI{25}{\kilo\meter}$ is the mountain half-width, $h_0 = \SI{3}{\kilo\meter}$ is the maximum mountain height, $\lambda = \SI{8}{\kilo\meter}$ is the wavelength, \(\alpha = \pi / \lambda\) and \(\beta = \pi / 2a\).
@@
 
-A tracer $\varphi$ is positioned upstream above the height of the terrain.  It has the shape
+A tracer, $\varphi$, is positioned upstream above the height of the terrain.  It has the shape
 \begin{align}
@@
 %
-\intertext{having radius $r$ given by}
+\intertext{having radius, $r$, given by}
 %
@@
 At $t = \SI{0}{\second}$, the anomaly is centred at $(x_0, z_0) = (\SI{-50}{\kilo\meter}, \SI{9}{\kilo\meter})$ so that the anomaly is upwind of the mountain and well above the maximum terrain height of \SI{3}{\kilo\meter}.
-Analytic solutions can be found by adjusting the anomaly centre such that $x_0 = ut$.
+Analytic solutions can be found by setting the anomaly centre such that $x_0 = ut$.
 
@@
 Tracer contours at \(t = \SI{0}{\second}, \SI{5000}{\second}\) and \(\SI{10000}{\second}\) are shown in Figure~\ref{fig:advection-tracer}.
-The result from \citet{schaer2002} using a fourth-order centred scheme and sigma coordinates (\ref{fig:advection-tracer}a) is compared with the upwind-biased cubic scheme on (\ref{fig:advection-tracer}b) the BTF grid, and (\ref{fig:advection-tracer}c) cut cell grid.
+The result from \citet{schaer2002} using a fourth-order centred scheme and sigma coordinates (\ref{fig:advection-tracer}a) is compared with the upwind-biased cubic scheme on the BTF grid (\ref{fig:advection-tracer}b), and cut cell grid (\ref{fig:advection-tracer}c).
 
@@
 The results of the cubic upwind-biased scheme on TF and regular grids are comparable with those for the fourth-order centred scheme from \citet{schaer2002}.
-Error is largest on the BTF grid with \(\ell^2 = \num{0.00758}\) but significant reduced on the SLEVE grid with \(\ell^2 = \num{0.00110}\).
+Error is largest on the BTF grid with \(\ell^2 = \num{0.00767}\) but significant reduced on the SLEVE grid with \(\ell^2 = \num{0.00108}\).
 The error is approximately halved by changing from the SLEVE grid to the cut cell grid.
@@
 Contour intervals are every 0.1.
-The terrain profile is also shown immediately above the ground.}
+The terrain profile is also shown immediately above the $x$ axis.}
 	\label{fig:advection-tracer}
@@
 We conclude from this test that accuracy depends upon alignment of the flow with the grid, and accuracy is not significantly reduced by grid distortions.
-Accuracy on the BTF grid is comparable with accuracy on the SLEVE grid in the horizontal advection test.
+Error on the BTF grid is comparable with the error on the SLEVE grid in the horizontal advection test.
 
 \subsection{Resting atmosphere}
-An idealised terrain profile is defined along with a stratified atmosphere at rest in hydrostatic balance.
+\label{sec:resting}
+
+An idealised terrain profile is defined along with a stably stratified atmosphere at rest in hydrostatic balance.
 The analytic solution is time-invariant, but numerical errors in calculating the horizontal pressure gradient can give rise to spurious velocities which become more severe over steeper terrain \citep{klemp2011}.
@@
 \begin{align}
-	h(x) = h_0 \exp \left( - \left( \frac{x}{a} \right)^2 \right) \cos^2 \left( \frac{\pi x}{\lambda} \right) \label{eqn:resting:mountain}
+	h(x) = h_0 \exp \left( - \left( \frac{x}{a} \right)^2 \right) \cos^2 \left( \alpha x \right) \label{eqn:resting:mountain}
 \end{align}
@@
 The sponge layer is only active on faces whose normal is vertical so that it damps vertical momentum only.
+\TODO{is this true of the inlet sponge as well?}
 
-Note that, while the domain itself is \SI{30}{\kilo\meter} in height, for the purposes of generating of BTF and SLEVE grids, the domain height is set to \SI{20}{\kilo\meter} because the sponge layer occupies the uppermost \SI{10}{\kilo\meter}.
+Note that, while the domain itself is \SI{30}{\kilo\meter} in height, for the purposes of generating BTF and SLEVE grids, the domain height is set to \SI{20}{\kilo\meter} because the sponge layer occupies the uppermost \SI{10}{\kilo\meter}.
 
@@
 %
-	\caption{Vertical cross section of vertical velocity contours in the gravity waves test after 5 hours on (a) the BTF grid compared with (b) the mass-conserving semi-implicit semi-Lagrangian solution from \citet{melvin2010}.
+	\caption{Vertical cross section of vertical velocity contours in the gravity waves test after 5 hours with mountain height $h_0 = \SI{250}{\meter}$ on (a) the BTF grid compared with (b) the mass-conserving semi-implicit semi-Lagrangian solution from \citet{melvin2010}.
 Contours are every \SI{5e-2}{\meter\per\second} with solid lines denoting ascent and dashed lines descent.}

### 2015-04-24 msc-publication.tex
Tweaks to abstract and intro from John and Hilary

@@
 %but are known to suffer from horizontal pressure gradient errors near steep terrain.
-\abstract{Terrain following coordinates are widely used in operational models but, more recently, the cut cell method has been proposed as an alternative that can more accurately represent the orography and associated atmospheric dynamics.
+\abstract{Terrain following coordinates are widely used in operational models but the cut cell method has been proposed as an alternative that can more accurately represent the orography and associated atmospheric dynamics.
 Because the type of grid is usually chosen during model implementation, it is typically necessary to use different models in order to compare the accuracy of different grids.
-On the contrary, we use a single model to enable a like-for-like comparison of terrain following and cut cell grids.
+On the contrary, we use a single C-grid finite volume model to enable a like-for-like comparison of terrain following and cut cell grids.
 A series of standard two-dimensional tests using idealised terrain are performed: tracer advection in a prescribed horizontal velocity field, a test of a stably stratified atmosphere at rest, and orographically induced gravity waves.
-In addition, we formulate a new tracer advection test having a velocity field that is tangential to the basic terrain following coordinate surfaces.
+In addition, we formulate a new tracer advection test having a velocity field that is everywhere tangential to the terrain following coordinate surfaces.
 The results of the advection tests demonstrate that tracer accuracy depends upon alignment of the flow with the grid.
 In the gravity waves test, results on all grids are in good agreement with existing results from the literature.
-However, we find that the Lorenz computational mode is manifested as a vertical zig--zag in potential temperature only on the cut cell grid.
-Although errors do vary between grids, overall, a high degree of accuracy is obtained for all tests on all grids.}
+However, we find that the Lorenz computational mode is excited as a vertical zig--zag in potential temperature on the cut cell grid.
+\TODO{this last sentence is weak: Although errors do vary between grids, overall, a high degree of accuracy is obtained for all tests on all grids.}}
 
@@
 \section{Introduction}
-Representing orography accurately in numerical weather prediction systems is necessary to model downslope winds and local precipitation patterns.
+Representing orography accurately in numerical weather prediction systems is necessary to model downslope winds and local precipitation.
 There are two main approaches to represent orography on a grid: terrain following layers and cut cells.
 Both methods align cells in vertical columns.
-Because most models are designed for a particular type of grid, existing studies of cut cells have compared results with terrain following grids using different models.
-Instead, this study uses a single model to enable a like-for-like comparison between terrain following and cut cell grids.
+Because most models are designed for a particular type of grid, existing studies of cut cell solutions have compared results with terrain following grid solutions implemented within different models.
+Instead, this study uses a single model to enable a like-for-like comparison between solutions using terrain following and cut cell grids.
 
-With increasing model resolution, terrain can become steeper, making accurate calculation of the horizontal pressure gradient more difficult \citep{gary1973,steppeler2002}.
+With increasing horizontal model resolution, the discrete representation of terrain can become steeper, making accurate calculation of the horizontal pressure gradient more difficult \citep{gary1973,steppeler2002}.
 Numerical errors in this calculation result in spurious winds and can cause numerical instability \citep{fast2003,webster2003}.
@@
 	%
-	\caption{Examples of (a) BTF, (b) SLEVE, and (c) a cut cell grid, all constructed with OpenFOAM.
-The two dimensional grids are \SI{20}{\kilo\meter} wide and \SI{2}{\kilo\meter} high.
+	\caption{Examples of (a) BTF, (b) SLEVE, and (c) a cut cell grid, showing cell edges in the lowest four layers.
+The two dimensional grids are \SI{20}{\kilo\meter} wide and \SI{20}{\kilo\meter} high.
 SLEVE parameters are specified in the resting atmosphere test in section~\ref{sec:results}\ref{sec:resting}.
@@
 \acknowledgments
-Start acknowledgments here.  \TODO{John, Terry, S-J Lock.  Assuming only Hilary is co-author...?}
+Start acknowledgments here.  \TODO{John, Terry.  Assuming only Hilary is co-author...?}
 

### 2015-04-24 msc-publication.tex
add streamfunction maths, model equations

@@
 
-%but are known to suffer from horizontal pressure gradient errors near steep terrain.
 \abstract{Terrain following coordinates are widely used in operational models but the cut cell method has been proposed as an alternative that can more accurately represent the orography and associated atmospheric dynamics.
@@
 
-Several studies have found that cut cells produce more accurate results when compared to TF coordinates.
+Several studies have shown examples where cut cells produce more accurate results when compared to TF coordinates.
 Spurious winds seen in TF coordinates are not present and errors do not increase with steeper terrain \citep{good2014}.
-\TODO{any other studies?}
+A comparison of TF and cut cells using real initial data by \citet{steppeler2013} found that precipitation and wind forecasts were more accurately in the cut cell model.
 
@@
 Section 2 presents the formulation of the terrain following and cut cell grids used in the experiments that follow.
-Section 3 analyses the results from two tracer advection tests, a test of a stably stratified atmosphere at rest, and orographically induced gravity waves.
-Concluding remarks are made in section 4.
+\TODO{section 3} Section 4 analyses the results from two tracer advection tests, a test of a stably stratified atmosphere initially at rest, and orographically induced gravity waves.
+Concluding remarks are made in section 5.
 
@@
 \end{equation}
-where, in two dimensions, \(z(x, z^\star)\) is the height of the coordinate surface at level \(z^\star\), \(H\) is the height of the domain, and \(h(x)\) is the height of the terrain surface.
+where, in two dimensions, \(z(x, z^\star)\) is the physical height of the coordinate surface at level \(z^\star\), \(H\) is the height of the domain, and \(h(x)\) is the height of the terrain surface.
 Using this coordinate, the terrain's influence decays linearly with height but disappears only at the top of the domain.
@@
 The sigma coordinate transform of \citet{phillips1957} is equivalent to the BTF coordinate transform since they both decay linearly.
-However, since they decay with pressure rather than height, sigma coordinates also change with horizontal variations in pressure.
+However, because \(\sigma = p/p_s\) decays with pressure rather than height, sigma coordinates also vary with surface pressure, \(p_s(x, t)\).
 
@@
 
+\section{Model equations}
+\label{sec:model}
+
+The C-grid finite volume model from \citet{weller-shahrokhi2014} discretises the fully-compressible Euler equations, given by
+\begin{subequations}
+\begin{align}
+	\text{Momentum} &\ &\  	\frac{\partial \rho \bm{u}}{\partial t} + \nabla \cdot \rho \bm{uu} &= \rho \bm{g} - c_p \rho \theta \nabla \Pi \label{eq:momentum} \\
+	\text{Continuity} &\ &\	\frac{\partial \rho}{\partial t} + \nabla \cdot \rho \bm{u} &= 0 \\
+	\text{$\theta$, flux form} &\ &\ \frac{\partial \rho \theta}{\partial t} + \nabla \cdot \rho \bm{u} \theta &= 0 \\
+	\text{$\theta$, advective form} &\ &\ \frac{\partial \theta}{\partial t} + \bm{u} \cdot \nabla \theta &= 0 \\
+	\text{Equation of state} &\ &\ \Pi^{(1 - \kappa)/\kappa} &= \frac{R \rho \theta}{p_0}
+\end{align}
+\end{subequations}
+where \(\rho\) is the density, \(\bm{u}\) is the velocity field, \(\bm{g}\) is the gravitational acceleration, \(c_p\) is the heat capacity at constant pressure, \(\theta = T \left(p_0/p\right)^\kappa\) is the potential temperature, \(T\) is the tempereature, \(p\) is the pressure, \(p_0\) is a reference pressure, \(\Pi = \left(p / p_0 \right)^\kappa\) is the Exner function of pressure, and \(\kappa = R/c_p\) is the gas constant to heat capacity ratio.
+
 \section{Results}
@@
 
-A series of standard, two-dimensional tests are performed over idealised orography.
+A series of two-dimensional tests are performed over idealised orography.
 For each test, results on the BTF, SLEVE and cut cell grid are compared.
@@
 Second, a new tracer advection test is formulated employing a terrain following velocity field to challenge the advection scheme on more orthogonal grids.
-The third test solves the Euler equations for a stably stratified atmosphere at rest, following \citet{klemp2011}.
+The third test solves the Euler equations for a stably stratified atmosphere initially at rest, following \citet{klemp2011}.
 Finally, as specified by \citet{schaer2002}, a test of orographically-induced gravity waves is performed.
@@
 We use a finite volume discretisation of the fully-compressible Euler equations described by \citet{weller-shahrokhi2014}.
-The model has a curl-free pressure gradient formulation, an upwind-biased cubic advection scheme, and uses a Lorenz staggering of thermodynamic variables.
+The model has a curl-free pressure gradient formulation and an upwind-biased cubic advection scheme.
+It uses a Lorenz staggering of thermodynamic variables such that $\theta$ is stored at cell centres and velocity at cell faces.
 No explicit diffusion is used in any of the tests.
@@
 \end{subequations}
-where $a = \SI{25}{\kilo\meter}$ is the mountain half-width, $h_0 = \SI{3}{\kilo\meter}$ is the maximum mountain height, $\lambda = \SI{8}{\kilo\meter}$ is the wavelength, \(\alpha = \pi / \lambda\) and \(\beta = \pi / 2a\).
+where $a = \SI{25}{\kilo\meter}$ is the mountain envelope half-width, $h_0 = \SI{3}{\kilo\meter}$ is the maximum mountain height, $\lambda = \SI{8}{\kilo\meter}$ is the wavelength, \(\alpha = \pi / \lambda\) and \(\beta = \pi / 2a\).
 On the SLEVE grid, the large-scale component $h_1$ is given by \(h_1(x) = h^\star(x) / 2\)
@@
 This results in a constant wind aloft, and zero flow at \SI{4}{\kilo\meter} and below.
-To ensure the discrete velocity field is non-divergent, velocities are prescribed at cell faces by summing the streamfunction at the centre of all face edges.
+
+To ensure the discrete velocity field is non-divergent, velocities are prescribed at cell faces by summing the streamfunction, \(\Psi\), at the centre of all face edges.
+Since \(u = \partial \Psi / \partial z\), the streamfunction is found by vertical integration of the velocity profile:
+\begin{align}
+	\Psi(z) &= \frac{u_0}{2} \left\{ \begin{array}{l l}
+		\left( 2z - z_1 - z_2 \right) & \enskip \text{if $z > z_2$} \\
+		z - z_1 - \frac{z_2 - z_1}{\pi} \sin \left(\pi \frac{z - z_1}{z_2-z_1}\right) & \enskip \text{if $z_1 < z \leq z_2$} \\
+		0 & \enskip \text{if $z \leq z_1$}
+	\end{array} \right.
+\end{align}
 
@@
 Tracer magnitude and shape are well-preserved on all grids, both above the mountain at \(t = \SI{5000}{\second}\) and past the mountain at \(t = \SI{10000}{\second}\).
-Advection is most accurate on the cut cell grid (figure~\ref{fig:advection-tracer}c) and regular grid (not shown).
+In this test, advection is most accurate on the cut cell grid (figure~\ref{fig:advection-tracer}c) and regular grid (not shown).
 As found by \citet{good2014}, the result is the same on both grids.
@@
 Hence, we find that \(x(t=\SI{10000}{\second}) = \SI{51577.4}{\meter}\).
-Because the velocity field is non-divergent, the flow accelerates over mountain ridges and the tracer travels \SI{1577.4}{\meter} further compared to advection in the horizontal velocity field.
-Tracer height is unchanged because advection is along the terrain following coordinate surface.
+Because the velocity field is non-divergent, the flow accelerates over mountain ridges and the tracer travels \SI{1577.4}{\meter} further compared to advection in the purely horizontal velocity field.
+Tracer height is unchanged downwind of the mountains because advection is along the terrain following coordinate surface.
 
@@
 
-\subsection{Resting atmosphere}
+\subsection{Stratified atmosphere initially at rest}
 \label{sec:resting}
@@
 Sponge layers are added to the upper \SI{10}{\kilo\meter} and leftmost \SI{10}{\kilo\meter} at the inlet boundary to damp the reflection of waves.
-The term $\mu \rho \bf{u}$ is subtracted from the momentum equation \TODO{include the equation set in this paper, or cite weller-shahrokhi2014 here?} where the damping function $\mu$ is adapted from \citet{melvin2010} such that
+The term $\mu \rho \bm{u}$ is subtracted from the momentum (equation \ref{eq:momentum}) where the damping function, \(\mu\), is adapted from \citet{melvin2010} such that
 \begin{align}
@@
 The sponge layer is only active on faces whose normal is vertical so that it damps vertical momentum only.
-\TODO{is this true of the inlet sponge as well?}
 
@@
 
-The model by \citet{weller-shahrokhi2014} has a Lorenz staggering of variables such that $\theta$ is stored at cell centres and velocity at cell faces.
 The inaccuracies in potential temperature become invisible when calculating momentum because $\theta$ is interpolated onto cell faces, so discrete hydrostatic balance is maintained.
@@
 To summarize, results of the gravity waves test on all grids are in good agreement with the reference solution from \citet{melvin2010}.
-The most prominent errors are found only on the cut cell grid, where the potential temperature errors near the ground are maintained by the Lorenz computational mode.
+The most prominent errors are found only on the cut cell grid, where the potential temperature errors near the ground result from exciting the Lorenz computational mode.
 
@@
 Across all tests, a high degree of accuracy was achieved for all grids.
-Even on the highly-distorted BTF grid, which is often found to give poor results \citet{schaer2002,klemp2011,good2014}, errors were often small in the tests presented here.
+Even on the highly-distorted BTF grid, which have previously been found to give poor results \citep{schaer2002,klemp2011,good2014}, errors were often small in the tests presented here.
 In the first two tests, tracers were advected in horizontal and terrain following velocity fields.
 We found that the accuracy of the upwind-biased cubic advection scheme depended upon alignment of the flow with the grid.
+
 Spurious vertical velocities were small in the resting atmosphere test, reaching a maximum of $\sim \SI{0.35}{\meter\per\second}$ on the BTF grid, compared to a maximum of $\sim \SI{10}{\meter\per\second}$ found by \citet{klemp2011}.
@@
 \acknowledgments
-Start acknowledgments here.  \TODO{John, Terry.  Assuming only Hilary is co-author...?}
+Start acknowledgments here.  \TODO{John, Terry.  NERC grant NE/H015698/1.  Assuming only Hilary is co-author...?}
 

### 2015-04-24 msc-publication.tex
Move more model description into its own section.

@@
 Both methods align cells in vertical columns.
-Because most models are designed for a particular type of grid, existing studies of cut cell solutions have compared results with terrain following grid solutions implemented within different models.
+Because most models are designed for a particular type of grid, existing studies of cut cell solutions have compared results with terrain following grid solutions implemented within different models, for example \citet{good2014}.
 Instead, this study uses a single model to enable a like-for-like comparison between solutions using terrain following and cut cell grids.
 
-With increasing horizontal model resolution, the discrete representation of terrain can become steeper, making accurate calculation of the horizontal pressure gradient more difficult \citep{gary1973,steppeler2002}.
+With increasing horizontal model resolution, the discrete representation of terrain can become steeper, making accurate calculation of the horizontal pressure gradient more difficult when using terrain following layers \citep{gary1973,steppeler2002}.
 Numerical errors in this calculation result in spurious winds and can cause numerical instability \citep{fast2003,webster2003}.
-Terrain following and cut cell methods seek to reduce the error that is associated with steep orography.
+Cut cell methods seek to reduce the error that is associated with steep orography.
 
@@
 The techniques proposed by \citet{klemp2011} and \cite{zaengl2012} both involve the calculation of the horizontal pressure gradient in the physical domain.
-They found that, by avoiding the calculation of metric terms associated with the computational domain, numerical errors were reduced.
+This gave them the flexibility to design more accurate horizontal pressure gradient discretizations using more appropriate stencils.
 
@@
 
-Cut cells is an alternative method in which cells that lie entirely below the terrain are removed, and those that intersect the surface are modified in shape so that they more closely fit the terrain.
+Cut cells is an alternative method in which the grid does not follow the terrain but, instead, cells that lie entirely below the terrain are removed, and those that intersect the surface are modified in shape so that they more closely fit the terrain.
 The resulting grid is orthogonal everywhere except near cells that have been cut.
@@
 
-\section{Model equations}
+\section{Model}
 \label{sec:model}
 
-The C-grid finite volume model from \citet{weller-shahrokhi2014} discretises the fully-compressible Euler equations, given by
+We use the finite volume model from \citet{weller-shahrokhi2014} which details a C-grid discretisation of the fully-compressible Euler equations, given by
 \begin{subequations}
@@
 
+The mode has a curl-free pressure gradient formulation and an upwind-biased cubic advection scheme.
+It uses a Lorenz staggering of thermodynamic variables such that $\theta$ is stored at cell centres and velocity at cell faces.
+
 \section{Results}
@@
 Finally, as specified by \citet{schaer2002}, a test of orographically-induced gravity waves is performed.
-
-We use a finite volume discretisation of the fully-compressible Euler equations described by \citet{weller-shahrokhi2014}.
-The model has a curl-free pressure gradient formulation and an upwind-biased cubic advection scheme.
-It uses a Lorenz staggering of thermodynamic variables such that $\theta$ is stored at cell centres and velocity at cell faces.
 No explicit diffusion is used in any of the tests.
@@
 This means that the grid is effectively regular when calculating horizontal and vertical gradients.
-This may account for the very small velocities found by \citet{good2014}.
+This would account for the very small velocities found by \citet{good2014}.
 
-In summary, spurious velocities in the resting atmosphere test were similar on both terrain following grids, with much lower errors compared to those from \citet{klemp2011}.
-The maximum vertical velocity was significantly decreased on the cut cell grid, so we conclude that non-orthogonality is a significant cause of numerical error in this test.
+In summary, spurious velocities in the resting atmosphere test were similar on both terrain following grids, with lower errors compared to those from \citet{klemp2011}.
+The maximum vertical velocity was significantly decreased on the cut cell grid, so we conclude that non-orthogonality, or lack of alignment of the grid with sufaces of constant gravitational potential are a significant cause of numerical error in this test.
 
@@
 The inaccuracies in potential temperature become invisible when calculating momentum because $\theta$ is interpolated onto cell faces, so discrete hydrostatic balance is maintained.
-This is a manifestation of the Lorenz computational mode, which is a disadvantage of the Lorenz grid that has been widely discussed in the literature \citep{arakawa-konor1996,holdaway2013b}.
+This is a manifestation of the Lorenz computational mode \citep{arakawa-konor1996,holdaway2013b}.
 In models that include moist processes, the Lorenz computational mode can disrupt clouds and generate spurious precipitation \citep{hollingsworth1995}.
@@
 To summarize, results of the gravity waves test on all grids are in good agreement with the reference solution from \citet{melvin2010}.
-The most prominent errors are found only on the cut cell grid, where the potential temperature errors near the ground result from exciting the Lorenz computational mode.
+The most prominent errors are found only on the cut cell grid, where the potential temperature errors near the ground excite the Lorenz computational mode.
 
@@
 Even on the highly-distorted BTF grid, which have previously been found to give poor results \citep{schaer2002,klemp2011,good2014}, errors were often small in the tests presented here.
-In the first two tests, tracers were advected in horizontal and terrain following velocity fields.
-We found that the accuracy of the upwind-biased cubic advection scheme depended upon alignment of the flow with the grid.
+In the first two tests, tracers were advected by horizontal and terrain following velocity fields.
+We found that the accuracy of the upwind-biased cubic advection scheme depended upon alignment of the flow with the grid rather than on grid distortions.
 
@@
 \acknowledgments
-Start acknowledgments here.  \TODO{John, Terry.  NERC grant NE/H015698/1.  Assuming only Hilary is co-author...?}
+\TODO{John, Terry. Assuming only Hilary is co-author...?} Weller is funded by NERC grant NE/H015698/1.
 

### 2015-04-27 fig-advection-tracer.tex
add subfigure labels, improve captioning
switch advection tests over to AtmosTests

@@
 \begin{document}
-\hspace{0.15in}\includegraphics[width=2.75in]{img/schaer-btf-4thorder.png}
+\begin{textblock}{1}(2.5,0.4)
+\normalsize
+(a)
+\end{textblock}
+\footnotesize
+\vspace*{-2.2em}
+\input{../advection-horizontal-linear-btf/tracer-contour}
 
-\footnotesize
-\input{../advectTracer-horizontal-btf-cubicUpwindCPCFit-divFree/tracer-contour-y-only}
+\begin{textblock}{1}(2.5,5.5)
+\normalsize
+(b)
+\end{textblock}
+\input{../advection-horizontal-cubicUpwind-btf/tracer-contour}
 
-\input{../advectTracer-horizontal-snapCol-cubicUpwindCPCFit-divFree/tracer-contour}
+\begin{textblock}{1}(2.5,10.3)
+\normalsize
+(c)
+\end{textblock}
+\input{../advection-horizontal-cubicUpwind-cutCell/tracer-contour}
 \end{document}

### 2015-04-27 msc-publication.tex
switch advection tests over to AtmosTests

@@
 Section 2 presents the formulation of the terrain following and cut cell grids used in the experiments that follow.
-\TODO{section 3} Section 4 analyses the results from two tracer advection tests, a test of a stably stratified atmosphere initially at rest, and orographically induced gravity waves.
+In section 3 we outline the model from \citet{weller-shahrokhi2014} and the governing equations.
+Section 4 analyses the results from two tracer advection tests, a test of a stably stratified atmosphere initially at rest, and orographically induced gravity waves.
 Concluding remarks are made in section 5.
@@
 
-Minimum and maximum tracer values and \(\ell^2\) error norms on the BTF, SLEVE, cut cell and regular grids are summarised in table~\ref{tab:advection}.
+Minimum and maximum tracer values and \(\ell^2\) error norms on the BTF, SLEVE, cut cell and regular grids are summarised in table~\ref{tab:advection}, where the \(\ell^2\) error norm is defined as 
+\begin{align}
+	\ell^2 &= \sqrt{\frac{\sum_c \left( \varphi - \varphi_{T} \right)^2 \mathcal{V}_c}{\sum_c \left( \varphi_T^2 \mathcal{V}_c \right)}}
+\end{align}
+where $\varphi$ is the numerical tracer value, $\varphi_T$ is the analytic value and $\mathcal{V}_c$ is the cell volume.
+
 The results of the cubic upwind-biased scheme on TF and regular grids are comparable with those for the fourth-order centred scheme from \citet{schaer2002}.

### 2015-04-27 msc-publication.tex
add more results to advection table

@@
 
-The test was executed on the BTF, SLEVE and cut cell grids, and on a regular grid with flat terrain.
+The test was executed on the BTF, SLEVE and cut cell grids, and on a regular grid with flat terrain using a centred linear scheme and the upwind-biased cubic scheme.
 Tracer contours at \(t = \SI{0}{\second}, \SI{5000}{\second}\) and \(\SI{10000}{\second}\) are shown in Figure~\ref{fig:advection-tracer}.
-The result from \citet{schaer2002} using a fourth-order centred scheme and sigma coordinates (\ref{fig:advection-tracer}a) is compared with the upwind-biased cubic scheme on the BTF grid (\ref{fig:advection-tracer}b), and cut cell grid (\ref{fig:advection-tracer}c).
+We do not compare with the results from \citet{schaer2002} because we were unable to reproduce their result using a centred linear scheme (their figure 6a), though similar results were obtained by halving the mountain height, \(h_0\).
 
-Tracer magnitude and shape are well-preserved on all grids, both above the mountain at \(t = \SI{5000}{\second}\) and past the mountain at \(t = \SI{10000}{\second}\).
+The result of the centred linear scheme on the BTF grid (\ref{fig:advection-tracer}a) is compared with the upwind-biased cubic scheme on the BTF grid (\ref{fig:advection-tracer}b), and cut cell grid (\ref{fig:advection-tracer}c).
+Using the upwind-biased cubic scheme, tracer magnitude and shape are well-preserved on all grids, both above the mountain at \(t = \SI{5000}{\second}\) and past the mountain at \(t = \SI{10000}{\second}\).
 In this test, advection is most accurate on the cut cell grid (figure~\ref{fig:advection-tracer}c) and regular grid (not shown).
@@
 This is to be expected since the wind is zero in the region of the ground and flow aloft is aligned with the grids.
-On the BTF grid, the tracer is less distorted by the cubic upwind-biased scheme (figure~\ref{fig:advection-tracer}b) compared to the fourth-order centred scheme from \citet{schaer2002} (figure~\ref{fig:advection-tracer}a).
+On the BTF grid, the tracer is less distorted by the cubic upwind-biased scheme (figure~\ref{fig:advection-tracer}b) compared to the centred linear scheme (figure~\ref{fig:advection-tracer}a).
 
@@
 
-The results of the cubic upwind-biased scheme on TF and regular grids are comparable with those for the fourth-order centred scheme from \citet{schaer2002}.
+The results of the cubic upwind-biased scheme on TF and regular grids are comparable with those for the fourth-order centred scheme from \citet{schaer2002} \TODO{we're not comparing with schaer any more}.
 Error is largest on the BTF grid with \(\ell^2 = \num{0.00767}\) but significant reduced on the SLEVE grid with \(\ell^2 = \num{0.00108}\).
@@
 %
-	\caption{Horizontally advected tracer contours at \(t = \SI{0}{\second}\), \SI{5000}{\second} and \SI{10000}{\second} using (a) the fourth-order centred difference scheme with sigma coordinates from \citet{schaer2002}, and the upwind-biased cubic scheme on (b) BTF grid, (c) cut cell grid.
+	\caption{Horizontally advected tracer contours at \(t = \SI{0}{\second}\), \SI{5000}{\second} and \SI{10000}{\second} using (a) centred linear scheme on BTF grid, and the upwind-biased cubic scheme on (b) BTF grid, (c) cut cell grid.
 Contour intervals are every 0.1.
@@
 	%
-	\caption{Errors in horizontal tracer advection at \(t = \SI{10000}{\second}\) using (a) the fourth-order centred difference scheme with sigma coordinates from \citet{schaer2002}, (b) the upwind-biased cubic scheme on a BTF grid.
+	\caption{Errors in horizontal tracer advection on the BTF grid at \(t = \SI{10000}{\second}\) using (a) centred linear scheme, (b) the upwind-biased cubic scheme.
 Contour intervals are every 0.01 with negative contours denoted by dashed lines.}
@@
 \begin{table*}[t]
-	\caption{Minimum and maximum tracer magnitudes and \(\ell^2\) error norms at \(t = \SI{10000}{\second}\) in the horizontal and terrain following tracer advection tests.
-In the horizontal advection test, results of the cubic upwind-biased scheme are compared with the fourth-order centred scheme from \citet{schaer2002}.}
+	\caption{Minimum and maximum tracer magnitudes and \(\ell^2\) error norms at \(t = \SI{10000}{\second}\) in the horizontal and terrain following tracer advection tests using centred linear and cubic upwind-biased schemes.}
 \label{tab:advection}
@@
 \footnotesize
-\begin{tabular}{ r @{\hspace{2em}} l l @{\hspace{1.5em}} l l l @{\hspace{2em}} l l l }
+\begin{tabular}{ r @{\hspace{1em}} l l l @{\hspace{0.5em}} l l l @{\hspace{1em}} l l l @{\hspace{0.5em}} l l l}
 \hline\hline
-& \multicolumn{5}{c}{Horizontal advection} & \multicolumn{3}{c}{Terrain following advection} \\
-& \multicolumn{2}{c}{Sch\"ar 4th order} & \multicolumn{3}{c}{Cubic upwind-biased} & \multicolumn{3}{c}{Cubic upwind-biased} \\
-& min & max & \(\ell^2\) error & min & max & \(\ell^2\) error & min & max \\
+& \multicolumn{6}{c}{Horizontal advection} & \multicolumn{6}{c}{Terrain following advection} \\
+& \multicolumn{3}{c}{Centred linear} & \multicolumn{3}{c}{Cubic upwind-biased} & \multicolumn{3}{c}{Centred linear} & \multicolumn{3}{c}{Cubic upwind-biased} \\
+& \(\ell^2\) error & min & max & \(\ell^2\) error & min & max & \(\ell^2\) error & min & max & \(\ell^2\) error & min & max \\
 \hline
-Analytic  & 0 & 0 & 1 & 0 & 1 & 0 & 0 & 1 \\
-BTF 	  & \num{-0.058} & \num{1.001} & \num{0.00767} & \num{-0.0433} & \num{0.928} & \num{0.00154} & \num{-0.0105} & \num{0.983} \\
-SLEVE 	  & \num{-0.002} & \num{0.984} & \num{0.00108} & \num{-0.0106} & \num{0.981} & \num{0.0122} & \num{-0.0273} & \num{0.864} \\
-Cut cell  & \multicolumn{1}{c}{---} & \multicolumn{1}{c}{---} & \num{0.000579} & \num{-0.00670} & \num{0.982} & \num{0.0136} & \num{-0.0297} & \num{0.850} \\
-No orography & \num{-0.002} & \num{0.984} & \num{0.000577} & \num{-0.00670} & \num{0.982} & \multicolumn{1}{c}{---} & \multicolumn{1}{c}{---} & \multicolumn{1}{c}{---} \\
+Analytic  & 0 & 0 & 1 & 0 & 0 & 1 & 0 & 0 & 1 & 0 & 0 & 1 \\
+BTF 	  & \num{0.0318} & \num{-0.442} & \num{0.849} & \num{0.00767} & \num{-0.0433} & \num{0.928} & \num{0.00251} & \num{-0.0244} & \num{0.985} & \num{0.00154} & \num{-0.0105} & \num{0.983} \\
+SLEVE 	  & \num{0.00233} & \num{-0.0253} & \num{0.945} & \num{0.00108} & \num{-0.0106} & \num{0.981} & \num{0.0298} & \num{-0.258} & \num{0.865} & \num{0.0122} & \num{-0.0273} & \num{0.864} \\
+Cut cell  & \num{0.00224} & \num{-0.0253} & \num{0.985} & \num{0.000579} & \num{-0.00670} & \num{0.982} & \num{0.0318} & \num{-0.270} & \num{0.846} & \num{0.0136} & \num{-0.0297} & \num{0.850} \\
+No orography & \num{0.00224} & \num{-0.0253} & \num{0.985} & \num{0.000577} & \num{-0.00670} & \num{0.982} & \multicolumn{1}{c}{---} & \multicolumn{1}{c}{---} & \multicolumn{1}{c}{---} & \multicolumn{1}{c}{---} & \multicolumn{1}{c}{---} & \multicolumn{1}{c}{---} \\
 \hline
@@
 In agreement with \citet{klemp2011}, vertical velocities are larger on more distorted grids.
-However, magnitudes are smaller comparing results on the terrain following grids with those from \citet{klemp2011}, with $w$ reaching a maximum of \(\sim \SI{0.35}{\meter\per\second}\) on the BTF grid in our test compared with a maximum of \(\sim \SI{10}{\meter\per\second}\) found by \citet{klemp2011}.
+However, magnitudes are smaller comparing results on the terrain following grids with those from \citet{klemp2011}, with $w$ reaching a maximum of \(\sim \SI{0.35}{\meter\per\second}\) on the BTF grid in our test compared with a maximum of \(\sim \SI{10}{\meter\per\second}\) found by \citet{klemp2011} \TODO{NO, compare with Klemp's improved results}.
 

### 2015-04-27 msc-publication.tex
make explicit what scheme we're discussing in advection test analysis

@@
 The result of the centred linear scheme on the BTF grid (\ref{fig:advection-tracer}a) is compared with the upwind-biased cubic scheme on the BTF grid (\ref{fig:advection-tracer}b), and cut cell grid (\ref{fig:advection-tracer}c).
+
+By \(t = \SI{10000}{\second}\), the tracer suffers from diffusion and distortion on the BTF grid using the centred linear scheme.
+Some artefacts remain above the mountain peak, and a computational mode is seen as a grid-scale oscillation that travels in the opposite direction to the wind.
+
 Using the upwind-biased cubic scheme, tracer magnitude and shape are well-preserved on all grids, both above the mountain at \(t = \SI{5000}{\second}\) and past the mountain at \(t = \SI{10000}{\second}\).
@@
 
-The results of the cubic upwind-biased scheme on TF and regular grids are comparable with those for the fourth-order centred scheme from \citet{schaer2002} \TODO{we're not comparing with schaer any more}.
-Error is largest on the BTF grid with \(\ell^2 = \num{0.00767}\) but significant reduced on the SLEVE grid with \(\ell^2 = \num{0.00108}\).
+For the cubic upwind-biased scheme, error is largest on the BTF grid with \(\ell^2 = \num{0.00767}\) but significant reduced on the SLEVE grid with \(\ell^2 = \num{0.00108}\).
 The error is approximately halved by changing from the SLEVE grid to the cut cell grid.
@@
 	%
-	\caption{Errors in horizontal tracer advection on the BTF grid at \(t = \SI{10000}{\second}\) using (a) centred linear scheme, (b) the upwind-biased cubic scheme.
+	\caption{Errors in horizontal tracer advection on the BTF grid at \(t = \SI{10000}{\second}\) using (a) a centred linear scheme, (b) the upwind-biased cubic scheme.
 Contour intervals are every 0.01 with negative contours denoted by dashed lines.}
@@
 \footnotesize
-\begin{tabular}{ r @{\hspace{1em}} l l l @{\hspace{0.5em}} l l l @{\hspace{1em}} l l l @{\hspace{0.5em}} l l l}
+\begin{tabular}{ r @{\hspace{0.5em}} l l l @{\hspace{0.2em}} l l l @{\hspace{0.5em}} l l l @{\hspace{0.2em}} l l l}
 \hline\hline
@@
 Cut cell  & \num{0.00224} & \num{-0.0253} & \num{0.985} & \num{0.000579} & \num{-0.00670} & \num{0.982} & \num{0.0318} & \num{-0.270} & \num{0.846} & \num{0.0136} & \num{-0.0297} & \num{0.850} \\
-No orography & \num{0.00224} & \num{-0.0253} & \num{0.985} & \num{0.000577} & \num{-0.00670} & \num{0.982} & \multicolumn{1}{c}{---} & \multicolumn{1}{c}{---} & \multicolumn{1}{c}{---} & \multicolumn{1}{c}{---} & \multicolumn{1}{c}{---} & \multicolumn{1}{c}{---} \\
+No terrain & \num{0.00224} & \num{-0.0253} & \num{0.985} & \num{0.000577} & \num{-0.00670} & \num{0.982} & \multicolumn{1}{c}{---} & \multicolumn{1}{c}{---} & \multicolumn{1}{c}{---} & \multicolumn{1}{c}{---} & \multicolumn{1}{c}{---} & \multicolumn{1}{c}{---} \\
 \hline
@@
 In the terrain following velocity field, tracer accuracy is greatest on the BTF grid.
-Errors are about ten times larger on the SLEVE and cut cell grids compared to the BTF grid.
+Using the cubic upwind-biased scheme, errors are about ten times larger on the SLEVE and cut cell grids compared to the BTF grid.
 

### 2015-04-28 msc-publication.tex
more corrections from John, swap Klemp2011 fig2 for fig4

@@
 
-The results of the tracer advection test show that, given an advection scheme of a sufficiently high order, distortions in the grid do not significantly distort the tracer.
+The results of the tracer advection test show that numerical errors are due to misalignment of the flow with the grid.
+Given an advection scheme of a sufficiently high order, distortions in the grid do not significantly distort the tracer.
 
@@
 The horizontal position of the tracer centre can be calculated by integrating along the trajectory to find $t$, the time taken to pass from one side of the mountain to the other:
-\begin{multline}
-t = \frac{x}{u_0} - \frac{h_0}{16 u_0 H} \left[ 4x + \frac{\sin 2 (\alpha + \beta) x}{\alpha + \beta} \right.+ \\
-\left. \frac{\sin 2(\alpha - \beta) x}{\alpha - \beta} + 2 \left( \frac{\sin 2\alpha x}{\alpha} + \frac{\sin 2\beta x}{\beta} \right) \right]
-\end{multline}
+\begin{align}
+	\mathrm{d}t &= \mathrm{d}x / u(x) \\
+	t &= \int_0^x \frac{H - h(x)}{u_0 H}\:\mathrm{d}x \\
+	t &= \frac{x}{u_0} - \frac{h_0}{16 u_0 H} \left[ 4x + \frac{\sin 2 (\alpha + \beta) x}{\alpha + \beta} \right.+ \nonumber \\
+   &\ \left. \frac{\sin 2(\alpha - \beta) x}{\alpha - \beta} + 2 \left( \frac{\sin 2\alpha x}{\alpha} + \frac{\sin 2\beta x}{\beta} \right) \right]
+\end{align}
 Hence, we find that \(x(t=\SI{10000}{\second}) = \SI{51577.4}{\meter}\).
@@
 \footnotesize
-\begin{tabular}{ r @{\hspace{0.5em}} l l l @{\hspace{0.2em}} l l l @{\hspace{0.5em}} l l l @{\hspace{0.2em}} l l l}
+\begin{tabular}{ r @{\hspace{0.5em}} l l l | @{\hspace{0.2em}} l l l | @{\hspace{0.5em}} l l l | @{\hspace{0.2em}} l l l}
 \hline\hline
@@
 Cut cell  & \num{0.00224} & \num{-0.0253} & \num{0.985} & \num{0.000579} & \num{-0.00670} & \num{0.982} & \num{0.0318} & \num{-0.270} & \num{0.846} & \num{0.0136} & \num{-0.0297} & \num{0.850} \\
-No terrain & \num{0.00224} & \num{-0.0253} & \num{0.985} & \num{0.000577} & \num{-0.00670} & \num{0.982} & \multicolumn{1}{c}{---} & \multicolumn{1}{c}{---} & \multicolumn{1}{c}{---} & \multicolumn{1}{c}{---} & \multicolumn{1}{c}{---} & \multicolumn{1}{c}{---} \\
+No terrain & \num{0.00224} & \num{-0.0253} & \num{0.985} & \num{0.000577} & \num{-0.00670} & \num{0.982} & \multicolumn{1}{c}{---} & \multicolumn{1}{c}{---} & {---} & \multicolumn{1}{c}{---} & \multicolumn{1}{c}{---} & \multicolumn{1}{c}{---} \\
 \hline
@@
 We conclude from this test that accuracy depends upon alignment of the flow with the grid, and accuracy is not significantly reduced by grid distortions.
-Error on the BTF grid is comparable with the error on the SLEVE grid in the horizontal advection test.
+Error on the BTF grid in the terrain following advection test is comparable with the error on the SLEVE grid in the horizontal advection test.
 
@@
 
-The initial thermodynamic conditions are in discrete hydrostatic balance, having a surface temperature of $\theta_0 = \SI{288}{\kelvin}$ and constant stability with Brunt-V\"ais\"al\"a frequency $N = \SI{0.01}{\per\second}$ everywhere, except for a more stable layer of $N = \SI{0.02}{\per\second}$ between $\SI{2}{\kilo\meter} \leq z \leq \SI{3}{\kilo\meter}$.
+The initial thermodynamic conditions are in discrete hydrostatic balance, having a reference potential temperature of $\theta(z = 0) = \SI{288}{\kelvin}$ and constant stability with Brunt-V\"ais\"al\"a frequency $N = \SI{0.01}{\per\second}$ everywhere, except for a more stable layer of $N = \SI{0.02}{\per\second}$ between $\SI{2}{\kilo\meter} \leq z \leq \SI{3}{\kilo\meter}$.
 Unlike \citet{klemp2011}, there is no eddy diffusion in the equation set.
@@
 In agreement with \citet{klemp2011}, vertical velocities are larger on more distorted grids.
-However, magnitudes are smaller comparing results on the terrain following grids with those from \citet{klemp2011}, with $w$ reaching a maximum of \(\sim \SI{0.35}{\meter\per\second}\) on the BTF grid in our test compared with a maximum of \(\sim \SI{10}{\meter\per\second}\) found by \citet{klemp2011} \TODO{NO, compare with Klemp's improved results}.
+However, magnitudes are smaller comparing results on the terrain following grids with those from \citet{klemp2011}.
+Using the model from \citet{weller-shahrokhi2014}, which includes a curl-free pressure gradient formulation, $w$ reaches a maximum of \(\sim \SI{0.35}{\meter\per\second}\) on the BTF grid, compared with a maximum of \(\sim \SI{7}{\meter\per\second}\) found by \citet{klemp2011} using their improved horizontal pressure gradient formulation.
 
@@
 %
-	\caption{Maximum spurious vertical velocity, \(w\) (\si{\meter\per\second}), in the resting atmosphere test with results on (a) BTF, SLEVE, Hybrid Terrain Following (HTF) and Smoothed Terrain Following (STF) coordinates from \citet{klemp2011}, (b) BTF, SLEVE, cut cell and regular grids using the model from \citet{weller-shahrokhi2014}.
+	\caption{Maximum spurious vertical velocity, \(w\) (\si{\meter\per\second}), in the resting atmosphere test with results on (a) BTF, SLEVE, Hybrid Terrain Following (HTF) and Smoothed Terrain Following (STF) coordinates from \citet{klemp2011} using their improved horizontal pressure gradient formulation, (b) BTF, SLEVE, cut cell and regular grids using the model from \citet{weller-shahrokhi2014} which includes a curl-free pressure gradient formulation.
 Note that vertical scales differ.}
@@
 
-The initial thermodynamic conditions have a surface temperature of $\theta_0 = \SI{288}{\kelvin}$ and constant stability with $N = \SI{0.01}{\per\second}$ everywhere.
+The initial thermodynamic conditions have a surface temperature of $\theta(z=0) = \SI{288}{\kelvin}$ and constant stability with $N = \SI{0.01}{\per\second}$ everywhere.
 A constant horizontal wind $u = \SI{10}{\meter\per\second}$ is prescribed at the inlet boundary.
@@
 
-No normal flow is imposed at the top and bottom boundaries and the outlet is zero gradient.
-For Exner, hydrostatic balance is prescribed on all boundaries.
+No normal flow is imposed at the top and bottom boundaries and the outlet has a zero gradient boundary condition.
+For the Exner function of pressure, hydrostatic balance is prescribed on all boundaries.
 The simulation is integrated forward by 5 hours with a timestep $\Delta t = \SI{8}{\second}$.

### 2015-04-28 msc-publication.tex
Describe RK2 scheme and other corrections

@@
 Representing orography accurately in numerical weather prediction systems is necessary to model downslope winds and local precipitation.
+Orography also exerts strong non-local influences by the atmospheric response to form drag and gravity wave drag.
 There are two main approaches to represent orography on a grid: terrain following layers and cut cells.
@@
 
+The hybrid terrain following (HTF) coordinates of \citet{simmons-burridge1981} improve upon BTF coordinates by using a decay function that allows the influence of the terrain to be removed at a specified height, producing flat model layers aloft.
+\TODO{I could mention that this is used operationally in UK MO (davies et al 2005) and ECMWF (good citation needed)}
+
 The smooth level vertical (SLEVE) coordinate proposed by \citet{schaer2002} achieves a more regular TF grid in the middle and top of the domain than the BTF coordinate.
@@
 Second, the surface faces were taken from the BTF grid and \texttt{snappyHexMesh} was used to intersect the surface with the grid.
-The tool removes cells whose centres are below the surface.
+The tool removes cells whose centres are below the surface and displaces boundary vertices so that they lie on the BTF surface.
 An example of the resulting grid is shown in figure~\ref{fig:grid}c.
@@
 There are two details of grid construction which mean that resulting cut cell grids can differ slightly from a typical cut cell grid created using a shaving method, as described by \citet{adcroft1997}.
-First, when \texttt{snappyHexMesh} moves points along the surface according to its heuristics, some points are moved horizontally.
+First, when \texttt{snappyHexMesh} moves vertices onto the terrain surface, some points are moved horizontally.
 Second, the utility does not create new points necessary for pentagonal cells.
@@
 The flux form of the advection equation, \(\partial \phi / \partial t + \nabla \cdot \left( \bm{u} \phi \right) = 0\), is solved using an upwind-biased cubic advection scheme which is non-monotonic and not flux corrected.
-The time derivative is solved using a second order Runge-Kutta scheme.
+The time derivative is solved using a three-stage, second order Runge-Kutta scheme:
+\begin{align}
+	\varphi^\star &= \varphi^{(n)} + \Delta t f(\phi^{(n)}) \\
+	\varphi^{\star\star} &= \varphi^{(n)} + \frac{\Delta t}{2} \left( f(\phi^{(n)}) + f(\phi^\star) \right) \\
+	\varphi^{(n+1)} &= \varphi^{(n)} + \frac{\Delta t}{2} \left( f(\phi^{(n)}) + f(\phi^{\star\star}) \right)
+\end{align}
+where \(f(\phi^{(n)}) = - \mathrm{div}(\bm{u} \varphi^{(n)})\) at time level \(n\).
 

### 2015-04-29 msc-publication.tex
Add citations for OpenFOAM & add2dMountain

@@
 In the gravity waves test, results on all grids are in good agreement with existing results from the literature.
-However, we find that the Lorenz computational mode is excited as a vertical zig--zag in potential temperature on the cut cell grid.
-\TODO{this last sentence is weak: Although errors do vary between grids, overall, a high degree of accuracy is obtained for all tests on all grids.}}
+However, we find that the Lorenz computational mode is excited as a vertical zig--zag in potential temperature on the cut cell grid.}
 
@@
 where, in two dimensions, \(z(x, z^\star)\) is the physical height of the coordinate surface at level \(z^\star\), \(H\) is the height of the domain, and \(h(x)\) is the height of the terrain surface.
+This formulation results in $z^\star/H$ ranging from 0 to 1.
 Using this coordinate, the terrain's influence decays linearly with height but disappears only at the top of the domain.
@@
 The hybrid terrain following (HTF) coordinates of \citet{simmons-burridge1981} improve upon BTF coordinates by using a decay function that allows the influence of the terrain to be removed at a specified height, producing flat model layers aloft.
-\TODO{I could mention that this is used operationally in UK MO (davies et al 2005) and ECMWF (good citation needed)}
+\TODO{include equation?}
 
@@
 First, a custom utility was used to move points beneath the surface up to the surface creating small cells near mountain peaks.
+The utility is available at \url{https://github.com/hertzsprung/AtmosFOAM/tree/2386ffb/applications/utilities/mesh/add2dMountain}.
 Second, the surface faces were taken from the BTF grid and \texttt{snappyHexMesh} was used to intersect the surface with the grid.
-The tool removes cells whose centres are below the surface and displaces boundary vertices so that they lie on the BTF surface.
+The tool removes cells whose centres are below the surface and displaces boundary vertices so that they are `snapped' to the BTF surface \citep{openfoam2015}.
 An example of the resulting grid is shown in figure~\ref{fig:grid}c.
@@
 
-The flux form of the advection equation, \(\partial \phi / \partial t + \nabla \cdot \left( \bm{u} \phi \right) = 0\), is solved using an upwind-biased cubic advection scheme which is non-monotonic and not flux corrected.
+The flux form of the advection equation, \(\partial \phi / \partial t + \nabla \cdot \left( \bm{u} \phi \right) = 0\), is solved using the upwind-biased cubic advection scheme described by \cite{weller-shahrokhi2014} which is non-monotonic and not flux corrected.
 The time derivative is solved using a three-stage, second order Runge-Kutta scheme:
@@
 \footnotesize
-\begin{tabular}{ r @{\hspace{0.5em}} l l l | @{\hspace{0.2em}} l l l | @{\hspace{0.5em}} l l l | @{\hspace{0.2em}} l l l}
+\begin{tabular}{ r @{\hspace{0.5em}} l l l | l l l | l l l | l l l}
 \hline\hline
@@
 	\caption{Vertical cross section of vertical velocity contours in the gravity waves test after 5 hours with mountain height $h_0 = \SI{250}{\meter}$ on (a) the BTF grid compared with (b) the mass-conserving semi-implicit semi-Lagrangian solution from \citet{melvin2010}.
-Contours are every \SI{5e-2}{\meter\per\second} with solid lines denoting ascent and dashed lines descent.}
+Contours are every \SI{5e-2}{\meter\per\second} with solid lines denoting ascent and dashed lines descent.
+Small oscillations in contour lines are due to contouring on an irregular grid.}
 	\label{fig:gw-w}

### 2015-04-29 msc-publication.tex
Fix table alignment to fit in margins
Avoid saying 'sufficiently high order'

@@
 The hybrid terrain following (HTF) coordinates of \citet{simmons-burridge1981} improve upon BTF coordinates by using a decay function that allows the influence of the terrain to be removed at a specified height, producing flat model layers aloft.
-\TODO{include equation?}
 
@@
 The results of the tracer advection test show that numerical errors are due to misalignment of the flow with the grid.
-Given an advection scheme of a sufficiently high order, distortions in the grid do not significantly distort the tracer.
+Using the upwind-biased cubic scheme, distortions in the grid do not significantly distort the tracer.
 
@@
 \footnotesize
-\begin{tabular}{ r @{\hspace{0.5em}} l l l | l l l | l l l | l l l}
+\begin{tabular}{ r @{\hspace{1.0em}} l @{\hspace{1.0em}} l @{\hspace{1.0em}} l | l @{\hspace{1.0em}} l @{\hspace{1.0em}} l | l @{\hspace{1.0em}} l @{\hspace{1.0em}} l | l @{\hspace{1.0em}} l @{\hspace{1.0em}} l}
 \hline\hline

### 2015-05-08 msc-publication.tex
fix streamfunction, add acknowledgements

@@
 
-\title{Comparison of Terrain Following and Cut Cell Grids with a Non-Hydrostatic Model}
-\authors{James Shaw\correspondingauthor{Department of Meteorology, University of Reading, Earley Gate, PO Box 243, Reading, RG6 6BB, UK.}}
-\affiliation{}
-\email{j.shaw@pgr.reading.ac.uk}
+\title{Comparison of Terrain Following and Cut Cell Grids using a Non-Hydrostatic Model}
+\authors{James Shaw\correspondingauthor{Department of Meteorology, University of Reading, Earley Gate, PO Box 243, Reading, RG6 6BB, UK.} and Hilary Weller}
+\affiliation{Department of Meteorology, University of Reading, Reading, United Kingdom}
+\email{js102@zepler.net}
 
@@
 
-The domain is \SI{300}{\kilo\meter} wide and \SI{25}{\kilo\meter} high, discretized onto a grid with \(\Delta x = \SI{1}{\kilo\meter}\) and \(\Delta z^\star = \SI{500}{\meter}\).
+\TODO{take care to explain domain and grid because of our previous differences with Schaer's results} The domain is \SI{300}{\kilo\meter} wide and \SI{25}{\kilo\meter} high, discretized onto a grid with \(\Delta x = \SI{1}{\kilo\meter}\) and \(\Delta z^\star = \SI{500}{\meter}\).
 The terrain is wave-shaped, specified by the surface height, \(h\), such that
@@
 To ensure the discrete velocity field is non-divergent, velocities are prescribed at cell faces by summing the streamfunction, \(\Psi\), at the centre of all face edges.
-Since \(u = \partial \Psi / \partial z\), the streamfunction is found by vertical integration of the velocity profile:
+Since \(u = - \partial \Psi / \partial z\), the streamfunction is found by vertical integration of the velocity profile:
 \begin{align}
@@
 Tracer contours at \(t = \SI{0}{\second}, \SI{5000}{\second}\) and \(\SI{10000}{\second}\) are shown in Figure~\ref{fig:advection-tracer}.
-We do not compare with the results from \citet{schaer2002} because we were unable to reproduce their result using a centred linear scheme (their figure 6a), though similar results were obtained by halving the mountain height, \(h_0\).
 
@@
 \begin{equation}
-	\Psi(x,z) = u_0 H \frac{z - h}{H - h}
+	\Psi(x,z) = - u_0 H \frac{z - h}{H - h}
 \end{equation}
@@
 \begin{align}
-	u &= \frac{\partial \Psi}{\partial z} = u_0 \frac{H}{H - h}, \quad w = -\frac{\partial \Psi}{\partial x} = u_0 H \frac{\mathrm{d} h}{\mathrm{d} x} \frac{H - z}{\left( H - h \right)^2}, \nonumber \\
+	u &= - \frac{\partial \Psi}{\partial z} = - u_0 \frac{H}{H - h}, \quad w = \frac{\partial \Psi}{\partial x} = u_0 H \frac{\mathrm{d} h}{\mathrm{d} x} \frac{z - H}{\left( H - h \right)^2}, \nonumber \\
 	\frac{\partial h}{\partial x} &= - h_0 \left[ 
@@
 Analytic  & 0 & 0 & 1 & 0 & 0 & 1 & 0 & 0 & 1 & 0 & 0 & 1 \\
-BTF 	  & \num{0.0318} & \num{-0.442} & \num{0.849} & \num{0.00767} & \num{-0.0433} & \num{0.928} & \num{0.00251} & \num{-0.0244} & \num{0.985} & \num{0.00154} & \num{-0.0105} & \num{0.983} \\
-SLEVE 	  & \num{0.00233} & \num{-0.0253} & \num{0.945} & \num{0.00108} & \num{-0.0106} & \num{0.981} & \num{0.0298} & \num{-0.258} & \num{0.865} & \num{0.0122} & \num{-0.0273} & \num{0.864} \\
-Cut cell  & \num{0.00224} & \num{-0.0253} & \num{0.985} & \num{0.000579} & \num{-0.00670} & \num{0.982} & \num{0.0318} & \num{-0.270} & \num{0.846} & \num{0.0136} & \num{-0.0297} & \num{0.850} \\
-No terrain & \num{0.00224} & \num{-0.0253} & \num{0.985} & \num{0.000577} & \num{-0.00670} & \num{0.982} & \multicolumn{1}{c}{---} & \multicolumn{1}{c}{---} & {---} & \multicolumn{1}{c}{---} & \multicolumn{1}{c}{---} & \multicolumn{1}{c}{---} \\
+BTF 	  & \num{999} & \num{999} & \num{999} & \num{999} & \num{999} & \num{9999} & \num{999} & \num{999} & \num{999} & \num{999} & \num{999} & \num{999} \\
+SLEVE & \num{999} & \num{999} & \num{999} & \num{999} & \num{999} & \num{9999} & \num{999} & \num{999} & \num{999} & \num{999} & \num{999} & \num{999} \\
+Cut cell & \num{999} & \num{999} & \num{999} & \num{999} & \num{999} & \num{9999} & \num{999} & \num{999} & \num{999} & \num{999} & \num{999} & \num{999} \\
+No terrain & \num{999} & \num{999} & \num{999} & \num{999} & \num{999} & \num{999} & \multicolumn{1}{c}{---} & \multicolumn{1}{c}{---} & {---} & \multicolumn{1}{c}{---} & \multicolumn{1}{c}{---} & \multicolumn{1}{c}{---} \\
 \hline
@@
 \acknowledgments
-\TODO{John, Terry. Assuming only Hilary is co-author...?} Weller is funded by NERC grant NE/H015698/1.
+I am grateful to my cosupervisors John Methven and Terry Davies for their valuable input, and to Christoph Sch\"{a}r for his assistance in reproducing his advection test results.
+I am thankful for the NERC studentship which helped fund this work.
+Weller is funded by NERC grant NE/H015698/1.
 

### 2015-05-15 msc-publication.tex
update table of advection stats, discuss domain differences and interp of z for calculating psi

@@
 
-\TODO{take care to explain domain and grid because of our previous differences with Schaer's results} The domain is \SI{300}{\kilo\meter} wide and \SI{25}{\kilo\meter} high, discretized onto a grid with \(\Delta x = \SI{1}{\kilo\meter}\) and \(\Delta z^\star = \SI{500}{\meter}\).
+The domain is \SI{301}{\kilo\meter} wide and \SI{25}{\kilo\meter} high, discretized onto a grid with \(\Delta x = \SI{1}{\kilo\meter}\) and \(\Delta z^\star = \SI{500}{\meter}\).
+The domain specified by \citet{schaer2002} is \SI{300}{\kilo\meter} between the outermost cell centres.
+In order to reproduce the result from \citet{schaer2002}, the domain has been extended horizontally by \(\Delta x/2\) \si{\meter} in both directions so that the distance between the outermost cell centres is still \SI{300}{\kilo\meter}.
+
 The terrain is wave-shaped, specified by the surface height, \(h\), such that
@@
 By \(t = \SI{10000}{\second}\), the tracer suffers from diffusion and distortion on the BTF grid using the centred linear scheme.
-Some artefacts remain above the mountain peak, and a computational mode is seen as a grid-scale oscillation that travels in the opposite direction to the wind.
+Some artefacts remain above the mountain peak, and the computational mode is seen as a grid-scale oscillation that travels in the opposite direction to the wind.
 
@@
 
-For the cubic upwind-biased scheme, error is largest on the BTF grid with \(\ell^2 = \num{0.00767}\) but significant reduced on the SLEVE grid with \(\ell^2 = \num{0.00108}\).
+The results of the cubic upwind-biased scheme on TF and regular grids are comparable with those for the fourth-order centred scheme from \citet{schaer2002}.
+Error is largest on the BTF grid with \(\ell^2 = \num{0.00791}\) but significant reduced on the SLEVE grid with \(\ell^2 = \num{0.00108}\).
 The error is approximately halved by changing from the SLEVE grid to the cut cell grid.
+
+Tracer minima and maxima are lower than values presented by \citet{schaer2002} because we do not interpolate the geometric height of cell faces from adjacent cell centres in order to calculate the streamfunction, \(\Psi(z)\).
+Removing the interpolation from Sch\"{a}r's code results in minima and maxima that are equal to those obtained from the OpenFOAM advection solver presented here.
+
+The centred linear scheme is sensitive to changes in the terrain profile: when the domain width is reduced to \SI{300}{\kilo\meter}, the $\ell^2$ error on the BTF grid rises from \num{0.0210} to \num{0.0318}.
+The upwind-biased cubic scheme is less sensitive to the same change, with the $\ell^2$ error decreasing from \num{0.00791} to \num{0.00767}.
 
@@
 Analytic   & 0 & 0 & 1 & 0 & 0 & 1 & 0 & 0 & 1 & 0 & 0 & 1 \\
-BTF 	  & \num{999} & \num{999} & \num{999} & \num{999} & \num{999} & \num{9999} & \num{999} & \num{999} & \num{999} & \num{999} & \num{999} & \num{999} \\
-SLEVE & \num{999} & \num{999} & \num{999} & \num{999} & \num{999} & \num{9999} & \num{999} & \num{999} & \num{999} & \num{999} & \num{999} & \num{999} \\
-Cut cell & \num{999} & \num{999} & \num{999} & \num{999} & \num{999} & \num{9999} & \num{999} & \num{999} & \num{999} & \num{999} & \num{999} & \num{999} \\
-No terrain & \num{999} & \num{999} & \num{999} & \num{999} & \num{999} & \num{999} & \multicolumn{1}{c}{---} & \multicolumn{1}{c}{---} & {---} & \multicolumn{1}{c}{---} & \multicolumn{1}{c}{---} & \multicolumn{1}{c}{---} \\
+% 	     l2		     min	     max	   l2		    min		     max           l2                        min                       max          l2                         min                       max
+BTF 	   & \num{0.0210}  & \num{-0.275}  & \num{0.925} & \num{0.00791}  & \num{-0.0446}  & \num{0.925} & \num{0.00251}           & \num{-0.0245}           & \num{0.985} & \num{0.00154}           & \num{-0.0110}           & \num{0.983} \\
+SLEVE      & \num{0.00233} & \num{-0.0252} & \num{0.985} & \num{0.00108}  & \num{-0.0106}  & \num{0.982} & \num{0.0173}            & \num{-0.120}            & \num{0.950} & \num{0.0120}            & \num{-0.0263}           & \num{0.865} \\
+Cut cell   & \num{0.00224} & \num{-0.0251} & \num{0.985} & \num{0.000577} & \num{-0.00674} & \num{0.983} & \num{13.6}              & \num{-623}              & \num{3480}  & \num{0.0134}            & \num{-0.028}            & \num{0.851} \\
+No terrain & \num{0.00223} & \num{-0.0251} & \num{0.985} & \num{0.000576} & \num{-0.00674} & \num{0.983} & \multicolumn{1}{c}{---} & \multicolumn{1}{c}{---} & {---}       & \multicolumn{1}{c}{---} & \multicolumn{1}{c}{---} & \multicolumn{1}{c}{---} \\
 \hline

### 2015-05-15 msc-publication.tex
Merge tracer and error contours into single figure
Added 4th order plots from schaer's code, just need to rearrange
everything to fit on a 3x2 grid of panels.

@@
 
-\begin{figure}
+\begin{figure*}
 	\centering
@@
 %
-	\caption{Horizontally advected tracer contours at \(t = \SI{0}{\second}\), \SI{5000}{\second} and \SI{10000}{\second} using (a) centred linear scheme on BTF grid, and the upwind-biased cubic scheme on (b) BTF grid, (c) cut cell grid.
-Contour intervals are every 0.1.
+	\caption{Horizontally advected tracer contours at \(t = \SI{0}{\second}\), \SI{5000}{\second} and \SI{10000}{\second} using (a) centred linear scheme on BTF grid, the upwind-biased cubic scheme on (b) cut cell grid, (d) BTF grid, and (c) the fourth order scheme from \citet{schaer2002}, with contour intervals every 0.1.
+Errors on the BTF grid at \(t = \SI{10000}{\second}\) are shown for (e) the fourth order scheme from \citet{schaer2002}, and (f) the upwind-biased cubic scheme, with contour intervals every 0.01.
+Negative contours denoted by dotted lines.
 The terrain profile is also shown immediately above the $x$ axis.}
 	\label{fig:advection-tracer}
-\end{figure}
-
-\begin{figure}
-	\centering
-	\includegraphics{../msc-publication-fig-advection-error/fig-advection-error.pdf}
-	%
-	\caption{Errors in horizontal tracer advection on the BTF grid at \(t = \SI{10000}{\second}\) using (a) a centred linear scheme, (b) the upwind-biased cubic scheme.
-Contour intervals are every 0.01 with negative contours denoted by dashed lines.}
-	\label{fig:advection-error}
-\end{figure}
+\end{figure*}
 

### 2015-05-18 fig-advection-tracer.tex
Merge tracer and error contours into single figure
Added 4th order plots from schaer's code, just need to rearrange
everything to fit on a 3x2 grid of panels.
Plot six panels for advection tracer/error contour

--- 2015-02-02 fig-advection-tracer.tex
@@
 \documentclass{article}
-\usepackage[paperwidth=3in,paperheight=4.6in,margin=0in]{geometry}
+\usepackage[paperwidth=6.4in,paperheight=4.6in,margin=0in]{geometry}
 \usepackage{mathptmx}%{times}
@@
 \begin{document}
-\hspace{0.15in}\includegraphics[width=2.75in]{img/schaer-btf-4thorder.png}
-
-\footnotesize
-\input{../advectTracer-horizontal-btf-cubicUpwindCPCFit-divFree/tracer-contour-y-only}
-
-\input{../advectTracer-horizontal-snapCol-cubicUpwindCPCFit-divFree/tracer-contour}
+\input{../advection-horizontal-contours/contours}
 \end{document}

### 2015-05-18 msc-publication.tex
Clarify tracer contour figure caption

@@
 %
-	\caption{Horizontally advected tracer contours at \(t = \SI{0}{\second}\), \SI{5000}{\second} and \SI{10000}{\second} using (a) centred linear scheme on BTF grid, the upwind-biased cubic scheme on (b) cut cell grid, (d) BTF grid, and (c) the fourth order scheme from \citet{schaer2002}, with contour intervals every 0.1.
+	\caption{Horizontally advected tracer contours at \(t = \SI{0}{\second}\), \SI{5000}{\second} and \SI{10000}{\second} using (a) centred linear scheme on a BTF grid, the upwind-biased cubic scheme on (b) the cut cell grid, (d) the BTF grid, and (c) the fourth order scheme from \citet{schaer2002} on the BTF grid, with contour intervals every 0.1.
 Errors on the BTF grid at \(t = \SI{10000}{\second}\) are shown for (e) the fourth order scheme from \citet{schaer2002}, and (f) the upwind-biased cubic scheme, with contour intervals every 0.01.

### 2015-05-18 msc-publication.tex
try splitting table into two tables

@@
 
+\begin{table*}[t]
+	\caption{Minimum and maximum tracer magnitudes and \(\ell^2\) error norms at \(t = \SI{10000}{\second}\) in the horizontal tracer advection tests using centred linear and cubic upwind-biased schemes, and the fourth order scheme from \citet{schaer2002}.}
+\label{tab:advection}
+%
+\centering
+\footnotesize
+\begin{tabular}{ r @{\hspace{1.0em}} l @{\hspace{1.0em}} l @{\hspace{1.0em}} l l @{\hspace{1.0em}} l @{\hspace{1.0em}} l}
+\hline\hline
+& \multicolumn{3}{c}{Centred linear} & \multicolumn{3}{c}{Cubic upwind-biased} \\
+& \(\ell^2\) error & min & max & \(\ell^2\) error & min & max \\
+\hline
+Analytic   & 0 & 0 & 1 & 0 & 0 & 1 \\
+% 	     l2		     min	     max	   l2		    min		     max        
+BTF 	   & \num{0.0210}  & \num{-0.275}  & \num{0.925} & \num{0.00791}  & \num{-0.0446}  & \num{0.925} \\
+SLEVE      & \num{0.00233} & \num{-0.0252} & \num{0.985} & \num{0.00108}  & \num{-0.0106}  & \num{0.982} \\
+Cut cell   & \num{0.00224} & \num{-0.0251} & \num{0.985} & \num{0.000577} & \num{-0.00674} & \num{0.983} \\
+No terrain & \num{0.00223} & \num{-0.0251} & \num{0.985} & \num{0.000576} & \num{-0.00674} & \num{0.983} \\
+\hline
+\end{tabular}
+\end{table*}
+
 \subsection{Terrain following advection}
@@
 \begin{table*}[t]
-	\caption{Minimum and maximum tracer magnitudes and \(\ell^2\) error norms at \(t = \SI{10000}{\second}\) in the horizontal and terrain following tracer advection tests using centred linear and cubic upwind-biased schemes.}
+	\caption{Minimum and maximum tracer magnitudes and \(\ell^2\) error norms at \(t = \SI{10000}{\second}\) in the terrain following tracer advection test using centred linear and cubic upwind-biased schemes.}
 \label{tab:advection}
@@
 \footnotesize
-\begin{tabular}{ r @{\hspace{1.0em}} l @{\hspace{1.0em}} l @{\hspace{1.0em}} l | l @{\hspace{1.0em}} l @{\hspace{1.0em}} l | l @{\hspace{1.0em}} l @{\hspace{1.0em}} l | l @{\hspace{1.0em}} l @{\hspace{1.0em}} l}
+\begin{tabular}{ r @{\hspace{1.0em}} l @{\hspace{1.0em}} l @{\hspace{1.0em}} l l @{\hspace{1.0em}} l @{\hspace{1.0em}} l}
 \hline\hline
-& \multicolumn{6}{c}{Horizontal advection} & \multicolumn{6}{c}{Terrain following advection} \\
-& \multicolumn{3}{c}{Centred linear} & \multicolumn{3}{c}{Cubic upwind-biased} & \multicolumn{3}{c}{Centred linear} & \multicolumn{3}{c}{Cubic upwind-biased} \\
-& \(\ell^2\) error & min & max & \(\ell^2\) error & min & max & \(\ell^2\) error & min & max & \(\ell^2\) error & min & max \\
+& \multicolumn{3}{c}{Centred linear} & \multicolumn{3}{c}{Cubic upwind-biased} \\
+& \(\ell^2\) error & min & max & \(\ell^2\) error & min & max \\
 \hline
-Analytic   & 0 & 0 & 1 & 0 & 0 & 1 & 0 & 0 & 1 & 0 & 0 & 1 \\
-% 	     l2		     min	     max	   l2		    min		     max           l2                        min                       max          l2                         min                       max
-BTF 	   & \num{0.0210}  & \num{-0.275}  & \num{0.925} & \num{0.00791}  & \num{-0.0446}  & \num{0.925} & \num{0.00251}           & \num{-0.0245}           & \num{0.985} & \num{0.00154}           & \num{-0.0110}           & \num{0.983} \\
-SLEVE      & \num{0.00233} & \num{-0.0252} & \num{0.985} & \num{0.00108}  & \num{-0.0106}  & \num{0.982} & \num{0.0173}            & \num{-0.120}            & \num{0.950} & \num{0.0120}            & \num{-0.0263}           & \num{0.865} \\
-Cut cell   & \num{0.00224} & \num{-0.0251} & \num{0.985} & \num{0.000577} & \num{-0.00674} & \num{0.983} & \num{13.6}              & \num{-623}              & \num{3480}  & \num{0.0134}            & \num{-0.028}            & \num{0.851} \\
-No terrain & \num{0.00223} & \num{-0.0251} & \num{0.985} & \num{0.000576} & \num{-0.00674} & \num{0.983} & \multicolumn{1}{c}{---} & \multicolumn{1}{c}{---} & {---}       & \multicolumn{1}{c}{---} & \multicolumn{1}{c}{---} & \multicolumn{1}{c}{---} \\
+Analytic   & 0 & 0 & 1 & 0 & 0 & 1 \\
+% 	     l2                        min                       max          l2                         min                       max
+BTF 	   & \num{0.00251}           & \num{-0.0245}           & \num{0.985} & \num{0.00154}           & \num{-0.0110}           & \num{0.983} \\
+SLEVE      & \num{0.0173}            & \num{-0.120}            & \num{0.950} & \num{0.0120}            & \num{-0.0263}           & \num{0.865} \\
+Cut cell   & \num{13.6}              & \num{-623}              & \num{3480}  & \num{0.0134}            & \num{-0.028}            & \num{0.851} \\
 \hline

### 2015-05-18 msc-publication.tex
Merge back into one table, rotate to make data fit

@@
 
-\begin{table*}[t]
-	\caption{Minimum and maximum tracer magnitudes and \(\ell^2\) error norms at \(t = \SI{10000}{\second}\) in the horizontal tracer advection tests using centred linear and cubic upwind-biased schemes, and the fourth order scheme from \citet{schaer2002}.}
-\label{tab:advection}
-%
-\centering
-\footnotesize
-\begin{tabular}{ r @{\hspace{1.0em}} l @{\hspace{1.0em}} l @{\hspace{1.0em}} l l @{\hspace{1.0em}} l @{\hspace{1.0em}} l}
-\hline\hline
-& \multicolumn{3}{c}{Centred linear} & \multicolumn{3}{c}{Cubic upwind-biased} \\
-& \(\ell^2\) error & min & max & \(\ell^2\) error & min & max \\
-\hline
-Analytic   & 0 & 0 & 1 & 0 & 0 & 1 \\
-% 	     l2		     min	     max	   l2		    min		     max        
-BTF 	   & \num{0.0210}  & \num{-0.275}  & \num{0.925} & \num{0.00791}  & \num{-0.0446}  & \num{0.925} \\
-SLEVE      & \num{0.00233} & \num{-0.0252} & \num{0.985} & \num{0.00108}  & \num{-0.0106}  & \num{0.982} \\
-Cut cell   & \num{0.00224} & \num{-0.0251} & \num{0.985} & \num{0.000577} & \num{-0.00674} & \num{0.983} \\
-No terrain & \num{0.00223} & \num{-0.0251} & \num{0.985} & \num{0.000576} & \num{-0.00674} & \num{0.983} \\
-\hline
-\end{tabular}
-\end{table*}
-
 \subsection{Terrain following advection}
@@
 \begin{table*}[t]
-	\caption{Minimum and maximum tracer magnitudes and \(\ell^2\) error norms at \(t = \SI{10000}{\second}\) in the terrain following tracer advection test using centred linear and cubic upwind-biased schemes.}
+	\caption{Minimum and maximum tracer magnitudes and \(\ell^2\) error norms at \(t = \SI{10000}{\second}\) in the horizontal and terrain following tracer advection tests using centred linear and cubic upwind-biased schemes and, for the horizontal advection test, the fourth order scheme from \citet{schaer2002}.}
 \label{tab:advection}
@@
 \footnotesize
-\begin{tabular}{ r @{\hspace{1.0em}} l @{\hspace{1.0em}} l @{\hspace{1.0em}} l l @{\hspace{1.0em}} l @{\hspace{1.0em}} l}
+\begin{tabular}{l l l l S S S S}
 \hline\hline
-& \multicolumn{3}{c}{Centred linear} & \multicolumn{3}{c}{Cubic upwind-biased} \\
-& \(\ell^2\) error & min & max & \(\ell^2\) error & min & max \\
+                  &                     &                  & Analytic & {BTF}   & {SLEVE}         & {Cut cell}     & {No terrain} \\
 \hline
-Analytic   & 0 & 0 & 1 & 0 & 0 & 1 \\
-% 	     l2                        min                       max          l2                         min                       max
-BTF 	   & \num{0.00251}           & \num{-0.0245}           & \num{0.985} & \num{0.00154}           & \num{-0.0110}           & \num{0.983} \\
-SLEVE      & \num{0.0173}            & \num{-0.120}            & \num{0.950} & \num{0.0120}            & \num{-0.0263}           & \num{0.865} \\
-Cut cell   & \num{13.6}              & \num{-623}              & \num{3480}  & \num{0.0134}            & \num{-0.028}            & \num{0.851} \\
+Horizontal        & Centred linear      & \(\ell^2\) error & 0        & 0.0210  & 0.00233         & 0.00224        & 0.00223     \\
+                  &                     & min              & 0        & -0.275  & -0.0252         & -0.0251        & -0.0251     \\
+                  &                     & max              & 1        & 0.925   & 0.985           & 0.985          & 0.985       \\
+                  & Fourth order        & \(\ell^2\) error & 0        & 999     & 999             & {---}          & 999         \\
+                  &                     & min              & 0        & 999     & 999             & {---}          & 999         \\
+                  &                     & max              & 1        & 999     & 999             & {---}          & 999         \\
+                  & Cubic upwind-biased & \(\ell^2\) error & 0        & 0.00791 & 0.00108         & 0.000577       & 0.000576    \\
+                  &                     & min              & 0        & -0.0446 & -0.0106         & -0.000674      & -0.00674    \\
+                  &                     & max              & 1        & 0.925   & 0.982           & 0.983          & 0.983       \\
+\hline
+Terrain following & Centred linear      & \(\ell^2\) error & 0        & 0.00251 & 0.0173          & 13.6           & {---}        \\
+		  &                     & min              & 0        & -0.0245 & -0.120          & -623           & {---}        \\
+                  &                     & max              & 1        & 0.985   & 0.950           & 3480           & {---}        \\
+		  & Cubic upwind-biased & \(\ell^2\) error & 0        & 0.00154 & 0.0120          & 0.0134         & {---}        \\
+                  &                     & min              & 0        & -0.0110 & -0.0263         & -0.028         & {---}        \\
+                  &                     & max              & 1        & 0.983   & 0.865           & 0.851          & {---}        \\
 \hline

### 2015-05-18 msc-publication.tex
add 4th order data

@@
                   &                     & max              & 1        & 0.925   & 0.985           & 0.985          & 0.985       \\
-                  & Fourth order        & \(\ell^2\) error & 0        & 999     & 999             & {---}          & 999         \\
-                  &                     & min              & 0        & 999     & 999             & {---}          & 999         \\
-                  &                     & max              & 1        & 999     & 999             & {---}          & 999         \\
+                  & Fourth order        & min              & 0        & -0.0926 & -0.00174        & {---}          & {---}       \\
+                  &                     & max              & 1        & 1.00    & 0.984           & {---}          & {---}       \\
                   & Cubic upwind-biased & \(\ell^2\) error & 0        & 0.00791 & 0.00108         & 0.000577       & 0.000576    \\

### 2015-05-19 msc-publication.tex
add titles to advection contours, add noOrography 4th order results

@@
 The test was executed on the BTF, SLEVE and cut cell grids, and on a regular grid with flat terrain using a centred linear scheme and the upwind-biased cubic scheme.
+Results were also obtained on BTF and SLEVE grids with the fourth order scheme from \citet{schaer2002} using a modified version of their Fortran code.
+The original version of the code interpolated the geometric height at cell vertices from values at adjacent half levels in order to calculate the streamfunction, \(\Psi(z)\).
+The modified version used here directly calculates the height at vertices to enable comparisons with the finite volume model solutions.
+
 Tracer contours at \(t = \SI{0}{\second}, \SI{5000}{\second}\) and \(\SI{10000}{\second}\) are shown in Figure~\ref{fig:advection-tracer}.
+The results are compared on the BTF grid for the centred linear scheme (\ref{fig:advection-tracer}a) and the fourth order scheme from \citet{schaer2002} (\ref{fig:advection-tracer}c), and the upwind-biased cubic scheme on the cut cell grid (\ref{fig:advection-tracer}b), and BTF grid (\ref{fig:advection-tracer}d).
+Tracer errors at \(t = \SI{10000}{\second}\) are shown for the fourth order and upwind-biased cubic schemes in figures \ref{fig:advection-tracer}e and \ref{fig:advection-tracer}f respectively.
 
-The result of the centred linear scheme on the BTF grid (\ref{fig:advection-tracer}a) is compared with the upwind-biased cubic scheme on the BTF grid (\ref{fig:advection-tracer}b), and cut cell grid (\ref{fig:advection-tracer}c).
-
-By \(t = \SI{10000}{\second}\), the tracer suffers from diffusion and distortion on the BTF grid using the centred linear scheme.
-Some artefacts remain above the mountain peak, and the computational mode is seen as a grid-scale oscillation that travels in the opposite direction to the wind.
+By \(t = \SI{10000}{\second}\), the tracer suffers from diffusion and distortion on the BTF grid using the centred linear scheme and some artefacts remain above the mountain peak.
+Distortions are reduced by using the fourth order scheme from \citet{schaer2002} (figure~\ref{fig:advection-tracer}c), but the computational mode is seen as a grid-scale oscillation that travels in the opposite direction to the wind (figure~\ref{fig:advection-tracer}e).
+The results from the centred linear and fourth order schemes are slightly worse than the respective results from \citet{schaer2002} (their figure 6a and figure 8) because gradients in $\Psi$ have not been removed by interpolation.
 
 Using the upwind-biased cubic scheme, tracer magnitude and shape are well-preserved on all grids, both above the mountain at \(t = \SI{5000}{\second}\) and past the mountain at \(t = \SI{10000}{\second}\).
-In this test, advection is most accurate on the cut cell grid (figure~\ref{fig:advection-tracer}c) and regular grid (not shown).
+In this test, advection is most accurate on the cut cell grid (figure~\ref{fig:advection-tracer}b) and regular grid (not shown).
 As found by \citet{good2014}, the result is the same on both grids.
 This is to be expected since the wind is zero in the region of the ground and flow aloft is aligned with the grids.
-On the BTF grid, the tracer is less distorted by the cubic upwind-biased scheme (figure~\ref{fig:advection-tracer}b) compared to the centred linear scheme (figure~\ref{fig:advection-tracer}a).
+On the BTF grid, the tracer is less distorted by the cubic upwind-biased scheme (figure~\ref{fig:advection-tracer}d) compared to the centred linear scheme (figure~\ref{fig:advection-tracer}a) or fourth order scheme (figure~\ref{fig:advection-tracer}c).
 
@@
 The error is approximately halved by changing from the SLEVE grid to the cut cell grid.
-
-Tracer minima and maxima are lower than values presented by \citet{schaer2002} because we do not interpolate the geometric height of cell faces from adjacent cell centres in order to calculate the streamfunction, \(\Psi(z)\).
-Removing the interpolation from Sch\"{a}r's code results in minima and maxima that are equal to those obtained from the OpenFOAM advection solver presented here.
+Tracer minima and maxima for the centred linear and fourth order schemes are lower than those presented by \citet{schaer2002} because no interpolation is used to calculate the streamfunction.
 
@@
                   &                     & max              & 1        & 0.925   & 0.985           & 0.985          & 0.985       \\
-                  & Fourth order        & min              & 0        & -0.0926 & -0.00174        & {---}          & {---}       \\
-                  &                     & max              & 1        & 1.00    & 0.984           & {---}          & {---}       \\
+                  & Fourth order        & min              & 0        & -0.0926 & -0.00174        & {---}          & -0.00178    \\
+                  &                     & max              & 1        & 1.00    & 0.984           & {---}          & 0.983       \\
                   & Cubic upwind-biased & \(\ell^2\) error & 0        & 0.00791 & 0.00108         & 0.000577       & 0.000576    \\

### 2015-05-20 msc-publication.tex
Incorport misc feedback from Hilary

@@
 In addition, we formulate a new tracer advection test having a velocity field that is everywhere tangential to the terrain following coordinate surfaces.
+This new test presents a challenge to cut cells in situations with a large Froude number and large Scorer parameter.
 The results of the advection tests demonstrate that tracer accuracy depends upon alignment of the flow with the grid.
+As expected, the cut cell grid maintains greater accuracy in the test of stationary, stratified flow.
 In the gravity waves test, results on all grids are in good agreement with existing results from the literature.
@@
 The exponent \(n\) was introduced by \citet{leuenberger2010} in order to increase cell thickness in the layers nearest the ground, allowing longer timesteps and permitting more accurate calculation of parameterised low-level heat and momentum fluxes.
-\citet{leuenberger2010} found the exponent has an optimal value of \(n = 1.35\).
+\TODO{how do thicker cells allow more accurate calculations?}  \citet{leuenberger2010} found the exponent has an optimal value of \(n = 1.35\).
 Choosing \(n = 1\) gives the decay functions used by \citet{schaer2002}.
@@
 The utility is available at \url{https://github.com/hertzsprung/AtmosFOAM/tree/2386ffb/applications/utilities/mesh/add2dMountain}.
-Second, the surface faces were taken from the BTF grid and \texttt{snappyHexMesh} was used to intersect the surface with the grid.
-The tool removes cells whose centres are below the surface and displaces boundary vertices so that they are `snapped' to the BTF surface \citep{openfoam2015}.
+\TODO{could just link to the whole of ExnerFoam/AtmosTests} Second, the surface faces were taken from the BTF grid and \texttt{snappyHexMesh} was used to intersect the surface with the grid.
+This tool removes cells whose centres are below the surface and displaces boundary vertices so that they are `snapped' to the BTF surface \citep{openfoam2015}.
 An example of the resulting grid is shown in figure~\ref{fig:grid}c.
@@
 	\text{Continuity} &\ &\	\frac{\partial \rho}{\partial t} + \nabla \cdot \rho \bm{u} &= 0 \\
-	\text{$\theta$, flux form} &\ &\ \frac{\partial \rho \theta}{\partial t} + \nabla \cdot \rho \bm{u} \theta &= 0 \\
-	\text{$\theta$, advective form} &\ &\ \frac{\partial \theta}{\partial t} + \bm{u} \cdot \nabla \theta &= 0 \\
+	\text{\(\rho \theta\) advection} &\ &\ \frac{\partial \rho \theta}{\partial t} + \nabla \cdot \rho \bm{u} \theta &= 0 \\
 	\text{Equation of state} &\ &\ \Pi^{(1 - \kappa)/\kappa} &= \frac{R \rho \theta}{p_0}
@@
 
-The mode has a curl-free pressure gradient formulation and an upwind-biased cubic advection scheme.
-It uses a Lorenz staggering of thermodynamic variables such that $\theta$ is stored at cell centres and velocity at cell faces.
+The model has a curl-free pressure gradient formulation and an upwind-biased cubic advection scheme.
+It uses a Lorenz C-grid staggering of thermodynamic variables such that $\theta$, $\rho$ and $\Pi$ are stored at cell centres and velocities at cell faces.
+
+For tracer advection tests, the advection equation in flux form, given by
+\begin{align}
+\partial \phi / \partial t + \nabla \cdot \left( \bm{u} \phi \right) = 0
+\end{align}
+is solved using the upwind-biased multidimensional cubic scheme from \citet{weller-shahrokhi2014} which is non-monotonic and not flux corrected.
+The time derivative is solved using a three-stage, second order Runge-Kutta scheme:
+\begin{subequations}
+\begin{align}
+	\varphi^\star &= \varphi^{(n)} + \Delta t f(\phi^{(n)}) \\
+	\varphi^{\star\star} &= \varphi^{(n)} + \frac{\Delta t}{2} \left( f(\phi^{(n)}) + f(\phi^\star) \right) \\
+	\varphi^{(n+1)} &= \varphi^{(n)} + \frac{\Delta t}{2} \left( f(\phi^{(n)}) + f(\phi^{\star\star}) \right)
+\end{align}
+\end{subequations}
+where \(f(\phi^{(n)}) = - \nabla \cdot (\bm{u} \varphi^{(n)})\) at time level \(n\).
 
@@
 The first test from \citet{schaer2002} advects a tracer in a horizontal velocity field.
-Second, a new tracer advection test is formulated employing a terrain following velocity field to challenge the advection scheme on more orthogonal grids.
+Second, a new tracer advection test is formulated employing a terrain following velocity field to challenge the advection scheme on cut cell grids.
 The third test solves the Euler equations for a stably stratified atmosphere initially at rest, following \citet{klemp2011}.
@@
 The domain is \SI{301}{\kilo\meter} wide and \SI{25}{\kilo\meter} high, discretized onto a grid with \(\Delta x = \SI{1}{\kilo\meter}\) and \(\Delta z^\star = \SI{500}{\meter}\).
-The domain specified by \citet{schaer2002} is \SI{300}{\kilo\meter} between the outermost cell centres.
+The domain specified by \citet{schaer2002} is \SI{300}{\kilo\meter} between the outermost cell centres where tracer values are specified.
 In order to reproduce the result from \citet{schaer2002}, the domain has been extended horizontally by \(\Delta x/2\) \si{\meter} in both directions so that the distance between the outermost cell centres is still \SI{300}{\kilo\meter}.
@@
 
-To ensure the discrete velocity field is non-divergent, velocities are prescribed at cell faces by summing the streamfunction, \(\Psi\), at the centre of all face edges.
+To ensure that the discrete velocity field is non-divergent, velocities are prescribed at cell faces by differencing the streamfunction, \(\Psi\), along the edges from \(\Psi\) stored at cell vertices.
 Since \(u = - \partial \Psi / \partial z\), the streamfunction is found by vertical integration of the velocity profile:
@@
 
-The flux form of the advection equation, \(\partial \phi / \partial t + \nabla \cdot \left( \bm{u} \phi \right) = 0\), is solved using the upwind-biased cubic advection scheme described by \cite{weller-shahrokhi2014} which is non-monotonic and not flux corrected.
-The time derivative is solved using a three-stage, second order Runge-Kutta scheme:
-\begin{align}
-	\varphi^\star &= \varphi^{(n)} + \Delta t f(\phi^{(n)}) \\
-	\varphi^{\star\star} &= \varphi^{(n)} + \frac{\Delta t}{2} \left( f(\phi^{(n)}) + f(\phi^\star) \right) \\
-	\varphi^{(n+1)} &= \varphi^{(n)} + \frac{\Delta t}{2} \left( f(\phi^{(n)}) + f(\phi^{\star\star}) \right)
-\end{align}
-where \(f(\phi^{(n)}) = - \mathrm{div}(\bm{u} \varphi^{(n)})\) at time level \(n\).
-
 Unlike \citet{schaer2002} who use periodic lateral boundaries, a fixed value of 0 is used at the inlet boundary and all other boundaries have zero gradient.
+Furthermore, a second order Runge-Kutta timestepping scheme is used here instead of the leapfrog scheme used by \citet{schaer2002}.
 Tests are integrated forward in time for \SI{10000}{\second} with a timestep of \(\Delta t = \SI{25}{\second}\).
@@
 
-By \(t = \SI{10000}{\second}\), the tracer suffers from diffusion and distortion on the BTF grid using the centred linear scheme and some artefacts remain above the mountain peak.
+By \(t = \SI{10000}{\second}\), the tracer suffers from distortion on the BTF grid using the centred linear scheme and some artefacts remain about the mountain peak.
+The tracer has spread vertically due to increased numerical errors when the tracer is transported between layers.
 Distortions are reduced by using the fourth order scheme from \citet{schaer2002} (figure~\ref{fig:advection-tracer}c), but the computational mode is seen as a grid-scale oscillation that travels in the opposite direction to the wind (figure~\ref{fig:advection-tracer}e).
@@
 
-The results of the tracer advection test show that numerical errors are due to misalignment of the flow with the grid.
+The results of the tracer advection test show that numerical errors are due primary to misalignment of the flow with the grid.
 Using the upwind-biased cubic scheme, distortions in the grid do not significantly distort the tracer.
@@
 Once again, the results are similar on all three grids, and results are shown on the BTF and cut cell grids in figures~\ref{fig:gw-theta}a and \ref{fig:gw-theta}c respectively.
-However, examining more closely the anomalies in the lee of the mountain, figure~\ref{fig:gw-theta}d shows that the bottommost layer is anomalously warm and the layer above it is anomalously cold.
+However, examining more closely the anomalies in the lee of the mountain for the cut cell grid, figure~\ref{fig:gw-theta}d shows that the bottommost layer is anomalously warm and the layer above it is anomalously cold.
 This feature is not present on the BTF grid (figure~\ref{fig:gw-theta}b) or the SLEVE grid (not shown).
@@
 Figure~\ref{fig:gw-exner-theta} shows vertical profiles of the Exner function of pressure and potential temperature in the lowest \SI{1}{\kilo\meter} in the lee of the mountain at \(x = \SI{50}{\kilo\meter}\).
-We find that the Exner function of pressure decreases linearly with height on all grids.
+We find that the decrease in the Exner function of pressure with height is close to linear on all grids.
 The potential temperature increases linearly with height on the BTF and SLEVE grids.

### 2015-05-21 msc-publication.tex
plot w contours from AtmosTests

@@
 with \(s_1\) and \(s_2\) are the scale heights of large-scale and small-scale terrain respectively.
-The exponent \(n\) was introduced by \citet{leuenberger2010} in order to increase cell thickness in the layers nearest the ground, allowing longer timesteps and permitting more accurate calculation of parameterised low-level heat and momentum fluxes.
-\TODO{how do thicker cells allow more accurate calculations?}  \citet{leuenberger2010} found the exponent has an optimal value of \(n = 1.35\).
+The exponent \(n\) was introduced by \citet{leuenberger2010} in order to increase cell thickness in the layers nearest the ground, allowing longer timesteps.
+\citet{leuenberger2010} found the exponent has an optimal value of \(n = 1.35\).
 Choosing \(n = 1\) gives the decay functions used by \citet{schaer2002}.
@@
 
-The test setup follows the specification by \cite{klemp2011}, but has a narrower domain that is \SI{20}{\kilo\meter} wide and \SI{20}{\kilo\meter} high in order to reduce simulation time, as used by \citet{weller-shahrokhi2014}.
-The grid resolution is \(\Delta x = \Delta z^\star = \SI{500}{\meter}\) as originally specified by \citet{klemp2011}.
-All boundary conditions are no normal flow.
+The test setup follows the specification by \cite{klemp2011}.  The grid resolution is \(\Delta x = \Delta z^\star = \SI{500}{\meter}\).  All boundary conditions are no normal flow.
 

### 2015-05-21 msc-publication.tex
Remove Klemp2011 figure.

@@
 The test was integrated forward by 5 hours on the BTF, SLEVE and cut cell grids, and a regular grid with flat terrain.
-Maximum vertical velocities are compared with the results from \citet{klemp2011} in figure~\ref{fig:resting} (note different vertical scales).
+Maximum vertical velocities are presented in figure~\ref{fig:resting}.
 In agreement with \citet{klemp2011}, vertical velocities are larger on more distorted grids.
@@
 %
-	\caption{Maximum spurious vertical velocity, \(w\) (\si{\meter\per\second}), in the resting atmosphere test with results on (a) BTF, SLEVE, Hybrid Terrain Following (HTF) and Smoothed Terrain Following (STF) coordinates from \citet{klemp2011} using their improved horizontal pressure gradient formulation, (b) BTF, SLEVE, cut cell and regular grids using the model from \citet{weller-shahrokhi2014} which includes a curl-free pressure gradient formulation.
-Note that vertical scales differ.}
+	\caption{Maximum spurious vertical velocity, \(w\) (\si{\meter\per\second}), in the resting atmosphere test with results on BTF, SLEVE, cut cell and regular grids using the model from \citet{weller-shahrokhi2014} which includes a curl-free pressure gradient formulation.}
 	\label{fig:resting}
-\end{figure*}
+\end{figure}
 

### 2015-05-24 src/tf-cutcell-comparison/fig-gravityWaves-theta.tex
move into src directory

--- 2015-03-05 fig-gravityWaves-theta.tex
@@
 
-\includegraphics[width=2.2in]{../gravityWaves-snapCol-250m/18000/thetaDiff.pdf}
+\includegraphics[width=2.2in]{../gravityWaves-cutCell-250m/18000/thetaDiff.pdf}
 \hspace{0.1in}
-\includegraphics[width=4in]{../gravityWaves-snapCol-250m/18000/thetaDiffZoom.pdf}
+\includegraphics[width=4in]{../gravityWaves-cutCell-250m/18000/thetaDiffZoom.pdf}
 \vspace*{0.1in}

### 2015-05-24 src/tf-cutcell-comparison/fig-meshes.tex
move into src directory

\documentclass{article}
\usepackage[paperwidth=6.3in,paperheight=2.1in,margin=0in]{geometry}
\usepackage{mathptmx}%{times}
\usepackage{graphicx}
\usepackage{siunitx}
\usepackage[absolute,overlay]{textpos}
\usepackage{color}

\setlength{\parindent}{0pt}
\begin{document}
\TPMargin{2pt}
\textblockcolour{white}
\begin{textblock}{0.5}(2.7,13.8)
\normalsize
(a)
\end{textblock}
\begin{textblock}{1}(8,13.8)
\normalsize
(b)
\end{textblock}
\begin{textblock}{1}(13.3,13.8)
\normalsize
(c)
\end{textblock}
\includegraphics{../mesh-btf-resting/constant/mesh.pdf}
\includegraphics{../mesh-sleve-resting/constant/mesh.pdf}
\includegraphics{../mesh-snapCol-resting/constant/mesh.pdf}
\end{document}

### 2015-05-24 src/tf-cutcell-comparison/tf-cutcell-comparison.tex
move into src directory

--- 2015-05-21 msc-publication.tex
@@
 Representing orography accurately in numerical weather prediction systems is necessary to model downslope winds and local precipitation.
-Orography also exerts strong non-local influences by the atmospheric response to form drag and gravity wave drag.
+Orography also exerts strong non-local influences from the heat release due to convection and by the atmospheric response to form drag and gravity wave drag.
 There are two main approaches to represent orography on a grid: terrain following layers and cut cells.
@@
 First, a custom utility was used to move points beneath the surface up to the surface creating small cells near mountain peaks.
-The utility is available at \url{https://github.com/hertzsprung/AtmosFOAM/tree/2386ffb/applications/utilities/mesh/add2dMountain}.
-\TODO{could just link to the whole of ExnerFoam/AtmosTests} Second, the surface faces were taken from the BTF grid and \texttt{snappyHexMesh} was used to intersect the surface with the grid.
+Second, the surface faces were taken from the BTF grid and \texttt{snappyHexMesh} was used to intersect the surface with the grid.
 This tool removes cells whose centres are below the surface and displaces boundary vertices so that they are `snapped' to the BTF surface \citep{openfoam2015}.
@@
 
-The model has a curl-free pressure gradient formulation and an upwind-biased cubic advection scheme.
-It uses a Lorenz C-grid staggering of thermodynamic variables such that $\theta$, $\rho$ and $\Pi$ are stored at cell centres and velocities at cell faces.
+The model uses a Lorenz C-grid staggering of thermodynamic variables such that $\theta$, $\rho$ and $\Pi$ are stored at cell centres and velocities at cell faces.
+
+A minor change has been made to the curl-free, non-orthogonal formulation as described in \citet{weller-shahrokhi2014} to improve stability on non-orthogonal grids.
+\citet{weller-shahrokhi2014} use a Hodge dual operator, \(H\), to map from co-variant to contra-variant velocity components:
+\begin{align}
+	U = HV
+\end{align}
+where \(U = \rho\bm{u}\cdot\bm{S}_f\) and \(V=\rho\bm{u}\cdot\bm{d}_f\).
+\(\bm{S}_f\) is the surface normal vector for a face, \(f\), which has a magnitude of the surface area, and \(\bm{d}_f\) is the vector between neighbouring cell centres.
+
+In order to create a semi-implicit formulation, \(H\) is split into diagonal and off-diagonal components and the off-diagonal components are treated explicitly in the Helmholtz equation.
+However, this was found to be unstable when using a large timestep on a grid with large non-orthogonality.
+Improved stability is achieved by splitting \(H\) into the orthogonal part and the non-orthogonal correction:
+\begin{align}
+	H = H_c + H_\mathrm{corr}
+\end{align}
+where the diagonal matrix \(H_c=\frac{|\bm{S}_f|}{|\bm{d}_f|}\) and the non-orthogonal correction is simply \(H_\mathrm{corr} = H - H_c\).
+The orthogonal part, \(H_c\), is treated implicitly in the Helmholtz equation and the non-orthogonal part, \(H_\mathrm{corr}\), is treated explicitly.
 
@@
 No explicit diffusion is used in any of the tests.
+
+The OpenFOAM implementation of the numerical model and grid generation utilities are available at \url{https://github.com/hertzsprung/AtmosFOAM/tree/shaw-weller-2015-mwr}.
+The test cases are located at \url{https://github.com/hertzsprung/AtmosTests/tree/shaw-weller-2015-mwr}.
 
@@
 Distortions are reduced by using the fourth order scheme from \citet{schaer2002} (figure~\ref{fig:advection-tracer}c), but the computational mode is seen as a grid-scale oscillation that travels in the opposite direction to the wind (figure~\ref{fig:advection-tracer}e).
-The results from the centred linear and fourth order schemes are slightly worse than the respective results from \citet{schaer2002} (their figure 6a and figure 8) because gradients in $\Psi$ have not been removed by interpolation.
+The results from the centred linear and fourth order schemes are slightly worse than the respective results from \citet{schaer2002} (their figure 6a and figure 8) because of the difference in the discretisation of $\Psi$.
 
@@
 
-The results of the tracer advection test show that numerical errors are due primary to misalignment of the flow with the grid.
+The results of the horizontal advection test show that numerical errors are due either to misalignment of the flow with the grid, or to grid distortions.
 Using the upwind-biased cubic scheme, distortions in the grid do not significantly distort the tracer.
@@
 Negative contours denoted by dotted lines.
-The terrain profile is also shown immediately above the $x$ axis.}
+The terrain profile is also shown immediately above the $x$ axis.
+Subfigures (c) and (e) produced using the modified version of the code from \citet{schaer2002}.}
 	\label{fig:advection-tracer}
@@
 In the horizontal advection test, results were more accurate where the flow was aligned with the grid layers, and distortions in the BTF grid led to increased errors.
-We formulate a new tracer advection test in which the velocity field is everywhere tangential to the basic terrain following coordinate surfaces.
+To determine the primary source of numerical error, we formulate a new tracer advection test in which the velocity field is everywhere tangential to the basic terrain following coordinate surfaces.
 In this new test, flow is aligned with the BTF grid layers.
@@
 \begin{table*}[t]
-	\caption{Minimum and maximum tracer magnitudes and \(\ell^2\) error norms at \(t = \SI{10000}{\second}\) in the horizontal and terrain following tracer advection tests using centred linear and cubic upwind-biased schemes and, for the horizontal advection test, the fourth order scheme from \citet{schaer2002}.}
+	\caption{Minimum and maximum tracer magnitudes and \(\ell^2\) error norms at \(t = \SI{10000}{\second}\) in the horizontal and terrain following tracer advection tests using centred linear and cubic upwind-biased schemes.
+For the horizontal advection test, minimum and maximum values are given for the fourth order scheme using the modified code from \citet{schaer2002}.}
 \label{tab:advection}
@@
 However, magnitudes are smaller comparing results on the terrain following grids with those from \citet{klemp2011}.
-Using the model from \citet{weller-shahrokhi2014}, which includes a curl-free pressure gradient formulation, $w$ reaches a maximum of \(\sim \SI{0.35}{\meter\per\second}\) on the BTF grid, compared with a maximum of \(\sim \SI{7}{\meter\per\second}\) found by \citet{klemp2011} using their improved horizontal pressure gradient formulation.
+Using the model from \citet{weller-shahrokhi2014}, which includes a curl-free pressure gradient formulation, $w$ reaches a maximum of \SI{0.33}{\meter\per\second} on the BTF grid, compared with a maximum of \(\sim \SI{7}{\meter\per\second}\) found by \citet{klemp2011} using their improved horizontal pressure gradient formulation.
 
 Unlike the result from \citet{klemp2011}, the SLEVE grid does not significantly reduce vertical velocities compared to the BTF grid.
-However, errors are three orders of magnitude smaller on the cut cell grid with vertical velocities of \(\sim \SI{1e-3}{\meter\per\second}\).
+However, errors are two orders of magnitude smaller on the cut cell grid with vertical velocities of \(\sim \SI{1e-4}{\meter\per\second}\).
 The smallest error of \(\sim \SI{1e-10}{\meter\per\second}\) is found on the regular grid.
@@
 Test results are compared between the BTF, SLEVE and cut cell grids.
-Vertical velocities on the BTF grid are shown in figure~\ref{fig:gw-w}a, which have few visible differences from results on the SLEVE and cut cell grids (not shown).
-Vertical velocities on all grids are in agreement with the high resolution solution from \citet{melvin2010}, shown in figure~\ref{fig:gw-w}b.
+Vertical velocities on the BTF, SLEVE and cut cell grids are visually indistinguishable (not shown).
+They agree with the high resolution mass-conserving semi-implicit semi-Lagrangian solution from \citet{melvin2010}.
 
@@
 The most prominent errors are found only on the cut cell grid, where the potential temperature errors near the ground excite the Lorenz computational mode.
-
-\begin{figure}
-	\centering
-	\includegraphics{../msc-publication-fig-gravityWaves/fig-gravityWaves.pdf}
-%
-	\caption{Vertical cross section of vertical velocity contours in the gravity waves test after 5 hours with mountain height $h_0 = \SI{250}{\meter}$ on (a) the BTF grid compared with (b) the mass-conserving semi-implicit semi-Lagrangian solution from \citet{melvin2010}.
-Contours are every \SI{5e-2}{\meter\per\second} with solid lines denoting ascent and dashed lines descent.
-Small oscillations in contour lines are due to contouring on an irregular grid.}
-	\label{fig:gw-w}
-\end{figure}
 

### 2015-05-24 src/tf-cutcell-comparison/tf-cutcell-comparison.tex
produce article without figures

@@
 \begin{document}
-\newcommand{\TODO}[1]{\textcolor{purple}{TODO: \emph{#1}}}
 
@@
 	\centering
-	\includegraphics{../msc-publication-fig-meshes/fig-meshes.pdf}
+%	\includegraphics{../msc-publication-fig-meshes/fig-meshes.pdf}
 	%
@@
 	\centering
-	\includegraphics{../msc-publication-fig-advection-tracer/fig-advection-tracer.pdf}
+%	\includegraphics{../msc-publication-fig-advection-tracer/fig-advection-tracer.pdf}
 %
@@
 	\centering
-	\includegraphics{../msc-publication-fig-resting/fig-resting.pdf}
+%	\includegraphics{../msc-publication-fig-resting/fig-resting.pdf}
 %
@@
 	\centering
-	\includegraphics{../msc-publication-fig-gravityWaves-theta/fig-gravityWaves-theta.pdf}
+%	\includegraphics{../msc-publication-fig-gravityWaves-theta/fig-gravityWaves-theta.pdf}
 %
@@
 	\centering
-	\includegraphics{../msc-publication-fig-gravityWaves-sampleLine/fig-gravityWaves-sampleLine.pdf}
+%	\includegraphics{../msc-publication-fig-gravityWaves-sampleLine/fig-gravityWaves-sampleLine.pdf}
 %

### 2015-05-24 src/tf-cutcell-comparison/fig-meshes.tex
plot GW sampleLine and meshes

@@
 \end{textblock}
-\includegraphics{../mesh-btf-resting/constant/mesh.pdf}
-\includegraphics{../mesh-sleve-resting/constant/mesh.pdf}
-\includegraphics{../mesh-snapCol-resting/constant/mesh.pdf}
+\includegraphics{../resting-mesh-btf/constant/mesh.pdf}
+\includegraphics{../resting-mesh-sleve/constant/mesh.pdf}
+\includegraphics{../resting-mesh-cutCell/constant/mesh.pdf}
 \end{document}

### 2015-05-25 src/tf-cutcell-comparison/tf-cutcell-comparison.tex
add advection contour figure
add resting-w plot
plot GW sampleLine and meshes
add thetaDiff plots
refer to new git repo

@@
 	\centering
-%	\includegraphics{../msc-publication-fig-meshes/fig-meshes.pdf}
+	\includegraphics{../fig-meshes/fig-meshes.pdf}
 	%
@@
 
-The OpenFOAM implementation of the numerical model and grid generation utilities are available at \url{https://github.com/hertzsprung/AtmosFOAM/tree/shaw-weller-2015-mwr}.
-The test cases are located at \url{https://github.com/hertzsprung/AtmosTests/tree/shaw-weller-2015-mwr}.
+The OpenFOAM implementation of the numerical model, grid generation utilities and test cases are available at \url{https://github.com/hertzsprung/tf-cutcell-comparison/tree/shaw-weller-2015-mwr}.
 
@@
 	\centering
-%	\includegraphics{../msc-publication-fig-advection-tracer/fig-advection-tracer.pdf}
+	\includegraphics{../fig-advection-tracer/fig-advection-tracer.pdf}
 %
@@
 	\centering
-%	\includegraphics{../msc-publication-fig-resting/fig-resting.pdf}
+	\includegraphics{../fig-resting/fig-resting.pdf}
 %
@@
 	\centering
-%	\includegraphics{../msc-publication-fig-gravityWaves-theta/fig-gravityWaves-theta.pdf}
+	\includegraphics{../fig-gravityWaves-theta/fig-gravityWaves-theta.pdf}
 %
@@
 	\centering
-%	\includegraphics{../msc-publication-fig-gravityWaves-sampleLine/fig-gravityWaves-sampleLine.pdf}
+	\includegraphics{../fig-gravityWaves-sampleLine/fig-gravityWaves-sampleLine.pdf}
 %

### 2015-06-08 src/tf-cutcell-comparison/tf-cutcell-comparison.tex
move figures and tables to end
add some minor changes from Hilary

@@
-\documentclass[twocol]{ametsoc}
-\journal{mwr}
+\documentclass{ametsoc}
 
@@
 
-This study uses the nonhydrostatic model from \citet{weller-shahrokhi2014} to enable a like-for-like comparison between terrain following and cut cell grids for idealised, two-dimensional test cases from the literature.
+This study uses a modification of the nonhydrostatic model from \citet{weller-shahrokhi2014} to enable a like-for-like comparison between terrain following and cut cell grids for idealised, two-dimensional test cases from the literature.
 Section 2 presents the formulation of the terrain following and cut cell grids used in the experiments that follow.
-In section 3 we outline the model from \citet{weller-shahrokhi2014} and the governing equations.
+In section 3 we give the governing equations, outline the model from \citet{weller-shahrokhi2014} and describe the modification which improves stability for long timesteps in the presence of steep orography.
 Section 4 analyses the results from two tracer advection tests, a test of a stably stratified atmosphere initially at rest, and orographically induced gravity waves.
 Concluding remarks are made in section 5.
-
-\begin{figure*}
-	\centering
-	\includegraphics{../fig-meshes/fig-meshes.pdf}
-	%
-	\caption{Examples of (a) BTF, (b) SLEVE, and (c) a cut cell grid, showing cell edges in the lowest four layers.
-The two dimensional grids are \SI{20}{\kilo\meter} wide and \SI{20}{\kilo\meter} high.
-SLEVE parameters are specified in the resting atmosphere test in section~\ref{sec:results}\ref{sec:resting}.
-The cut cell grid was created by intersecting the terrain surface with a regular grid as described in section~\ref{sec:grid}.}
-	\label{fig:grid}
-\end{figure*}
 
@@
 
-The model uses a Lorenz C-grid staggering of thermodynamic variables such that $\theta$, $\rho$ and $\Pi$ are stored at cell centres and velocities at cell faces.
+The model uses a Lorenz C-grid staggering of thermodynamic variables such that $\theta$, $\rho$ and $\Pi$ are stored at cell centres and the covariant component of velocity at cell faces.
 
@@
 
-\begin{figure*}
-	\centering
-	\includegraphics{../fig-advection-tracer/fig-advection-tracer.pdf}
-%
-	\caption{Horizontally advected tracer contours at \(t = \SI{0}{\second}\), \SI{5000}{\second} and \SI{10000}{\second} using (a) centred linear scheme on a BTF grid, the upwind-biased cubic scheme on (b) the cut cell grid, (d) the BTF grid, and (c) the fourth order scheme from \citet{schaer2002} on the BTF grid, with contour intervals every 0.1.
-Errors on the BTF grid at \(t = \SI{10000}{\second}\) are shown for (e) the fourth order scheme from \citet{schaer2002}, and (f) the upwind-biased cubic scheme, with contour intervals every 0.01.
-Negative contours denoted by dotted lines.
-The terrain profile is also shown immediately above the $x$ axis.
-Subfigures (c) and (e) produced using the modified version of the code from \citet{schaer2002}.}
-	\label{fig:advection-tracer}
-\end{figure*}
-
 \subsection{Terrain following advection}
@@
 
-\begin{table*}[t]
-	\caption{Minimum and maximum tracer magnitudes and \(\ell^2\) error norms at \(t = \SI{10000}{\second}\) in the horizontal and terrain following tracer advection tests using centred linear and cubic upwind-biased schemes.
-For the horizontal advection test, minimum and maximum values are given for the fourth order scheme using the modified code from \citet{schaer2002}.}
-\label{tab:advection}
-%
-\centering
-\footnotesize
-\begin{tabular}{l l l l S S S S}
-\hline\hline
-                  &                     &                  & Analytic & {BTF}   & {SLEVE}         & {Cut cell}     & {No terrain} \\
-\hline
-Horizontal        & Centred linear      & \(\ell^2\) error & 0        & 0.0210  & 0.00233         & 0.00224        & 0.00223     \\
-                  &                     & min              & 0        & -0.275  & -0.0252         & -0.0251        & -0.0251     \\
-                  &                     & max              & 1        & 0.925   & 0.985           & 0.985          & 0.985       \\
-                  & Fourth order        & min              & 0        & -0.0926 & -0.00174        & {---}          & -0.00178    \\
-                  &                     & max              & 1        & 1.00    & 0.984           & {---}          & 0.983       \\
-                  & Cubic upwind-biased & \(\ell^2\) error & 0        & 0.00791 & 0.00108         & 0.000577       & 0.000576    \\
-                  &                     & min              & 0        & -0.0446 & -0.0106         & -0.000674      & -0.00674    \\
-                  &                     & max              & 1        & 0.925   & 0.982           & 0.983          & 0.983       \\
-\hline
-Terrain following & Centred linear      & \(\ell^2\) error & 0        & 0.00251 & 0.0173          & 13.6           & {---}        \\
-		  &                     & min              & 0        & -0.0245 & -0.120          & -623           & {---}        \\
-                  &                     & max              & 1        & 0.985   & 0.950           & 3480           & {---}        \\
-		  & Cubic upwind-biased & \(\ell^2\) error & 0        & 0.00154 & 0.0120          & 0.0134         & {---}        \\
-                  &                     & min              & 0        & -0.0110 & -0.0263         & -0.028         & {---}        \\
-                  &                     & max              & 1        & 0.983   & 0.865           & 0.851          & {---}        \\
-\hline
-\end{tabular}
-\end{table*}
-
 $\ell^2$ errors and tracer extrema for this test are compared with the horizontal advection results in table~\ref{tab:advection}.
@@
 However, magnitudes are smaller comparing results on the terrain following grids with those from \citet{klemp2011}.
-Using the model from \citet{weller-shahrokhi2014}, which includes a curl-free pressure gradient formulation, $w$ reaches a maximum of \SI{0.33}{\meter\per\second} on the BTF grid, compared with a maximum of \(\sim \SI{7}{\meter\per\second}\) found by \citet{klemp2011} using their improved horizontal pressure gradient formulation.
+The results presented in figure~\ref{fig:resting}, which use a curl-free pressure gradient, have maximum spurious values of $w$ of \SI{0.33}{\meter\per\second} on the BTF grid, compared with a maximum of \(\sim \SI{7}{\meter\per\second}\) found by \citet{klemp2011} using their improved horizontal pressure gradient formulation.
+The results in figure~\ref{fig:resting} have the same maximum errors as \citet{weller-shahrokhi2014} but, due to the more stable split into implicitly and explicitly treated terms, the errors decay over time due to the dissipative nature of the advection scheme.
 
@@
 The maximum vertical velocity was significantly decreased on the cut cell grid, so we conclude that non-orthogonality, or lack of alignment of the grid with sufaces of constant gravitational potential are a significant cause of numerical error in this test.
-
-\begin{figure}
-	\centering
-	\includegraphics{../fig-resting/fig-resting.pdf}
-%
-	\caption{Maximum spurious vertical velocity, \(w\) (\si{\meter\per\second}), in the resting atmosphere test with results on BTF, SLEVE, cut cell and regular grids using the model from \citet{weller-shahrokhi2014} which includes a curl-free pressure gradient formulation.}
-	\label{fig:resting}
-\end{figure}
 
@@
 
-\begin{figure*}
-	\centering
-	\includegraphics{../fig-gravityWaves-theta/fig-gravityWaves-theta.pdf}
-%
-	\caption{Anomalies in potential temperature in the gravity waves test after 5 hours with a mountain height, \(h_0 = \SI{250}{\meter}\).
-The central domain in the lowest \SI{12}{\kilo\meter} is shown on (a) the BTF grid, and (c) the cut cell grid.
-The four lowest layers of each grid are shown for (b) BTF, and (d) cut cell grids, using a narrower potential temperature scale.
-The results on the SLEVE grid (not shown) are qualitatively identical to results on the BTF grid.}
-	\label{fig:gw-theta}
-\end{figure*}
-
-\begin{figure}
-	\centering
-	\includegraphics{../fig-gravityWaves-sampleLine/fig-gravityWaves-sampleLine.pdf}
-%
-	\caption{Vertical profiles of the Exner function of pressure, \(\Pi\), and potential temperature, \(\theta\), in the gravity waves test with a mountain height of \(h_0 = \SI{500}{\meter}\).
-Exner profile is visually identical on all grids for both mountain heights; for clarity, the Exner profile is only plotted for the BTF grid.
-The computational mode is manifested as a zig-zag in potential temperature on the cut cell grid which.
-Results on the SLEVE grid (not shown) are qualitatively the same as those on the BTF grid.
-The thermal profile with a lesser mountain height of \(h_0 = \SI{250}{\meter}\) (not shown) exhibits a computational mode with smaller amplitude.}
-	\label{fig:gw-exner-theta}
-\end{figure}
-
 \section{Conclusions}
@@
 % TABLES
+\begin{table*}
+	\caption{Minimum and maximum tracer magnitudes and \(\ell^2\) error norms at \(t = \SI{10000}{\second}\) in the horizontal and terrain following tracer advection tests using centred linear and cubic upwind-biased schemes.
+For the horizontal advection test, minimum and maximum values are given for the fourth order scheme using the modified code from \citet{schaer2002}.}
+\label{tab:advection}
+%
+\centering
+\footnotesize
+\begin{tabular}{l l l l S S S S}
+\hline\hline
+                  &                     &                  & Analytic & {BTF}   & {SLEVE}         & {Cut cell}     & {No terrain} \\
+\hline
+Horizontal        & Centred linear      & \(\ell^2\) error & 0        & 0.0210  & 0.00233         & 0.00224        & 0.00223     \\
+                  &                     & min              & 0        & -0.275  & -0.0252         & -0.0251        & -0.0251     \\
+                  &                     & max              & 1        & 0.925   & 0.985           & 0.985          & 0.985       \\
+                  & Fourth order        & min              & 0        & -0.0926 & -0.00174        & {---}          & -0.00178    \\
+                  &                     & max              & 1        & 1.00    & 0.984           & {---}          & 0.983       \\
+                  & Cubic upwind-biased & \(\ell^2\) error & 0        & 0.00791 & 0.00108         & 0.000577       & 0.000576    \\
+                  &                     & min              & 0        & -0.0446 & -0.0106         & -0.000674      & -0.00674    \\
+                  &                     & max              & 1        & 0.925   & 0.982           & 0.983          & 0.983       \\
+\hline
+Terrain following & Centred linear      & \(\ell^2\) error & 0        & 0.00251 & 0.0173          & 13.6           & {---}        \\
+		  &                     & min              & 0        & -0.0245 & -0.120          & -623           & {---}        \\
+                  &                     & max              & 1        & 0.985   & 0.950           & 3480           & {---}        \\
+		  & Cubic upwind-biased & \(\ell^2\) error & 0        & 0.00154 & 0.0120          & 0.0134         & {---}        \\
+                  &                     & min              & 0        & -0.0110 & -0.0263         & -0.028         & {---}        \\
+                  &                     & max              & 1        & 0.983   & 0.865           & 0.851          & {---}        \\
+\hline
+\end{tabular}
+\end{table*}
 
 % FIGURES
+\begin{figure*}
+	\centering
+	\includegraphics{../fig-meshes/fig-meshes.pdf}
+	%
+	\caption{Examples of (a) BTF, (b) SLEVE, and (c) a cut cell grid, showing cell edges in the lowest four layers.
+The two dimensional grids are \SI{20}{\kilo\meter} wide and \SI{20}{\kilo\meter} high.
+SLEVE parameters are specified in the resting atmosphere test in section~\ref{sec:results}\ref{sec:resting}.
+The cut cell grid was created by intersecting the terrain surface with a regular grid as described in section~\ref{sec:grid}.}
+	\label{fig:grid}
+\end{figure*}
+
+\begin{figure*}
+	\centering
+	\includegraphics{../fig-advection-tracer/fig-advection-tracer.pdf}
+%
+	\caption{Horizontally advected tracer contours at \(t = \SI{0}{\second}\), \SI{5000}{\second} and \SI{10000}{\second} using (a) centred linear scheme on a BTF grid, the upwind-biased cubic scheme on (b) the cut cell grid, (d) the BTF grid, and (c) the fourth order scheme from \citet{schaer2002} on the BTF grid, with contour intervals every 0.1.
+Errors on the BTF grid at \(t = \SI{10000}{\second}\) are shown for (e) the fourth order scheme from \citet{schaer2002}, and (f) the upwind-biased cubic scheme, with contour intervals every 0.01.
+Negative contours denoted by dotted lines.
+The terrain profile is also shown immediately above the $x$ axis.
+Subfigures (c) and (e) produced using the modified version of the code from \citet{schaer2002}.}
+	\label{fig:advection-tracer}
+\end{figure*}
+
+\begin{figure}
+	\centering
+	\includegraphics{../fig-resting/fig-resting.pdf}
+%
+	\caption{Maximum spurious vertical velocity, \(w\) (\si{\meter\per\second}), in the resting atmosphere test with results on BTF, SLEVE, cut cell and regular grids using the model from \citet{weller-shahrokhi2014} which includes a curl-free pressure gradient formulation.}
+	\label{fig:resting}
+\end{figure}
+
+\begin{figure*}
+	\centering
+	\includegraphics{../fig-gravityWaves-theta/fig-gravityWaves-theta.pdf}
+%
+	\caption{Anomalies in potential temperature in the gravity waves test after 5 hours with a mountain height, \(h_0 = \SI{250}{\meter}\).
+The central domain in the lowest \SI{12}{\kilo\meter} is shown on (a) the BTF grid, and (c) the cut cell grid.
+The four lowest layers of each grid are shown for (b) BTF, and (d) cut cell grids, using a narrower potential temperature scale.
+The results on the SLEVE grid (not shown) are qualitatively identical to results on the BTF grid.}
+	\label{fig:gw-theta}
+\end{figure*}
+
+\begin{figure}
+	\centering
+	\includegraphics{../fig-gravityWaves-sampleLine/fig-gravityWaves-sampleLine.pdf}
+%
+	\caption{Vertical profiles of the Exner function of pressure, \(\Pi\), and potential temperature, \(\theta\), in the gravity waves test with a mountain height of \(h_0 = \SI{500}{\meter}\).
+Exner profile is visually identical on all grids for both mountain heights; for clarity, the Exner profile is only plotted for the BTF grid.
+The computational mode is manifested as a zig-zag in potential temperature on the cut cell grid which.
+Results on the SLEVE grid (not shown) are qualitatively the same as those on the BTF grid.
+The thermal profile with a lesser mountain height of \(h_0 = \SI{250}{\meter}\) (not shown) exhibits a computational mode with smaller amplitude.}
+	\label{fig:gw-exner-theta}
+\end{figure}
 

### 2015-06-08 src/tf-cutcell-comparison/tf-cutcell-comparison.tex
Mention planetary waves, fix max w

@@
 Representing orography accurately in numerical weather prediction systems is necessary to model downslope winds and local precipitation.
-Orography also exerts strong non-local influences from the heat release due to convection and by the atmospheric response to form drag and gravity wave drag.
+Orography also exerts strong non-local influences: from the latent heat release due to convection, by directly forcing planetary waves, and by the atmospheric response to form drag and gravity wave drag.
 There are two main approaches to represent orography on a grid: terrain following layers and cut cells.
@@
 
-In order to create a semi-implicit formulation, \(H\) is split into diagonal and off-diagonal components and the off-diagonal components are treated explicitly in the Helmholtz equation.
+In order to create a semi-implicit formulation, \cite{weller-shahrokhi2014} split \(H\) into diagonal and off-diagonal components and the off-diagonal components are treated explicitly in the Helmholtz equation.
 However, this was found to be unstable when using a large timestep on a grid with large non-orthogonality.
@@
 \end{align}
-where the diagonal matrix \(H_c=\frac{|\bm{S}_f|}{|\bm{d}_f|}\) and the non-orthogonal correction is simply \(H_\mathrm{corr} = H - H_c\).
+where the diagonal matrix \(H_c=|\bm{S}_f|/|\bm{d}_f|\) and the non-orthogonal correction is simply \(H_\mathrm{corr} = H - H_c\).
 The orthogonal part, \(H_c\), is treated implicitly in the Helmholtz equation and the non-orthogonal part, \(H_\mathrm{corr}\), is treated explicitly.
@@
 
-Spurious vertical velocities were small in the resting atmosphere test, reaching a maximum of $\sim \SI{0.35}{\meter\per\second}$ on the BTF grid, compared to a maximum of $\sim \SI{10}{\meter\per\second}$ found by \citet{klemp2011}.
+Spurious vertical velocities were small in the resting atmosphere test, reaching a maximum of $\sim \SI{0.35}{\meter\per\second}$ on the BTF grid, compared to a maximum of $\sim \SI{7}{\meter\per\second}$ found by \citet{klemp2011}.
 In the gravity waves test, vertical velocities were in good agreement with the reference solution from \citet{melvin2010} across all grids.

### 2015-06-08 src/tf-cutcell-comparison/tf-cutcell-comparison.tex
Minor fixes from John's review

@@
 
-\abstract{Terrain following coordinates are widely used in operational models but the cut cell method has been proposed as an alternative that can more accurately represent the orography and associated atmospheric dynamics.
+\abstract{Terrain following coordinates are widely used in operational models but the cut cell method has been proposed as an alternative that can more accurately represent atmospheric dynamics over steep orography.
 Because the type of grid is usually chosen during model implementation, it is typically necessary to use different models in order to compare the accuracy of different grids.
-On the contrary, we use a single C-grid finite volume model to enable a like-for-like comparison of terrain following and cut cell grids.
-A series of standard two-dimensional tests using idealised terrain are performed: tracer advection in a prescribed horizontal velocity field, a test of a stably stratified atmosphere at rest, and orographically induced gravity waves.
-In addition, we formulate a new tracer advection test having a velocity field that is everywhere tangential to the terrain following coordinate surfaces.
+In contrast, here a single C-grid finite volume model is used to enable a like-for-like comparison of terrain following and cut cell grids.
+A series of standard two-dimensional tests using idealised terrain are performed: tracer advection in a prescribed horizontal velocity field, a test starting from stably stratified initial conditions, and orographically induced gravity waves described by nonhydrostatic dynamics.
+In addition, a new tracer advection test is formulated having a velocity field that is everywhere tangential to the terrain following coordinate surfaces.
 This new test presents a challenge to cut cells in situations with a large Froude number and large Scorer parameter.
@@
 In the gravity waves test, results on all grids are in good agreement with existing results from the literature.
-However, we find that the Lorenz computational mode is excited as a vertical zig--zag in potential temperature on the cut cell grid.}
+However, it is found that the Lorenz computational mode is excited as a vertical zig--zag in potential temperature on the cut cell grid.}
 
@@
 Despite their associated numerical errors, TF layers are in widespread operational use \citep{steppeler2003}.
-They are attractive because their rectangular structure is simple to process by computer, boundary layer resolution can be increased with variable spacing of vertical layers \citep{schaer2002}, and cell sizes remain almost constant \citep{jebens2011}.
+They are attractive because their rectangular structure is simple to process by computer and link with parameterisations, boundary layer resolution can be increased with variable spacing of vertical layers \citep{schaer2002}, and cell sizes remain almost constant \citep{jebens2011}.
 
@@
 Several studies have shown examples where cut cells produce more accurate results when compared to TF coordinates.
-Spurious winds seen in TF coordinates are not present and errors do not increase with steeper terrain \citep{good2014}.
-A comparison of TF and cut cells using real initial data by \citet{steppeler2013} found that precipitation and wind forecasts were more accurately in the cut cell model.
+Spurious winds seen in TF coordinates are not present with cut cells and errors do not increase with steeper terrain \citep{good2014}.
+A comparison of TF and cut cells using real initial data by \citet{steppeler2013} found that five-day forecasts of precipitation and wind over Asia in January 1989 were more accurate in the cut cell model.
 
-This study uses a modification of the nonhydrostatic model from \citet{weller-shahrokhi2014} to enable a like-for-like comparison between terrain following and cut cell grids for idealised, two-dimensional test cases from the literature.
+This study uses a modification of the fully-compressible model from \citet{weller-shahrokhi2014} to enable a like-for-like comparison between terrain following and cut cell grids for idealised, two-dimensional test cases from the literature.
 Section 2 presents the formulation of the terrain following and cut cell grids used in the experiments that follow.
@@
 where, in two dimensions, \(z(x, z^\star)\) is the physical height of the coordinate surface at level \(z^\star\), \(H\) is the height of the domain, and \(h(x)\) is the height of the terrain surface.
-This formulation results in $z^\star/H$ ranging from 0 to 1.
+In this formulation $z^\star/H$ ranges from 0 to 1.
 Using this coordinate, the terrain's influence decays linearly with height but disappears only at the top of the domain.
@@
 
-The hybrid terrain following (HTF) coordinates of \citet{simmons-burridge1981} improve upon BTF coordinates by using a decay function that allows the influence of the terrain to be removed at a specified height, producing flat model layers aloft.
+The hybrid terrain following (HTF) coordinates of \citet{simmons-burridge1981} improve upon BTF coordinates by using a decay function that allows the influence of the terrain to be removed at a specified height, producing model layers aloft that have constant pressure.
 
@@
 	\text{Continuity} &\ &\	\frac{\partial \rho}{\partial t} + \nabla \cdot \rho \bm{u} &= 0 \\
-	\text{\(\rho \theta\) advection} &\ &\ \frac{\partial \rho \theta}{\partial t} + \nabla \cdot \rho \bm{u} \theta &= 0 \\
-	\text{Equation of state} &\ &\ \Pi^{(1 - \kappa)/\kappa} &= \frac{R \rho \theta}{p_0}
+	\text{Thermodynamic equation} &\ &\ \frac{\partial \rho \theta}{\partial t} + \nabla \cdot \rho \bm{u} \theta &= 0 \\
+	\text{Ideal gas law} &\ &\ \Pi^{(1 - \kappa)/\kappa} &= \frac{R \rho \theta}{p_0}
 \end{align}
@@
 
-In order to create a semi-implicit formulation, \cite{weller-shahrokhi2014} split \(H\) into diagonal and off-diagonal components and the off-diagonal components are treated explicitly in the Helmholtz equation.
+In order to create a semi-implicit formulation, \cite{weller-shahrokhi2014} split \(H\) into diagonal and off-diagonal components and the off-diagonal components are treated explicitly in the Helmholtz equation for \(\Pi\).
 However, this was found to be unstable when using a large timestep on a grid with large non-orthogonality.
@@
 \begin{align}
-	\ell^2 &= \sqrt{\frac{\sum_c \left( \varphi - \varphi_{T} \right)^2 \mathcal{V}_c}{\sum_c \left( \varphi_T^2 \mathcal{V}_c \right)}}
+	\ell^2 &= \sqrt{\frac{\sum_c \left( \varphi - \varphi_{T} \right)^2 \mathcal{V}_c}{\sum_c \left( \varphi_T^2 \mathcal{V}_c \right)}} \label{eqn:l2-error}
 \end{align}
@@
 \begin{equation}
-	\Psi(x,z) = - u_0 H \frac{z - h}{H - h}
+	\Psi(x,z) = u_0 H \frac{z - h}{H - h}
 \end{equation}
@@
 \begin{align}
-	u &= - \frac{\partial \Psi}{\partial z} = - u_0 \frac{H}{H - h}, \quad w = \frac{\partial \Psi}{\partial x} = u_0 H \frac{\mathrm{d} h}{\mathrm{d} x} \frac{z - H}{\left( H - h \right)^2}, \nonumber \\
-	\frac{\partial h}{\partial x} &= - h_0 \left[ 
+	u &= \frac{\partial \Psi}{\partial z} = u_0 \frac{H}{H - h}, \quad w = -\frac{\partial \Psi}{\partial x} = u_0 H \frac{\mathrm{d} h}{\mathrm{d} x} \frac{z - h}{\left( H - h \right)^2}, \nonumber \\
+	\frac{\mathrm{d} h}{\mathrm{d} x} &= - h_0 \left[ 
 		\beta \cos^2 \left( \alpha x \right) \sin \left( 2 \beta x \right) +
@@
 Using the cubic upwind-biased scheme, errors are about ten times larger on the SLEVE and cut cell grids compared to the BTF grid.
+Large errors are generated using the centred linear scheme on the cut cell grid due to large courant numbers near the ground.
 
@@
 The initial thermodynamic conditions have a surface temperature of $\theta(z=0) = \SI{288}{\kelvin}$ and constant stability with $N = \SI{0.01}{\per\second}$ everywhere.
-A constant horizontal wind $u = \SI{10}{\meter\per\second}$ is prescribed at the inlet boundary.
+A uniform horizontal wind $u = \SI{10}{\meter\per\second}$ is prescribed at the inlet boundary.
 
@@
 
-Spurious vertical velocities were small in the resting atmosphere test, reaching a maximum of $\sim \SI{0.35}{\meter\per\second}$ on the BTF grid, compared to a maximum of $\sim \SI{7}{\meter\per\second}$ found by \citet{klemp2011}.
+Spurious vertical velocities were small in the resting initial state test, reaching a maximum of $\sim \SI{0.35}{\meter\per\second}$ on the BTF grid, compared to a maximum of $\sim \SI{7}{\meter\per\second}$ found by \citet{klemp2011}.
 In the gravity waves test, vertical velocities were in good agreement with the reference solution from \citet{melvin2010} across all grids.
@@
 \acknowledgments
-I am grateful to my cosupervisors John Methven and Terry Davies for their valuable input, and to Christoph Sch\"{a}r for his assistance in reproducing his advection test results.
+I am grateful to my co-supervisors John Methven and Terry Davies for their valuable input, and to Christoph Sch\"{a}r (ETH) for his assistance in reproducing his advection test results.
 I am thankful for the NERC studentship which helped fund this work.
@@
 \begin{table*}
-	\caption{Minimum and maximum tracer magnitudes and \(\ell^2\) error norms at \(t = \SI{10000}{\second}\) in the horizontal and terrain following tracer advection tests using centred linear and cubic upwind-biased schemes.
-For the horizontal advection test, minimum and maximum values are given for the fourth order scheme using the modified code from \citet{schaer2002}.}
+	\caption{Minimum and maximum tracer magnitudes and \(\ell^2\) error norms (defined by equation~\ref{eqn:l2-error}) at \(t = \SI{10000}{\second}\) in the horizontal and terrain following tracer advection tests using centred linear and cubic upwind-biased schemes.
+For the horizontal advection test, minimum and maximum values are given for the fourth order scheme using the modified code from \citet{schaer2002}.
+\(\ell^2\) error norms were not calculated by this code.}
 \label{tab:advection}
@@
 	\caption{Examples of (a) BTF, (b) SLEVE, and (c) a cut cell grid, showing cell edges in the lowest four layers.
-The two dimensional grids are \SI{20}{\kilo\meter} wide and \SI{20}{\kilo\meter} high.
+The full two dimensional grids are \SI{20}{\kilo\meter} wide and \SI{20}{\kilo\meter} high.
 SLEVE parameters are specified in the resting atmosphere test in section~\ref{sec:results}\ref{sec:resting}.
-The cut cell grid was created by intersecting the terrain surface with a regular grid as described in section~\ref{sec:grid}.}
+The cut cell grid was created by intersecting the terrain surface with a regular grid as described in section~\ref{sec:grid}.
+Axes are in units of \si{\meter}.}
 	\label{fig:grid}
@@
 %
-	\caption{Horizontally advected tracer contours at \(t = \SI{0}{\second}\), \SI{5000}{\second} and \SI{10000}{\second} using (a) centred linear scheme on a BTF grid, the upwind-biased cubic scheme on (b) the cut cell grid, (d) the BTF grid, and (c) the fourth order scheme from \citet{schaer2002} on the BTF grid, with contour intervals every 0.1.
+	\caption{Horizontally advected tracer contours at \(t = \SI{0}{\second}\), \SI{5000}{\second} and \SI{10000}{\second} using (a) centred linear scheme on the BTF grid, (b) the upwind-biased cubic scheme on the cut cell grid, (c) the fourth order scheme from \citet{schaer2002} on the BTF grid, and (d) the upwind-biased cubic scheme on the BTF grid with contour intervals every 0.1.
 Errors on the BTF grid at \(t = \SI{10000}{\second}\) are shown for (e) the fourth order scheme from \citet{schaer2002}, and (f) the upwind-biased cubic scheme, with contour intervals every 0.01.
@@
 The four lowest layers of each grid are shown for (b) BTF, and (d) cut cell grids, using a narrower potential temperature scale.
-The results on the SLEVE grid (not shown) are qualitatively identical to results on the BTF grid.}
+The results on the SLEVE grid (not shown) are qualitatively identical to results on the BTF grid.
+Axes are in units of \si{\meter}.}
 	\label{fig:gw-theta}
@@
 	\caption{Vertical profiles of the Exner function of pressure, \(\Pi\), and potential temperature, \(\theta\), in the gravity waves test with a mountain height of \(h_0 = \SI{500}{\meter}\).
-Exner profile is visually identical on all grids for both mountain heights; for clarity, the Exner profile is only plotted for the BTF grid.
-The computational mode is manifested as a zig-zag in potential temperature on the cut cell grid which.
+The Exner profile is visually identical on all grids for both mountain heights; for clarity, the Exner profile is only plotted for the BTF grid.
+The Lorenz computational mode is manifested as a zig-zag in potential temperature on the cut cell grid.
 Results on the SLEVE grid (not shown) are qualitatively the same as those on the BTF grid.

### 2015-06-10 src/tf-cutcell-comparison/tf-cutcell-comparison.tex
describe eta coords

@@
-\documentclass{ametsoc}
+\documentclass[twocol]{ametsoc}
+\journal{mwr}
 
@@
 A comparison of TF and cut cells using real initial data by \citet{steppeler2013} found that five-day forecasts of precipitation and wind over Asia in January 1989 were more accurate in the cut cell model.
+
+Another alternative method is the eta coordinate, described by \citet{mesinger1988}.
+This transformation, expressed in pressure coordinates, quantises the surface pressure at each grid box using prescribed geometric heights.
+This results in terrain profiles having a staircase pattern which is known as `step' orography.
+The eta coordinate improves the accuracy of the horizontal pressure gradient calculation compared to the sigma coordinate \citep{mesinger1988}.
+
+In an experiment of orographically induced gravity waves, \citet{gallus-klemp2000} found that horizontal flow along the lee slope was artificially weak.
+\citet{mesinger2012} offers an explanation for this behaviour: air flowing along the lee slope cannot travel diagonally downwards but must first travel horizontally, then vertically downward.
+However, lee slope winds are weakened because some of the air continues to be transported horizontally aloft.
+
+\citet{mesinger2012} refined the formulation to allow diagonal transport of momentum and temperature immediately above sloping terrain.
+This arrangement is similar to the finite volume cut cell method.
+The new method improved test results, increasing lee slope winds by \SIrange{4}{5}{\meter\per\second} \citep{mesinger2012}.
 
@@
 \begin{align}
-	u &= \frac{\partial \Psi}{\partial z} = u_0 \frac{H}{H - h}, \quad w = -\frac{\partial \Psi}{\partial x} = u_0 H \frac{\mathrm{d} h}{\mathrm{d} x} \frac{z - h}{\left( H - h \right)^2}, \nonumber \\
+	u &= \frac{\partial \Psi}{\partial z} = u_0 \frac{H}{H - h}, \quad w = -\frac{\partial \Psi}{\partial x} = u_0 H \frac{\mathrm{d} h}{\mathrm{d} x} \frac{z - H}{\left( H - h \right)^2}, \nonumber \\
 	\frac{\mathrm{d} h}{\mathrm{d} x} &= - h_0 \left[ 
@@
 Using the cubic upwind-biased scheme, errors are about ten times larger on the SLEVE and cut cell grids compared to the BTF grid.
-Large errors are generated using the centred linear scheme on the cut cell grid due to large courant numbers near the ground.
+
+Using the centred linear scheme on the cut cell grid, numerical instability is generated by flow through small cells: when interpolating \(\phi\) onto faces, flux from a large cell into a small cell means that the downwind cell is given greater weighting.
+The maximum courant number is less than one in all of the advection tests, hence it is not the source of this instability.
 

### 2015-06-10 src/tf-cutcell-comparison/tf-cutcell-comparison.tex
remove twocol layout
redundant sentence for TF flow

@@
 In an experiment of orographically induced gravity waves, \citet{gallus-klemp2000} found that horizontal flow along the lee slope was artificially weak.
-\citet{mesinger2012} offers an explanation for this behaviour: air flowing along the lee slope cannot travel diagonally downwards but must first travel horizontally, then vertically downward.
+\citet{mesinger2012} offer an explanation for this behaviour: air flowing along the lee slope cannot travel diagonally downwards but must first travel horizontally, then vertically downward.
 However, lee slope winds are weakened because some of the air continues to be transported horizontally aloft.
@@
 To determine the primary source of numerical error, we formulate a new tracer advection test in which the velocity field is everywhere tangential to the basic terrain following coordinate surfaces.
-In this new test, flow is aligned with the BTF grid layers.
 Misalignment on the SLEVE and cut cell grids is designed to challenge the advection scheme.

### 2015-07-03 src/tf-cutcell-comparison/tf-cutcell-comparison.tex
use l2 norm values, not L2

@@
 The results of the cubic upwind-biased scheme on TF and regular grids are comparable with those for the fourth-order centred scheme from \citet{schaer2002}.
-Error is largest on the BTF grid with \(\ell^2 = \num{0.00791}\) but significant reduced on the SLEVE grid with \(\ell^2 = \num{0.00108}\).
+Error is largest on the BTF grid with \(\ell^2 = \num{0.107}\) but is significantly reduced on the SLEVE grid with \(\ell^2 = \num{0.0146}\).
 The error is approximately halved by changing from the SLEVE grid to the cut cell grid.
@@
 
-The centred linear scheme is sensitive to changes in the terrain profile: when the domain width is reduced to \SI{300}{\kilo\meter}, the $\ell^2$ error on the BTF grid rises from \num{0.0210} to \num{0.0318}.
-The upwind-biased cubic scheme is less sensitive to the same change, with the $\ell^2$ error decreasing from \num{0.00791} to \num{0.00767}.
+The centred linear scheme is sensitive to changes in the terrain profile: when the domain width is reduced to \SI{300}{\kilo\meter}, the $\ell^2$ error on the BTF grid rises from \num{0.284} to \num{0.432}.
+The upwind-biased cubic scheme is less sensitive to the same change, with the $\ell^2$ error decreasing from \num{0.107} to \num{0.104}.
 
@@
 	\caption{Minimum and maximum tracer magnitudes and \(\ell^2\) error norms (defined by equation~\ref{eqn:l2-error}) at \(t = \SI{10000}{\second}\) in the horizontal and terrain following tracer advection tests using centred linear and cubic upwind-biased schemes.
-For the horizontal advection test, minimum and maximum values are given for the fourth order scheme using the modified code from \citet{schaer2002}.
-\(\ell^2\) error norms were not calculated by this code.}
+For the horizontal advection test, \(\ell^2\) error norms, minimum and maximum values are given for the fourth order scheme using the modified code from \citet{schaer2002}.}
 \label{tab:advection}
@@
 \hline
-Horizontal        & Centred linear      & \(\ell^2\) error & 0        & 0.0210  & 0.00233         & 0.00224        & 0.00223     \\
+Horizontal        & Centred linear      & \(\ell^2\) error & 0        & 0.284   & 0.0316          & 0.0304         & 0.0304      \\
                   &                     & min              & 0        & -0.275  & -0.0252         & -0.0251        & -0.0251     \\
                   &                     & max              & 1        & 0.925   & 0.985           & 0.985          & 0.985       \\
-                  & Fourth order        & min              & 0        & -0.0926 & -0.00174        & {---}          & -0.00178    \\
+                  & Fourth order        & \(\ell^2\) error & 0        & 0.0938  & 0.00244         & {---}          & 0.00234     \\
+                  &                     & min              & 0        & -0.0926 & -0.00174        & {---}          & -0.00178    \\
                   &                     & max              & 1        & 1.00    & 0.984           & {---}          & 0.983       \\
-                  & Cubic upwind-biased & \(\ell^2\) error & 0        & 0.00791 & 0.00108         & 0.000577       & 0.000576    \\
+                  & Cubic upwind-biased & \(\ell^2\) error & 0        & 0.107   & 0.0146          & 0.00784        & 0.00784     \\
                   &                     & min              & 0        & -0.0446 & -0.0106         & -0.000674      & -0.00674    \\
@@
 \hline
-Terrain following & Centred linear      & \(\ell^2\) error & 0        & 0.00251 & 0.0173          & 13.6           & {---}        \\
+Terrain following & Centred linear      & \(\ell^2\) error & 0        & 0.0341  & 0.235           & 184            & {---}        \\
 		  &                     & min              & 0        & -0.0245 & -0.120          & -623           & {---}        \\
                   &                     & max              & 1        & 0.985   & 0.950           & 3480           & {---}        \\
-		  & Cubic upwind-biased & \(\ell^2\) error & 0        & 0.00154 & 0.0120          & 0.0134         & {---}        \\
+		  & Cubic upwind-biased & \(\ell^2\) error & 0        & 0.0209  & 0.162           & 0.181          & {---}        \\
                   &                     & min              & 0        & -0.0110 & -0.0263         & -0.028         & {---}        \\

### 2015-08-20 src/tf-cutcell-comparison/tf-cutcell-comparison.tex
minor corrections and revisions from reviewer

--- 2015-06-10 src/tf-cutcell-comparison/tf-cutcell-comparison.tex
@@
 \bibpunct{(}{)}{;}{a}{}{,}
+\newcommand{\TODO}[1]{\textcolor{purple}{TODO: \emph{#1}}}
 
@@
 This new test presents a challenge to cut cells in situations with a large Froude number and large Scorer parameter.
+\TODO{discuss Froude number/Scorer parameter in body.}
 The results of the advection tests demonstrate that tracer accuracy depends upon alignment of the flow with the grid.
@@
 Orography also exerts strong non-local influences: from the latent heat release due to convection, by directly forcing planetary waves, and by the atmospheric response to form drag and gravity wave drag.
-There are two main approaches to represent orography on a grid: terrain following layers and cut cells.
-Both methods align cells in vertical columns.
+There are two main approaches to representing orography on a grid: terrain following layers and cut cells.
+\TODO{another way of treating orography are 'immersed boundary methods' or 'embedded boundary methods'}  Both \TODO{all?} methods align cells in vertical columns.
 Because most models are designed for a particular type of grid, existing studies of cut cell solutions have compared results with terrain following grid solutions implemented within different models, for example \citet{good2014}.
@@
 TF layers are usually implemented using a coordinate transformation onto a rectangular computational domain, which introduces metric terms into the equations of motion.
-The techniques proposed by \citet{klemp2011} and \cite{zaengl2012} both involve the calculation of the horizontal pressure gradient in the physical domain.
+The techniques proposed by \citet{klemp2011} and \cite{zaengl2012} both involve the calculation of the horizontal pressure gradient in the physical domain \TODO{interpolate to z-levels in order to compute the horizontal pressure gradient}.
 This gave them the flexibility to design more accurate horizontal pressure gradient discretizations using more appropriate stencils.
@@
 They are attractive because their rectangular structure is simple to process by computer and link with parameterisations, boundary layer resolution can be increased with variable spacing of vertical layers \citep{schaer2002}, and cell sizes remain almost constant \citep{jebens2011}.
+\TODO{clarify; constant cell sizes but variable spacing of vertical layers}
 
@@
 
-The cut cell method can create some very small cells which reduce computational efficiency \citep{klein2009}, and several approaches have been tried to alleviate the problem \citep{steppeler2002,yamazaki-satomura2010,jebens2011}.
+The cut cell method can create some very small cells which reduce computational efficiency \citep{klein2009}, and several approaches have been tried to alleviate the problem.
+\citet{yamazaki-satomura2010} combine small cells with horizontally or vertically adjacent cells.
+\citet{steppeler2002} employ a thin-wall appoximation to increase the computational volume of small cells without altering the terrain.
+\citet{jebens2011} avoid the timestep restriction associated with explicit schemes by using an implicit method for cut cells and a semi-explicit method elsewhere.
 
@@
 
-In an experiment of orographically induced gravity waves, \citet{gallus-klemp2000} found that horizontal flow along the lee slope was artificially weak.
+In an experiment of orographically induced gravity waves, \citet{gallus-klemp2000} found that horizontal flow along the lee slope was artificially weak in the Eta model.
 \citet{mesinger2012} offer an explanation for this behaviour: air flowing along the lee slope cannot travel diagonally downwards but must first travel horizontally, then vertically downward.
@@
 \end{equation}
-where, in two dimensions, \(z(x, z^\star)\) is the physical height of the coordinate surface at level \(z^\star\), \(H\) is the height of the domain, and \(h(x)\) is the height of the terrain surface.
-In this formulation $z^\star/H$ ranges from 0 to 1.
+where, in two dimensions, \(z(x, z^\star)\) is the physical height of the Cartesian coordinate surface at the model level with transformed height \(z^\star\), \(H\) is the height of the domain, and \(h(x)\) is the height of the terrain surface.
+In this formulation $z$ varies between $h$ and $H$ and $z^\star$ ranges from 0 to $H$.
 Using this coordinate, the terrain's influence decays linearly with height but disappears only at the top of the domain.
 An example is shown in figure~\ref{fig:grid}a.
-
-The sigma coordinate transform of \citet{phillips1957} is equivalent to the BTF coordinate transform since they both decay linearly.
-However, because \(\sigma = p/p_s\) decays with pressure rather than height, sigma coordinates also vary with surface pressure, \(p_s(x, t)\).
-
-The hybrid terrain following (HTF) coordinates of \citet{simmons-burridge1981} improve upon BTF coordinates by using a decay function that allows the influence of the terrain to be removed at a specified height, producing model layers aloft that have constant pressure.
 
@@
 Most implementations of terrain following layers use a coordinate system that makes the domain rectangular, but introduces metric terms into the equations of motion.
-Instead, the model employed in this study uses Cartesian coordinates and unstructured grids.
+Instead, the model employed in this study uses Cartesian coordinates and unstructured grids \TODO{they're very structured!}.
 By doing so, results from the same model can be compared between terrain following and cut cell grids without modifying the equation set or discretisation.
@@
 \label{sec:model}
+\TODO{Use model in 2D x-z slice configuration without Coriolis forces}
+\TODO{Add info about the different numerical options that are assessed in combination with the orography treatment}
 
-We use the finite volume model from \citet{weller-shahrokhi2014} which details a C-grid discretisation of the fully-compressible Euler equations, given by
+We use the finite volume model from \citet{weller-shahrokhi2014} which details a discretisation of the fully-compressible Euler equations, given by
 \begin{subequations}
@@
 \end{subequations}
-where \(\rho\) is the density, \(\bm{u}\) is the velocity field, \(\bm{g}\) is the gravitational acceleration, \(c_p\) is the heat capacity at constant pressure, \(\theta = T \left(p_0/p\right)^\kappa\) is the potential temperature, \(T\) is the tempereature, \(p\) is the pressure, \(p_0\) is a reference pressure, \(\Pi = \left(p / p_0 \right)^\kappa\) is the Exner function of pressure, and \(\kappa = R/c_p\) is the gas constant to heat capacity ratio.
+where \(\rho\) is the density, \(\bm{u}\) is the velocity field, \(\bm{g}\) is the gravitational acceleration, \(c_p\) is the heat capacity at constant pressure, \(\theta = T \left(p_0/p\right)^\kappa\) is the potential temperature, \(T\) is the temperature, \(p\) is the pressure, \(p_0\) is a reference pressure, \(\Pi = \left(p / p_0 \right)^\kappa\) is the Exner function of pressure, and \(\kappa = R/c_p\) is the gas constant to heat capacity ratio.
 
-The model uses a Lorenz C-grid staggering of thermodynamic variables such that $\theta$, $\rho$ and $\Pi$ are stored at cell centres and the covariant component of velocity at cell faces.
+The model uses C-grid staggering in the horizontal and Lorenz staggering in the vertical such that $\theta$, $\rho$ and $\Pi$ are stored at cell centres and the covariant component of velocity at cell faces.
 
+\TODO{move this to appendix and clarify}
 A minor change has been made to the curl-free, non-orthogonal formulation as described in \citet{weller-shahrokhi2014} to improve stability on non-orthogonal grids.
@@
 
-For tracer advection tests, the advection equation in flux form, given by
+For tracer advection tests, the advection equation is solved in flux form, given by
 \begin{align}
@@
 \end{align}
-is solved using the upwind-biased multidimensional cubic scheme from \citet{weller-shahrokhi2014} which is non-monotonic and not flux corrected.
+where the tracer density, $\phi$, is interpolated onto cell faces using the upwind-biased multidimensional cubic scheme from \citet{weller-shahrokhi2014} which is non-monotonic and not flux corrected.
 The time derivative is solved using a three-stage, second order Runge-Kutta scheme:
@@
 \begin{align}
-	\varphi^\star &= \varphi^{(n)} + \Delta t f(\phi^{(n)}) \\
-	\varphi^{\star\star} &= \varphi^{(n)} + \frac{\Delta t}{2} \left( f(\phi^{(n)}) + f(\phi^\star) \right) \\
-	\varphi^{(n+1)} &= \varphi^{(n)} + \frac{\Delta t}{2} \left( f(\phi^{(n)}) + f(\phi^{\star\star}) \right)
+	\phi^\star &= \phi^{(n)} + \Delta t f(\phi^{(n)}) \\
+	\phi^{\star\star} &= \phi^{(n)} + \frac{\Delta t}{2} \left( f(\phi^{(n)}) + f(\phi^\star) \right) \\
+	\phi^{(n+1)} &= \phi^{(n)} + \frac{\Delta t}{2} \left( f(\phi^{(n)}) + f(\phi^{\star\star}) \right)
 \end{align}
 \end{subequations}
-where \(f(\phi^{(n)}) = - \nabla \cdot (\bm{u} \varphi^{(n)})\) at time level \(n\).
+where \(f(\phi^{(n)}) = - \nabla \cdot (\bm{u} \phi^{(n)})\) at time level \(n\).
 
@@
 \end{subequations}
-where $a = \SI{25}{\kilo\meter}$ is the mountain envelope half-width, $h_0 = \SI{3}{\kilo\meter}$ is the maximum mountain height, $\lambda = \SI{8}{\kilo\meter}$ is the wavelength, \(\alpha = \pi / \lambda\) and \(\beta = \pi / 2a\).
+where $a = \SI{25}{\kilo\meter}$ is the mountain envelope half-width, $h_0 = \SI{3}{\kilo\meter}$ is the maximum mountain height, $\lambda = \SI{8}{\kilo\meter}$ is the wavelength, \(\alpha = \pi / \lambda\) and \(\beta = \pi / (2a)\).
 On the SLEVE grid, the large-scale component $h_1$ is given by \(h_1(x) = h^\star(x) / 2\)
@@
 where $u_0 = \SI{10}{\meter\per\second}$, $z_1 = \SI{4}{\kilo\meter}$ and $z_2 = \SI{5}{\kilo\meter}$.
-This results in a constant wind aloft, and zero flow at \SI{4}{\kilo\meter} and below.
+This results in a constant wind above $z_2$, and zero flow at \SI{4}{\kilo\meter} and below.
 
+\TODO{double check signs of all streamfunctions (in horizontal and TF advection test cases)}
 To ensure that the discrete velocity field is non-divergent, velocities are prescribed at cell faces by differencing the streamfunction, \(\Psi\), along the edges from \(\Psi\) stored at cell vertices.
-Since \(u = - \partial \Psi / \partial z\), the streamfunction is found by vertical integration of the velocity profile:
+Since \(u = \partial \Psi / \partial z\), the streamfunction is found by vertical integration of the velocity profile:
 \begin{align}
@@
 
-A tracer, $\varphi$, is positioned upstream above the height of the terrain.  It has the shape
+A tracer with density $\phi$ is positioned upstream above the height of the terrain.  It has the shape
 \begin{align}
-	\varphi(x, z) &= \varphi_0 \left\{ \begin{array}{l l}
+	\phi(x, z) &= \phi_0 \left\{ \begin{array}{l l}
 		\cos^2 \left( \frac{\pi r}{2} \right) & \quad \text{if $r \leq 1$} \\
@@
 \end{align}
-where $A_x = \SI{25}{\kilo\meter}$, $A_z = \SI{3}{\kilo\meter}$ are the horizontal and vertical half-widths respectively, and $\varphi_0 = 1$ is the maximum magnitude of the anomaly.
-At $t = \SI{0}{\second}$, the anomaly is centred at $(x_0, z_0) = (\SI{-50}{\kilo\meter}, \SI{9}{\kilo\meter})$ so that the anomaly is upwind of the mountain and well above the maximum terrain height of \SI{3}{\kilo\meter}.
-Analytic solutions can be found by setting the anomaly centre such that $x_0 = ut$.
+where $A_x = \SI{25}{\kilo\meter}$, $A_z = \SI{3}{\kilo\meter}$ are the horizontal and vertical half-widths respectively, and $\phi_0 = 1$ is the maximum density of the tracer.
+At $t = \SI{0}{\second}$, the tracer is centred at $(x_0, z_0) = (\SI{-50}{\kilo\meter}, \SI{9}{\kilo\meter})$ so that the tracer is upwind of the mountain and well above the maximum terrain height of \SI{3}{\kilo\meter}.
+Analytic solutions can be found by setting the tracer centre such that $x_0 = ut$.
 
@@
 The modified version used here directly calculates the height at vertices to enable comparisons with the finite volume model solutions.
+\TODO{move this detail into section 3 and add more info and references}
 
@@
 
-By \(t = \SI{10000}{\second}\), the tracer suffers from distortion on the BTF grid using the centred linear scheme and some artefacts remain about the mountain peak.
+By \(t = \SI{10000}{\second}\), the tracer suffers from distortion on the BTF grid using the centred linear scheme and some artifacts remain about the mountain peak.
 The tracer has spread vertically due to increased numerical errors when the tracer is transported between layers.
@@
 
-Minimum and maximum tracer values and \(\ell^2\) error norms on the BTF, SLEVE, cut cell and regular grids are summarised in table~\ref{tab:advection}, where the \(\ell^2\) error norm is defined as 
+Minimum and maximum tracer values and \(\ell_2\) error norms on the BTF, SLEVE, cut cell and regular grids are summarised in table~\ref{tab:advection}, where the \(\ell_2\) error norm is defined as 
 \begin{align}
-	\ell^2 &= \sqrt{\frac{\sum_c \left( \varphi - \varphi_{T} \right)^2 \mathcal{V}_c}{\sum_c \left( \varphi_T^2 \mathcal{V}_c \right)}} \label{eqn:l2-error}
+	\ell_2 &= \sqrt{\frac{\sum_c \left( \phi - \phi_{T} \right)^2 \mathcal{V}_c}{\sum_c \left( \phi_T^2 \mathcal{V}_c \right)}} \label{eqn:l2-error}
 \end{align}
-where $\varphi$ is the numerical tracer value, $\varphi_T$ is the analytic value and $\mathcal{V}_c$ is the cell volume.
+where $\phi$ is the numerical tracer value, $\phi_T$ is the analytic value and $\mathcal{V}_c$ is the cell volume.
 
 The results of the cubic upwind-biased scheme on TF and regular grids are comparable with those for the fourth-order centred scheme from \citet{schaer2002}.
-Error is largest on the BTF grid with \(\ell^2 = \num{0.00791}\) but significant reduced on the SLEVE grid with \(\ell^2 = \num{0.00108}\).
+Error is largest on the BTF grid with \(\ell_2 = \num{0.107}\) but is significantly reduced on the SLEVE grid with \(\ell_2 = \num{0.0146}\).
 The error is approximately halved by changing from the SLEVE grid to the cut cell grid.
@@
 
-The centred linear scheme is sensitive to changes in the terrain profile: when the domain width is reduced to \SI{300}{\kilo\meter}, the $\ell^2$ error on the BTF grid rises from \num{0.0210} to \num{0.0318}.
-The upwind-biased cubic scheme is less sensitive to the same change, with the $\ell^2$ error decreasing from \num{0.00791} to \num{0.00767}.
+The centred linear scheme is sensitive to changes in the terrain profile: when the domain width is reduced to \SI{300}{\kilo\meter}, the $\ell_2$ error on the BTF grid rises from \num{0.284} to \num{0.432}.
+The upwind-biased cubic scheme is less sensitive to the same change, with the $\ell_2$ error decreasing from \num{0.107} to \num{0.104}.
 
@@
 
-$\ell^2$ errors and tracer extrema for this test are compared with the horizontal advection results in table~\ref{tab:advection}.
+$\ell_2$ errors and tracer extrema for this test are compared with the horizontal advection results in table~\ref{tab:advection}.
 In the terrain following velocity field, tracer accuracy is greatest on the BTF grid.
@@
 Using the centred linear scheme on the cut cell grid, numerical instability is generated by flow through small cells: when interpolating \(\phi\) onto faces, flux from a large cell into a small cell means that the downwind cell is given greater weighting.
-The maximum courant number is less than one in all of the advection tests, hence it is not the source of this instability.
+The maximum Courant number is less than one in all of the advection tests, hence it is not the source of this instability.
 
@@
 
-The test setup follows the specification by \cite{klemp2011}.  The grid resolution is \(\Delta x = \Delta z^\star = \SI{500}{\meter}\).  All boundary conditions are no normal flow.
+The test setup follows the specification by \cite{klemp2011}.
+\TODO{how large is the domain?
+do I say this elsewhere?
+if not, say it here} The grid resolution is \(\Delta x = \Delta z^\star = \SI{500}{\meter}\).
+All boundary conditions are no normal flow.
 
@@
 Following \citet{melvin2010}, the domain is \SI{300}{\kilo\meter} wide and \SI{30}{\kilo\meter} high.
-The mountain profile has the same form as equation~\ref{eqn:resting:mountain}.
+\TODO{state grid spacing} The mountain profile has the same form as equation~\ref{eqn:resting:mountain}.
 Tests are performed with mountain heights of $h_0 = \SI{250}{\meter}$ and \(h_0 = \SI{500}{\meter}\).
@@
 \begin{table*}
-	\caption{Minimum and maximum tracer magnitudes and \(\ell^2\) error norms (defined by equation~\ref{eqn:l2-error}) at \(t = \SI{10000}{\second}\) in the horizontal and terrain following tracer advection tests using centred linear and cubic upwind-biased schemes.
-For the horizontal advection test, minimum and maximum values are given for the fourth order scheme using the modified code from \citet{schaer2002}.
-\(\ell^2\) error norms were not calculated by this code.}
+	\caption{Minimum and maximum tracer magnitudes and \(\ell_2\) error norms (defined by equation~\ref{eqn:l2-error}) at \(t = \SI{10000}{\second}\) in the horizontal and terrain following tracer advection tests using centred linear and cubic upwind-biased schemes.
+For the horizontal advection test, \(\ell_2\) error norms, minimum and maximum values are given for the fourth order scheme using the modified code from \citet{schaer2002}.}
 \label{tab:advection}
@@
 \hline
-Horizontal        & Centred linear      & \(\ell^2\) error & 0        & 0.0210  & 0.00233         & 0.00224        & 0.00223     \\
+Horizontal        & Centred linear      & \(\ell_2\) error & 0        & 0.284   & 0.0316          & 0.0304         & 0.0304      \\
                   &                     & min              & 0        & -0.275  & -0.0252         & -0.0251        & -0.0251     \\
                   &                     & max              & 1        & 0.925   & 0.985           & 0.985          & 0.985       \\
-                  & Fourth order        & min              & 0        & -0.0926 & -0.00174        & {---}          & -0.00178    \\
+                  & Fourth order        & \(\ell_2\) error & 0        & 0.0938  & 0.00244         & {---}          & 0.00234     \\
+                  &                     & min              & 0        & -0.0926 & -0.00174        & {---}          & -0.00178    \\
                   &                     & max              & 1        & 1.00    & 0.984           & {---}          & 0.983       \\
-                  & Cubic upwind-biased & \(\ell^2\) error & 0        & 0.00791 & 0.00108         & 0.000577       & 0.000576    \\
+                  & Cubic upwind-biased & \(\ell_2\) error & 0        & 0.107   & 0.0146          & 0.00784        & 0.00784     \\
                   &                     & min              & 0        & -0.0446 & -0.0106         & -0.000674      & -0.00674    \\
@@
 \hline
-Terrain following & Centred linear      & \(\ell^2\) error & 0        & 0.00251 & 0.0173          & 13.6           & {---}        \\
-		  &                     & min              & 0        & -0.0245 & -0.120          & -623           & {---}        \\
-                  &                     & max              & 1        & 0.985   & 0.950           & 3480           & {---}        \\
-		  & Cubic upwind-biased & \(\ell^2\) error & 0        & 0.00154 & 0.0120          & 0.0134         & {---}        \\
+Terrain following & Centred linear      & \(\ell_2\) error & 0        & 0.0341  & 0.235           & {unstable}     & {---}        \\
+	          &                     & min              & 0        & -0.0245 & -0.120          & {unstable}     & {---}        \\
+		  &                     & max              & 1        & 0.985   & 0.950           & {unstable}     & {---}        \\
+		  & Cubic upwind-biased & \(\ell_2\) error & 0        & 0.0209  & 0.162           & 0.181          & {---}        \\
                   &                     & min              & 0        & -0.0110 & -0.0263         & -0.028         & {---}        \\

### 2015-08-21 src/tf-cutcell-comparison/tf-cutcell-comparison.tex
move nonortho description into appendix, include mu in momentum eqn

@@
 There are two main approaches to representing orography on a grid: terrain following layers and cut cells.
-\TODO{another way of treating orography are 'immersed boundary methods' or 'embedded boundary methods'}  Both \TODO{all?} methods align cells in vertical columns.
+Another way of treating orography is the immersed (or embedded) boundary method \citep{simon2012}.
+All methods align cells in vertical columns.
 Because most models are designed for a particular type of grid, existing studies of cut cell solutions have compared results with terrain following grid solutions implemented within different models, for example \citet{good2014}.
@@
 \begin{align}
-	\text{Momentum} &\ &\  	\frac{\partial \rho \bm{u}}{\partial t} + \nabla \cdot \rho \bm{uu} &= \rho \bm{g} - c_p \rho \theta \nabla \Pi \label{eq:momentum} \\
+	\text{Momentum} &\ &\  	\frac{\partial \rho \bm{u}}{\partial t} + \nabla \cdot \rho \bm{uu} &= \rho \bm{g} - c_p \rho \theta \nabla \Pi - \mu \rho \bm{u} \label{eq:momentum} \\
 	\text{Continuity} &\ &\	\frac{\partial \rho}{\partial t} + \nabla \cdot \rho \bm{u} &= 0 \\
@@
 where \(\rho\) is the density, \(\bm{u}\) is the velocity field, \(\bm{g}\) is the gravitational acceleration, \(c_p\) is the heat capacity at constant pressure, \(\theta = T \left(p_0/p\right)^\kappa\) is the potential temperature, \(T\) is the temperature, \(p\) is the pressure, \(p_0\) is a reference pressure, \(\Pi = \left(p / p_0 \right)^\kappa\) is the Exner function of pressure, and \(\kappa = R/c_p\) is the gas constant to heat capacity ratio.
+\(\mu\) is a dimensionless damping function used for the sponge layer in the gravity waves test in section~\ref{sec:results}\ref{sec:gw}.
 
 The model uses C-grid staggering in the horizontal and Lorenz staggering in the vertical such that $\theta$, $\rho$ and $\Pi$ are stored at cell centres and the covariant component of velocity at cell faces.
-
-\TODO{move this to appendix and clarify}
-A minor change has been made to the curl-free, non-orthogonal formulation as described in \citet{weller-shahrokhi2014} to improve stability on non-orthogonal grids.
-\citet{weller-shahrokhi2014} use a Hodge dual operator, \(H\), to map from co-variant to contra-variant velocity components:
-\begin{align}
-	U = HV
-\end{align}
-where \(U = \rho\bm{u}\cdot\bm{S}_f\) and \(V=\rho\bm{u}\cdot\bm{d}_f\).
-\(\bm{S}_f\) is the surface normal vector for a face, \(f\), which has a magnitude of the surface area, and \(\bm{d}_f\) is the vector between neighbouring cell centres.
-
-In order to create a semi-implicit formulation, \cite{weller-shahrokhi2014} split \(H\) into diagonal and off-diagonal components and the off-diagonal components are treated explicitly in the Helmholtz equation for \(\Pi\).
-However, this was found to be unstable when using a large timestep on a grid with large non-orthogonality.
-Improved stability is achieved by splitting \(H\) into the orthogonal part and the non-orthogonal correction:
-\begin{align}
-	H = H_c + H_\mathrm{corr}
-\end{align}
-where the diagonal matrix \(H_c=|\bm{S}_f|/|\bm{d}_f|\) and the non-orthogonal correction is simply \(H_\mathrm{corr} = H - H_c\).
-The orthogonal part, \(H_c\), is treated implicitly in the Helmholtz equation and the non-orthogonal part, \(H_\mathrm{corr}\), is treated explicitly.
+A minor change has been made to the non-orthogonal formulation as described in \citet{weller-shahrokhi2014} to improve stability on non-orthogonal grid, and this is detailed in appendix A.
 
@@
 The initial thermodynamic conditions are in discrete hydrostatic balance, having a reference potential temperature of $\theta(z = 0) = \SI{288}{\kelvin}$ and constant stability with Brunt-V\"ais\"al\"a frequency $N = \SI{0.01}{\per\second}$ everywhere, except for a more stable layer of $N = \SI{0.02}{\per\second}$ between $\SI{2}{\kilo\meter} \leq z \leq \SI{3}{\kilo\meter}$.
+The damping function, \(\mu\), is set to zero.
 Unlike \citet{klemp2011}, there is no eddy diffusion in the equation set.
@@
 \subsection{Gravity waves}
+\label{sec:gw}
 The test originally specified by \citet{schaer2002} prescribes flow over terrain with small-scale and large-scale undulations which induces propagating and evanescent gravity waves.
@@
 Sponge layers are added to the upper \SI{10}{\kilo\meter} and leftmost \SI{10}{\kilo\meter} at the inlet boundary to damp the reflection of waves.
-The term $\mu \rho \bm{u}$ is subtracted from the momentum (equation \ref{eq:momentum}) where the damping function, \(\mu\), is adapted from \citet{melvin2010} such that
+The damping function, \(\mu\), is adapted from \citet{melvin2010} such that
 \begin{align}
@@
 Weller is funded by NERC grant NE/H015698/1.
+
+% APPENDICES
+\appendix[A]
+\appendixtitle{Non-orthogonal formulation}
+
+\TODO{Hilary will redraft this appendix}
+A minor change has been made to the curl-free, non-orthogonal formulation as described in \citet{weller-shahrokhi2014} to improve stability on non-orthogonal grids.
+\citet{weller-shahrokhi2014} use a Hodge dual operator, \(H\), to map from co-variant to contra-variant velocity components:
+\begin{align}
+	U = HV
+\end{align}
+where \(U = \rho\bm{u}\cdot\bm{S}_f\) and \(V=\rho\bm{u}\cdot\bm{d}_f\).
+\(\bm{S}_f\) is the surface normal vector for a face, \(f\), which has a magnitude of the surface area, and \(\bm{d}_f\) is the vector between neighbouring cell centres.
+
+In order to create a semi-implicit formulation, \cite{weller-shahrokhi2014} split \(H\) into diagonal and off-diagonal components and the off-diagonal components are treated explicitly in the Helmholtz equation for \(\Pi\).
+However, this was found to be unstable when using a large timestep on a grid with large non-orthogonality.
+Improved stability is achieved by splitting \(H\) into the orthogonal part and the non-orthogonal correction:
+\begin{align}
+	H = H_c + H_\mathrm{corr}
+\end{align}
+where the diagonal matrix \(H_c=|\bm{S}_f|/|\bm{d}_f|\) and the non-orthogonal correction is simply \(H_\mathrm{corr} = H - H_c\).
+The orthogonal part, \(H_c\), is treated implicitly in the Helmholtz equation and the non-orthogonal part, \(H_\mathrm{corr}\), is treated explicitly.
 

### 2015-08-24 src/tf-cutcell-comparison/tf-cutcell-comparison.tex
further minor additions and clarifications

@@
 TF layers are usually implemented using a coordinate transformation onto a rectangular computational domain, which introduces metric terms into the equations of motion.
-The techniques proposed by \citet{klemp2011} and \cite{zaengl2012} both involve the calculation of the horizontal pressure gradient in the physical domain \TODO{interpolate to z-levels in order to compute the horizontal pressure gradient}.
+The techniques proposed by \citet{klemp2011} and \cite{zaengl2012} both involve interpolation onto $z$-levels in order to calculate the horizontal pressure gradient.
 This gave them the flexibility to design more accurate horizontal pressure gradient discretizations using more appropriate stencils.
@@
 Despite their associated numerical errors, TF layers are in widespread operational use \citep{steppeler2003}.
-They are attractive because their rectangular structure is simple to process by computer and link with parameterisations, boundary layer resolution can be increased with variable spacing of vertical layers \citep{schaer2002}, and cell sizes remain almost constant \citep{jebens2011}.
-\TODO{clarify; constant cell sizes but variable spacing of vertical layers}
+They are attractive because their rectangular structure is simple to process by computer and link with parameterisations, and boundary layer resolution can be increased with variable spacing of vertical layers \citep{schaer2002}.
 
@@
 
-Most implementations of terrain following layers use a coordinate system that makes the domain rectangular, but introduces metric terms into the equations of motion.
-Instead, the model employed in this study uses Cartesian coordinates and unstructured grids \TODO{they're very structured!}.
+Most implementations of terrain following layers use a coordinate system that makes the computational domain rectangular, but introduces metric terms into the equations of motion.
+Instead, the model employed in this study uses Cartesian coordinates and non-orthogonal grids.
 By doing so, results from the same model can be compared between terrain following and cut cell grids without modifying the equation set or discretisation.
@@
 
-\TODO{double check signs of all streamfunctions (in horizontal and TF advection test cases)}
 To ensure that the discrete velocity field is non-divergent, velocities are prescribed at cell faces by differencing the streamfunction, \(\Psi\), along the edges from \(\Psi\) stored at cell vertices.
@@
 \begin{align}
-	u &= \frac{\partial \Psi}{\partial z} = u_0 \frac{H}{H - h}, \quad w = -\frac{\partial \Psi}{\partial x} = u_0 H \frac{\mathrm{d} h}{\mathrm{d} x} \frac{z - H}{\left( H - h \right)^2}, \nonumber \\
+	u &= \frac{\partial \Psi}{\partial z} = u_0 \frac{H}{H - h}, \quad w = -\frac{\partial \Psi}{\partial x} = u_0 H \frac{\mathrm{d} h}{\mathrm{d} x} \frac{H - z}{\left( H - h \right)^2}, \nonumber \\
 	\frac{\mathrm{d} h}{\mathrm{d} x} &= - h_0 \left[ 
@@
 The test setup follows the specification by \cite{klemp2011}.
-\TODO{how large is the domain?
-do I say this elsewhere?
-if not, say it here} The grid resolution is \(\Delta x = \Delta z^\star = \SI{500}{\meter}\).
+The domain is \SI{200}{\kilo\meter} wide and \SI{20}{\kilo\meter} high, and the grid resolution is \(\Delta x = \Delta z^\star = \SI{500}{\meter}\).
 All boundary conditions are no normal flow.
@@
 
-Following \citet{melvin2010}, the domain is \SI{300}{\kilo\meter} wide and \SI{30}{\kilo\meter} high.
-\TODO{state grid spacing} The mountain profile has the same form as equation~\ref{eqn:resting:mountain}.
+Following \citet{melvin2010}, the domain is \SI{300}{\kilo\meter} wide and \SI{30}{\kilo\meter} high with a grid resolution of \(\Delta x = \SI{500}{\meter}\) and \(\Delta z^\star = \SI{300}{\meter}\).
+\TODO{we want to consider dz=250m and dz=200m, too} The mountain profile has the same form as equation~\ref{eqn:resting:mountain}.
 Tests are performed with mountain heights of $h_0 = \SI{250}{\meter}$ and \(h_0 = \SI{500}{\meter}\).

### 2015-09-01 src/tf-cutcell-comparison/tf-cutcell-comparison.tex
say model is 2D x-z slice, no Coriolis

@@
 \label{sec:model}
-\TODO{Use model in 2D x-z slice configuration without Coriolis forces}
 \TODO{Add info about the different numerical options that are assessed in combination with the orography treatment}
@@
 
-The model uses C-grid staggering in the horizontal and Lorenz staggering in the vertical such that $\theta$, $\rho$ and $\Pi$ are stored at cell centres and the covariant component of velocity at cell faces.
+The model is operated in a two-dimensional $x$--$z$ slice configuration without Coriolis forces.
+The model uses the C-grid staggering in the horizontal and the Lorenz staggering in the vertical such that $\theta$, $\rho$ and $\Pi$ are stored at cell centres and the covariant component of velocity at cell faces.
 A minor change has been made to the non-orthogonal formulation as described in \citet{weller-shahrokhi2014} to improve stability on non-orthogonal grid, and this is detailed in appendix A.

### 2015-09-17 src/tf-cutcell-comparison/tf-cutcell-comparison.tex
clarify cell centres are centroids

@@
 \email{js102@zepler.net}
+
+% Every term in the discretised equations of motion is a potential source of numerical error.
+% Errors in calculating the horizontal pressure gradient generate spurious circulations and so much effort has been put into improving this calculation.
+This is often achieved by making the grid more orthogonal, flattening terrain following coordinate surfaces [citations] or using a cut cell approach [citations].
+% TF citations: mahrer1984, schaer2002, klemp2011
+% cut cell citations: good2014, yamazaki2015 puts this clearly but not yet published
+% In a test of a stably stratified atmosphere initially at rest, we find that spurious velocities are no more than \SI{0.1}{\meter\per\second} on any grid.
+[TODO: and this is much better than klemp2011]
+% Advection terms also contribute to numerical error.
+We show that advection is most accurate when the flow is aligned with the grid layers.
+Near the surface, flow tends to follow the terrain which results in larger errors when more orthogonal grids are used.
+% In a test of orographically induced gravity waves, we find that the advection of potential temperature generates grid-scale errors on the cut cell grid.
+The errors persist because the Lorenz staggering of variables renders them invisible to the model.
+[TODO: I don't think we have enough evidence to support this final sentence]
 
@@
 The model is operated in a two-dimensional $x$--$z$ slice configuration without Coriolis forces.
-The model uses the C-grid staggering in the horizontal and the Lorenz staggering in the vertical such that $\theta$, $\rho$ and $\Pi$ are stored at cell centres and the covariant component of velocity at cell faces.
+The model uses the C-grid staggering in the horizontal and the Lorenz staggering in the vertical such that $\theta$, $\rho$ and $\Pi$ are stored at cell centroids and the covariant component of velocity at cell faces.
 A minor change has been made to the non-orthogonal formulation as described in \citet{weller-shahrokhi2014} to improve stability on non-orthogonal grid, and this is detailed in appendix A.
@@
 \subsection{Terrain following advection}
+\TODO{make it clearer that we're trying to determine if non-orthogonality or misalignment of the flow is the primary source of error}
+
 In the horizontal advection test, results were more accurate where the flow was aligned with the grid layers, and distortions in the BTF grid led to increased errors.
@@
 
-\citet{good2014} found the maximum vertical velocity in their cut cell model was \SI{1e-12}{\meter\per\second}, which is better than any result obtained using the model by \citet{weller-shahrokhi2014}.
-It is worth noting that, in the model used by \citet{good2014}, cell centres are in the centre of the uncut cell, resulting in the centre of some cut cells being below the ground (S.-J.
+\citet{good2014} found the maximum vertical velocity in their cut cell model was \SI{1e-12}{\meter\per\second}, which is better than any result obtained here.
+It is worth noting that the model by \citet{weller-shahrokhi2014} stores values at the geometric centre of cut cells.
+However, in the model used by \citet{good2014}, cell centres are in the centre of the uncut cell, resulting in the centre of some cut cells being below the ground (S.-J.
 Lock 2014, personal communication).

### 2015-09-17 src/tf-cutcell-comparison/tf-cutcell-comparison.tex
remove sampleLine diagram and discussion of it, remove discussion of h0=500m

@@
 This feature is not present on the BTF grid (figure~\ref{fig:gw-theta}b) or the SLEVE grid (not shown).
-
-In a further test, the mountain height is doubled from \SI{250}{\meter} to \SI{500}{\meter} with all other parameter values unchanged.
-The same spurious anomaly in potential temperature is again present on the cut cell grid but its amplitude increases.
-Figure~\ref{fig:gw-exner-theta} shows vertical profiles of the Exner function of pressure and potential temperature in the lowest \SI{1}{\kilo\meter} in the lee of the mountain at \(x = \SI{50}{\kilo\meter}\).
-We find that the decrease in the Exner function of pressure with height is close to linear on all grids.
-The potential temperature increases linearly with height on the BTF and SLEVE grids.
-On the cut cell grid, the potential temperature anomalies seen in figure~\ref{fig:gw-theta}d appear as a zig--zag in figure~\ref{fig:gw-exner-theta}.
-
-The inaccuracies in potential temperature become invisible when calculating momentum because $\theta$ is interpolated onto cell faces, so discrete hydrostatic balance is maintained.
-This is a manifestation of the Lorenz computational mode \citep{arakawa-konor1996,holdaway2013b}.
-In models that include moist processes, the Lorenz computational mode can disrupt clouds and generate spurious precipitation \citep{hollingsworth1995}.
+\TODO{not present?
+I think it is, just much smaller amplitude}
 
 To summarize, results of the gravity waves test on all grids are in good agreement with the reference solution from \citet{melvin2010}.
-The most prominent errors are found only on the cut cell grid, where the potential temperature errors near the ground excite the Lorenz computational mode.
 
@@
 
-\begin{figure}
-	\centering
-	\includegraphics{../fig-gravityWaves-sampleLine/fig-gravityWaves-sampleLine.pdf}
-%
-	\caption{Vertical profiles of the Exner function of pressure, \(\Pi\), and potential temperature, \(\theta\), in the gravity waves test with a mountain height of \(h_0 = \SI{500}{\meter}\).
-The Exner profile is visually identical on all grids for both mountain heights; for clarity, the Exner profile is only plotted for the BTF grid.
-The Lorenz computational mode is manifested as a zig-zag in potential temperature on the cut cell grid.
-Results on the SLEVE grid (not shown) are qualitatively the same as those on the BTF grid.
-The thermal profile with a lesser mountain height of \(h_0 = \SI{250}{\meter}\) (not shown) exhibits a computational mode with smaller amplitude.}
-	\label{fig:gw-exner-theta}
-\end{figure}
-
 \end{document}

### 2015-09-17 src/tf-cutcell-comparison/tf-cutcell-comparison.tex
remove conclusions about Lorenz computational mode

@@
 % In a test of orographically induced gravity waves, we find that the advection of potential temperature generates grid-scale errors on the cut cell grid.
-The errors persist because the Lorenz staggering of variables renders them invisible to the model.
-[TODO: I don't think we have enough evidence to support this final sentence]
 
@@
 As expected, the cut cell grid maintains greater accuracy in the test of stationary, stratified flow.
-In the gravity waves test, results on all grids are in good agreement with existing results from the literature.
-However, it is found that the Lorenz computational mode is excited as a vertical zig--zag in potential temperature on the cut cell grid.}
+In the gravity waves test, results on all grids are in good agreement with existing results from the literature.}
 
@@
 Conversely, in the terrain following advection test, errors were large on the SLEVE and cut cell grids where velocities were misaligned with the grids.
-In the gravity waves test, the Lorenz computational mode was manifested as a zig--zag in potential temperature in the lowest layers in the lee of the mountain.
-This spurious error was excited only on the cut cell grid.
-This test motivates further work to formulate a Charney--Phillips staggering of variables on cut cell grids.
 

### 2015-09-17 src/tf-cutcell-comparison/WellerAppendix.tex
I am using a draft document class so that I can compile it without errors
I have changed $\bm{u}$ to $\mathbf{u}$ (I think this style makes the vectors stand out more. But you can change back if you disagree)

I have changed $\mathbf{uu}$ to $\mathbf{u}\otimes\mathbf{u}$ for clarity

I have removed your ending to the Model section and replaced it with one which cites the appendix

I have removed your appendix and replaced it with mine, in a different file to avoid conflicts. I will now continue to work on the appendix and hopefully won't make many more changes to the main file

\section{Semi-implicit treatment of the Hodge operator\label{appx:SI}}

In order to ensure curl-free pressure gradients, following \citet{WS14},
the momentum at the cell face in the direction between cell centres
is used as the prognostic variable for velocity:
\begin{equation}
V_{f}=\rho_{f}\mathbf{u}_{f}\cdot\mathbf{d}_{f}
\end{equation}
where $\mathbf{d}_{f}$ is the vector between cell centres and subscript
$f$ means ``at face $f$''. Fluxes across faces are diagnostic
variables:
\begin{equation}
U_{f}=\rho_{f}\mathbf{u}_{f}\cdot\mathbf{S}_{f}
\end{equation}
where $\mathbf{S}_{f}$ is the outward pointing normal vector to face
$f$ with magnitude equal to the area of the face. $U$ is the vector
of all values of $U_{f}$ and $V$ is the vector of all values of
$V_{f}$ then we can define the Hodge operator: a matrix to transform
$V$ to $U$:
\begin{equation}
U=HV.
\end{equation}
For energy conservation, \citet{TC12} showed that the Hodge operator
must be symmetric and positive definite. We define a symmetric $H$
suitable for arbitrary 3D meshes:
\begin{equation}
U_{f}=(\rho\mathbf{u})_{F}\cdot\mathbf{S}_{f}
\end{equation}
where subscript $F$ implies mid-point interpolation from two surrounding
cell values onto face $f$:
\begin{equation}
(\rho\mathbf{u})_{F}=\frac{1}{2}\sum_{c\in f}(\rho\mathbf{u})_{C}
\end{equation}
where $c\in f$ means the two cells either side of cell $c$ and where
$(\rho\mathbf{u})_{C}$ is the consistent cell centre reconstruction
of $\rho\mathbf{u}$ from surrounding values of $V_{f}$:
\[
(\rho\mathbf{u})_{C}=\left(\sum_{f^{\prime}\in c}\mathbf{d}_{f^{\prime}}\otimes\mathbf{d}_{f^{\prime}}^{T}\right)^{-1}\sum_{f^{\prime}\in c}\mathbf{d}_{f^{\prime}}V_{f^{\prime}}
\]
where $\mathbf{d}_{f^{\prime}}\otimes\mathbf{d}_{f^{\prime}}^{T}$
is a $3\times3$ tensor and so the inversion of the tensor sum is
a local operation which can be calculated once for each cell of the
grid before time-stepping begins.

### 2015-09-17 src/tf-cutcell-comparison/model.tex
I am using a draft document class so that I can compile it without errors
I have changed $\bm{u}$ to $\mathbf{u}$ (I think this style makes the vectors stand out more. But you can change back if you disagree)

I have changed $\mathbf{uu}$ to $\mathbf{u}\otimes\mathbf{u}$ for clarity

I have removed your ending to the Model section and replaced it with one which cites the appendix

I have removed your appendix and replaced it with mine, in a different file to avoid conflicts. I will now continue to work on the appendix and hopefully won't make many more changes to the main file

Acoustic and gravity waves are treated implicitly and advection is treated explicitly. The trapezoidal implicit treatment of fast waves and the Hodge operator suitable for non-orthogonal grids are described in appendix \ref{appx:SI}. To avoid time-splitting errors between the advection and the fast waves, the advection is time-stepped using a three-stage, second-order Runge-Kutta scheme. When considering only the linear advection of a tracer, $\phi$:
\begin{align}
\partial \phi / \partial t + \nabla \cdot \left( \mathbf{u} \phi \right) = 0
\label{eq:advect}
\end{align}
the Runge-Kutta scheme is defined as:
\begin{subequations}
\begin{align}
	\phi^\star &= \phi^{(n)} + \Delta t f(\phi^{(n)}) \\
	\phi^{\star\star} &= \phi^{(n)} + \frac{\Delta t}{2} \left( f(\phi^{(n)}) + f(\phi^\star) \right) \\
	\phi^{(n+1)} &= \phi^{(n)} + \frac{\Delta t}{2} \left( f(\phi^{(n)}) + f(\phi^{\star\star}) \right)
\end{align}
\end{subequations}
where \(f(\phi^{(n)}) = - \nabla \cdot (\mathbf{u} \phi^{(n)})\) at time level \(n\). The advection terms of the momentum and $\theta$ equations, (\ref{eq:momentum}) and (\ref{eq:theta}) and of the linear advection equation, (\ref{eq:advect}) are discretised in flux form using the upwind-biased, multi-dimensional cubic scheme from \citet{weller-shahrokhi2014} which is non-monotonic and not flux corrected.

### 2015-09-17 src/tf-cutcell-comparison/tf-cutcell-comparison.tex
I am using a draft document class so that I can compile it without errors
I have changed $\bm{u}$ to $\mathbf{u}$ (I think this style makes the vectors stand out more. But you can change back if you disagree)

I have changed $\mathbf{uu}$ to $\mathbf{u}\otimes\mathbf{u}$ for clarity

I have removed your ending to the Model section and replaced it with one which cites the appendix

I have removed your appendix and replaced it with mine, in a different file to avoid conflicts. I will now continue to work on the appendix and hopefully won't make many more changes to the main file

@@
-\documentclass{ametsoc}
+\documentclass[draft]{ametsoc}
 
@@
 \usepackage{siunitx}
+\usepackage[title]{appendix}
 
@@
 \begin{align}
-	\text{Momentum} &\ &\  	\frac{\partial \rho \bm{u}}{\partial t} + \nabla \cdot \rho \bm{uu} &= \rho \bm{g} - c_p \rho \theta \nabla \Pi - \mu \rho \bm{u} \label{eq:momentum} \\
-	\text{Continuity} &\ &\	\frac{\partial \rho}{\partial t} + \nabla \cdot \rho \bm{u} &= 0 \\
-	\text{Thermodynamic equation} &\ &\ \frac{\partial \rho \theta}{\partial t} + \nabla \cdot \rho \bm{u} \theta &= 0 \\
+	\text{Momentum} &\ &\  	\frac{\partial \rho \mathbf{u}}{\partial t} + \nabla \cdot \rho \mathbf{u}\otimes\mathbf{u} &= \rho \mathbf{g} - c_p \rho \theta \nabla \Pi - \mu \rho \mathbf{u} \label{eq:momentum} \\
+	\text{Continuity} &\ &\	\frac{\partial \rho}{\partial t} + \nabla \cdot \rho \mathbf{u} &= 0 \\
+	\text{Thermodynamic equation} &\ &\ \frac{\partial \rho \theta}{\partial t} + \nabla \cdot \rho \mathbf{u} \theta &= 0 \label{eq:theta} \\
 	\text{Ideal gas law} &\ &\ \Pi^{(1 - \kappa)/\kappa} &= \frac{R \rho \theta}{p_0}
@@
 \end{subequations}
-where \(\rho\) is the density, \(\bm{u}\) is the velocity field, \(\bm{g}\) is the gravitational acceleration, \(c_p\) is the heat capacity at constant pressure, \(\theta = T \left(p_0/p\right)^\kappa\) is the potential temperature, \(T\) is the temperature, \(p\) is the pressure, \(p_0\) is a reference pressure, \(\Pi = \left(p / p_0 \right)^\kappa\) is the Exner function of pressure, and \(\kappa = R/c_p\) is the gas constant to heat capacity ratio.
+where \(\rho\) is the density, \(\mathbf{u}\) is the velocity field, \(\mathbf{g}\) is the gravitational acceleration, \(c_p\) is the heat capacity at constant pressure, \(\theta = T \left(p_0/p\right)^\kappa\) is the potential temperature, \(T\) is the temperature, \(p\) is the pressure, \(p_0\) is a reference pressure, \(\Pi = \left(p / p_0 \right)^\kappa\) is the Exner function of pressure, and \(\kappa = R/c_p\) is the gas constant to heat capacity ratio.
 \(\mu\) is a dimensionless damping function used for the sponge layer in the gravity waves test in section~\ref{sec:results}\ref{sec:gw}.
@@
 The model uses the C-grid staggering in the horizontal and the Lorenz staggering in the vertical such that $\theta$, $\rho$ and $\Pi$ are stored at cell centroids and the covariant component of velocity at cell faces.
-A minor change has been made to the non-orthogonal formulation as described in \citet{weller-shahrokhi2014} to improve stability on non-orthogonal grid, and this is detailed in appendix A.
+\input{model}
+%A minor change has been made to the non-orthogonal formulation as described in \citet{weller-shahrokhi2014} to improve stability on non-orthogonal grid, and this is detailed in appendix A.
 
-For tracer advection tests, the advection equation is solved in flux form, given by
-\begin{align}
-\partial \phi / \partial t + \nabla \cdot \left( \bm{u} \phi \right) = 0
-\end{align}
-where the tracer density, $\phi$, is interpolated onto cell faces using the upwind-biased multidimensional cubic scheme from \citet{weller-shahrokhi2014} which is non-monotonic and not flux corrected.
-The time derivative is solved using a three-stage, second order Runge-Kutta scheme:
-\begin{subequations}
-\begin{align}
-	\phi^\star &= \phi^{(n)} + \Delta t f(\phi^{(n)}) \\
-	\phi^{\star\star} &= \phi^{(n)} + \frac{\Delta t}{2} \left( f(\phi^{(n)}) + f(\phi^\star) \right) \\
-	\phi^{(n+1)} &= \phi^{(n)} + \frac{\Delta t}{2} \left( f(\phi^{(n)}) + f(\phi^{\star\star}) \right)
-\end{align}
-\end{subequations}
-where \(f(\phi^{(n)}) = - \nabla \cdot (\bm{u} \phi^{(n)})\) at time level \(n\).
+%For tracer advection tests, the advection equation is solved in flux form, given by
+%\begin{align}
+%\partial \phi / \partial t + \nabla \cdot \left( \mathbf{u} \phi \right) = 0
+%\end{align}
+%where the tracer density, $\phi$, is interpolated onto cell faces using the upwind-biased multidimensional cubic scheme from \citet{weller-shahrokhi2014} which is non-monotonic and not flux corrected.
+%The time derivative is solved using a three-stage, second order Runge-Kutta scheme:
+%\begin{subequations}
+%\begin{align}
+%	\phi^\star &= \phi^{(n)} + \Delta t f(\phi^{(n)}) \\
+%	\phi^{\star\star} &= \phi^{(n)} + \frac{\Delta t}{2} \left( f(\phi^{(n)}) + f(\phi^\star) \right) \\
+%	\phi^{(n+1)} &= \phi^{(n)} + \frac{\Delta t}{2} \left( f(\phi^{(n)}) + f(\phi^{\star\star}) \right)
+%\end{align}
+%\end{subequations}
+%where \(f(\phi^{(n)}) = - \nabla \cdot (\mathbf{u} \phi^{(n)})\) at time level \(n\).
 
@@
 
-% APPENDICES
-\appendix[A]
-\appendixtitle{Non-orthogonal formulation}
+\begin{appendices}
+\input{WellerAppendix}
+\end{appendices}
 
-\TODO{Hilary will redraft this appendix}
-A minor change has been made to the curl-free, non-orthogonal formulation as described in \citet{weller-shahrokhi2014} to improve stability on non-orthogonal grids.
+%% APPENDICES
+%\appendix[A]
+%\appendixtitle{Non-orthogonal formulation}
+
+%\TODO{Hilary will redraft this appendix}
+%A minor change has been made to the curl-free, non-orthogonal formulation as described in \citet{weller-shahrokhi2014} to improve stability on non-orthogonal grids.
 \citet{weller-shahrokhi2014} use a Hodge dual operator, \(H\), to map from co-variant to contra-variant velocity components:
-\begin{align}
-	U = HV
-\end{align}
-where \(U = \rho\bm{u}\cdot\bm{S}_f\) and \(V=\rho\bm{u}\cdot\bm{d}_f\).
-\(\bm{S}_f\) is the surface normal vector for a face, \(f\), which has a magnitude of the surface area, and \(\bm{d}_f\) is the vector between neighbouring cell centres.
+%\begin{align}
+%	U = HV
+%\end{align}
+%where \(U = \rho\mathbf{u}\cdot\mathbf{S}_f\) and \(V=\rho\mathbf{u}\cdot\mathbf{d}_f\).
+\(\mathbf{S}_f\) is the surface normal vector for a face, \(f\), which has a magnitude of the surface area, and \(\mathbf{d}_f\) is the vector between neighbouring cell centres.
 
-In order to create a semi-implicit formulation, \cite{weller-shahrokhi2014} split \(H\) into diagonal and off-diagonal components and the off-diagonal components are treated explicitly in the Helmholtz equation for \(\Pi\).
+%In order to create a semi-implicit formulation, \cite{weller-shahrokhi2014} split \(H\) into diagonal and off-diagonal components and the off-diagonal components are treated explicitly in the Helmholtz equation for \(\Pi\).
 However, this was found to be unstable when using a large timestep on a grid with large non-orthogonality.
 Improved stability is achieved by splitting \(H\) into the orthogonal part and the non-orthogonal correction:
-\begin{align}
-	H = H_c + H_\mathrm{corr}
-\end{align}
-where the diagonal matrix \(H_c=|\bm{S}_f|/|\bm{d}_f|\) and the non-orthogonal correction is simply \(H_\mathrm{corr} = H - H_c\).
+%\begin{align}
+%	H = H_c + H_\mathrm{corr}
+%\end{align}
+%where the diagonal matrix \(H_c=|\mathbf{S}_f|/|\mathbf{d}_f|\) and the non-orthogonal correction is simply \(H_\mathrm{corr} = H - H_c\).
 The orthogonal part, \(H_c\), is treated implicitly in the Helmholtz equation and the non-orthogonal part, \(H_\mathrm{corr}\), is treated explicitly.

### 2015-09-17 src/tf-cutcell-comparison/WellerAppendix.tex
First draft of appendix on the Hodge operator and the semi-implicit method

@@
 
-In order to ensure curl-free pressure gradients, following \citet{WS14},
-the momentum at the cell face in the direction between cell centres
-is used as the prognostic variable for velocity:
+In order to ensure curl-free pressure gradients, following \citet{weller-shahrokhi2014},
+the co-variant momentum component, that is the momentum at the cell
+face in the direction between cell centres is used as the prognostic
+variable for velocity:
 \begin{equation}
@@
 where $\mathbf{d}_{f}$ is the vector between cell centres and subscript
-$f$ means ``at face $f$''. Fluxes across faces are diagnostic
-variables:
+$f$ means ``at face $f$''. Fluxes across faces, or the contravariant
+component, are diagnostic variables:
 \begin{equation}
@@
 \end{equation}
-where $c\in f$ means the two cells either side of cell $c$ and where
-$(\rho\mathbf{u})_{C}$ is the consistent cell centre reconstruction
-of $\rho\mathbf{u}$ from surrounding values of $V_{f}$:
+where $c\in f$ means the two cells either side of cell $c$. $(\rho\mathbf{u})_{C}$
+is the consistent cell centre reconstruction of $\rho\mathbf{u}$
+from surrounding values of $V_{f}$:
 \[
@@
 a local operation which can be calculated once for each cell of the
-grid before time-stepping begins.
+grid before time-stepping begins. The $H$ implied by this reconstruction
+of $U$ is likely to be positive definite for meshes with sufficiently
+low non-orthogonality. However this has not been proved.
+
+The semi-implicit technique involves combining the momentum (\ref{eq:momentum}),
+continuity (\ref{eq:cont}) and $\theta$ (\ref{eq:theta}) equations
+and the equation of state (\ref{eq:state}) to form a Helmholtz equation
+to be solved implicitly (see \citet{weller-shahrokhi2014}). The semi-implicit
+solution technique with a Hodge operator can be defined by considering
+only a discretised from of the continuity equation:
+\begin{equation}
+\frac{\phi^{(n+1)}-\rho^{(n)}}{\Delta t}+\frac{1}{2}\left\{ \nabla\cdot(HV)^{(n)}+\nabla\cdot(HV)^{(n+1)}\right\} =0.\label{eq:contD}
+\end{equation}
+The divergence is discretised using Gauss' divergence theorem so that:
+\begin{equation}
+\nabla\cdot\left(HV\right)=\frac{1}{V_{c}}\sum_{f\in c}n_{f}(HV)_{f}\label{eq:GaussDiv}
+\end{equation}
+where $V_{c}$ is the volume of cell $c$ and $n_{f}=1$ if $\mathbf{d}_{f}$
+points outwards from the cell and $n_{f}=$-1 otherwise. Equation
+(\ref{eq:GaussDiv}) is now a sum over a sum since $(HV)_{f}$ is
+one element of a matrix-vector multiply. In order to simplify the
+construction of the matrix for the Helmholtz problem, only the diagonal
+terms of $HV$ are treated implicitly. We therefore separate $H$
+into a diagonal and off-diagonal matrix:
+\begin{equation}
+H=H_{d}+H_{off}.
+\end{equation}
+Eqn (\ref{eq:contD}) can now be approximated by:
+\begin{equation}
+\frac{\phi^{(n+1)}-\rho^{(n)}}{\Delta t}+\frac{1}{2}\left\{ \nabla\cdot(HV)^{(n)}+\nabla\cdot(H_{d}V)^{(n+1)}+\nabla\cdot(H_{off}V)^{\ell}\right\} =0\label{eq:contDd}
+\end{equation}
+where superscript $\ell$ means lagged values taken from a previous
+iteration or from a previous stage of a Runge-Kutta scheme. This was
+the approach taken by \citet{weller-shahrokhi2014}. However, numerical
+solution of eqn (\ref{eq:contDd}) turns out to be unstable when using
+a large time-step on a highly non-orthogonal mesh associated with
+terrain following coordinates over steep orography. Improved stability
+and energy conservation can be achieved by splitting $H$ into a diagonal
+component which would be correct on an orthogonal grid and a non-orthogonal
+correction:
+\begin{equation}
+H=H_{c}+H_{corr}
+\end{equation}
+where the diagonal matrix $H_{c}=|\mathbf{S}_{f}|/|\mathbf{d}_{f}|$
+and the non-orthogonal correction is $H_{corr}=H-H_{c}$. The orthogonal
+part, $H_{c}$, can be treated implicitly in the Helmholtz equation:
+\begin{equation}
+\frac{\phi^{(n+1)}-\rho^{(n)}}{\Delta t}+\frac{1}{2}\left\{ \nabla\cdot(HV)^{(n)}+\nabla\cdot(H_{c}V)^{(n+1)}+\nabla\cdot(H_{corr}V)^{\ell}\right\} =0.\label{eq:contDc}
+\end{equation}
+This form is used for the solutions of the Euler equations in this
+paper and is stable, with good energy conservation for all of the
+tests presented.

### 2015-09-18 src/tf-cutcell-comparison/WellerAppendix.tex
Minor grammar and formatting tweaks to appendix.

@@
 the co-variant momentum component, that is the momentum at the cell
-face in the direction between cell centres is used as the prognostic
+face in the direction between cell centres, is used as the prognostic
 variable for velocity:
@@
 where $\mathbf{d}_{f}$ is the vector between cell centres and subscript
-$f$ means ``at face $f$''. Fluxes across faces, or the contravariant
-component, are diagnostic variables:
+$f$ means ``at face $f$''. The contravariant momentum component,
+that is the flux across faces, is a diagnostic variable:
 \begin{equation}
@@
 \end{equation}
-where $\mathbf{S}_{f}$ is the outward pointing normal vector to face
-$f$ with magnitude equal to the area of the face. $U$ is the vector
-of all values of $U_{f}$ and $V$ is the vector of all values of
-$V_{f}$ then we can define the Hodge operator: a matrix to transform
-$V$ to $U$:
+where $\mathbf{S}_{f}$ is the outward-pointing normal vector to face
+$f$ with magnitude equal to the area of the face. If $U$ is the
+vector of all values of $U_{f}$ and $V$ is the vector of all values
+of $V_{f}$ then we can define the Hodge operator as a matrix that
+transforms $V$ to $U$:
 \begin{equation}
@@
 \end{equation}
-where subscript $F$ implies mid-point interpolation from two surrounding
+where subscript $F$ denotes mid-point interpolation from two surrounding
 cell values onto face $f$:
@@
 \end{equation}
-where $c\in f$ means the two cells either side of cell $c$. $(\rho\mathbf{u})_{C}$
+where $c\in f$ denotes the two cells sharing face $f$. $(\rho\mathbf{u})_{C}$
 is the consistent cell centre reconstruction of $\rho\mathbf{u}$
@@
 is a $3\times3$ tensor and so the inversion of the tensor sum is
-a local operation which can be calculated once for each cell of the
+a local operation which can be calculated once for every cell in the
 grid before time-stepping begins. The $H$ implied by this reconstruction
 of $U$ is likely to be positive definite for meshes with sufficiently
-low non-orthogonality. However this has not been proved.
+low non-orthogonality, although this has not been proved.
 
-The semi-implicit technique involves combining the momentum (\ref{eq:momentum}),
-continuity (\ref{eq:cont}) and $\theta$ (\ref{eq:theta}) equations
-and the equation of state (\ref{eq:state}) to form a Helmholtz equation
-to be solved implicitly (see \citet{weller-shahrokhi2014}). The semi-implicit
-solution technique with a Hodge operator can be defined by considering
-only a discretised from of the continuity equation:
+The semi-implicit technique involves combining the momentum \ref{eq:momentum},
+continuity \ref{eq:cont} and $\theta$ \ref{eq:theta} equations
+and the equation of state \ref{eq:state} to form a Helmholtz equation
+to be solved implicitly, as described by \citet{weller-shahrokhi2014}.
+The semi-implicit solution technique with a Hodge operator can be
+defined by considering only a discretised form of the continuity equation:
 \begin{equation}
@@
 \begin{equation}
-\nabla\cdot\left(HV\right)=\frac{1}{V_{c}}\sum_{f\in c}n_{f}(HV)_{f}\label{eq:GaussDiv}
+\nabla\cdot\left(HV\right)=\frac{1}{\mathcal{V}_{c}}\sum_{f\in c}n_{f}(HV)_{f}\label{eq:GaussDiv}
 \end{equation}
-where $V_{c}$ is the volume of cell $c$ and $n_{f}=1$ if $\mathbf{d}_{f}$
-points outwards from the cell and $n_{f}=$-1 otherwise. Equation
-(\ref{eq:GaussDiv}) is now a sum over a sum since $(HV)_{f}$ is
-one element of a matrix-vector multiply. In order to simplify the
-construction of the matrix for the Helmholtz problem, only the diagonal
-terms of $HV$ are treated implicitly. We therefore separate $H$
-into a diagonal and off-diagonal matrix:
+where $\mathcal{V}_{c}$ is the volume of cell $c$, $f\in c$ denotes
+the faces of cell $c$, and $n_{f}=1$ if $\mathbf{d}_{f}$ points
+outwards from the cell and $n_{f}=$-1 otherwise. Equation \ref{eq:GaussDiv}
+is now a sum over a sum since $(HV)_{f}$ is one element of a matrix-vector
+multiply. In order to simplify the construction of the matrix for
+the Helmholtz problem, only the diagonal terms of $HV$ are treated
+implicitly. Therefore, $H$ is separated into a diagonal and off-diagonal
+matrix:
 \begin{equation}
@@
 \end{equation}
-Eqn (\ref{eq:contD}) can now be approximated by:
+Equation \ref{eq:contD} can now be approximated by:
 \begin{equation}
@@
 \end{equation}
-where superscript $\ell$ means lagged values taken from a previous
+where superscript $\ell$ denotes lagged values taken from a previous
 iteration or from a previous stage of a Runge-Kutta scheme. This was
-the approach taken by \citet{weller-shahrokhi2014}. However, numerical
-solution of eqn (\ref{eq:contDd}) turns out to be unstable when using
-a large time-step on a highly non-orthogonal mesh associated with
-terrain following coordinates over steep orography. Improved stability
+the approach taken by \citet{weller-shahrokhi2014}. However, the
+numerical solution of equation \ref{eq:contDd} turns out to be unstable
+when using a large time-step on highly non-orthogonal meshes associated
+with terrain following layers over steep orography. Improved stability
 and energy conservation can be achieved by splitting $H$ into a diagonal

### 2015-09-18 src/tf-cutcell-comparison/tf-cutcell-comparison.tex
First draft of appendix on the Hodge operator and the semi-implicit method
describe test setup of thermal advection

@@
 % Every term in the discretised equations of motion is a potential source of numerical error.
-% Errors in calculating the horizontal pressure gradient generate spurious circulations and so much effort has been put into improving this calculation.
-This is often achieved by making the grid more orthogonal, flattening terrain following coordinate surfaces [citations] or using a cut cell approach [citations].
+% Errors in calculating the horizontal pressure gradient generate spurious circulations and much effort has been put into improving this calculation.
+This is often achieved by making the grid more orthogonal: flattening terrain following coordinate surfaces [citations] or using a cut cell approach [citations].
 % TF citations: mahrer1984, schaer2002, klemp2011
@@
 	\text{Momentum} &\ &\  	\frac{\partial \rho \mathbf{u}}{\partial t} + \nabla \cdot \rho \mathbf{u}\otimes\mathbf{u} &= \rho \mathbf{g} - c_p \rho \theta \nabla \Pi - \mu \rho \mathbf{u} \label{eq:momentum} \\
-	\text{Continuity} &\ &\	\frac{\partial \rho}{\partial t} + \nabla \cdot \rho \mathbf{u} &= 0 \\
+	\text{Continuity} &\ &\	\frac{\partial \rho}{\partial t} + \nabla \cdot \rho \mathbf{u} &= 0 \label{eq:cont} \\
 	\text{Thermodynamic equation} &\ &\ \frac{\partial \rho \theta}{\partial t} + \nabla \cdot \rho \mathbf{u} \theta &= 0 \label{eq:theta} \\
-	\text{Ideal gas law} &\ &\ \Pi^{(1 - \kappa)/\kappa} &= \frac{R \rho \theta}{p_0}
+	\text{Ideal gas law} &\ &\ \Pi^{(1 - \kappa)/\kappa} &= \frac{R \rho \theta}{p_0} \label{eq:state}
 \end{align}
@@
 \begin{equation}
-	\Psi(x,z) = u_0 H \frac{z - h}{H - h}
+	\Psi(x,z) = u_0 H \frac{z - h}{H - h} \label{eqn:streamfunc-btf}
 \end{equation}
@@
 \begin{align}
-	u &= \frac{\partial \Psi}{\partial z} = u_0 \frac{H}{H - h}, \quad w = -\frac{\partial \Psi}{\partial x} = u_0 H \frac{\mathrm{d} h}{\mathrm{d} x} \frac{H - z}{\left( H - h \right)^2}, \nonumber \\
+	u &= \frac{\partial \Psi}{\partial z} = u_0 \frac{H}{H - h}, \quad w = -\frac{\partial \Psi}{\partial x} = u_0 H \frac{\mathrm{d} h}{\mathrm{d} x} \frac{H - z}{\left( H - h \right)^2} \label{eqn:uw-btf} \\
 	\frac{\mathrm{d} h}{\mathrm{d} x} &= - h_0 \left[ 
@@
 
+\subsection{Terrain following advection of thermal profile}
+The spatial domain, mountain profile, grid resolutions and timesteps are the same as those in the gravity waves test in section~\ref{sec:results}\ref{sec:gw}.
+The terrain following velocity field is defined by the streamfunction:
+\begin{align}
+	\Psi(x,z) = u_0 \left\{ \begin{array}{l l}
+			H_\mathrm{TF} \frac{z - h}{H_\mathrm{TF} - h} \label{eqn:streamfunc-btf} & \enskip \text{if $z \leq H_\mathrm{TF}$} \\
+			z & \enskip \text{if $z > H_\mathrm{TF}$}
+	\end{array} \right.
+\end{align}
+where \(H_\mathrm{TF} = \SI{20}{\kilo\meter}\) is the level at which the terrain following layers become flat; the domain height is \(\SI{30}{\kilo\meter}\).
+For $z \leq H$, the $u$ and $w$ components of velocity are given by equation~\ref{eqn:uw-btf}, but $h(x)$ has the same form as equation~\ref{eqn:resting:mountain}, hence the derivative is:
+\begin{align}
+	\frac{\mathrm{d} h}{\mathrm{d} x} = -h_0 \exp \left(- \left( \frac{x}{a} \right)^2 \right) \left[ \frac{\pi}{\lambda} \sin \left( \frac{2 \pi x}{\lambda} \right) - \frac{2x}{a^2} \cos^2\left( \frac{\pi x}{\lambda} \right) \right]
+\end{align}
+For $z > H$, $u = u_0$ and $w = 0$.
+
+The tracer field, \(\phi\), is the same as the initial potential temperature field in the gravity waves test:
+\begin{align}
+	\phi(z) = \phi_0 \exp \left( \frac{N^2}{g} z \right) \label{eqn:thermal-tracer}
+\end{align}
+where \(\phi_0 = 288\) and \(N = \SI{0.01}{\per\second}\).
+The inlet boundary has fixed values prescribed by equation~\ref{eqn:thermal-tracer} and a zero gradient boundary condition is applied at the outlet.
+
 \section{Conclusions}
@@
 
+%% APPENDICES
 \begin{appendices}
@@
 \end{appendices}
-
-%% APPENDICES
-%\appendix[A]
-%\appendixtitle{Non-orthogonal formulation}
-
-%\TODO{Hilary will redraft this appendix}
-%A minor change has been made to the curl-free, non-orthogonal formulation as described in \citet{weller-shahrokhi2014} to improve stability on non-orthogonal grids.
-\citet{weller-shahrokhi2014} use a Hodge dual operator, \(H\), to map from co-variant to contra-variant velocity components:
-%\begin{align}
-%	U = HV
-%\end{align}
-%where \(U = \rho\mathbf{u}\cdot\mathbf{S}_f\) and \(V=\rho\mathbf{u}\cdot\mathbf{d}_f\).
-\(\mathbf{S}_f\) is the surface normal vector for a face, \(f\), which has a magnitude of the surface area, and \(\mathbf{d}_f\) is the vector between neighbouring cell centres.
-
-%In order to create a semi-implicit formulation, \cite{weller-shahrokhi2014} split \(H\) into diagonal and off-diagonal components and the off-diagonal components are treated explicitly in the Helmholtz equation for \(\Pi\).
-However, this was found to be unstable when using a large timestep on a grid with large non-orthogonality.
-Improved stability is achieved by splitting \(H\) into the orthogonal part and the non-orthogonal correction:
-%\begin{align}
-%	H = H_c + H_\mathrm{corr}
-%\end{align}
-%where the diagonal matrix \(H_c=|\mathbf{S}_f|/|\mathbf{d}_f|\) and the non-orthogonal correction is simply \(H_\mathrm{corr} = H - H_c\).
-The orthogonal part, \(H_c\), is treated implicitly in the Helmholtz equation and the non-orthogonal part, \(H_\mathrm{corr}\), is treated explicitly.
 

### 2015-09-18 src/tf-cutcell-comparison/tf-cutcell-comparison.tex
make the motivation of terrain following advection test clearer

@@
 \subsection{Terrain following advection}
-\TODO{make it clearer that we're trying to determine if non-orthogonality or misalignment of the flow is the primary source of error}
-
-In the horizontal advection test, results were more accurate where the flow was aligned with the grid layers, and distortions in the BTF grid led to increased errors.
-To determine the primary source of numerical error, we formulate a new tracer advection test in which the velocity field is everywhere tangential to the basic terrain following coordinate surfaces.
-Misalignment on the SLEVE and cut cell grids is designed to challenge the advection scheme.
+In the horizontal advection test, results were least accurate on the BTF grid, where the grid was most non-orthogonal and flow was misaligned with the grid layers.
+Here, we formulate a new tracer advection test in which the velocity field is everywhere tangential to the basic terrain following coordinate surfaces.
+This test determines whether the primary source of numerical error is due to non-orthogonality or misalignment of the flow with grid layers.
 

### 2015-09-18 src/tf-cutcell-comparison/model.tex
more description of models in Model section, allows us to remove some detail from advection results

@@
 To avoid time-splitting errors between the advection and the fast waves, the advection is time-stepped using a three-stage, second-order Runge-Kutta scheme.
-When considering only the linear advection of a tracer, $\phi$:
-\begin{align}
-\partial \phi / \partial t + \nabla \cdot \left( \mathbf{u} \phi \right) = 0
-\label{eq:advect}
-\end{align}
-the Runge-Kutta scheme is defined as:
-\begin{subequations}
-\begin{align}
-	\phi^\star &= \phi^{(n)} + \Delta t f(\phi^{(n)}) \\
-	\phi^{\star\star} &= \phi^{(n)} + \frac{\Delta t}{2} \left( f(\phi^{(n)}) + f(\phi^\star) \right) \\
-	\phi^{(n+1)} &= \phi^{(n)} + \frac{\Delta t}{2} \left( f(\phi^{(n)}) + f(\phi^{\star\star}) \right)
-\end{align}
-\end{subequations}
-where \(f(\phi^{(n)}) = - \nabla \cdot (\mathbf{u} \phi^{(n)})\) at time level \(n\).
-The advection terms of the momentum and $\theta$ equations, (\ref{eq:momentum}) and (\ref{eq:theta}) and of the linear advection equation, (\ref{eq:advect}) are discretised in flux form using the upwind-biased, multi-dimensional cubic scheme from \citet{weller-shahrokhi2014} which is non-monotonic and not flux corrected.
+The advection terms of the momentum and $\theta$ equations, (\ref{eq:momentum}) and (\ref{eq:theta}) are discretised in flux form using the upwind-biased, multi-dimensional cubic scheme from \citet{weller-shahrokhi2014}.

### 2015-09-18 src/tf-cutcell-comparison/tf-cutcell-comparison.tex
more description of models in Model section, allows us to remove some detail from advection results

@@
 \label{sec:model}
-\TODO{Add info about the different numerical options that are assessed in combination with the orography treatment}
+Three models are used for the test cases in this study, and all are operated in a two-dimensional $x$--$z$ slice configuration.  The first model discretises the linear advection equation in flux form:
+\begin{align}
+\partial \phi / \partial t + \nabla \cdot \left( \mathbf{u} \phi \right) = 0
+\label{eq:advect}
+\end{align}
+where the tracer density, $\phi$, is interpolated onto cell faces using one of two schemes: first, the centred linear scheme which takes the average of the two neighbouring cell values; second, the upwind-biased multidimensional cubic scheme from \citet{weller-shahrokhi2014} which is non-monotonic and not flux corrected.
+The time derivative is solved using a three-stage, second order Runge-Kutta scheme defined as:
+\begin{subequations}
+\begin{align}
+	\phi^\star &= \phi^{(n)} + \Delta t f(\phi^{(n)}) \\
+	\phi^{\star\star} &= \phi^{(n)} + \frac{\Delta t}{2} \left( f(\phi^{(n)}) + f(\phi^\star) \right) \\
+	\phi^{(n+1)} &= \phi^{(n)} + \frac{\Delta t}{2} \left( f(\phi^{(n)}) + f(\phi^{\star\star}) \right)
+\end{align}
+\end{subequations}
+where \(f(\phi^{(n)}) = - \nabla \cdot (\mathbf{u} \phi^{(n)})\) at time level \(n\). 
+To ensure that the discrete velocity field is non-divergent, velocities are prescribed at cell faces by differencing the streamfunction, \(\Psi(x, z)\), along the edges from \(\Psi\) stored at cell vertices.
 
-We use the finite volume model from \citet{weller-shahrokhi2014} which details a discretisation of the fully-compressible Euler equations, given by
+The second model is a modified version of the linear advection finite difference model first used by \citet{schaer2002}.
+It uses terrain following coordinates and it is configured with leapfrog timestepping and a fourth-order centred difference scheme: \TODO{this wasn't properly documented in schaer2002 but we can document it using the Fortran code}
+\begin{align}
+	TODO
+\end{align}
+
+Once again, velocity fields are prescribed using a streamfunction defined at cell vertices (referred to as double staggered grid points by \citet{schaer2002}).
+The original version of the code interpolated the geometric height, $z$, at doubly staggered points from values at adjacent half levels in order to calculate the streamfunction.
+The modified version used here directly calculates the height at vertices to enable comparisons with the finite volume model solutions.
+
+The third model is taken from \citet{weller-shahrokhi2014} which details a discretisation of the fully-compressible Euler equations, given by
 \begin{subequations}
@@
 
-The model is operated in a two-dimensional $x$--$z$ slice configuration without Coriolis forces.
-The model uses the C-grid staggering in the horizontal and the Lorenz staggering in the vertical such that $\theta$, $\rho$ and $\Pi$ are stored at cell centroids and the covariant component of velocity at cell faces.
+The fully-compressible model uses the C-grid staggering in the horizontal and the Lorenz staggering in the vertical such that $\theta$, $\rho$ and $\Pi$ are stored at cell centroids and the covariant component of velocity at cell faces.
+The model is configured without Coriolis forces.
+
 \input{model}
-%A minor change has been made to the non-orthogonal formulation as described in \citet{weller-shahrokhi2014} to improve stability on non-orthogonal grid, and this is detailed in appendix A.
-
-%For tracer advection tests, the advection equation is solved in flux form, given by
-%\begin{align}
-%\partial \phi / \partial t + \nabla \cdot \left( \mathbf{u} \phi \right) = 0
-%\end{align}
-%where the tracer density, $\phi$, is interpolated onto cell faces using the upwind-biased multidimensional cubic scheme from \citet{weller-shahrokhi2014} which is non-monotonic and not flux corrected.
-%The time derivative is solved using a three-stage, second order Runge-Kutta scheme:
-%\begin{subequations}
-%\begin{align}
-%	\phi^\star &= \phi^{(n)} + \Delta t f(\phi^{(n)}) \\
-%	\phi^{\star\star} &= \phi^{(n)} + \frac{\Delta t}{2} \left( f(\phi^{(n)}) + f(\phi^\star) \right) \\
-%	\phi^{(n+1)} &= \phi^{(n)} + \frac{\Delta t}{2} \left( f(\phi^{(n)}) + f(\phi^{\star\star}) \right)
-%\end{align}
-%\end{subequations}
-%where \(f(\phi^{(n)}) = - \nabla \cdot (\mathbf{u} \phi^{(n)})\) at time level \(n\).
 
@@
 
-To ensure that the discrete velocity field is non-divergent, velocities are prescribed at cell faces by differencing the streamfunction, \(\Psi\), along the edges from \(\Psi\) stored at cell vertices.
-Since \(u = \partial \Psi / \partial z\), the streamfunction is found by vertical integration of the velocity profile:
+\TODO{Schaer defines u=-dPsi/dz, w=dPsi/dx --- check all this streamfunction stuff AGAIN}
+
+The discrete velocity field is defined using a streamfunction, \(\Psi\).  Since \(u = \partial \Psi / \partial z\), the streamfunction is found by vertical integration of the velocity profile:
 \begin{align}
@@
 The test was executed on the BTF, SLEVE and cut cell grids, and on a regular grid with flat terrain using a centred linear scheme and the upwind-biased cubic scheme.
-Results were also obtained on BTF and SLEVE grids with the fourth order scheme from \citet{schaer2002} using a modified version of their Fortran code.
-The original version of the code interpolated the geometric height at cell vertices from values at adjacent half levels in order to calculate the streamfunction, \(\Psi(z)\).
-The modified version used here directly calculates the height at vertices to enable comparisons with the finite volume model solutions.
-\TODO{move this detail into section 3 and add more info and references}
+Results were also obtained on BTF and SLEVE grids with the fourth order scheme from \citet{schaer2002} using a modified version of their code.
 
@@
 	\Psi(x,z) = u_0 \left\{ \begin{array}{l l}
-			H_\mathrm{TF} \frac{z - h}{H_\mathrm{TF} - h} \label{eqn:streamfunc-btf} & \enskip \text{if $z \leq H_\mathrm{TF}$} \\
+			H_\mathrm{TF} \frac{z - h}{H_\mathrm{TF} - h} & \enskip \text{if $z \leq H_\mathrm{TF}$} \\
 			z & \enskip \text{if $z > H_\mathrm{TF}$}

### 2015-09-21 src/tf-cutcell-comparison/tf-cutcell-comparison.tex
correct streamfunction signs again, describe 4th order difference, TV analysis for nonorthogonality

@@
 
-\section{Model}
+\section{Models}
 \label{sec:model}
-Three models are used for the test cases in this study, and all are operated in a two-dimensional $x$--$z$ slice configuration.  The first model discretises the linear advection equation in flux form:
+Three models are used for the test cases in this study: two linear advection models and a model of the fully-compressible Euler equations.
+All are operated in a two-dimensional $x$--$z$ slice configuration.
+
+\subsection{Finite volume linear advection model}
+The first model discretises the linear advection equation in flux form:
 \begin{align}
@@
 \end{align}
-where the tracer density, $\phi$, is interpolated onto cell faces using one of two schemes: first, the centred linear scheme which takes the average of the two neighbouring cell values; second, the upwind-biased multidimensional cubic scheme from \citet{weller-shahrokhi2014} which is non-monotonic and not flux corrected.
+where $\mathbf{u} = (u, w)$ is a prescribed velocity field and the tracer density, $\phi$, is interpolated onto cell faces using one of two schemes: first, the centred linear scheme which takes the average of the two neighbouring cell values; second, the upwind-biased multidimensional cubic scheme from \citet{weller-shahrokhi2014} which is non-monotonic and not flux corrected.
 The time derivative is solved using a three-stage, second order Runge-Kutta scheme defined as:
@@
 
-The second model is a modified version of the linear advection finite difference model first used by \citet{schaer2002}.
-It uses terrain following coordinates and it is configured with leapfrog timestepping and a fourth-order centred difference scheme: \TODO{this wasn't properly documented in schaer2002 but we can document it using the Fortran code}
+\subsection{Finite difference linear advection model}
+The second model is a modified version of the linear advection model first used by \citet{schaer2002}.
+It uses terrain following coordinates and it is configured with leapfrog timestepping and a fourth-order centred difference scheme:
 \begin{align}
-	TODO
+	\frac{\partial \phi_i}{\partial x} \approx \frac{1}{\Delta x} \left( \frac{1}{12} \phi_{i-2} - \frac{2}{3} \phi_{i-1} + \frac{2}{3} \phi_{i+1} - \frac{1}{12} \phi_{i+2} \right)
 \end{align}
@@
 
+\subsection{Fully-compressible model}
 The third model is taken from \citet{weller-shahrokhi2014} which details a discretisation of the fully-compressible Euler equations, given by
@@
 Finally, as specified by \citet{schaer2002}, a test of orographically-induced gravity waves is performed.
-No explicit diffusion is used in any of the tests.
+\TODO{advection of thermal profile} No explicit diffusion is used in any of the tests.
 
@@
 
-\TODO{Schaer defines u=-dPsi/dz, w=dPsi/dx --- check all this streamfunction stuff AGAIN}
-
-The discrete velocity field is defined using a streamfunction, \(\Psi\).  Since \(u = \partial \Psi / \partial z\), the streamfunction is found by vertical integration of the velocity profile:
+The discrete velocity field is defined using a streamfunction, \(\Psi\).  Given that \(u = -\partial \Psi / \partial z\), the streamfunction is found by vertical integration of the velocity profile:
 \begin{align}
-	\Psi(z) &= \frac{u_0}{2} \left\{ \begin{array}{l l}
+	\Psi(z) &= -\frac{u_0}{2} \left\{ \begin{array}{l l}
 		\left( 2z - z_1 - z_2 \right) & \enskip \text{if $z > z_2$} \\
@@
 The centred linear scheme is sensitive to changes in the terrain profile: when the domain width is reduced to \SI{300}{\kilo\meter}, the $\ell_2$ error on the BTF grid rises from \num{0.284} to \num{0.432}.
+The same sensitivity is found using terrain following coordinates and centred differences in space and time, as implemented by \citet{schaer2002}.
+This sensitivity can be explained by realising that increased non-orthogonality or flow misalignment tends to increase errors.
+To measure non-orthogonality, consider the total variation, $TV$, of the discrete terrain profile:
+\begin{align}
+	TV = \sum_i | h(x_{i+1}) - h(x_i) |
+\end{align}
+where $x_i$ is the horizontal location of a vertex on the lower boundary.
+For a domain width of \SI{301}{\kilo\meter}, $TV = \SI{18095}{\meter}$, but for a width of \SI{300}{\kilo\meter}, $TV = \SI{18964}{\meter}$.
 The upwind-biased cubic scheme is less sensitive to the same change, with the $\ell_2$ error decreasing from \num{0.107} to \num{0.104}.
@@
 \begin{equation}
-	\Psi(x,z) = u_0 H \frac{z - h}{H - h} \label{eqn:streamfunc-btf}
+	\Psi(x,z) = -u_0 H \frac{z - h}{H - h} \label{eqn:streamfunc-btf}
 \end{equation}
@@
 \begin{align}
-	u &= \frac{\partial \Psi}{\partial z} = u_0 \frac{H}{H - h}, \quad w = -\frac{\partial \Psi}{\partial x} = u_0 H \frac{\mathrm{d} h}{\mathrm{d} x} \frac{H - z}{\left( H - h \right)^2} \label{eqn:uw-btf} \\
+	u &= -\frac{\partial \Psi}{\partial z} = u_0 \frac{H}{H - h}, \quad w = \frac{\partial \Psi}{\partial x} = u_0 H \frac{\mathrm{d} h}{\mathrm{d} x} \frac{H - z}{\left( H - h \right)^2} \label{eqn:uw-btf} \\
 	\frac{\mathrm{d} h}{\mathrm{d} x} &= - h_0 \left[ 
@@
 \begin{align}
-	\Psi(x,z) = u_0 \left\{ \begin{array}{l l}
+	\Psi(x,z) = -u_0 \left\{ \begin{array}{l l}
 			H_\mathrm{TF} \frac{z - h}{H_\mathrm{TF} - h} & \enskip \text{if $z \leq H_\mathrm{TF}$} \\

### 2015-09-22 src/tf-cutcell-comparison/tf-cutcell-comparison.tex
add example of snappyHexMesh behaviour including diagram

@@
-\documentclass[draft]{ametsoc}
+\documentclass[twocol]{ametsoc}
+\journal{mwr}
 
@@
 First, when \texttt{snappyHexMesh} moves vertices onto the terrain surface, some points are moved horizontally.
-Second, the utility does not create new points necessary for pentagonal cells.
+Second, it does not create new points necessary for pentagonal cells.
+An example of these effects are shown in figure~\ref{fig:snappyHexMesh} in which vertex $v$ has been moved vertically and horizontally to avoid creating a small cell.
+The resultant mesh, shown by thick dashed lines, contains triangular and quadrangular cells.
+With a typical shaving method, cell $c$ would be pentagonal, as shown by the dotted line.
+However, \texttt{snappyHexMesh} creates a quadrilateral cell.
 
@@
 \end{align}
+and similarly for \(\partial \phi_i / \partial z\).
 
@@
 The same sensitivity is found using terrain following coordinates and centred differences in space and time, as implemented by \citet{schaer2002}.
-This sensitivity can be explained by realising that increased non-orthogonality or flow misalignment tends to increase errors.
-To measure non-orthogonality, consider the total variation, $TV$, of the discrete terrain profile:
-\begin{align}
-	TV = \sum_i | h(x_{i+1}) - h(x_i) |
-\end{align}
-where $x_i$ is the horizontal location of a vertex on the lower boundary.
-For a domain width of \SI{301}{\kilo\meter}, $TV = \SI{18095}{\meter}$, but for a width of \SI{300}{\kilo\meter}, $TV = \SI{18964}{\meter}$.
 The upwind-biased cubic scheme is less sensitive to the same change, with the $\ell_2$ error decreasing from \num{0.107} to \num{0.104}.
+It is likely that changes in the domain width affect the wave power spectrum of the discrete terrain profile.
+High-frequency oscillations persist in the centred linear scheme but are damped by the upwind-biased scheme.
+This is an area for further study.
 
@@
 	\centering
+	\includegraphics{../fig-snappyHexMesh/fig-snappyHexMesh.pdf}
+	%
+	\caption{Example behaviour of the \texttt{snappyHexMesh} tool in cut cell grid creation.
+A terrain surface denoted by a thick grey line is intersected with an orthogonal grid shown by thin solid lines.
+Vertex $v$ is `snapped' to the surface, moving from the position marked by the open circle to that of the closed circle.
+The resultant cut cell grid is shown by thick dashed lines.
+With a typical cut cell technique, cell $c$ would be a pentagon, as shown by the dotted line, but \texttt{snappyHexMesh} produces a quadrangular cell.}
+	\label{fig:snappyHexMesh}
+\end{figure*}
+
+\begin{figure*}
+	\centering
 	\includegraphics{../fig-advection-tracer/fig-advection-tracer.pdf}

### 2015-09-28 src/tf-cutcell-comparison/tf-cutcell-comparison.tex
tweak caption for cut cell mesh generation figure

@@
 The OpenFOAM utility \texttt{snappyHexMesh} was used to create a grid that approximates the cut cell method.
-First, a custom utility was used to move points beneath the surface up to the surface creating small cells near mountain peaks.
-Second, the surface faces were taken from the BTF grid and \texttt{snappyHexMesh} was used to intersect the surface with the grid.
-This tool removes cells whose centres are below the surface and displaces boundary vertices so that they are `snapped' to the BTF surface \citep{openfoam2015}.
+First, a custom utility modifies an orthogonal grid, moving points beneath the surface upwards so that they lie just beneath the surface, creating layers of thin cells.
+Second, the surface faces were taken from the BTF grid and \texttt{snappyHexMesh} was used to intersect the surface with the grid produced by the custom utility.
+This tool removes cells whose centres are below the surface, including the layers of thin cells produced in the first step.
+In addition, \texttt{snappyHexMesh} displaces boundary vertices so that they are `snapped' to the BTF surface \citep{openfoam2015}.
+It was found that, by using the custom utility to preprocess the orthogonal grid, \texttt{snappyHexMesh} performed fewer, smaller vertex displacements, resulting in grids that are closer to those produced by typical shaving methods, as described by \citet{adcroft1997}.
 An example of the resulting grid is shown in figure~\ref{fig:grid}c.
 
-There are two details of grid construction which mean that resulting cut cell grids can differ slightly from a typical cut cell grid created using a shaving method, as described by \citet{adcroft1997}.
+There are two details of grid construction which mean that resulting cut cell grids can differ slightly from a typical cut cell grid.
 First, when \texttt{snappyHexMesh} moves vertices onto the terrain surface, some points are moved horizontally.
 Second, it does not create new points necessary for pentagonal cells.
-An example of these effects are shown in figure~\ref{fig:snappyHexMesh} in which vertex $v$ has been moved vertically and horizontally to avoid creating a small cell.
+An example of these effects is shown in figure~\ref{fig:snappyHexMesh} in which vertex $v$ has been moved vertically and horizontally to avoid creating a small cell.
 The resultant mesh, shown by thick dashed lines, contains triangular and quadrangular cells.
@@
 Following \citet{melvin2010}, the domain is \SI{300}{\kilo\meter} wide and \SI{30}{\kilo\meter} high with a grid resolution of \(\Delta x = \SI{500}{\meter}\) and \(\Delta z^\star = \SI{300}{\meter}\).
-\TODO{we want to consider dz=250m and dz=200m, too} The mountain profile has the same form as equation~\ref{eqn:resting:mountain}.
+\TODO{we want to consider much higher resolutions, too} The mountain profile has the same form as equation~\ref{eqn:resting:mountain}.
 Tests are performed with mountain heights of $h_0 = \SI{250}{\meter}$ and \(h_0 = \SI{500}{\meter}\).
@@
 	%
-	\caption{Example behaviour of the \texttt{snappyHexMesh} tool in cut cell grid creation.
-A terrain surface denoted by a thick grey line is intersected with an orthogonal grid shown by thin solid lines.
-Vertex $v$ is `snapped' to the surface, moving from the position marked by the open circle to that of the closed circle.
+	\caption{Illustrative behaviour of the \texttt{snappyHexMesh} tool in cut cell grid creation.
+The terrain surface, denoted by a thick grey line, is taken from the lower boundary of the BTF grid.
+The surface is intersected with an orthogonal grid shown by thin solid lines.
+\TODO{I think this example is clearer showing the difference between an orthogonal grid and the resulting cut cell grid.
+In fact, we modify the orthogonal grid with a custom utility before it is passed to snappyHexMesh}  Vertex $v$ is `snapped' to the surface, moving from the position marked by the open circle to that of the closed circle.
 The resultant cut cell grid is shown by thick dashed lines.

### 2015-09-28 src/tf-cutcell-comparison/tf-cutcell-comparison.tex
introduce thermal advection test

@@
 It was found that, by using the custom utility to preprocess the orthogonal grid, \texttt{snappyHexMesh} performed fewer, smaller vertex displacements, resulting in grids that are closer to those produced by typical shaving methods, as described by \citet{adcroft1997}.
-An example of the resulting grid is shown in figure~\ref{fig:grid}c.
 
@@
 However, \texttt{snappyHexMesh} creates a quadrilateral cell.
+
+\TODO{give multiple examples of cut cell grids used in subsequent test cases.  acknowledge that we can still get thin cells in some instances.}
 
@@
 \subsection{Terrain following advection of thermal profile}
+The potential temperature errors in the gravity waves test varied between terrain following and cut cell grids of the same resolution.
+This variation may be due to differences in the wind fields between grids, or errors in the advection of potential temperature, amongst other possible causes.
+To help establish the primary source of error, a new advection test is formulated in which the initial potential temperature field from the gravity waves test is used as a tracer density field.
+To eliminate any differences in wind fields, the tracer is advected in a fixed, terrain-following velocity field that mimics the flow in the gravity waves test.
+
 The spatial domain, mountain profile, grid resolutions and timesteps are the same as those in the gravity waves test in section~\ref{sec:results}\ref{sec:gw}.
@@
 
+Columns of low tracer density are seen above the mountain peaks which have been advected upwards from the surface on the windward mountain slope, with the lowest tracer densities over the highest central peak.
+
+\TODO{analyse results of test, draw conclusions}
+
 \section{Conclusions}
@@
 Conversely, in the terrain following advection test, errors were large on the SLEVE and cut cell grids where velocities were misaligned with the grids.
+
+\TODO{do we want to say more about how cubicUpwindCPCFit is designed to cope with small cells?
+this approach differs from other existing approaches.
+the technique will be described in future publication.
+(although an implementation of the technique is already available online)}
 

### 2015-09-30 src/tf-cutcell-comparison/tf-cutcell-comparison.tex
remove old gravityWaves plot, add some more description of theta errors

@@
 
-Following \citet{melvin2010}, the domain is \SI{300}{\kilo\meter} wide and \SI{30}{\kilo\meter} high with a grid resolution of \(\Delta x = \SI{500}{\meter}\) and \(\Delta z^\star = \SI{300}{\meter}\).
-\TODO{we want to consider much higher resolutions, too} The mountain profile has the same form as equation~\ref{eqn:resting:mountain}.
-Tests are performed with mountain heights of $h_0 = \SI{250}{\meter}$ and \(h_0 = \SI{500}{\meter}\).
+Following \citet{melvin2010}, the domain is \SI{300}{\kilo\meter} wide and \SI{30}{\kilo\meter} high.
+\TODO{we want to consider much higher resolutions, too} The mountain profile has the same form as equation~\ref{eqn:resting:mountain}, but the gravity waves tests have a mountain height of $h_0 = \SI{250}{\meter}$.
 As in the resting atmosphere test, $a = \SI{5}{\kilo\meter}$ is the mountain half-width and $\lambda = \SI{4}{\kilo\meter}$ is the wavelength.
@@
 For the Exner function of pressure, hydrostatic balance is prescribed on all boundaries.
-The simulation is integrated forward by 5 hours with a timestep $\Delta t = \SI{8}{\second}$.
 
-Test results are compared between the BTF, SLEVE and cut cell grids.
-Vertical velocities on the BTF, SLEVE and cut cell grids are visually indistinguishable (not shown).
+Test results are compared between the BTF, SLEVE and cut cell grids at several resolutions.
+The spatial and temporal resolutions tested are shown in table~\ref{tab:gw-resolutions}.
+The lowest resolution is the same as that used by \citet{schaer2002}, and higher resolutions preserve the same aspect ratio.
+The vertical resolution is chosen to test various configurations of cut cell grid.
+At $\Delta z = \SI{300}{\meter}$, the mountain lies entirely within the lowest layer of cells, while at $\Delta z = \SI{250}{\meter}$ and $\Delta z = \SI{125}{\meter}$ the mountain peak is aligned with the interface between layers.
+With increasing resolutions up to $\Delta z = \SI{50}{\meter}$, the orography intersects more layers and becomes better resolved.
+\TODO{a reviewer (and Terry) suggested these tests, but we haven't found anything notable with any particular cut cell grid.
+so I don't see much value in some of this description.
+also not sure how much value there is in having a figure showing several cut cell grids (e.g.
+300dz, 250dz, 200dz)}
+The simulation is integrated forward by 5 hours and the timestep, $\Delta t = 8 \Delta z / \SI{300}{\second}$, is chosen so that it scales linearly with spatial resolution and, following the original test specified by \citet{schaer2002}, $\Delta t = \SI{8}{\second}$ when $\Delta z = \SI{300}{\meter}$.
+
+At $\Delta z = \SI{300}{\meter}$, vertical velocities on the BTF, SLEVE and cut cell grids are visually indistinguishable (not shown).
 They agree with the high resolution mass-conserving semi-implicit semi-Lagrangian solution from \citet{melvin2010}.
+\TODO{perhaps I should spot check some of the grids at higher resolutions but, given that the theta fields converge overall, I don't see why vertical velocities contours would do anything funky}
 
 The initial thermal profile is subtracted from the potential temperature field at the end of the integration to reveal the structure of thermal anomalies.
-Once again, the results are similar on all three grids, and results are shown on the BTF and cut cell grids in figures~\ref{fig:gw-theta}a and \ref{fig:gw-theta}c respectively.
-However, examining more closely the anomalies in the lee of the mountain for the cut cell grid, figure~\ref{fig:gw-theta}d shows that the bottommost layer is anomalously warm and the layer above it is anomalously cold.
-This feature is not present on the BTF grid (figure~\ref{fig:gw-theta}b) or the SLEVE grid (not shown).
-\TODO{not present?
-I think it is, just much smaller amplitude}
+Results are similar across all grid types and resolutions, but examining the anomalies in the lee of the mountain errors are found in the lowest few layers: the bottommost layer is anomalously warm and the layer above it is anomalously cold.
+\TODO{this isn't true at all resolutions, e.g.
+50dz is cold at the bottom}  At all resolutions, the amplitude of the errors is greater on the cut cell grid than the BTF \TODO{or SLEVE?} grid.
+As resolution increases, the anomalously cold layers first increase in amplitude, then begin to span multiple layers, which suggests that the errors are controlled by the mountain height.
+\TODO{that conclusion might be stronger with tests at different mountain heights?
+oh no, not more tests!
+:-P}
 
@@
 
-Cut cell grids reduced errors in two of the four tests.
+Cut cell grids reduced errors in two of the five tests.
 First, in the horizontal advection test, tracer accuracy on the cut cell grid was almost as good as accuracy on a regular grid with no mountain.
@@
 
+\begin{table}
+	\caption{Spatial and temporal resolutions used in the gravity waves test.
+The lowest resolution has the same parameters as the original test case specified by \citet{schaer2002}.
+At higher resolutions, the vertical resolution is chosen, and horizontal and temporal resolutions are calculated to preserve the same aspect ratio.
+\TODO{don't need to include all these resolutions, just those need to do suitable analysis/conclusions}}
+	\label{tab:gw-resolutions}
+%
+\centering
+\begin{tabular}{S S S}
+\hline\hline
+{$\Delta z$ (\si{\meter})} & {$\Delta x$ (\si{\meter})} & {$\Delta t$ (\si{\second})} \\
+\hline
+300	& 500	& 8 \\
+250	& 416.7 & 6.667 \\
+200	& 333.3 & 5.333 \\
+150	& 250	& 4 \\
+125	& 208.3 & 3.333 \\
+100	& 166.7 & 2.667 \\
+75	& 125	& 2 \\
+50	& 83.33 & 1.333 \\
+\hline
+\end{tabular}
+\end{table}
+
 % FIGURES
@@
 
-\begin{figure*}
-	\centering
-	\includegraphics{../fig-gravityWaves-theta/fig-gravityWaves-theta.pdf}
-%
-	\caption{Anomalies in potential temperature in the gravity waves test after 5 hours with a mountain height, \(h_0 = \SI{250}{\meter}\).
-The central domain in the lowest \SI{12}{\kilo\meter} is shown on (a) the BTF grid, and (c) the cut cell grid.
-The four lowest layers of each grid are shown for (b) BTF, and (d) cut cell grids, using a narrower potential temperature scale.
-The results on the SLEVE grid (not shown) are qualitatively identical to results on the BTF grid.
-Axes are in units of \si{\meter}.}
-	\label{fig:gw-theta}
-\end{figure*}
-
 \end{document}

### 2015-10-01 src/tf-cutcell-comparison/fig-gravityWaves.tex
plot more gravityWave results and further analyse them

\documentclass{article}
\usepackage{showframe}
\usepackage[paperwidth=6.4in,paperheight=8.3in,margin=0in]{geometry}
\usepackage{mathptmx}%{times}
\usepackage{graphicx}
\usepackage{siunitx}
\usepackage[absolute,overlay]{textpos}
\usepackage{color}

\setlength{\parindent}{0pt}
\begin{document}
\centering
\includegraphics[width=6in]{../gravityWaves-btf-250m-300dz/18000/thetaDiffOverview.pdf}
\vspace*{0.1in}
\includegraphics[width=2in]{../gravityWaves-btf-250m-300dz/18000/thetaDiffMountain.pdf}
\includegraphics[width=1in]{../gravityWaves-btf-250m-300dz/18000/thetaDiffLee.pdf}
\hspace{0.2in}
\includegraphics[width=2in]{../gravityWaves-cutCell-250m-300dz/18000/thetaDiffMountain.pdf}
\includegraphics[width=1in]{../gravityWaves-cutCell-250m-300dz/18000/thetaDiffLee.pdf}
\vspace*{0.1in}
\includegraphics[width=2in]{../gravityWaves-btf-250m-250dz/18000/thetaDiffMountain.pdf}
\includegraphics[width=1in]{../gravityWaves-btf-250m-250dz/18000/thetaDiffLee.pdf}
\hspace{0.2in}
\includegraphics[width=2in]{../gravityWaves-cutCell-250m-250dz/18000/thetaDiffMountain.pdf}
\includegraphics[width=1in]{../gravityWaves-cutCell-250m-250dz/18000/thetaDiffLee.pdf}
\vspace*{0.1in}
\includegraphics[width=2in]{../gravityWaves-btf-250m-300dz/18000/thetaDiffMountain.pdf}
\includegraphics[width=1in]{../gravityWaves-btf-250m-300dz/18000/thetaDiffLee.pdf}
\hspace{0.2in}
\includegraphics[width=2in]{../gravityWaves-btf-250m-300dz/18000/thetaDiffMountain.pdf}
\includegraphics[width=1in]{../gravityWaves-btf-250m-300dz/18000/thetaDiffLee.pdf}
\vspace*{0.1in}
\includegraphics[width=2in]{../gravityWaves-btf-250m-300dz/18000/thetaDiffMountain.pdf}
\includegraphics[width=1in]{../gravityWaves-btf-250m-300dz/18000/thetaDiffLee.pdf}
\hspace{0.2in}
\includegraphics[width=2in]{../gravityWaves-btf-250m-300dz/18000/thetaDiffMountain.pdf}
\includegraphics[width=1in]{../gravityWaves-btf-250m-300dz/18000/thetaDiffLee.pdf}
\end{document}

### 2015-10-01 src/tf-cutcell-comparison/tf-cutcell-comparison.tex
plot more gravityWave results and further analyse them

@@
 The initial thermal profile is subtracted from the potential temperature field at the end of the integration to reveal the structure of thermal anomalies.
-Results are similar across all grid types and resolutions, but examining the anomalies in the lee of the mountain errors are found in the lowest few layers: the bottommost layer is anomalously warm and the layer above it is anomalously cold.
-\TODO{this isn't true at all resolutions, e.g.
-50dz is cold at the bottom}  At all resolutions, the amplitude of the errors is greater on the cut cell grid than the BTF \TODO{or SLEVE?} grid.
-As resolution increases, the anomalously cold layers first increase in amplitude, then begin to span multiple layers, which suggests that the errors are controlled by the mountain height.
-\TODO{that conclusion might be stronger with tests at different mountain heights?
-oh no, not more tests!
-:-P}
+Results are similar across all grid types and resolutions but, examining the anomalies in the lee of the mountain, errors are found in the lowest few layers.
+Errors are generated on the lee slopes of the mountain and are advected to the outlet boundary, forming striped anomalies.
+At resolution of $\Delta z > \SI{50}{\meter}$, the lowest layer is anomalously warm and the layer above is anomalously cold.
+As resolution increases, the anomalously cold layer first increases in amplitude, then the cold anomaly begins to span multiple layers.
+At $\Delta z = \SI{50}{\meter}$, the structure is reversed, having lower layers that are too cold and layers above that are too warm.
+At all resolutions, the amplitude and height of the errors are greater on the cut cell grid compared to the BTF \TODO{or SLEVE?} grid.
 
 To summarize, results of the gravity waves test on all grids are in good agreement with the reference solution from \citet{melvin2010}.
+Stripes of potential temperature anomalies are found in the lee of the mountain, and these errors are most severe on the cut cell grid.
+\TODO{say we want to find out what's causing them}
 
@@
 
+\begin{figure*}
+	\centering
+	\includegraphics[width=6in]{../fig-gravityWaves/fig-gravityWaves.pdf}
+%
+	\caption{\TODO{}}
+	\label{fig:gw}
+\end{figure*}
+
 \end{document}

### 2015-10-01 src/tf-cutcell-comparison/fig-gravityWaves-theta.tex
don't plot heatmap for lee of mountain, will plot this as a line plot instead

--- 2015-02-10 fig-gravityWaves-theta.tex
@@
 \documentclass{article}
-\usepackage[paperwidth=6.4in,paperheight=4.1in,margin=0in]{geometry}
+\usepackage{showframe}
+\usepackage[paperwidth=6.4in,paperheight=8.3in,margin=0in]{geometry}
 \usepackage{mathptmx}%{times}
@@
 \begin{document}
-\includegraphics[height=2.2in,angle=270]{../gravityWaves-btf/18000/thetaDiff.eps}
-\hspace{0.1in}
-\includegraphics[height=4in,angle=270]{../gravityWaves-btf/18000/thetaDiffZoom.eps}
+\centering
+\includegraphics[width=6in]{../gravityWaves-btf-250m-300dz/18000/thetaDiffOverview.pdf}
 \vspace*{0.1in}
-
-\includegraphics[height=2.2in,angle=270]{../gravityWaves-snapCol/18000/thetaDiff.eps}
-\hspace{0.1in}
-\includegraphics[height=4in,angle=270]{../gravityWaves-snapCol/18000/thetaDiffZoom.eps}
+\includegraphics[width=3in]{../gravityWaves-btf-250m-300dz/18000/thetaDiffMountain.pdf}
+\includegraphics[width=3in]{../gravityWaves-cutCell-250m-300dz/18000/thetaDiffMountain.pdf}
 \vspace*{0.1in}
-
-\includegraphics[height=2.2in,angle=270]{../gravityWaves-diagnostics/thetaDiffLegend.eps}
-\hspace{0.2in}
-\includegraphics[height=3.9in,angle=270]{../gravityWaves-diagnostics/thetaDiffZoomLegend.eps}
+\includegraphics[width=3in]{../gravityWaves-btf-250m-250dz/18000/thetaDiffMountain.pdf}
+\includegraphics[width=3in]{../gravityWaves-cutCell-250m-250dz/18000/thetaDiffMountain.pdf}
+\vspace*{0.1in}
+\includegraphics[width=3in]{../gravityWaves-btf-250m-300dz/18000/thetaDiffMountain.pdf}
+\includegraphics[width=3in]{../gravityWaves-btf-250m-300dz/18000/thetaDiffMountain.pdf}
+\vspace*{0.1in}
+\includegraphics[width=3in]{../gravityWaves-btf-250m-300dz/18000/thetaDiffMountain.pdf}
+\includegraphics[width=3in]{../gravityWaves-btf-250m-300dz/18000/thetaDiffMountain.pdf}
 \end{document}

### 2015-10-01 src/tf-cutcell-comparison/tf-cutcell-comparison.tex
don't plot heatmap for lee of mountain, will plot this as a line plot instead

@@
 As in the resting atmosphere test, $a = \SI{5}{\kilo\meter}$ is the mountain half-width and $\lambda = \SI{4}{\kilo\meter}$ is the wavelength.
-On the optimised SLEVE grid, $s_1 = \SI{5}{\kilo\meter}$ is the large scale height, $s_2 = \SI{2}{\kilo\meter}$ is the small scale height and the optimal exponent value $n = 1.35$ is used.
 
@@
 
-Note that, while the domain itself is \SI{30}{\kilo\meter} in height, for the purposes of generating BTF and SLEVE grids, the domain height is set to \SI{20}{\kilo\meter} because the sponge layer occupies the uppermost \SI{10}{\kilo\meter}.
+Note that, while the domain itself is \SI{30}{\kilo\meter} in height, for the purposes of generating BTF grids, the domain height is set to \SI{20}{\kilo\meter} because the sponge layer occupies the uppermost \SI{10}{\kilo\meter}.
 
@@
 
-Test results are compared between the BTF, SLEVE and cut cell grids at several resolutions.
+Test results are compared between the BTF and cut cell grids at several resolutions.
 The spatial and temporal resolutions tested are shown in table~\ref{tab:gw-resolutions}.
@@
 
-At $\Delta z = \SI{300}{\meter}$, vertical velocities on the BTF, SLEVE and cut cell grids are visually indistinguishable (not shown).
+At $\Delta z = \SI{300}{\meter}$, vertical velocities on the BTF and cut cell grids are visually indistinguishable (not shown).
 They agree with the high resolution mass-conserving semi-implicit semi-Lagrangian solution from \citet{melvin2010}.
-\TODO{perhaps I should spot check some of the grids at higher resolutions but, given that the theta fields converge overall, I don't see why vertical velocities contours would do anything funky}
 
@@
 At $\Delta z = \SI{50}{\meter}$, the structure is reversed, having lower layers that are too cold and layers above that are too warm.
-At all resolutions, the amplitude and height of the errors are greater on the cut cell grid compared to the BTF \TODO{or SLEVE?} grid.
+At all resolutions, the amplitude and height of the errors are greater on the cut cell grid compared to the BTF grid.
 
@@
 	\centering
-	\includegraphics[width=6in]{../fig-gravityWaves/fig-gravityWaves.pdf}
+	\includegraphics[width=6in]{../fig-gravityWaves-theta/fig-gravityWaves-theta.pdf}
 %
 	\caption{\TODO{}}
-	\label{fig:gw}
+	\label{fig:gw-theta}
 \end{figure*}

### 2015-10-02 src/tf-cutcell-comparison/fig-gravityWaves-theta.tex
add sampleLine plot

@@
 \vspace*{0.1in}
-\includegraphics[width=3in]{../gravityWaves-btf-250m-300dz/18000/thetaDiffMountain.pdf}
-\includegraphics[width=3in]{../gravityWaves-btf-250m-300dz/18000/thetaDiffMountain.pdf}
+\includegraphics[width=3in]{../gravityWaves-btf-250m-200dz/18000/thetaDiffMountain.pdf}
+\includegraphics[width=3in]{../gravityWaves-cutCell-250m-200dz/18000/thetaDiffMountain.pdf}
 \vspace*{0.1in}
-\includegraphics[width=3in]{../gravityWaves-btf-250m-300dz/18000/thetaDiffMountain.pdf}
-\includegraphics[width=3in]{../gravityWaves-btf-250m-300dz/18000/thetaDiffMountain.pdf}
+\includegraphics[width=3in]{../gravityWaves-btf-250m-150dz/18000/thetaDiffMountain.pdf}
+\includegraphics[width=3in]{../gravityWaves-cutCell-250m-150dz/18000/thetaDiffMountain.pdf}
 \end{document}

### 2015-10-11 src/tf-cutcell-comparison/fig-gravityWaves-theta.tex
add more gravityWaves plots (just for comparison)

@@
 \includegraphics[width=3in]{../gravityWaves-cutCell-250m-150dz/18000/thetaDiffMountain.pdf}
+\vspace*{0.1in}
+\includegraphics[width=3in]{../gravityWaves-btf-250m-100dz/18000/thetaDiffMountain.pdf}
+\includegraphics[width=3in]{../gravityWaves-cutCell-250m-100dz/18000/thetaDiffMountain.pdf}
+\vspace*{0.1in}
+\includegraphics[width=3in]{../gravityWaves-btf-250m-75dz/18000/thetaDiffMountain.pdf}
+\includegraphics[width=3in]{../gravityWaves-cutCell-250m-75dz/18000/thetaDiffMountain.pdf}
 \end{document}

### 2015-10-14 src/tf-cutcell-comparison/fig-gravityWaves-theta.tex

--- 2015-10-01 src/tf-cutcell-comparison/fig-gravityWaves-theta.tex
@@
 \includegraphics[width=3in]{../gravityWaves-cutCell-250m-300dz/18000/thetaDiffMountain.pdf}
+300m
 \vspace*{0.1in}
@@
 \includegraphics[width=3in]{../gravityWaves-cutCell-250m-250dz/18000/thetaDiffMountain.pdf}
+250m
 \vspace*{0.1in}
-\includegraphics[width=3in]{../gravityWaves-btf-250m-300dz/18000/thetaDiffMountain.pdf}
-\includegraphics[width=3in]{../gravityWaves-btf-250m-300dz/18000/thetaDiffMountain.pdf}
+\includegraphics[width=3in]{../gravityWaves-btf-250m-200dz/18000/thetaDiffMountain.pdf}
+\includegraphics[width=3in]{../gravityWaves-cutCell-250m-200dz/18000/thetaDiffMountain.pdf}
+200m
 \vspace*{0.1in}
-\includegraphics[width=3in]{../gravityWaves-btf-250m-300dz/18000/thetaDiffMountain.pdf}
-\includegraphics[width=3in]{../gravityWaves-btf-250m-300dz/18000/thetaDiffMountain.pdf}
+\includegraphics[width=3in]{../gravityWaves-btf-250m-150dz/18000/thetaDiffMountain.pdf}
+\includegraphics[width=3in]{../gravityWaves-cutCell-250m-150dz/18000/thetaDiffMountain.pdf}
+150m
+\vspace*{0.1in}
+%\includegraphics[width=3in]{../gravityWaves-btf-250m-100dz/18000/thetaDiffMountain.pdf}
+%\includegraphics[width=3in]{../gravityWaves-cutCell-250m-100dz/18000/thetaDiffMountain.pdf}
+%100m (OLD)
+%\vspace*{0.1in}
+%\includegraphics[width=3in]{../gravityWaves-btf-250m-75dz/18000/thetaDiffMountain.pdf}
+%\includegraphics[width=3in]{../gravityWaves-cutCell-250m-75dz/18000/thetaDiffMountain.pdf}
+%75m (OLD)
 \end{document}

### 2015-10-14 src/tf-cutcell-comparison/tf-cutcell-comparison.tex
add sampleLine plot
use alpha to tidy up dh/dx for thermal advection

@@
 
-% Every term in the discretised equations of motion is a potential source of numerical error.
-% Errors in calculating the horizontal pressure gradient generate spurious circulations and much effort has been put into improving this calculation.
-This is often achieved by making the grid more orthogonal: flattening terrain following coordinate surfaces [citations] or using a cut cell approach [citations].
-% TF citations: mahrer1984, schaer2002, klemp2011
-% cut cell citations: good2014, yamazaki2015 puts this clearly but not yet published
-% In a test of a stably stratified atmosphere initially at rest, we find that spurious velocities are no more than \SI{0.1}{\meter\per\second} on any grid.
-[TODO: and this is much better than klemp2011]
-% Advection terms also contribute to numerical error.
-We show that advection is most accurate when the flow is aligned with the grid layers.
-Near the surface, flow tends to follow the terrain which results in larger errors when more orthogonal grids are used.
-% In a test of orographically induced gravity waves, we find that the advection of potential temperature generates grid-scale errors on the cut cell grid.
-
 \abstract{Terrain following coordinates are widely used in operational models but the cut cell method has been proposed as an alternative that can more accurately represent atmospheric dynamics over steep orography.
@@
 \maketitle
+\TODO{this is preliminary, but somewhere in the body I want to highlight the tension between horizontal pressure gradient/advection errors and the choice of grid:}
+
+\TODO{Every term in the discretised equations of motion is a potential source of numerical error.
+ Errors in calculating the horizontal pressure gradient generate spurious circulations and much effort has been put into improving this calculation.
+This is often achieved by making the grid more orthogonal: flattening terrain following coordinate surfaces [citations] or using a cut cell approach [citations].
+ TF citations: mahrer1984, schaer2002, klemp2011
+ cut cell citations: good2014, yamazaki2015 puts this clearly but not yet published
+ In a test of a stably stratified atmosphere initially at rest, we find that spurious velocities are no more than \SI{0.1}{\meter\per\second} on any grid.
+[TODO: and this is much better than klemp2011]
+ Advection terms also contribute to numerical error.
+We show that advection is most accurate when the flow is aligned with the grid layers.
+Near the surface, flow tends to follow the terrain which results in larger errors when more orthogonal grids are used.
+ In a test of orographically induced gravity waves, we find that the advection of potential temperature generates grid-scale errors on the cut cell grid.}
 
@@
 Following \citet{melvin2010}, the domain is \SI{300}{\kilo\meter} wide and \SI{30}{\kilo\meter} high.
-\TODO{we want to consider much higher resolutions, too} The mountain profile has the same form as equation~\ref{eqn:resting:mountain}, but the gravity waves tests have a mountain height of $h_0 = \SI{250}{\meter}$.
+The mountain profile has the same form as equation~\ref{eqn:resting:mountain}, but the gravity waves tests have a mountain height of $h_0 = \SI{250}{\meter}$.
 As in the resting atmosphere test, $a = \SI{5}{\kilo\meter}$ is the mountain half-width and $\lambda = \SI{4}{\kilo\meter}$ is the wavelength.
 
-The initial thermodynamic conditions have a surface temperature of $\theta(z=0) = \SI{288}{\kelvin}$ and constant stability with $N = \SI{0.01}{\per\second}$ everywhere.
-A uniform horizontal wind $u = \SI{10}{\meter\per\second}$ is prescribed at the inlet boundary.
+A uniform horizontal wind $u = \SI{10}{\meter\per\second}$ is prescribed in the interior domain and at the inlet boundary.
+No normal flow is imposed at the top and bottom boundaries and the velocity field has a zero gradient outlet boundary condition.
+
+The initial thermodynamic conditions have a temperature at $z=0$ of $\theta_0 = \SI{288}{\kelvin}$ and constant static stability with $N = \SI{0.01}{\per\second}$ everywhere, such that
+\begin{align}
+	\theta(z) = \theta_0 \exp \left( \frac{N^2}{g} z \right) \label{eqn:thermal-profile}
+\end{align}
+for $z > 0$.
+Potential temperature values are prescribed at the inlet and upper boundary using equation~\ref{eqn:thermal-profile}, and a zero gradient boundary condition is applied at the outlet.
+At the ground, fixed gradients are imposed by calculating the component of $\nabla \theta$ normal to each face using the vertical derivative of equation~\ref{eqn:thermal-profile}.
+For the Exner function of pressure, hydrostatic balance is prescribed on all boundaries.
 
@@
 
-No normal flow is imposed at the top and bottom boundaries and the outlet has a zero gradient boundary condition.
-For the Exner function of pressure, hydrostatic balance is prescribed on all boundaries.
-
 Test results are compared between the BTF and cut cell grids at several resolutions.
@@
 With increasing resolutions up to $\Delta z = \SI{50}{\meter}$, the orography intersects more layers and becomes better resolved.
-\TODO{a reviewer (and Terry) suggested these tests, but we haven't found anything notable with any particular cut cell grid.
-so I don't see much value in some of this description.
-also not sure how much value there is in having a figure showing several cut cell grids (e.g.
-300dz, 250dz, 200dz)}
+
 The simulation is integrated forward by 5 hours and the timestep, $\Delta t = 8 \Delta z / \SI{300}{\second}$, is chosen so that it scales linearly with spatial resolution and, following the original test specified by \citet{schaer2002}, $\Delta t = \SI{8}{\second}$ when $\Delta z = \SI{300}{\meter}$.
@@
 \begin{align}
-	\frac{\mathrm{d} h}{\mathrm{d} x} = -h_0 \exp \left(- \left( \frac{x}{a} \right)^2 \right) \left[ \frac{\pi}{\lambda} \sin \left( \frac{2 \pi x}{\lambda} \right) - \frac{2x}{a^2} \cos^2\left( \frac{\pi x}{\lambda} \right) \right]
+	\frac{\mathrm{d} h}{\mathrm{d} x} = -h_0 \exp \left(- \left( \frac{x}{a} \right)^2 \right) \left[ \alpha \sin \left( 2 \alpha x \right) - \frac{2x}{a^2} \cos^2\left( \alpha x \right) \right]
 \end{align}
@@
 
-The tracer field, \(\phi\), is the same as the initial potential temperature field in the gravity waves test:
+The initial tracer density field, \(\phi\), and its boundary conditions, are the same as those of the initial potential temperature field in the gravity waves test.
+Following the gravity waves test, the simulation is integrated forward by \SI{18000}{\second}, by which time the tracer initially upwind of the mountain will have cleared the mountain range.
+Hence, the analytic solution, $\phi_T$, can be found by considering the vertical displacement of the tracer profile by the terrain following velocity field.
 \begin{align}
-	\phi(z) = \phi_0 \exp \left( \frac{N^2}{g} z \right) \label{eqn:thermal-tracer}
+	\phi_T(x, z) = \phi_0 \exp \left( \frac{N^2}{g} z^\star(x, z) \right) 
 \end{align}
-where \(\phi_0 = 288\) and \(N = \SI{0.01}{\per\second}\).
-The inlet boundary has fixed values prescribed by equation~\ref{eqn:thermal-tracer} and a zero gradient boundary condition is applied at the outlet.
-
-Columns of low tracer density are seen above the mountain peaks which have been advected upwards from the surface on the windward mountain slope, with the lowest tracer densities over the highest central peak.
+where the tracer density at $z = 0$, $\phi_0 = 288$, and the transform, $z^\star$, is given by rearranging equation~\ref{eqn:btf}.
+\TODO{is this clear enough?
+my point is I don't need to worry about an analytic solution where only /part/ of the tracer has been advected over the mountain range.
+the analytic solution at t=18000s is symmetric about x=0.}
 
@@
 At higher resolutions, the vertical resolution is chosen, and horizontal and temporal resolutions are calculated to preserve the same aspect ratio.
-\TODO{don't need to include all these resolutions, just those need to do suitable analysis/conclusions}}
+\TODO{don't necessary need to include all these resolutions, just those need to do suitable analysis/conclusions}.
+\TODO{*** might have to be smaller for cut cell grid due to small cell problem}}
 	\label{tab:gw-resolutions}
@@
 150	& 250	& 4 \\
-125	& 208.3 & 3.333 \\
+125	& 208.3 & 3.333 \TODO{***} \\
 100	& 166.7 & 2.667 \\
@@
 %
-	\caption{\TODO{}}
+	\caption{\TODO{This will eventually include a large panel showing BTF dz=50m (the best "reference" solution we have).
+Below that I intend to compare BTF/cut cell results at dz=50m, between gravity waves and thermal advection tests.
+These plots will zoom into a small domain covering the mountain peaks and a small part of the lee side.}}
 	\label{fig:gw-theta}
@@
 
+\begin{figure}
+	\centering
+	\includegraphics{../fig-gravityWaves-sampleLine/fig-gravityWaves-sampleLine.pdf}
+%
+	\caption{\TODO{Vertical sample through x=50km (quite far to the east of the mountain) in gravity waves tests.
+BTF results I have so far are in good agreement.
+The cut cell results are not.
+I am not yet sure if it will be worth taking sample lines through thermal advection tracer diffs because not all of them agree with the gravity waves results.
+This can, perhaps, be forgiven because the wind fields differ between the advection and gravity waves tests.}}
+	\label{fig:gw-sampleLine}
+\end{figure}
+
 \end{document}

### 2015-10-16 src/tf-cutcell-comparison/tf-cutcell-comparison.tex
update advection stats to reflect new advection fix (277963c)

@@
 
-The initial thermodynamic conditions have a temperature at $z=0$ of $\theta_0 = \SI{288}{\kelvin}$ and constant static stability with $N = \SI{0.01}{\per\second}$ everywhere, such that
+The initial thermodynamic conditions have constant static stability with $N = \SI{0.01}{\per\second}$ everywhere, such that
 \begin{align}
@@
 \end{align}
-for $z > 0$.
+where the temperature at $z=0$ is $\theta_0 = \SI{288}{\kelvin}$.
 Potential temperature values are prescribed at the inlet and upper boundary using equation~\ref{eqn:thermal-profile}, and a zero gradient boundary condition is applied at the outlet.
@@
 % TABLES
+\begin{table}
+	\caption{\TODO{mesh statistics.  BTF max/min cell area ratio is very close to 1 irrespective of resolution.}}
+	\label{tab:gw-meshes}
+%
+\centering
+\begin{tabular}{S S S S}
+\hline\hline
+
+{$\Delta z$ (\si{\meter})} & {cut cell max/min cell area ratio} & {BTF nonortho} & {cut cell nonortho} \\
+\hline
+300	& 4.325 & 10.2	& 7.3 \\
+250	& 5.422 & 10.8	& 20.9 \\
+200	& 25.36 & 10.8	& 31.9 \\
+150	& 10.03 & 11.0	& 29.4 \\
+125	& 10.11 & 11.1	& 24.37 \\
+100	& 36.27 & 11.1	& 32.2 \\
+75	& 52.89 & 11.2	& 34.2 \\
+50	& 75.25 & 11.2	& 31.7 \\
+\hline
+\end{tabular}
+\end{table}
+
 \begin{table*}
@@
                   &                     & max              & 1        & 1.00    & 0.984           & {---}          & 0.983       \\
-                  & Cubic upwind-biased & \(\ell_2\) error & 0        & 0.107   & 0.0146          & 0.00784        & 0.00784     \\
-                  &                     & min              & 0        & -0.0446 & -0.0106         & -0.000674      & -0.00674    \\
-                  &                     & max              & 1        & 0.925   & 0.982           & 0.983          & 0.983       \\
+                  & Cubic upwind-biased & \(\ell_2\) error & 0        & 0.112   & 0.0146          & 0.00784        & 0.00784     \\
+                  &                     & min              & 0        & -0.0464 & -0.0106         & -0.000674      & -0.00674    \\
+                  &                     & max              & 1        & 0.922   & 0.982           & 0.983          & 0.983       \\
 \hline
@@
 		  & Cubic upwind-biased & \(\ell_2\) error & 0        & 0.0209  & 0.162           & 0.181          & {---}        \\
-                  &                     & min              & 0        & -0.0110 & -0.0263         & -0.028         & {---}        \\
+                  &                     & min              & 0        & -0.0110 & -0.0263         & -0.0284        & {---}        \\
                   &                     & max              & 1        & 0.983   & 0.865           & 0.851          & {---}        \\

### 2015-10-19 src/tf-cutcell-comparison/tf-cutcell-comparison.tex
add feedback from Hilary, start cleaning up tables and figure captions

@@
 In addition, a new tracer advection test is formulated having a velocity field that is everywhere tangential to the terrain following coordinate surfaces.
-This new test presents a challenge to cut cells in situations with a large Froude number and large Scorer parameter.
-\TODO{discuss Froude number/Scorer parameter in body.}
+This new test presents a challenge to the advection scheme on cut cell grids.
 The results of the advection tests demonstrate that tracer accuracy depends upon alignment of the flow with the grid.
-As expected, the cut cell grid maintains greater accuracy in the test of stationary, stratified flow.
-In the gravity waves test, results on all grids are in good agreement with existing results from the literature.}
+As expected, the cut cell grid maintains greater accuracy for horizontal advection and in the test of stationary, stratified flow.
+In the gravity waves test, results on all grids are in good agreement with existing results from the literature, although terrain following velocity fields lead to errors on cut cell grids.}
 
@@
 Representing orography accurately in numerical weather prediction systems is necessary to model downslope winds and local precipitation.
-Orography also exerts strong non-local influences: from the latent heat release due to convection, by directly forcing planetary waves, and by the atmospheric response to form drag and gravity wave drag.
-There are two main approaches to representing orography on a grid: terrain following layers and cut cells.
-Another way of treating orography is the immersed (or embedded) boundary method \citep{simon2012}.
+Orography also exerts strong non-local influences: from the latent heat release due to convection, by directly forcing gravity waves and planetary waves, and by the atmospheric response to form drag and gravity wave drag.
+There are two main approaches to representing orography on a grid: terrain following layers and cut cells, with the immersed (or embedded) boundary method \citep{simon2012} being similar to a cut cell approach.
 All methods align cells in vertical columns.
@@
 Spurious winds seen in TF coordinates are not present with cut cells and errors do not increase with steeper terrain \citep{good2014}.
-A comparison of TF and cut cells using real initial data by \citet{steppeler2013} found that five-day forecasts of precipitation and wind over Asia in January 1989 were more accurate in the cut cell model.
+A comparison of TF and cut cells using real initial data by \citet{steppeler2013} found that five-day forecasts of precipitation and wind over Asia in January 1989 were more accurate in the cut cell model, although this result was dependent on using an old version of a model.
 
@@
 The OpenFOAM utility \texttt{snappyHexMesh} was used to create a grid that approximates the cut cell method.
-First, a custom utility modifies an orthogonal grid, moving points beneath the surface upwards so that they lie just beneath the surface, creating layers of thin cells.
-Second, the surface faces were taken from the BTF grid and \texttt{snappyHexMesh} was used to intersect the surface with the grid produced by the custom utility.
-This tool removes cells whose centres are below the surface, including the layers of thin cells produced in the first step.
+The utility intersects the surface faces from the BTF grid with a rectangular grid, removing cells whose centres lie below the surface.
 In addition, \texttt{snappyHexMesh} displaces boundary vertices so that they are `snapped' to the BTF surface \citep{openfoam2015}.
-It was found that, by using the custom utility to preprocess the orthogonal grid, \texttt{snappyHexMesh} performed fewer, smaller vertex displacements, resulting in grids that are closer to those produced by typical shaving methods, as described by \citet{adcroft1997}.
+The grid that \texttt{snappyHexMesh} accepts is generated by a custom utility so that \texttt{snappyHexMesh} performs fewer, smaller vertex displacements, resulting in grids that are closer to those produced by typical shaving methods, as described by \citet{adcroft1997}.
 
@@
 where \(f(\phi^{(n)}) = - \nabla \cdot (\mathbf{u} \phi^{(n)})\) at time level \(n\).
+This time-stepping scheme is used for consistency with the trapezoidal implicit scheme used for the fully-compressible model, described in section~\ref{sec:model}\ref{sec:euler-model}.
 To ensure that the discrete velocity field is non-divergent, velocities are prescribed at cell faces by differencing the streamfunction, \(\Psi(x, z)\), along the edges from \(\Psi\) stored at cell vertices.
@@
 The second model is a modified version of the linear advection model first used by \citet{schaer2002}.
-It uses terrain following coordinates and it is configured with leapfrog timestepping and a fourth-order centred difference scheme:
+It uses terrain following coordinates and it is configured with leapfrog timestepping and either second-order centred differences, or a fourth-order centred difference scheme given by:
 \begin{align}
@@
 Once again, velocity fields are prescribed using a streamfunction defined at cell vertices (referred to as double staggered grid points by \citet{schaer2002}).
-The original version of the code interpolated the geometric height, $z$, at doubly staggered points from values at adjacent half levels in order to calculate the streamfunction.
+The original version of the code interpolated the geometric height, $z$, at doubly staggered points from values at adjacent half levels in order to calculate the streamfunction, effectively smoothing the orography.
 The modified version used here directly calculates the height at vertices to enable comparisons with the finite volume model solutions.
@@
 \subsection{Fully-compressible model}
+\label{sec:euler-model}
 The third model is taken from \citet{weller-shahrokhi2014} which details a discretisation of the fully-compressible Euler equations, given by
@@
 The third test solves the Euler equations for a stably stratified atmosphere initially at rest, following \citet{klemp2011}.
-Finally, as specified by \citet{schaer2002}, a test of orographically-induced gravity waves is performed.
-\TODO{advection of thermal profile} No explicit diffusion is used in any of the tests.
+Fourth, as specified by \citet{schaer2002}, a test of orographically-induced gravity waves is performed.
+Finally, another tracer advection test is formulated that transports a stably stratified thermal profile in a terrain following velocity field.
+No explicit diffusion is used in any of the tests.
 
@@
 Analytic solutions can be found by setting the tracer centre such that $x_0 = ut$.
-
 Unlike \citet{schaer2002} who use periodic lateral boundaries, a fixed value of 0 is used at the inlet boundary and all other boundaries have zero gradient.
@@
 The test was executed on the BTF, SLEVE and cut cell grids, and on a regular grid with flat terrain using a centred linear scheme and the upwind-biased cubic scheme.
-Results were also obtained on BTF and SLEVE grids with the fourth order scheme from \citet{schaer2002} using a modified version of their code.
+Results were also obtained on BTF and SLEVE grids with the fourth order scheme from \citet{schaer2002} using the modified version of their code.
 
@@
 By \(t = \SI{10000}{\second}\), the tracer suffers from distortion on the BTF grid using the centred linear scheme and some artifacts remain about the mountain peak.
+The results are visually identical to those using the modified Sch\"{a}r model with centred linear differencing.
 The tracer has spread vertically due to increased numerical errors when the tracer is transported between layers.
-Distortions are reduced by using the fourth order scheme from \citet{schaer2002} (figure~\ref{fig:advection-tracer}c), but the computational mode is seen as a grid-scale oscillation that travels in the opposite direction to the wind (figure~\ref{fig:advection-tracer}e).
-The results from the centred linear and fourth order schemes are slightly worse than the respective results from \citet{schaer2002} (their figure 6a and figure 8) because of the difference in the discretisation of $\Psi$.
+Distortions are reduced by using the fourth order scheme from \citet{schaer2002} (figure~\ref{fig:advection-tracer}c), but dispersion errors remain with grid-scale oscillations that travel in the opposite direction to the wind (figure~\ref{fig:advection-tracer}e).
+The results from the centred linear and fourth order schemes are slightly worse than the respective results from \citet{schaer2002} (their figure 6a and figure 8) because of the difference in the discretisation of the streamfunction, $\Psi$.
 
@@
 
-The centred linear scheme is sensitive to changes in the terrain profile: when the domain width is reduced to \SI{300}{\kilo\meter}, the $\ell_2$ error on the BTF grid rises from \num{0.284} to \num{0.432}.
+The centred linear scheme is sensitive to grid-scale changes in the terrain profile: when the domain width is reduced to \SI{300}{\kilo\meter}, the $\ell_2$ error on the BTF grid rises from \num{0.284} to \num{0.432}.
 The same sensitivity is found using terrain following coordinates and centred differences in space and time, as implemented by \citet{schaer2002}.
@@
 It is likely that changes in the domain width affect the wave power spectrum of the discrete terrain profile.
-High-frequency oscillations persist in the centred linear scheme but are damped by the upwind-biased scheme.
-This is an area for further study.
+Grid-scale oscillations persist in the centred linear scheme but are damped by the upwind-biased scheme.
 
@@
 
-Using the centred linear scheme on the cut cell grid, numerical instability is generated by flow through small cells: when interpolating \(\phi\) onto faces, flux from a large cell into a small cell means that the downwind cell is given greater weighting.
+Using the centred linear scheme on the cut cell grid, numerical instability is generated by flow through small cells: when interpolating \(\phi\) onto faces, flux from a large cell into a small cell means that the downwind cell is given greater weighting, effectively making the scheme downwind-biased rather than upwind-biased.
 The maximum Courant number is less than one in all of the advection tests, hence it is not the source of this instability.
@@
 At the ground, fixed gradients are imposed by calculating the component of $\nabla \theta$ normal to each face using the vertical derivative of equation~\ref{eqn:thermal-profile}.
-For the Exner function of pressure, hydrostatic balance is prescribed on all boundaries.
+For the Exner function of pressure, hydrostatic balance is prescribed on top and bottom boundaries and the inlet and outlet are zero normal gradient.
 
@@
 With increasing resolutions up to $\Delta z = \SI{50}{\meter}$, the orography intersects more layers and becomes better resolved.
+\TODO{These grids are shown in figure x.}
 
@@
 Stripes of potential temperature anomalies are found in the lee of the mountain, and these errors are most severe on the cut cell grid.
-\TODO{say we want to find out what's causing them}
 
@@
 where the tracer density at $z = 0$, $\phi_0 = 288$, and the transform, $z^\star$, is given by rearranging equation~\ref{eqn:btf}.
-\TODO{is this clear enough?
-my point is I don't need to worry about an analytic solution where only /part/ of the tracer has been advected over the mountain range.
-the analytic solution at t=18000s is symmetric about x=0.}
 
@@
 Across all tests, a high degree of accuracy was achieved for all grids.
-Even on the highly-distorted BTF grid, which have previously been found to give poor results \citep{schaer2002,klemp2011,good2014}, errors were often small in the tests presented here.
+Even on the highly-distorted BTF grid, which has previously been found to give poor results \citep{schaer2002,klemp2011,good2014}, errors were often small in the tests presented here.
 In the first two tests, tracers were advected by horizontal and terrain following velocity fields.
@@
 % TABLES
-\begin{table}
-	\caption{\TODO{mesh statistics.  BTF max/min cell area ratio is very close to 1 irrespective of resolution.}}
-	\label{tab:gw-meshes}
-%
-\centering
-\begin{tabular}{S S S S}
-\hline\hline
-
-{$\Delta z$ (\si{\meter})} & {cut cell max/min cell area ratio} & {BTF nonortho} & {cut cell nonortho} \\
-\hline
-300	& 4.325 & 10.2	& 7.3 \\
-250	& 5.422 & 10.8	& 20.9 \\
-200	& 25.36 & 10.8	& 31.9 \\
-150	& 10.03 & 11.0	& 29.4 \\
-125	& 10.11 & 11.1	& 24.37 \\
-100	& 36.27 & 11.1	& 32.2 \\
-75	& 52.89 & 11.2	& 34.2 \\
-50	& 75.25 & 11.2	& 31.7 \\
-\hline
-\end{tabular}
-\end{table}
 
@@
 	\caption{Spatial and temporal resolutions used in the gravity waves test.
-The lowest resolution has the same parameters as the original test case specified by \citet{schaer2002}.
-At higher resolutions, the vertical resolution is chosen, and horizontal and temporal resolutions are calculated to preserve the same aspect ratio.
-\TODO{don't necessary need to include all these resolutions, just those need to do suitable analysis/conclusions}.
-\TODO{*** might have to be smaller for cut cell grid due to small cell problem}}
+The resolution of $\Delta z = \SI{300}{\meter}$ has the same parameters as the original test case specified by \citet{schaer2002}.
+At other resolutions, the vertical resolution is prescribed, and horizontal and temporal resolutions are calculated to preserve the same aspect ratio as the original test case.}
 	\label{tab:gw-resolutions}
@@
 \hline
+500	& 833.3 & 13.33 \\
 300	& 500	& 8 \\
@@
 150	& 250	& 4 \\
-125	& 208.3 & 3.333 \TODO{***} \\
+125	& 208.3 & 3.333 \\
 100	& 166.7 & 2.667 \\
@@
 50	& 83.33 & 1.333 \\
+\hline
+\end{tabular}
+\end{table}
+
+\begin{table}
+	\caption{\TODO{DOUBLE-CHECK THESE VALUES.}}
+	\label{tab:gw-meshes}
+%
+\centering
+\begin{tabular}{S S S S S}
+\hline\hline
+
+{$\Delta z$ (\si{\meter})} & \multicolumn{2}{c}{max/min cell area ratio} & \multicolumn{2}{c}{Nonorthogonality (\si{\degree})} \\
+ & {BTF} & {Cut cell} & {BTF} & {Cut cell} \\
+\hline
+500	& 1.01 &  999  & 999    & 999 \\
+300	& 1.01 &  4.33 & 10.2	& 7.3 \\
+250	& 1.01 &  5.42 & 10.8	& 20.9 \\
+200	& 1.01 &  25.4 & 10.8	& 31.9 \\
+150	& 1.01 &  10.0 & 11.0	& 29.4 \\
+125	& 1.01 &  10.1 & 11.1	& 24.37 \\
+100	& 1.01 &  36.3 & 11.1	& 32.2 \\
+75	& 1.01 &  52.9 & 11.2	& 34.2 \\
+50	& 1.01 &  75.3 & 11.2	& 31.7 \\
 \hline
@@
 The terrain surface, denoted by a thick grey line, is taken from the lower boundary of the BTF grid.
-The surface is intersected with an orthogonal grid shown by thin solid lines.
-\TODO{I think this example is clearer showing the difference between an orthogonal grid and the resulting cut cell grid.
-In fact, we modify the orthogonal grid with a custom utility before it is passed to snappyHexMesh}  Vertex $v$ is `snapped' to the surface, moving from the position marked by the open circle to that of the closed circle.
+The surface is intersected with a rectangular grid shown by thin solid lines.
+Vertex $v$ is `snapped' to the surface, moving from the position marked by the open circle to that of the closed circle.
 The resultant cut cell grid is shown by thick dashed lines.
@@
 %
-	\caption{\TODO{This will eventually include a large panel showing BTF dz=50m (the best "reference" solution we have).
-Below that I intend to compare BTF/cut cell results at dz=50m, between gravity waves and thermal advection tests.
-These plots will zoom into a small domain covering the mountain peaks and a small part of the lee side.}}
+	\caption{Differences in potential temperature between the start and end of the gravity waves test with $\Delta z = \SI{50}{\meter}$ \TODO{if we have the results in time!} showing (a) waves in potential temperature on the BTF grid.
+Enlargements of the same domain near the mountain on (b) the BTF grid, and (c) the cut cell grid \TODO{revealing errors near the ground in the lee of the mountain}.
+The gravity waves test results are compared with differences in tracer between the numeric and analytic solutions at $t = \SI{18000}{\second}$ on (d) the BTF grid, and (e) the cut cell grid, in which errors are seen which are similar to those in (c).}
 	\label{fig:gw-theta}

### 2015-10-19 src/tf-cutcell-comparison/tf-cutcell-comparison.tex
discuss mesh statistics

@@
 
+The simulation is integrated forward by 5 hours and the timestep, $\Delta t = 8 \Delta z / \SI{300}{\second}$, is chosen so that it scales linearly with spatial resolution and, following the original test specified by \citet{schaer2002}, $\Delta t = \SI{8}{\second}$ when $\Delta z = \SI{300}{\meter}$.
 Test results are compared between the BTF and cut cell grids at several resolutions.
@@
 
-The simulation is integrated forward by 5 hours and the timestep, $\Delta t = 8 \Delta z / \SI{300}{\second}$, is chosen so that it scales linearly with spatial resolution and, following the original test specified by \citet{schaer2002}, $\Delta t = \SI{8}{\second}$ when $\Delta z = \SI{300}{\meter}$.
+Some statistics of the grids are presented in table~\ref{tab:gw-meshes}.
+The ratio of minimum and maximum cell areas provides an indication of size of the smallest cut cells.
+As expected, there is almost no variation in cell sizes on the BTF grids.
+Small cells are generated on cut cell grids at resolutions finer than $\Delta z = \SI{300}{\meter}$ in which the terrain intersects grid layers.
+Nonorthogonality is measured by the angle between the vector connecting adjacent cell centres and the normal vector of the shared cell face, ranging from \SI{0}{\degree} on an orthogonal grid, to a maximum of \SI{90}{\degree}.
+On the BTF grid, nonorthogonality gradually increases from \TODO{\SI{999}{\degree}} to \SI{11.1}{\degree} as increasing horizontal resolution leads to steeper terrain.
+Nonorthogonality is higher on almost all cut cell grids.
 
@@
 They agree with the high resolution mass-conserving semi-implicit semi-Lagrangian solution from \citet{melvin2010}.
-
 The initial thermal profile is subtracted from the potential temperature field at the end of the integration to reveal the structure of thermal anomalies.
@@
 I am grateful to my co-supervisors John Methven and Terry Davies for their valuable input, and to Christoph Sch\"{a}r (ETH) for his assistance in reproducing his advection test results.
+The authors thank the two anonymous reviewers whose comments greatly improved this manuscript.
 I am thankful for the NERC studentship which helped fund this work.
@@
 Enlargements of the same domain near the mountain on (b) the BTF grid, and (c) the cut cell grid \TODO{revealing errors near the ground in the lee of the mountain}.
-The gravity waves test results are compared with differences in tracer between the numeric and analytic solutions at $t = \SI{18000}{\second}$ on (d) the BTF grid, and (e) the cut cell grid, in which errors are seen which are similar to those in (c).}
+The gravity waves test results are compared with thermal advection test results, showing differences in tracer density between the numeric and analytic solutions at $t = \SI{18000}{\second}$ on (d) the BTF grid, and (e) the cut cell grid, in which errors are seen which are similar to those in (c).
+Axes are in units of \si{\meter}.}
 	\label{fig:gw-theta}
@@
 %
-	\caption{\TODO{Vertical sample through x=50km (quite far to the east of the mountain) in gravity waves tests.
-BTF results I have so far are in good agreement.
-The cut cell results are not.
-I am not yet sure if it will be worth taking sample lines through thermal advection tracer diffs because not all of them agree with the gravity waves results.
-This can, perhaps, be forgiven because the wind fields differ between the advection and gravity waves tests.}}
+	\caption{Vertical profiles of potential temperature differences between the start and end of the gravity waves test on (a) the BTF grid, and (b) the cut cell grid.
+Results are compared with thermal advection tests results, showing differences in tracer density between the numeric and analytic solutions at $t = \SI{18000}{\second}$ on (c) the BTF grid, and (d) the cut cell grid.
+\TODO{a little bit of analysis: explain why BTF GW increases temperature with height; say that BTF results in good agreement in both GW and adv tests; errors in cut cell results in both tests in the lowest layers }}
 	\label{fig:gw-sampleLine}

### 2015-10-19 src/tf-cutcell-comparison/fig-gravityWaves-theta.tex
annotate gravityWave figures

--- 2015-10-02 src/tf-cutcell-comparison/fig-gravityWaves-theta.tex
@@
 \usepackage{showframe}
-\usepackage[paperwidth=6.4in,paperheight=8.3in,margin=0in]{geometry}
+\usepackage[paperwidth=6.2in,paperheight=5.6in,margin=0in]{geometry}
 \usepackage{mathptmx}%{times}
@@
 \begin{document}
+\TPMargin{3pt}
+\textblockcolour{white}
+\begin{textblock}{0.75}(1.4,0.5)
+	\normalsize
+	\centering
+	(a)
+\end{textblock}
+\begin{textblock}{0.75}(1.2,6.5)
+	\normalsize
+	\centering
+	(b)
+\end{textblock}
+\begin{textblock}{0.75}(9.05,6.5)
+	\normalsize
+	\centering
+	(c)
+\end{textblock}
+\begin{textblock}{0.75}(1.2,11.3)
+	\normalsize
+	\centering
+	(d)
+\end{textblock}
+\begin{textblock}{0.75}(9.05,11.3)
+	\normalsize
+	\centering
+	(e)
+\end{textblock}
 \centering
@@
 \includegraphics[width=3in]{../gravityWaves-cutCell-250m-300dz/18000/thetaDiffMountain.pdf}
-\vspace*{0.1in}
+\includegraphics[width=3in]{../thermalAdvection-btf-250m-300dz/18000/tracerDiffMountain.pdf}
+\includegraphics[width=3in]{../thermalAdvection-cutCell-250m-300dz/18000/tracerDiffMountain.pdf}
+
+\newpage
+
+250m GW
+
 \includegraphics[width=3in]{../gravityWaves-btf-250m-250dz/18000/thetaDiffMountain.pdf}
 \includegraphics[width=3in]{../gravityWaves-cutCell-250m-250dz/18000/thetaDiffMountain.pdf}
-\vspace*{0.1in}
+
+250m adv
+
+\includegraphics[width=3in]{../thermalAdvection-btf-250m-250dz/18000/tracerDiffMountain.pdf}
+\includegraphics[width=3in]{../thermalAdvection-cutCell-250m-250dz/18000/tracerDiffMountain.pdf}
+
+\newpage
+200m GW
+
 \includegraphics[width=3in]{../gravityWaves-btf-250m-200dz/18000/thetaDiffMountain.pdf}
 \includegraphics[width=3in]{../gravityWaves-cutCell-250m-200dz/18000/thetaDiffMountain.pdf}
-\vspace*{0.1in}
+
+200m adv
+
+\includegraphics[width=3in]{../thermalAdvection-btf-250m-200dz/18000/tracerDiffMountain.pdf}
+\includegraphics[width=3in]{../thermalAdvection-cutCell-250m-200dz/18000/tracerDiffMountain.pdf}
+
+\newpage
+
+150m GW
+
 \includegraphics[width=3in]{../gravityWaves-btf-250m-150dz/18000/thetaDiffMountain.pdf}
 \includegraphics[width=3in]{../gravityWaves-cutCell-250m-150dz/18000/thetaDiffMountain.pdf}
+
+150m adv
+
+\includegraphics[width=3in]{../thermalAdvection-btf-250m-150dz/18000/tracerDiffMountain.pdf}
+\includegraphics[width=3in]{../thermalAdvection-cutCell-250m-150dz/18000/tracerDiffMountain.pdf}
 \end{document}

### 2015-10-20 src/tf-cutcell-comparison/fig-gravityWaves-theta.tex
errors are smaller now so revise analysis in GWs

@@
 \usepackage{showframe}
-\usepackage[paperwidth=6.2in,paperheight=5.6in,margin=0in]{geometry}
+\usepackage[paperwidth=6.2in,paperheight=2.3in,margin=0in]{geometry}
 \usepackage{mathptmx}%{times}
@@
 \begin{document}
-\TPMargin{3pt}
-\textblockcolour{white}
-\begin{textblock}{0.75}(1.4,0.5)
-	\normalsize
-	\centering
-	(a)
-\end{textblock}
-\begin{textblock}{0.75}(1.2,6.5)
-	\normalsize
-	\centering
-	(b)
-\end{textblock}
-\begin{textblock}{0.75}(9.05,6.5)
-	\normalsize
-	\centering
-	(c)
-\end{textblock}
-\begin{textblock}{0.75}(1.2,11.3)
-	\normalsize
-	\centering
-	(d)
-\end{textblock}
-\begin{textblock}{0.75}(9.05,11.3)
-	\normalsize
-	\centering
-	(e)
-\end{textblock}
 \centering
@@
 \vspace*{0.1in}
+
+\newpage
+500m GW
+
+\includegraphics[width=3in]{../gravityWaves-btf-250m-500dz/18000/thetaDiffMountain.pdf}
+\includegraphics[width=3in]{../gravityWaves-cutCell-250m-500dz/18000/thetaDiffMountain.pdf}
+
+500m adv
+
+\includegraphics[width=3in]{../thermalAdvection-btf-250m-500dz/18000/tracerDiffMountain.pdf}
+\includegraphics[width=3in]{../thermalAdvection-cutCell-250m-500dz/18000/tracerDiffMountain.pdf}
+
+\newpage
+300m GW
+
 \includegraphics[width=3in]{../gravityWaves-btf-250m-300dz/18000/thetaDiffMountain.pdf}
 \includegraphics[width=3in]{../gravityWaves-cutCell-250m-300dz/18000/thetaDiffMountain.pdf}
+
+300m adv
+
 \includegraphics[width=3in]{../thermalAdvection-btf-250m-300dz/18000/tracerDiffMountain.pdf}
 \includegraphics[width=3in]{../thermalAdvection-cutCell-250m-300dz/18000/tracerDiffMountain.pdf}
-
-\newpage
-
-250m GW
-
-\includegraphics[width=3in]{../gravityWaves-btf-250m-250dz/18000/thetaDiffMountain.pdf}
-\includegraphics[width=3in]{../gravityWaves-cutCell-250m-250dz/18000/thetaDiffMountain.pdf}
-
-250m adv
-
-\includegraphics[width=3in]{../thermalAdvection-btf-250m-250dz/18000/tracerDiffMountain.pdf}
-\includegraphics[width=3in]{../thermalAdvection-cutCell-250m-250dz/18000/tracerDiffMountain.pdf}
 
@@
 \newpage
-
 150m GW
@@
 \includegraphics[width=3in]{../thermalAdvection-cutCell-250m-150dz/18000/tracerDiffMountain.pdf}
+
+\newpage
+
+125m GW
+
+\includegraphics[width=3in]{../gravityWaves-btf-250m-125dz/18000/thetaDiffMountain.pdf}
+\includegraphics[width=3in]{../gravityWaves-cutCell-250m-125dz/18000/thetaDiffMountain.pdf}
+
+125m adv
+
+\includegraphics[width=3in]{../thermalAdvection-btf-250m-125dz/18000/tracerDiffMountain.pdf}
+\includegraphics[width=3in]{../thermalAdvection-cutCell-250m-125dz/18000/tracerDiffMountain.pdf}
+
+\newpage
+
+100m GW
+
+\includegraphics[width=3in]{../gravityWaves-btf-250m-100dz/18000/thetaDiffMountain.pdf}
+\includegraphics[width=3in]{../gravityWaves-cutCell-250m-100dz/18000/thetaDiffMountain.pdf}
+
+100m adv
+
+\includegraphics[width=3in]{../thermalAdvection-btf-250m-100dz/18000/tracerDiffMountain.pdf}
+\includegraphics[width=3in]{../thermalAdvection-cutCell-250m-100dz/18000/tracerDiffMountain.pdf}
+
+\newpage
+
+50m adv
+
+\includegraphics[width=3in]{../thermalAdvection-btf-250m-50dz/18000/tracerDiffMountain.pdf}
+\includegraphics[width=3in]{../thermalAdvection-cutCell-250m-50dz/18000/tracerDiffMountain.pdf}
+
 \end{document}

### 2015-10-20 src/tf-cutcell-comparison/tf-cutcell-comparison.tex
errors are smaller now so revise analysis in GWs

@@
 Nonorthogonality is measured by the angle between the vector connecting adjacent cell centres and the normal vector of the shared cell face, ranging from \SI{0}{\degree} on an orthogonal grid, to a maximum of \SI{90}{\degree}.
-On the BTF grid, nonorthogonality gradually increases from \TODO{\SI{999}{\degree}} to \SI{11.1}{\degree} as increasing horizontal resolution leads to steeper terrain.
-Nonorthogonality is higher on almost all cut cell grids.
+On the BTF grid, maximum nonorthogonality gradually increases from \TODO{\SI{999}{\degree}} to \SI{11.1}{\degree} as increasing horizontal resolution leads to steeper terrain.
+Maximum nonorthogonality is higher on almost all cut cell grids.
+\TODO{so what?
+if we want to talk about nonortho we need a reason for doing so...}
 
@@
 The initial thermal profile is subtracted from the potential temperature field at the end of the integration to reveal the structure of thermal anomalies.
-Results are similar across all grid types and resolutions but, examining the anomalies in the lee of the mountain, errors are found in the lowest few layers.
-Errors are generated on the lee slopes of the mountain and are advected to the outlet boundary, forming striped anomalies.
-At resolution of $\Delta z > \SI{50}{\meter}$, the lowest layer is anomalously warm and the layer above is anomalously cold.
-As resolution increases, the anomalously cold layer first increases in amplitude, then the cold anomaly begins to span multiple layers.
-At $\Delta z = \SI{50}{\meter}$, the structure is reversed, having lower layers that are too cold and layers above that are too warm.
-At all resolutions, the amplitude and height of the errors are greater on the cut cell grid compared to the BTF grid.
+The anomalies on the BTF grid with $\Delta z = \SI{50}{\meter}$ is shown in figure~\ref{fig:gw-theta}a.
+A vertical profile is taken at $x = \SI{50}{\kilo\meter}$, marked by the dashed line in figure~\ref{fig:gw-theta}a, with results shown for the BTF grids in figure~\ref{fig:gw-sampleLine}a and on the cut cell grids in figure~\ref{fig:gw-sampleLine}b.
+On all grids, potential temperature increases with height due to the presence of a gravity wave at this location.
+Results are seen to converge on all grids, with two exceptions.
+First, the result on the cut cell grid with $\Delta z = \SI{500}{\meter}$ differs from results on all other grids, though at this coarse resolution the mountain is poorly resolved.
+Second, small errors are seen in the lowest layers on the cut cell grids.
 
 To summarize, results of the gravity waves test on all grids are in good agreement with the reference solution from \citet{melvin2010}.
-Stripes of potential temperature anomalies are found in the lee of the mountain, and these errors are most severe on the cut cell grid.
+The potential temperature field converges with increasing resolution, though errors are found in the lowest layers on the cut cell grids.
 
@@
 \begin{table}
-	\caption{\TODO{DOUBLE-CHECK THESE VALUES.}}
+	\caption{Cell area ratios and nonorthogonality of BTF and cut cell grids used in the gravity waves and thermal advection tests.
+Cell sizes are almost uniform on BTF grids, but for the cut cell grids the cell area ratio gives an indication of the smallest cell sizes.
+\TODO{describe nonorthogonality}  \TODO{DOUBLE-CHECK THESE VALUES.}}
 	\label{tab:gw-meshes}
@@
 
-{$\Delta z$ (\si{\meter})} & \multicolumn{2}{c}{max/min cell area ratio} & \multicolumn{2}{c}{Nonorthogonality (\si{\degree})} \\
+{$\Delta z$ (\si{\meter})} & \multicolumn{2}{c}{max/min cell area ratio} & \multicolumn{2}{c}{Maximum nonorthogonality (\si{\degree})} \\
  & {BTF} & {Cut cell} & {BTF} & {Cut cell} \\
@@
 %
-	\caption{Differences in potential temperature between the start and end of the gravity waves test with $\Delta z = \SI{50}{\meter}$ \TODO{if we have the results in time!} showing (a) waves in potential temperature on the BTF grid.
-Enlargements of the same domain near the mountain on (b) the BTF grid, and (c) the cut cell grid \TODO{revealing errors near the ground in the lee of the mountain}.
-The gravity waves test results are compared with thermal advection test results, showing differences in tracer density between the numeric and analytic solutions at $t = \SI{18000}{\second}$ on (d) the BTF grid, and (e) the cut cell grid, in which errors are seen which are similar to those in (c).
-Axes are in units of \si{\meter}.}
+	\caption{Differences in potential temperature between the start and end of the gravity waves test on the BTF grid with $\Delta z = \SI{50}{\meter}$ \TODO{if we have the results in time!}  Axes are in units of \si{\meter}.}
 	\label{fig:gw-theta}

### 2015-10-20 src/tf-cutcell-comparison/fig-thermalAdvection.tex
split thermalAdvection plot into separate figure

--- 2015-10-14 src/tf-cutcell-comparison/fig-gravityWaves-theta.tex
@@
 \usepackage{showframe}
-\usepackage[paperwidth=6.4in,paperheight=8.3in,margin=0in]{geometry}
+\usepackage[paperwidth=3.1in,paperheight=3.2in,margin=0in]{geometry}
 \usepackage{mathptmx}%{times}
@@
 \begin{document}
+\TPMargin{3pt}
+\textblockcolour{white}
+\begin{textblock}{0.75}(2.2,0.4)
+\normalsize
 \centering
-\includegraphics[width=6in]{../gravityWaves-btf-250m-300dz/18000/thetaDiffOverview.pdf}
+(a)
+\end{textblock}
+\begin{textblock}{0.75}(2.2,8.4)
+\normalsize
+\centering
+(b)
+\end{textblock}
+\centering
+\includegraphics[width=3in]{../thermalAdvection-btf-250m-50dz/18000/tracerDiffMountain.pdf} \\
 \vspace*{0.1in}
-\includegraphics[width=3in]{../gravityWaves-btf-250m-300dz/18000/thetaDiffMountain.pdf}
-\includegraphics[width=3in]{../gravityWaves-cutCell-250m-300dz/18000/thetaDiffMountain.pdf}
-300m
-\vspace*{0.1in}
-\includegraphics[width=3in]{../gravityWaves-btf-250m-250dz/18000/thetaDiffMountain.pdf}
-\includegraphics[width=3in]{../gravityWaves-cutCell-250m-250dz/18000/thetaDiffMountain.pdf}
-250m
-\vspace*{0.1in}
-\includegraphics[width=3in]{../gravityWaves-btf-250m-200dz/18000/thetaDiffMountain.pdf}
-\includegraphics[width=3in]{../gravityWaves-cutCell-250m-200dz/18000/thetaDiffMountain.pdf}
-200m
-\vspace*{0.1in}
-\includegraphics[width=3in]{../gravityWaves-btf-250m-150dz/18000/thetaDiffMountain.pdf}
-\includegraphics[width=3in]{../gravityWaves-cutCell-250m-150dz/18000/thetaDiffMountain.pdf}
-150m
-\vspace*{0.1in}
-%\includegraphics[width=3in]{../gravityWaves-btf-250m-100dz/18000/thetaDiffMountain.pdf}
-%\includegraphics[width=3in]{../gravityWaves-cutCell-250m-100dz/18000/thetaDiffMountain.pdf}
-%100m (OLD)
-%\vspace*{0.1in}
-%\includegraphics[width=3in]{../gravityWaves-btf-250m-75dz/18000/thetaDiffMountain.pdf}
-%\includegraphics[width=3in]{../gravityWaves-cutCell-250m-75dz/18000/thetaDiffMountain.pdf}
-%75m (OLD)
+\includegraphics[width=3in]{../thermalAdvection-cutCell-250m-50dz/18000/tracerDiffMountain.pdf}
 \end{document}

### 2015-10-21 src/tf-cutcell-comparison/fig-gravityWaves-meshes.tex
plot gravity waves cut cell meshes

--- 2015-05-24 src/tf-cutcell-comparison/fig-meshes.tex
@@
 \end{textblock}
-\includegraphics{../resting-mesh-btf/constant/mesh.pdf}
-\includegraphics{../resting-mesh-sleve/constant/mesh.pdf}
-\includegraphics{../resting-mesh-cutCell/constant/mesh.pdf}
+\includegraphics{../gravityWaves-mesh-cutCell-250m-300dz/constant/mesh.pdf}
+\includegraphics{../gravityWaves-mesh-cutCell-250m-250dz/constant/mesh.pdf}
+\includegraphics{../gravityWaves-mesh-cutCell-250m-200dz/constant/mesh.pdf}
 \end{document}

### 2015-10-21 src/tf-cutcell-comparison/fig-gravityWaves-theta.tex
plot gravity waves cut cell meshes

@@
 \documentclass{article}
-\usepackage{showframe}
-\usepackage[paperwidth=6.2in,paperheight=2.3in,margin=0in]{geometry}
+\usepackage[paperwidth=6.2in,paperheight=2.6in,margin=0in]{geometry}
 \usepackage{mathptmx}%{times}
@@
 \centering
-\includegraphics[width=6in]{../gravityWaves-btf-250m-300dz/18000/thetaDiffOverview.pdf}
-\vspace*{0.1in}
+\includegraphics[width=6in]{../gravityWaves-btf-250m-50dz/18000/thetaDiffOverview.pdf} \\
+\includegraphics[height=5in,angle=270]{thetaDiffOverview_theta_diff.eps}
 

### 2015-10-21 src/tf-cutcell-comparison/fig-thermalAdvection.tex
plot gravity waves cut cell meshes

@@
 \documentclass{article}
-\usepackage{showframe}
-\usepackage[paperwidth=3.1in,paperheight=3.2in,margin=0in]{geometry}
+\usepackage[paperwidth=3.1in,paperheight=3.5in,margin=0in]{geometry}
 \usepackage{mathptmx}%{times}
@@
 \end{textblock}
+\begin{textblock}{2}(9,0.4)
+\normalsize
+$\ell^2 = \num{5.86e-8}$
+\end{textblock}
 \begin{textblock}{0.75}(2.2,8.4)
@@
 \end{textblock}
+\begin{textblock}{2}(9,8.4)
+\normalsize
+$\ell^2 = \num{5.94e-6}$
+\end{textblock}
 \centering
@@
 \vspace*{0.1in}
-\includegraphics[width=3in]{../thermalAdvection-cutCell-250m-50dz/18000/tracerDiffMountain.pdf}
+\includegraphics[width=3in]{../thermalAdvection-cutCell-250m-50dz/18000/tracerDiffMountain.pdf} \\
+\includegraphics[height=3in,angle=270]{tracerDiffMountain_T_diff.eps}
 \end{document}

### 2015-10-21 src/tf-cutcell-comparison/tf-cutcell-comparison.tex
split thermalAdvection plot into separate figure
plot gravity waves cut cell meshes

@@
 With increasing resolutions up to $\Delta z = \SI{50}{\meter}$, the orography intersects more layers and becomes better resolved.
-\TODO{These grids are shown in figure x.}
+Three of the cut cell grids are shown in figure~\ref{fig:gw-meshes} at $\Delta z = \SI{300}{\meter}$, $\SI{250}{\meter}$ and $\SI{200}{\meter}$.
+Small cells are visible on the $\SI{200}{\meter}$ grid but, on the $\SI{250}{\meter}$ grid, the grid generator has merged the small cells with those in the layer above.
 
@@
 Nonorthogonality is measured by the angle between the vector connecting adjacent cell centres and the normal vector of the shared cell face, ranging from \SI{0}{\degree} on an orthogonal grid, to a maximum of \SI{90}{\degree}.
-On the BTF grid, maximum nonorthogonality gradually increases from \TODO{\SI{999}{\degree}} to \SI{11.1}{\degree} as increasing horizontal resolution leads to steeper terrain.
+On the BTF grid, maximum nonorthogonality gradually increases from \SI{9.2}{\degree} to \SI{11.2}{\degree} as increasing horizontal resolution leads to steeper terrain.
 Maximum nonorthogonality is higher on almost all cut cell grids.
@@
 To summarize, results of the gravity waves test on all grids are in good agreement with the reference solution from \citet{melvin2010}.
-The potential temperature field converges with increasing resolution, though errors are found in the lowest layers on the cut cell grids.
+The potential temperature field converges, though errors are found in the lowest layers on the cut cell grids.
 
@@
 
-\TODO{analyse results of test, draw conclusions}
+Enlargements of the error field near the mountain is shown in figure~\ref{fig:thermalAdvection}.
+\TODO{Errors are negligible on the BTF grids with a maximum error in tracer density of \num{123}.}
+Errors are also small on the cut cell grids at $\Delta z = \SI{500}{\meter}$ and $\Delta z = \SI{300}{\meter}$.
+However, at resolutions where the terrain intersects layers of the cut cell grids, advection errors are apparent around mountain peaks, with small cells having the largest errors.
+These errors are advected horizontally along the lee side, forming stripes.
+
+For comparison with the potential temperature anomalies in the gravity waves test, vertical profiles of tracer density error are taken at $x = \SI{50}{\kilo\meter}$.
+As seen in figure~\ref{fig:gw-theta}c, errors are negligible on the BTF grids, but figure~\ref{fig:gw-theta}d reveals significant errors in the lowest layers of the cut cell grids that were advected from the mountain peaks.
+The magnitude of error is about ten times larger than the potential temperature anomalies found on the cut cell grids in the gravity waves test seen in figure~\ref{fig:gw-theta}b.
+We offer two explanations for this difference: first, there are no restoring forces in the linear advection model and, second, the wind fields in the two tests are different.
+
+While the magnitude and structure of tracer error on the cut cell grids differs from potential temperature anomalies in the gravity waves test, results on the BTF grids are in close agreement in both tests but not on the cut cell grids.
+Therefore, it is likely that anomalies on the cut cell grids in the gravity waves test are primarily due to errors in the advection of potential temperature.
 
@@
 First, in the horizontal advection test, tracer accuracy on the cut cell grid was almost as good as accuracy on a regular grid with no mountain.
-Second, in the resting atmosphere test, spurious vertical velocities were two orders of magnitude smaller on the cut cell grid compared with the terrain folllowing grids.
+Second, in the resting atmosphere test, spurious vertical velocities were two orders of magnitude smaller on the cut cell grid compared with the terrain following grids.
 
-Conversely, in the terrain following advection test, errors were large on the SLEVE and cut cell grids where velocities were misaligned with the grids.
+Conversely, in the first terrain following tracer advection test, errors were large on the SLEVE and cut cell grids where velocities were misaligned with the grids.
+Errors were also large on the cut cell grids in the second terrain following tracer advection test.
+This result suggests that, in the gravity waves test, that potential temperature anomalies in the cut cell grids are primarily due to advection errors.
 
-\TODO{do we want to say more about how cubicUpwindCPCFit is designed to cope with small cells?
-this approach differs from other existing approaches.
+\TODO{the cubicUpwindCPCFit advection scheme offers an approach for treating small cells that differs from other existing approaches.
 the technique will be described in future publication.
-(although an implementation of the technique is already available online)}
+(although an implementation of the technique is already available online)  I haven't talked about this anywhere in the body of the paper yet...
+I need to!}
 
@@
 \hline
-500	& 1.01 &  999  & 999    & 999 \\
+500	& 1.01 &  1.68  & 9.2   & 4.1 \\
 300	& 1.01 &  4.33 & 10.2	& 7.3 \\
@@
 	\centering
+	\includegraphics{../fig-gravityWaves-meshes/fig-gravityWaves-meshes.pdf}
+	%
+	\caption{Cut cell grids used for the gravity waves and thermal advection tests at (a) $\Delta z = \SI{300}{\meter}$, (b) $\Delta z = \SI{250}{\meter}$, and (c) $\Delta z = \SI{200}{\meter}$.
+The mountain peak $h_0 = \SI{250}{\meter}$.
+At $\Delta z = \SI{250}{\meter}$, the grid creation process has merged small cells with the cells in the layer above but, at $\Delta z = \SI{200}{\meter}$, small cells have been retained.
+The full two dimensional grids are \SI{300}{\kilo\meter} wide and \SI{30}{\kilo\meter} high.
+Axes are in units of \si{\meter}.}
+	\label{fig:gw-meshes}
+\end{figure*}
+
+\begin{figure*}
+	\centering
 	\includegraphics[width=6in]{../fig-gravityWaves-theta/fig-gravityWaves-theta.pdf}
 %
-	\caption{Differences in potential temperature between the start and end of the gravity waves test on the BTF grid with $\Delta z = \SI{50}{\meter}$ \TODO{if we have the results in time!}  Axes are in units of \si{\meter}.}
+	\caption{Differences in potential temperature between the start and end of the gravity waves test on the BTF grid with $\Delta z = \SI{50}{\meter}$.
+Axes are in units of \si{\meter}.
+\TODO{draw a dashed line at x=50km if possible}}
 	\label{fig:gw-theta}
@@
 Results are compared with thermal advection tests results, showing differences in tracer density between the numeric and analytic solutions at $t = \SI{18000}{\second}$ on (c) the BTF grid, and (d) the cut cell grid.
-\TODO{a little bit of analysis: explain why BTF GW increases temperature with height; say that BTF results in good agreement in both GW and adv tests; errors in cut cell results in both tests in the lowest layers }}
+The results are convergent, with the exception of the cut cell grid at $\Delta z = \SI{500}{\meter}$, and errors are found in the lowest layers on the cut cell grids.}
 	\label{fig:gw-sampleLine}
+\end{figure*}
+
+\begin{figure}
+	\centering
+	\includegraphics{../fig-thermalAdvection/fig-thermalAdvection.pdf}
+%
+	\caption{Error in tracer density in the thermal advection test at $\Delta z = \SI{50}{\meter}$ on (a) the BTF grid, and (b) the cut cell grid.  \TODO{a sentence of analysis}}
+	\label{fig:thermalAdvection}
 \end{figure}

### 2015-10-21 src/tf-cutcell-comparison/tf-cutcell-comparison.tex
discuss small cell treatment by cubicUpwindCPCFit, tweaks to figures

@@
 
-\TODO{give multiple examples of cut cell grids used in subsequent test cases.  acknowledge that we can still get thin cells in some instances.}
-
 \section{Models}
@@
 All are operated in a two-dimensional $x$--$z$ slice configuration.
+
+The two finite volume models make use of the upwind-biased multidimensional cubic advection scheme from \citet{weller-shahrokhi2014} which is non-monotonic and not flux corrected.  
+The scheme uses a least squares fit of a multidimensional polynomial using an upwind-biased stencil.
+Weightings are iteratively adjusted in the least squares fit procedure to ensure that advection remains upwind-biased where the stencil includes small cut cells.
+\TODO{I can see reviewers saying that this isn't a sufficiently detailed explanation...?}
 
@@
 \end{align}
-where $\mathbf{u} = (u, w)$ is a prescribed velocity field and the tracer density, $\phi$, is interpolated onto cell faces using one of two schemes: first, the centred linear scheme which takes the average of the two neighbouring cell values; second, the upwind-biased multidimensional cubic scheme from \citet{weller-shahrokhi2014} which is non-monotonic and not flux corrected.
+where $\mathbf{u} = (u, w)$ is a prescribed velocity field and the tracer density, $\phi$, is interpolated onto cell faces using one of two schemes: first, the centred linear scheme which takes the average of the two neighbouring cell values; second, the upwind-biased cubic scheme.
 The time derivative is solved using a three-stage, second order Runge-Kutta scheme defined as:
@@
 
-\subsection{Fully-compressible model}
+\subsection{Finite volume fully-compressible model}
 \label{sec:euler-model}
@@
 Nonorthogonality is measured by the angle between the vector connecting adjacent cell centres and the normal vector of the shared cell face, ranging from \SI{0}{\degree} on an orthogonal grid, to a maximum of \SI{90}{\degree}.
+The maximum nonorthogonality is the largest value of any face.
 On the BTF grid, maximum nonorthogonality gradually increases from \SI{9.2}{\degree} to \SI{11.2}{\degree} as increasing horizontal resolution leads to steeper terrain.
-Maximum nonorthogonality is higher on almost all cut cell grids.
+Maximum nonorthogonality is higher on almost all cut cell grids at about \SI{30}{\degree}.
 \TODO{so what?
@@
 
-Enlargements of the error field near the mountain is shown in figure~\ref{fig:thermalAdvection}.
-\TODO{Errors are negligible on the BTF grids with a maximum error in tracer density of \num{123}.}
+Errors are small on the BTF grids with a maximum error in tracer density of \num{0.0376} at $\Delta z = \SI{500}{\meter}$ and errors are 10 to 100 times smaller on higher resolution BTF grids.
 Errors are also small on the cut cell grids at $\Delta z = \SI{500}{\meter}$ and $\Delta z = \SI{300}{\meter}$.
@@
 These errors are advected horizontally along the lee side, forming stripes.
+Enlargements of the error field near the mountain are shown in figure~\ref{fig:thermalAdvection} at $\Delta z = \SI{50}{\meter}$.
 
@@
 
-\TODO{the cubicUpwindCPCFit advection scheme offers an approach for treating small cells that differs from other existing approaches.
-the technique will be described in future publication.
-(although an implementation of the technique is already available online)  I haven't talked about this anywhere in the body of the paper yet...
-I need to!}
+The cubic upwind-biased advection scheme takes an approach for treating small cut cells that differs from other existing approaches by adjusting weightings to ensure that advection remains upwind-biased near small cells.
+The technique is implemented in OpenFOAM and will be described in a future publication.
 
@@
 Cell sizes are almost uniform on BTF grids, but for the cut cell grids the cell area ratio gives an indication of the smallest cell sizes.
-\TODO{describe nonorthogonality}  \TODO{DOUBLE-CHECK THESE VALUES.}}
+Nonorthogonality is the angle between the surface normal vector and the vector joining the two adjacent cell centres.}
 	\label{tab:gw-meshes}
@@
 	\caption{Differences in potential temperature between the start and end of the gravity waves test on the BTF grid with $\Delta z = \SI{50}{\meter}$.
-Axes are in units of \si{\meter}.
-\TODO{draw a dashed line at x=50km if possible}}
+The dashed line at $x = \SI{50}{\kilo\meter}$ marks the position of the vertical profile in figure~\ref{fig:gw-sampleLine}.
+Axes are in units of \si{\meter}.}
 	\label{fig:gw-theta}
@@
 %
-	\caption{Error in tracer density in the thermal advection test at $\Delta z = \SI{50}{\meter}$ on (a) the BTF grid, and (b) the cut cell grid.  \TODO{a sentence of analysis}}
+	\caption{Error in tracer density in the thermal advection test at a resolution of $\Delta z = \SI{50}{\meter}$ on (a) the BTF grid, and (b) the cut cell grid.
+Errors are negligible on the BTF grid, but on the cut cell grid errors are generated near mountainous terrain and are advected horizontally on the lee side.
+Axes are in units of \si{\meter}.}
 	\label{fig:thermalAdvection}

### 2015-10-22 src/tf-cutcell-comparison/model.tex
wording tweaks

@@
 To avoid time-splitting errors between the advection and the fast waves, the advection is time-stepped using a three-stage, second-order Runge-Kutta scheme.
-The advection terms of the momentum and $\theta$ equations, (\ref{eq:momentum}) and (\ref{eq:theta}) are discretised in flux form using the upwind-biased, multi-dimensional cubic scheme from \citet{weller-shahrokhi2014}.
+The advection terms of the momentum and $\theta$ equations, (\ref{eq:momentum}) and (\ref{eq:theta}), are discretised in flux form using the upwind-biased cubic scheme.

### 2015-10-22 src/tf-cutcell-comparison/tf-cutcell-comparison.tex
wording tweaks

@@
 A series of standard two-dimensional tests using idealised terrain are performed: tracer advection in a prescribed horizontal velocity field, a test starting from stably stratified initial conditions, and orographically induced gravity waves described by nonhydrostatic dynamics.
-In addition, a new tracer advection test is formulated having a velocity field that is everywhere tangential to the terrain following coordinate surfaces.
-This new test presents a challenge to the advection scheme on cut cell grids.
+In addition, two new tracer advection tests are formulated having a velocity field that is everywhere tangential to the terrain following coordinate surfaces.
+These new tests present a challenge to the advection scheme on cut cell grids.
 The results of the advection tests demonstrate that tracer accuracy depends upon alignment of the flow with the grid.
@@
 \maketitle
-\TODO{this is preliminary, but somewhere in the body I want to highlight the tension between horizontal pressure gradient/advection errors and the choice of grid:}
-
-\TODO{Every term in the discretised equations of motion is a potential source of numerical error.
- Errors in calculating the horizontal pressure gradient generate spurious circulations and much effort has been put into improving this calculation.
-This is often achieved by making the grid more orthogonal: flattening terrain following coordinate surfaces [citations] or using a cut cell approach [citations].
- TF citations: mahrer1984, schaer2002, klemp2011
- cut cell citations: good2014, yamazaki2015 puts this clearly but not yet published
- In a test of a stably stratified atmosphere initially at rest, we find that spurious velocities are no more than \SI{0.1}{\meter\per\second} on any grid.
-[TODO: and this is much better than klemp2011]
- Advection terms also contribute to numerical error.
-We show that advection is most accurate when the flow is aligned with the grid layers.
-Near the surface, flow tends to follow the terrain which results in larger errors when more orthogonal grids are used.
- In a test of orographically induced gravity waves, we find that the advection of potential temperature generates grid-scale errors on the cut cell grid.}
 
@@
 
-This study uses a modification of the fully-compressible model from \citet{weller-shahrokhi2014} to enable a like-for-like comparison between terrain following and cut cell grids for idealised, two-dimensional test cases from the literature.
+This study uses a modified version of the fully-compressible model from \citet{weller-shahrokhi2014} to enable a like-for-like comparison between terrain following and cut cell grids for idealised, two-dimensional test cases from the literature.
 Section 2 presents the formulation of the terrain following and cut cell grids used in the experiments that follow.
 In section 3 we give the governing equations, outline the model from \citet{weller-shahrokhi2014} and describe the modification which improves stability for long timesteps in the presence of steep orography.
-Section 4 analyses the results from two tracer advection tests, a test of a stably stratified atmosphere initially at rest, and orographically induced gravity waves.
+Section 4 analyses the results from three tracer advection tests, a test of a stably stratified atmosphere initially at rest, and orographically induced gravity waves.
 Concluding remarks are made in section 5.
@@
 
-The OpenFOAM utility \texttt{snappyHexMesh} was used to create a grid that approximates the cut cell method.
+The OpenFOAM utility \texttt{snappyHexMesh} was used to create grids that approximate the cut cell method.
 The utility intersects the surface faces from the BTF grid with a rectangular grid, removing cells whose centres lie below the surface.
 In addition, \texttt{snappyHexMesh} displaces boundary vertices so that they are `snapped' to the BTF surface \citep{openfoam2015}.
-The grid that \texttt{snappyHexMesh} accepts is generated by a custom utility so that \texttt{snappyHexMesh} performs fewer, smaller vertex displacements, resulting in grids that are closer to those produced by typical shaving methods, as described by \citet{adcroft1997}.
+The grid that \texttt{snappyHexMesh} accepts as input is generated by a custom utility so that \texttt{snappyHexMesh} performs fewer, smaller vertex displacements, resulting in grids that are closer to those produced by typical shaving methods, as described by \citet{adcroft1997}.
 
@@
 Once again, velocity fields are prescribed using a streamfunction defined at cell vertices (referred to as double staggered grid points by \citet{schaer2002}).
-The original version of the code interpolated the geometric height, $z$, at doubly staggered points from values at adjacent half levels in order to calculate the streamfunction, effectively smoothing the orography.
+The original version of the code effectively smoothed the orography, interpolating the geometric height, $z$, at doubly staggered points from values at adjacent half levels in order to calculate the streamfunction.
 The modified version used here directly calculates the height at vertices to enable comparisons with the finite volume model solutions.
@@
 A series of two-dimensional tests are performed over idealised orography.
-For each test, results on the BTF, SLEVE and cut cell grid are compared.
+For each test, results on terrain following and cut cell grids are compared.
 The first test from \citet{schaer2002} advects a tracer in a horizontal velocity field.
@@
 The spatial domain, mountain profile, initial tracer profile and discretisation are the same as those in the horizontal tracer advection test.
-The velocity field is defined using a streamfunction, $\Psi$, so that the continuous velocity field is non-divergent and follows the BTF coordinate surfaces given by equation~\ref{eqn:btf} such that
+The velocity field is defined using a streamfunction, $\Psi$, so that the discrete velocity field is non-divergent and follows the BTF coordinate surfaces given by equation~\ref{eqn:btf} such that
 \begin{equation}
@@
 In summary, spurious velocities in the resting atmosphere test were similar on both terrain following grids, with lower errors compared to those from \citet{klemp2011}.
-The maximum vertical velocity was significantly decreased on the cut cell grid, so we conclude that non-orthogonality, or lack of alignment of the grid with sufaces of constant gravitational potential are a significant cause of numerical error in this test.
+The maximum vertical velocity was significantly decreased on the cut cell grid, so we conclude that non-orthogonality, or lack of alignment of the grid with surfaces of constant gravitational potential are a significant cause of numerical error in this test.
 
@@
 Maximum nonorthogonality is higher on almost all cut cell grids at about \SI{30}{\degree}.
-\TODO{so what?
-if we want to talk about nonortho we need a reason for doing so...}
+\TODO{what do the nonortho stats mean for us?
+we might expected larger HPG errors with greater nonortho (and this was indeed the case looking at max nonortho stats in the resting atmosphere, although I've not discussed those in the paper as it stands)}
 
@@
 The initial thermal profile is subtracted from the potential temperature field at the end of the integration to reveal the structure of thermal anomalies.
-The anomalies on the BTF grid with $\Delta z = \SI{50}{\meter}$ is shown in figure~\ref{fig:gw-theta}a.
-A vertical profile is taken at $x = \SI{50}{\kilo\meter}$, marked by the dashed line in figure~\ref{fig:gw-theta}a, with results shown for the BTF grids in figure~\ref{fig:gw-sampleLine}a and on the cut cell grids in figure~\ref{fig:gw-sampleLine}b.
+The anomalies on the BTF grid with $\Delta z = \SI{50}{\meter}$ is shown in figure~\ref{fig:gw-theta}.
+A vertical profile is taken at $x = \SI{50}{\kilo\meter}$, marked by the dashed line in figure~\ref{fig:gw-theta}, with results shown for the BTF grids in figure~\ref{fig:gw-sampleLine}a and on the cut cell grids in figure~\ref{fig:gw-sampleLine}b.
 On all grids, potential temperature increases with height due to the presence of a gravity wave at this location.
@@
 \subsection{Terrain following advection of thermal profile}
-The potential temperature errors in the gravity waves test varied between terrain following and cut cell grids of the same resolution.
+The potential temperature anomalies in the gravity waves test varied between terrain following and cut cell grids of the same resolution.
 This variation may be due to differences in the wind fields between grids, or errors in the advection of potential temperature, amongst other possible causes.
@@
 where \(H_\mathrm{TF} = \SI{20}{\kilo\meter}\) is the level at which the terrain following layers become flat; the domain height is \(\SI{30}{\kilo\meter}\).
-For $z \leq H$, the $u$ and $w$ components of velocity are given by equation~\ref{eqn:uw-btf}, but $h(x)$ has the same form as equation~\ref{eqn:resting:mountain}, hence the derivative is:
+For $z \leq H_\mathrm{TF}$, the $u$ and $w$ components of velocity are given by equation~\ref{eqn:uw-btf}, but $h(x)$ has the same form as equation~\ref{eqn:resting:mountain}, hence the derivative is:
 \begin{align}
@@
 \end{align}
-For $z > H$, $u = u_0$ and $w = 0$.
+For $z > H_\mathrm{TF}$, $u = u_0$ and $w = 0$.
 
@@
 
+The tests we have presented demonstrate the trade-off between errors in calculating pressure gradient and advection terms: pressure gradient errors can be reduced by using a cut cell grid but, when flow follows the terrain, advection errors are reduced using a terrain following grid.
+
 The cubic upwind-biased advection scheme takes an approach for treating small cut cells that differs from other existing approaches by adjusting weightings to ensure that advection remains upwind-biased near small cells.
-The technique is implemented in OpenFOAM and will be described in a future publication.
+The technique is already implemented in OpenFOAM and will be described in a future publication.
 
@@
 \centering
+\footnotesize
 \begin{tabular}{S S S}
@@
 Cell sizes are almost uniform on BTF grids, but for the cut cell grids the cell area ratio gives an indication of the smallest cell sizes.
-Nonorthogonality is the angle between the surface normal vector and the vector joining the two adjacent cell centres.}
+Nonorthogonality is the angle between the surface normal vector and the vector joining the two adjacent cell centres.
+Except at coarse resolutions, cut cell grids have greater maximum nonorthogonality than BTF grids.}
 	\label{tab:gw-meshes}
@@
 \centering
+\footnotesize
 \begin{tabular}{S S S S S}

### 2015-10-23 src/tf-cutcell-comparison/tf-cutcell-comparison.tex
add more description of advection scheme, minor tweaks from HW feedback

@@
 These new tests present a challenge to the advection scheme on cut cell grids.
-The results of the advection tests demonstrate that tracer accuracy depends upon alignment of the flow with the grid.
+The results of the advection tests demonstrate that tracer accuracy depends primarily upon alignment of the flow with the grid rather than the grid orthogonality.
 As expected, the cut cell grid maintains greater accuracy for horizontal advection and in the test of stationary, stratified flow.
-In the gravity waves test, results on all grids are in good agreement with existing results from the literature, although terrain following velocity fields lead to errors on cut cell grids.}
+In the gravity waves test, results on all grids are in good agreement with existing results from the literature, although terrain following velocity fields lead to errors on cut cell grids.
+Due to semi-implicit timestepping and an upwind-biased, explicit advection scheme, there were no timestep restrictions associated with small cut cells.}
 
@@
 All methods align cells in vertical columns.
-Because most models are designed for a particular type of grid, existing studies of cut cell solutions have compared results with terrain following grid solutions implemented within different models, for example \citet{good2014}.
+Most models are designed for a particular type of grid, and the study by \citet{good2014} compared cut cell results with terrain following solutions implemented within different models.
 Instead, this study uses a single model to enable a like-for-like comparison between solutions using terrain following and cut cell grids.
@@
 This gave them the flexibility to design more accurate horizontal pressure gradient discretizations using more appropriate stencils.
+The technique proposed by \citet{weller-shahrokhi2014} involved calculating pressure gradients in the direction aligned with the grid, thus ensuring curl-free pressure gradients and improving accuracy.
 
@@
 
-Several studies have shown examples where cut cells produce more accurate results when compared to TF coordinates.
+Some studies have shown examples where cut cells produce more accurate results when compared to TF coordinates.
 Spurious winds seen in TF coordinates are not present with cut cells and errors do not increase with steeper terrain \citep{good2014}.
@@
 
-The OpenFOAM utility \texttt{snappyHexMesh} was used to create grids that approximate the cut cell method.
-The utility intersects the surface faces from the BTF grid with a rectangular grid, removing cells whose centres lie below the surface.
-In addition, \texttt{snappyHexMesh} displaces boundary vertices so that they are `snapped' to the BTF surface \citep{openfoam2015}.
-The grid that \texttt{snappyHexMesh} accepts as input is generated by a custom utility so that \texttt{snappyHexMesh} performs fewer, smaller vertex displacements, resulting in grids that are closer to those produced by typical shaving methods, as described by \citet{adcroft1997}.
+Cut cell grids were created by starting from a uniform grid, moving all cell vertices lying underneath the orography up to the surface.
+Then, the OpenFOAM utility \texttt{snappyHexMesh} was used to remove cells whose centres lay below the surface.
+In addition, the utility displaces boundary vertices so that they are `snapped' to the surface \citep{openfoam2015}.  An example of the resulting cut cell grid is shown in figure~\ref{fig:grid}c.
 
-There are two details of grid construction which mean that resulting cut cell grids can differ slightly from a typical cut cell grid.
+There are two details of grid construction which mean that resulting cut cell grids can differ slightly from a typical cut cell grid as described by \citet{adcroft1997}.
 First, when \texttt{snappyHexMesh} moves vertices onto the terrain surface, some points are moved horizontally.
@@
 An example of these effects is shown in figure~\ref{fig:snappyHexMesh} in which vertex $v$ has been moved vertically and horizontally to avoid creating a small cell.
-The resultant mesh, shown by thick dashed lines, contains triangular and quadrangular cells.
+The resultant grid, shown by thick dashed lines, contains triangular and quadrangular cells.
 With a typical shaving method, cell $c$ would be pentagonal, as shown by the dotted line.
 However, \texttt{snappyHexMesh} creates a quadrilateral cell.
+
+Some thin cells are generated but, unlike most cut cell grids, cells are typically made smaller in height but their width is unaltered.
+A grid that has these thin cells can be seen in figure~\ref{fig:gw-meshes}c.
+This technique has the advantage that cells are not shortened in the direction of flow and so there should be no additional constraints on the advective Courant number.
 
@@
 The two finite volume models make use of the upwind-biased multidimensional cubic advection scheme from \citet{weller-shahrokhi2014} which is non-monotonic and not flux corrected.  
-The scheme uses a least squares fit of a multidimensional polynomial using an upwind-biased stencil.
-Weightings are iteratively adjusted in the least squares fit procedure to ensure that advection remains upwind-biased where the stencil includes small cut cells.
-\TODO{I can see reviewers saying that this isn't a sufficiently detailed explanation...?}
+The scheme uses a least-squares approach to fit a multidimensional polynomial over an upwind-biased stencil which contains more cells than the number of polynomial coefficients.
+This fit is used to interpolate cell values onto face values for discretisation of the advection term using Guass's divergence theorem.
+Following \cite{lashley2002} and \cite{weller2009},  the two cells either side of the face we are interpolating onto are weighted in the least squares fit so that the fit goes nearly exactly through these cell centres but does not go exactly through the other points.
+This method worked well when used for terrain following meshes by \citet{weller-shahrokhi2014} but can be unstable in the presence of very small cut cells.
+This is because the least squares fit can generate a larger interpolation weight for the downwind cell than the upwind cell.
+In order to overcome this problem, wherever a large downwind cell interpolation weight is calculated by the least-squares fit, the weighting of the upwind cell is increased for the least-squares fitting and the fit is re-calculated.
+This procedure is repeated until the interpolation weight of the upwind cell is greater than the interpolation weight of the downwind cell.
+More details of this approach and a study of its behaviour is the subject of a future publication.
 
@@
 Analytic solutions can be found by setting the tracer centre such that $x_0 = ut$.
-Unlike \citet{schaer2002} who use periodic lateral boundaries, a fixed value of 0 is used at the inlet boundary and all other boundaries have zero gradient.
-Furthermore, a second order Runge-Kutta timestepping scheme is used here instead of the leapfrog scheme used by \citet{schaer2002}.
 Tests are integrated forward in time for \SI{10000}{\second} with a timestep of \(\Delta t = \SI{25}{\second}\).
@@
 
-The initial thermodynamic conditions are in discrete hydrostatic balance, having a reference potential temperature of $\theta(z = 0) = \SI{288}{\kelvin}$ and constant stability with Brunt-V\"ais\"al\"a frequency $N = \SI{0.01}{\per\second}$ everywhere, except for a more stable layer of $N = \SI{0.02}{\per\second}$ between $\SI{2}{\kilo\meter} \leq z \leq \SI{3}{\kilo\meter}$.
+The initial thermodynamic conditions are in discrete hydrostatic balance in the vertical direction, having a reference potential temperature of $\theta(z = 0) = \SI{288}{\kelvin}$ and constant stability with Brunt-V\"ais\"al\"a frequency $N = \SI{0.01}{\per\second}$ everywhere, except for a more stable layer of $N = \SI{0.02}{\per\second}$ between $\SI{2}{\kilo\meter} \leq z \leq \SI{3}{\kilo\meter}$.
 The damping function, \(\mu\), is set to zero.
@@
 The results presented in figure~\ref{fig:resting}, which use a curl-free pressure gradient, have maximum spurious values of $w$ of \SI{0.33}{\meter\per\second} on the BTF grid, compared with a maximum of \(\sim \SI{7}{\meter\per\second}\) found by \citet{klemp2011} using their improved horizontal pressure gradient formulation.
-The results in figure~\ref{fig:resting} have the same maximum errors as \citet{weller-shahrokhi2014} but, due to the more stable split into implicitly and explicitly treated terms, the errors decay over time due to the dissipative nature of the advection scheme.
+The results in figure~\ref{fig:resting} have the same maximum errors as \citet{weller-shahrokhi2014} but, due to the more stable split into implicitly and explicitly treated terms (described in the appendix), the errors decay over time due to the dissipative nature of the advection scheme.
 
@@
 \citet{good2014} found the maximum vertical velocity in their cut cell model was \SI{1e-12}{\meter\per\second}, which is better than any result obtained here.
-It is worth noting that the model by \citet{weller-shahrokhi2014} stores values at the geometric centre of cut cells.
-However, in the model used by \citet{good2014}, cell centres are in the centre of the uncut cell, resulting in the centre of some cut cells being below the ground (S.-J.
+It is worth noting that our model stores values at the geometric centre of cut cells, whereas the model used by \citet{good2014} has cell centres at the centre of the uncut cell, resulting in the centre of some cut cells being below the ground (S.-J.
 Lock 2014, personal communication).
@@
 
-Some statistics of the grids are presented in table~\ref{tab:gw-meshes}.
-The ratio of minimum and maximum cell areas provides an indication of size of the smallest cut cells.
+The ratio of minimum and maximum cell areas in the various grids is shown in table~\ref{tab:gw-meshes}, providing an indication of size of the smallest cut cells.
 As expected, there is almost no variation in cell sizes on the BTF grids.
 Small cells are generated on cut cell grids at resolutions finer than $\Delta z = \SI{300}{\meter}$ in which the terrain intersects grid layers.
-Nonorthogonality is measured by the angle between the vector connecting adjacent cell centres and the normal vector of the shared cell face, ranging from \SI{0}{\degree} on an orthogonal grid, to a maximum of \SI{90}{\degree}.
-The maximum nonorthogonality is the largest value of any face.
-On the BTF grid, maximum nonorthogonality gradually increases from \SI{9.2}{\degree} to \SI{11.2}{\degree} as increasing horizontal resolution leads to steeper terrain.
-Maximum nonorthogonality is higher on almost all cut cell grids at about \SI{30}{\degree}.
-\TODO{what do the nonortho stats mean for us?
-we might expected larger HPG errors with greater nonortho (and this was indeed the case looking at max nonortho stats in the resting atmosphere, although I've not discussed those in the paper as it stands)}
 
@@
 A vertical profile is taken at $x = \SI{50}{\kilo\meter}$, marked by the dashed line in figure~\ref{fig:gw-theta}, with results shown for the BTF grids in figure~\ref{fig:gw-sampleLine}a and on the cut cell grids in figure~\ref{fig:gw-sampleLine}b.
-On all grids, potential temperature increases with height due to the presence of a gravity wave at this location.
-Results are seen to converge on all grids, with two exceptions.
-First, the result on the cut cell grid with $\Delta z = \SI{500}{\meter}$ differs from results on all other grids, though at this coarse resolution the mountain is poorly resolved.
-Second, small errors are seen in the lowest layers on the cut cell grids.
+On all grids, potential temperature differences increase with height in the lowest \SI{1200}{\meter} at $x = \SI{50}{\kilo\meter}$, in agreement with the results seen in figure~\ref{fig:gw-theta}.
+Results are seen to converge on all grids, with the exception of small errors in the lowest layers on the cut cell grids.
 
@@
 The potential temperature field converges, though errors are found in the lowest layers on the cut cell grids.
+The source of the errors in the cut cell grids will be investigated further with an advection test in the following subsection.
 
@@
 
-Errors are small on the BTF grids with a maximum error in tracer density of \num{0.0376} at $\Delta z = \SI{500}{\meter}$ and errors are 10 to 100 times smaller on higher resolution BTF grids.
-Errors are also small on the cut cell grids at $\Delta z = \SI{500}{\meter}$ and $\Delta z = \SI{300}{\meter}$.
-However, at resolutions where the terrain intersects layers of the cut cell grids, advection errors are apparent around mountain peaks, with small cells having the largest errors.
-These errors are advected horizontally along the lee side, forming stripes.
-Enlargements of the error field near the mountain are shown in figure~\ref{fig:thermalAdvection} at $\Delta z = \SI{50}{\meter}$.
+Enlargements of the error field near the mountain are shown in figure~\ref{fig:thermalAdvection} at $\Delta z = \SI{50}{\meter}$ with contours in tracer density overlayed.
+Errors are too small to be visible on the BTF grid with a maximum error of \num{4.26e-4}.
+However, on the cut cell grid, the maximum error is \num{0.302}.
+Advection errors are apparent around mountainous terrain, with small cells having the largest errors.
+These errors are advected horizontally along the lee slope, forming stripes.
+The same error structure is present on all cut cell grids in which the terrain intersects the grid layers.
 
@@
 While the magnitude and structure of tracer error on the cut cell grids differs from potential temperature anomalies in the gravity waves test, results on the BTF grids are in close agreement in both tests but not on the cut cell grids.
-Therefore, it is likely that anomalies on the cut cell grids in the gravity waves test are primarily due to errors in the advection of potential temperature.
+Therefore, it is likely that anomalies on the cut cell grids in the gravity waves test are primarily due to errors in the advection of potential temperature through cut cells.
 
@@
 The cubic upwind-biased advection scheme takes an approach for treating small cut cells that differs from other existing approaches by adjusting weightings to ensure that advection remains upwind-biased near small cells.
-The technique is already implemented in OpenFOAM and will be described in a future publication.
+The technique is already implemented in OpenFOAM and will be described in greater detail a future publication.
+Combined with semi-implicit timestepping and a grid generator that preserves cell length in the direction of the flow, small cells did not impose additional timestep constraints.
 
@@
 \begin{table}
-	\caption{Cell area ratios and nonorthogonality of BTF and cut cell grids used in the gravity waves and thermal advection tests.
-Cell sizes are almost uniform on BTF grids, but for the cut cell grids the cell area ratio gives an indication of the smallest cell sizes.
-Nonorthogonality is the angle between the surface normal vector and the vector joining the two adjacent cell centres.
-Except at coarse resolutions, cut cell grids have greater maximum nonorthogonality than BTF grids.}
+	\caption{Cell area ratios of BTF and cut cell grids used in the gravity waves and thermal advection tests.
+Cell sizes are almost uniform on BTF grids, but for the cut cell grids the cell area ratio gives an indication of the smallest cell sizes.}
 	\label{tab:gw-meshes}
@@
 \footnotesize
-\begin{tabular}{S S S S S}
+\begin{tabular}{S S S}
 \hline\hline
 
-{$\Delta z$ (\si{\meter})} & \multicolumn{2}{c}{max/min cell area ratio} & \multicolumn{2}{c}{Maximum nonorthogonality (\si{\degree})} \\
- & {BTF} & {Cut cell} & {BTF} & {Cut cell} \\
+{$\Delta z$ (\si{\meter})} & \multicolumn{2}{c}{max/min cell area ratio} \\
+ & {BTF} & {Cut cell} \\
 \hline
-500	& 1.01 &  1.68  & 9.2   & 4.1 \\
-300	& 1.01 &  4.33 & 10.2	& 7.3 \\
-250	& 1.01 &  5.42 & 10.8	& 20.9 \\
-200	& 1.01 &  25.4 & 10.8	& 31.9 \\
-150	& 1.01 &  10.0 & 11.0	& 29.4 \\
-125	& 1.01 &  10.1 & 11.1	& 24.37 \\
-100	& 1.01 &  36.3 & 11.1	& 32.2 \\
-75	& 1.01 &  52.9 & 11.2	& 34.2 \\
-50	& 1.01 &  75.3 & 11.2	& 31.7 \\
+500	& 1.01 &  1.68 \\
+300	& 1.01 &  4.33 \\
+250	& 1.01 &  5.42 \\
+200	& 1.01 &  25.4 \\
+150	& 1.01 &  10.0 \\
+125	& 1.01 &  10.1 \\
+100	& 1.01 &  36.3 \\
+75	& 1.01 &  52.9 \\
+50	& 1.01 &  75.3 \\
 \hline
@@
 Results are compared with thermal advection tests results, showing differences in tracer density between the numeric and analytic solutions at $t = \SI{18000}{\second}$ on (c) the BTF grid, and (d) the cut cell grid.
-The results are convergent, with the exception of the cut cell grid at $\Delta z = \SI{500}{\meter}$, and errors are found in the lowest layers on the cut cell grids.}
+The results are convergent, with the exception of the cut cell grid at $\Delta z = \SI{500}{\meter}$, and errors are found in the lowest layers on the cut cell grids.
+\TODO{add GW cut cell dz=50m once the test has completed}}
 	\label{fig:gw-sampleLine}
@@
 Errors are negligible on the BTF grid, but on the cut cell grid errors are generated near mountainous terrain and are advected horizontally on the lee side.
+Contours in tracer density are overlayed.
 Axes are in units of \si{\meter}.}

### 2015-10-23 src/tf-cutcell-comparison/tf-cutcell-comparison.tex
parenthesise equation refs

@@
 The spatial domain, mountain profile, initial tracer profile and discretisation are the same as those in the horizontal tracer advection test.
-The velocity field is defined using a streamfunction, $\Psi$, so that the discrete velocity field is non-divergent and follows the BTF coordinate surfaces given by equation~\ref{eqn:btf} such that
+The velocity field is defined using a streamfunction, $\Psi$, so that the discrete velocity field is non-divergent and follows the BTF coordinate surfaces given by equation~(\ref{eqn:btf}) such that
 \begin{equation}
@@
 Following \citet{melvin2010}, the domain is \SI{300}{\kilo\meter} wide and \SI{30}{\kilo\meter} high.
-The mountain profile has the same form as equation~\ref{eqn:resting:mountain}, but the gravity waves tests have a mountain height of $h_0 = \SI{250}{\meter}$.
+The mountain profile has the same form as equation~(\ref{eqn:resting:mountain}), but the gravity waves tests have a mountain height of $h_0 = \SI{250}{\meter}$.
 As in the resting atmosphere test, $a = \SI{5}{\kilo\meter}$ is the mountain half-width and $\lambda = \SI{4}{\kilo\meter}$ is the wavelength.
@@
 where the temperature at $z=0$ is $\theta_0 = \SI{288}{\kelvin}$.
-Potential temperature values are prescribed at the inlet and upper boundary using equation~\ref{eqn:thermal-profile}, and a zero gradient boundary condition is applied at the outlet.
-At the ground, fixed gradients are imposed by calculating the component of $\nabla \theta$ normal to each face using the vertical derivative of equation~\ref{eqn:thermal-profile}.
+Potential temperature values are prescribed at the inlet and upper boundary using equation~(\ref{eqn:thermal-profile}), and a zero gradient boundary condition is applied at the outlet.
+At the ground, fixed gradients are imposed by calculating the component of $\nabla \theta$ normal to each face using the vertical derivative of equation~(\ref{eqn:thermal-profile}).
 For the Exner function of pressure, hydrostatic balance is prescribed on top and bottom boundaries and the inlet and outlet are zero normal gradient.
@@
 where \(H_\mathrm{TF} = \SI{20}{\kilo\meter}\) is the level at which the terrain following layers become flat; the domain height is \(\SI{30}{\kilo\meter}\).
-For $z \leq H_\mathrm{TF}$, the $u$ and $w$ components of velocity are given by equation~\ref{eqn:uw-btf}, but $h(x)$ has the same form as equation~\ref{eqn:resting:mountain}, hence the derivative is:
+For $z \leq H_\mathrm{TF}$, the $u$ and $w$ components of velocity are given by equation~(\ref{eqn:uw-btf}), but $h(x)$ has the same form as equation~(\ref{eqn:resting:mountain}), hence the derivative is:
 \begin{align}
@@
 \end{align}
-where the tracer density at $z = 0$, $\phi_0 = 288$, and the transform, $z^\star$, is given by rearranging equation~\ref{eqn:btf}.
+where the tracer density at $z = 0$, $\phi_0 = 288$, and the transform, $z^\star$, is given by rearranging equation~(\ref{eqn:btf}).
 
@@
 \begin{table*}
-	\caption{Minimum and maximum tracer magnitudes and \(\ell_2\) error norms (defined by equation~\ref{eqn:l2-error}) at \(t = \SI{10000}{\second}\) in the horizontal and terrain following tracer advection tests using centred linear and cubic upwind-biased schemes.
+	\caption{Minimum and maximum tracer magnitudes and \(\ell_2\) error norms, defined by equation~(\ref{eqn:l2-error}), at \(t = \SI{10000}{\second}\) in the horizontal and terrain following tracer advection tests using centred linear and cubic upwind-biased schemes.
 For the horizontal advection test, \(\ell_2\) error norms, minimum and maximum values are given for the fourth order scheme using the modified code from \citet{schaer2002}.}

### 2015-10-26 src/tf-cutcell-comparison/responses.tex
collate responses into single document

\documentclass{article}
\usepackage{url}
\usepackage{natbib}
\usepackage{enumitem}
\newlist{comment}{enumerate}{1}
\setlist[comment,1]{label={\arabic{commenti}.},leftmargin=*,resume=comment}

\begin{document}
\section*{Review \#1}
\begin{quotation}
In this paper, the authors compare the behavior of idealized
test-case results for a terrain-following vertical coordinate and a
cut-cell coordinate, where both grid structures are configured within
the same non hydrostatic finite-volume model. The 2-D test cases
considered are tracer advection within a fixed wind field, a resting
atmosphere simulation with terrain, and a mountain wave simulation
following \citet{schaer2002}. The  results in testing this model
with terrain-following coordinates for the resting atmosphere and
mountain wave cases have already been well documented by \citet{weller-shahrokhi2014}. The application of this model in the cut-cell
configuration appears to be the new and interesting advancement, but
I feel further description and testing of this cut-cell
implementation is needed to demonstrate that it is a viable
alternative to terrain following coordinates.
\end{quotation}
We are not trying to prove that cut cells are a viable alternative. We are trying to do a fair, clean comparison between cut cells and terrain following on a range of test cases.

In the revised version, we present further description of the cut cell grids and of the numerics necessary to run on a cut cell grid with small cells. We also present results of an additional test case in comparison to the previous draft.

\begin{quotation}
\noindent
\textbf{Specific comments}

The tracer advection test is not really a test of the dynamic core
with terrain-following or cut-cell coordinates per se. Rather, it is
a test of the accuracy of advection in a fixed wind field that is
either parallel to the coordinate surfaces (zero covariant normal
component of wind) or non-parallel with a varying covariant
component. Schar's (2002) original test case clearly showed that the
advection numerics are more accurate as the cross-coordinate flow is
reduced. The authors' results for uniform flow reproduce this
behavior. Specifying a wind field that is everywhere parallel to the
terrain-following coordinate surfaces again just confirms this
behavior, with the terrain-following configuration (now having zero
covariant flow) being more accurate. I don't see where much new has
been learned from this test.
\end{quotation}
The terrain following configuration is designed to find out whether
error is primarily due to a non-orthogonal grid or misalignment of the
flow.  Our results demonstrate that the latter is the primary source of
error.  The manuscript has been revised to make this clearer.

The terrain following velocity test case is proposed as a challenge for those developing cut cell models. Advection schemes must be accurate to advect tracers using terrain following winds on cut cell grids.

We have also now developed this test case further to advect stratified temperature in order to explain the discrepancies in the gravity wave test case.

\begin{quotation}
Verifying that the cut-cell implementation produces much smaller 
perturbations than the terrain-following version is a relevant
sanity check. As the authors' point out, defining variables in the
centers of uncut cells essentially removes the perturbations to
machine roundoff.
\end{quotation}
It is Good et al. 2014 that use centres of uncut cells.  We use geometric cell centres.  This has been clarified in the revised manuscript.

\begin{quotation}
Thus the small vertical variations along a coordinate surface in the 
present implementation should (and do) produce much smaller 
perturbations than in the terrain-following version (which, except
for a minor change in the implicit formulation, has already been
documented in Weller and Shahrokhi (2014)). However, this is not a
challenging test case for a cut-cell coordinate model.
\end{quotation}
We are not specifically trying to challenge the cut cell model. We are trying to do a clean comparison between cut cells and terrain following. This test case is an essential part of such a comparison. It is also of relevance that errors considerably larger than round-off are produced by the model on the cut cell grid because pressure gradients are not exactly horizontal between the cut and uncut cells because we are using the geometric cell centre of the cut cells.

\begin{quotation}
The mountain wave test case represents a realistic flow that can 
challenge the numerics of the cut-cell implementation. These
challenges are known to arise, for example, when the terrain cuts
through horizontal coordinate surfaces and when cells are present
that have a very small volume. In the test case shown, with a
vertical grid spacing of 300 m and a maximum terrain height of 250 m,
the terrain does not reach the first grid level, so the cut-cell
numerics are not fundamentally different from those in the
terrain-following formulation and should (and do) produce similar
results. The potential temperature anomalies at low levels in the lee
of the mountain do suggest, however, that there is already some
distortion caused by the cut-cell numerics. The authors attribute
these inaccuracies to presence of the Lorenz computational mode. As
evidence, they refer to a "zig-zag" structure in the vertical
potential temperature profile in figure 5, which is not very apparent
to me. It's also quite possible that the slightly warmer air at level
1 and cooler air at levels 2 and 3 are produced in the vicinity of
the cut cells above the terrain and swept downstream. In any event,
the terrain-following version also uses Lorenz staggering and does
not exhibit this anomalous behavior.
\end{quotation}
Having analysed test results at increasing vertical resolution (as
suggested below), we agree that errors are primarily due to inaccuracies
in the advection of theta.  An additional advection test has been
performed which provides evidence for this hypothesis.  Errors on cut
cell grids have been reduced since the submission of the previous manuscript due to improvements to the advection scheme and due to using fixed gradient rather than fixed values boundary conditions.

The numerics are identical when using cut cell or terrain following. It is the different grids leading to different truncation errors.

\begin{quotation}
As mentioned above, I believe this work would have stronger
scientific merit if it were refocused, with more emphasis on the
construction and discriminative testing of the cut-cell numerics, and
less emphasis on cases that provide little challenge for these
numerics. To begin, it would be helpful to expand the description of
the grid construction presently given on lines 127-137. A number of
questions immediately come to  mind, for example: How are the BTF
grid surfaces actually used to modify the cut-cell grid. Why do grid
points need to be moved horizontally,
\end{quotation}
This section has been expanded to further explain and justify how cut
cell grids are generated.  An illustrative example has been included in
a new figure demonstrating the movement of grid points and the lack of
pentagonal cells.

\begin{quotation}
and where are the cell centers used for the numerics actually
located.
\end{quotation}
Clarified in section 2 that values are stored at cell centroids.

\begin{quotation}
What are the most problematic cells that might arise with this
technique, and is there anything that limits the minimum cell size?
Perhaps illustrating the cut-cell grid structures for several types
of terrain situations would be helpful. In testing the cut-cell 
numerics, the mountain wave case provides a realistic flow with
known solutions. If its desired to stay with the same terrain height,
then evaluating the sensitivity of results to the vertical grid size
would be informative. Setting the nominal dz to 200, 250, and 300 m
would span a range of grid configurations relative to the terrain,
from nonintersecting for dz=300m, to just intersecting the terrain
crest for dz=250 m, to more fully intersecting the terrain for
dz=200m. Setting dz=125 m could also be interesting as it would
presumably correspond to the same relative grid cells as shown it
figure 1c. One could also consider a higher terrain (say 500 m) with
similar relative grid spacings. I believe such tests could be quite
illuminating in evaluating the viability of this interesting cut-cell
implementation.
\end{quotation}
Further tests have been performed at dz=500m, 250m, 200m, 150m, 125m,
100m, 75m and 50m, keeping the mountain height, $h_0$=250m constant.  The
same grids have also been used in the thermal advection test.  A figure
has been added showing the 300m, 250m and 200m cut cell grids. Agreed, these new tests are illuminating.

\section*{Review \#2}
\begin{quotation}
\noindent
\textbf{Summary}

\noindent The paper uses a non-hydrostatic 2D (x-z) slice model to compare two terrain-following (TF) vertical coordinates (the BTF grid by \citet{galchen-somerville1975} and the SLEVE coordinate by \citet{schaer2002}) to the `cut cell' method via idealized test cases.
This reveals the accuracy of the two modelling paradigms for tracer advection experiments with prescribed flow fields, a steady-state test in the presence of topography and nonlinear flows with orographically-induced gravity waves.  The new aspects of the manuscript are that the TF and cut cell methods are compared in an identical model framework, and that a new advection test is proposed that prescribes the flow along sloping TF coordinates.  This reveals the differences between TF and cut cell methods more objectively.
Another theme of the paper is that the comparison also involves different numerical methods.

In general, the results are interesting.  It is found that the accuracy of the methods strongly depends on the alignment of the flow field with the model levels, and that the cut cell method can trigger a computational mode that is present with the vertical Lorenz staggering.
In addition, it is interesting that the curl-free formulation of the non-hydrostatic model significantly improves the results of the steady-state test case.
Even when using the TF coordinates this mimetic property diminishes the spurious vertical velocities that have been documented in various other publications for TF coordinates.

There are a couple of weaknesses in the paper that need to be addressed:
\begin{enumerate}
\item A sign error in the derivation of the new advection test case in section 4b.
	Maybe it is just a typo in the manuscript, but if this wrong formulation has indeed been used
	in the example results, they all need to be rerun and corrected (see the detailed information below).
\item Inconsistent sign convention for the stream function and the derived velocities.  The sign is switched in sections 4a and 4b, which needs to be unified.
\item Model description in section 3: Since this paper not only compares the orography treatment but also the accuracy of some numerical methods it needs to provide more information about these numerical variants.  E.g. there is no information about the `centered linear' finite-volume scheme, not even a reference.  Is the linear scheme 2nd-order accurate and the cubic scheme 4th-order accurate?  In addition, it is unclear what method the nonlinear steady-state test at rest and the gravity wave test used (the cubic option?).  Include a short description of the linear scheme, the cubic scheme and the fourth-order scheme from \cite{schaer2002} since the manuscript emphasises these aspects in combination with the orography treatment.

The description of the Hodge dual operator is incomprehensible in its current form and does not seem to be crucial for the assessment of the orography analysis.  I suggest moving this description into an appendix and extending it somewhat to make it self-explanatory.  Currently, it is very confusing to read about a semi-implicit formulation, and see the explicit Runge-Kutta time stepping algorithm in the next paragraph.
\end{enumerate}
\end{quotation}
New `Model' section added, each model is described separately.  Hodge dual operator moved to appendix.

\begin{quotation}
\noindent
\textbf{Detailed comments}

\begin{comment}
\item Line 21: The Scorer parameter and Froude number are mentioned in the abstract, but never again in the manuscript.  Either remove this sentence from the abstract, or add the missing discussion why larger Scorer parameters and Froude numbers are a challenge for the cut cell method.  This is not obvious.
\end{comment}
\end{quotation}
Removed.

\begin{quotation}
\begin{comment}
\item Line 33: another way of treating orography are `immersed boundary methods' or `embedded boundary methods'.  Please add this information including a reference.
\end{comment}
\end{quotation}
Done, cited \citet{simon2012}.

\begin{quotation}
\begin{comment}
\item Line 33, typo: `\ldots approaches to respresenting \ldots'
\end{comment}
\end{quotation}
Fixed.

\begin{quotation}
\begin{comment}
\item Line 53/54: do you mean interpolations to z-levels in order to compute the pressure gradient?
\end{comment}
\end{quotation}
Yes, rephrased this sentence to say so.

\begin{quotation}
\begin{comment}
\item Line 58/59: conflicting information: you highlight the variable spacing and then the fact that cell sizes remain almost constant
\end{comment}
\end{quotation}
Removed the latter statement, since cell sizes are not constant even with constant layer spacing, especially over steep terrain.

\begin{quotation}
\begin{comment}
\item Line 67: provide some brief information about the `several approaches'.  A since sentence does not justify a paragraph.
\end{comment}
\end{quotation}
Added description of each approach.

\begin{quotation}
\begin{comment}
\item Line 80: add `\ldots artificially weak in the Eta model'.
\end{comment}
\end{quotation}
Done.

\begin{quotation}
\begin{comment}
\item Line 102: be more explicit: `\ldots of the Cartesian coordinate surface at the model level with transformed height $z^\star$'.  Spell out explicitly that $z$ varies between $h$ and $H$ and that $z^\star$ varies between 0 and $H$.
\end{comment}
\end{quotation}
Done.

\begin{quotation}
\begin{comment}
\item Line 106--111: This manuscript does not discuss or use pressure-based vertical coordinates at all, and the information about sigma and HTF is irrelevant.  Furthermore, the sigma definition is only correct for models with zero pressure at the model top.  Remove these lines.
\end{comment}
\end{quotation}
Removed discussion of sigma and HTF.

\begin{quotation}
\begin{comment}
\item Line 124: It is not clear what you mean by `unstructured grids'.  The grids seem to be very structured.  Clarify.
\end{comment}
\end{quotation}
Changed to `non-orthogonal grids'.  We hope this clarifies that the issue is the treatment of non-orthogonality: metric terms or the grid.

\begin{quotation}
\begin{comment}
\item Section 3: It would be helpful to point out up front that you use the model in a 2D x-z slice configuration without Coriolis forces.
\end{comment}
\end{quotation}
Added.

\begin{quotation}
\begin{comment}
\item Line 142: typo `\ldots temperature, \ldots'
\end{comment}
\end{quotation}
Fixed.

\begin{quotation}
\begin{comment}
\item Line 145: There is no `Lorenz C grid staggering', the model uses the C-grid staggering in the horizontal and the Lorenz grid in the vertical.
\end{comment}
\end{quotation}
Amended.

\begin{quotation}
\begin{comment}
\item Lines 145--160: This information is disconnected from the whole paper.  Expand the explanations to make them self-explanatory and move the paragraph into an appendix.  Clarify that this treatment is only needed for nonlinear flows and that the advection uses an explicit time-stepping scheme.
\end{comment}
\end{quotation}
Description expanded and moved to the appendix.

\begin{quotation}
\begin{comment}
\item Eq. (7): The tracer symbol $\phi$ is undefined from a physical viewpoints.  Do you imply the tracer density?
\end{comment}
\end{quotation}
Yes, now says, `\ldots where the tracer density, $\phi$, \ldots'.

\begin{quotation}
Later, in Eqs. (8) and (12) symbol $\varphi$ or a mix of $\phi$ and $\varphi$ are used.  Unify the presentation.
\end{quotation}
$\varphi$ removed, now using $\phi$ everywhere.

\begin{quotation}
\begin{comment}
\item Section 3: Add information about the different numerical options that are assessed in combination with the orography treatment.
\end{comment}
\end{quotation}
Included clearer description of models, including advection schemes.

\begin{quotation}
\begin{comment}
\item Line 189: the term 2a needs be in parenthesis
\end{comment}
\end{quotation}
Done.

\begin{quotation}
\begin{comment}
\item Line 194: be more precise: `\ldots constant wind above $z_2$ \ldots'
\end{comment}
\end{quotation}
Done.

\begin{quotation}
\begin{comment}
\item Line 198 and Eq. (16): Two different sign conventions for the stream function are used with flipped signs that either lead to a negative (line 198) or positive (Eq. (16)) sign in the definition of the zonal velocity $u$.  Unify the presentation.
\end{comment}
\end{quotation}
We now follow \citet{schaer2002} and use $u=-\partial \Psi/\partial z$ everywhere.  Signs corrected as appropriate to ensure horizontal velocities are positive.

\begin{quotation}
\begin{comment}
\item Line 199: rephrase `A tracer with density $\phi$ is \ldots'
\end{comment}
\end{quotation}
Done.

\begin{quotation}
Why do you highlight a tracer `anomaly' in lines 202, 203 and 204?  Is there a background tracer distribution?  If not, rephrase the `anomaly'.
\end{quotation}
There is no background tracer distribution.  Changed `anomaly' to `tracer'.

\begin{quotation}
\begin{comment}
\item Line 211: The information about different numerical schemes is very sudden here without enough information.  Add a description and references for all numerical options in section 2.
\end{comment}
\end{quotation}
Moved to section 3, more details of schemes added.

\begin{quotation}
\begin{comment}
\item Line 223: typo: artifact
\end{comment}
\end{quotation}
Done.

\begin{quotation}
\begin{comment}
\item Eq. 14: It is highly unusual to use a superscript (instead of a subscript) for the error norm
\end{comment}
\end{quotation}
Changed to subscript everywhere.

\begin{quotation}
\begin{comment}
\item Line 247: The result that the increase of the model domain from 300 to 301km leads to a 50\% increase in the $\ell_2$ error seems to be highly questionable.  This needs to be double-checked and explained in greater detail.
\end{comment}
\end{quotation}
Verified in two models, more detail included in the revised manuscript.

\begin{quotation}
\begin{comment}
\item Eq. (16): see comment 19 concerning the different sign conventions and unify the presentation.  In addition, the formulation for $w$ is wrong and has a sign error.  The vertical velocity $w$ needs to be multiplied with ($-1$) to correct it in its current form.  In case you used the wrong formulation for the results, they all need to be rerun.
\end{comment}
\end{quotation}
Sign error in $w$ equation fixed, but the code was already correct\footnote{\url{https://github.com/hertzsprung/AtmosFOAM/blob/master/src/orography/BtfVelocityProfile.C#L25}}.

\begin{quotation}
\begin{comment}
\item Line 281: `\ldots Courant number \ldots'
\end{comment}
\end{quotation}
Done.

\begin{quotation}
\begin{comment}
\item Section 3d: what is the grid spacing for this test case?
\end{comment}
\end{quotation}
The original test case was $\Delta x = 500m$, $\Delta z = 300m$.  The other reviewer suggested additional tests at higher resolutions.

\begin{quotation}
\begin{comment}
\item Line 344: The physical unit of $\mu$ is missing.
\end{comment}
\end{quotation}
$\mu$ is dimensionless.  It is now included in the momentum eqn 4a and noted that it is only used in the gravity waves test.  We also note that $\mu = 0$ for the resting atmosphere test in section 4c.

\begin{quotation}
\begin{comment}
\item Line 414: `\ldots Charney--Phillips \ldots'
\end{comment}
\end{quotation}
Fixed.

\begin{quotation}
\begin{comment}
\item line 414: I assume that `eta' need to be capitalized here, correct?
\end{comment}
\end{quotation}
Yes, it is now `Meso-Eta'.

\begin{quotation}
\begin{comment}
\item Line 428: reference is incomplete
\end{comment}
\end{quotation}
Now includes type, number and address.

\begin{quotation}
\begin{comment}
\item Line 443: `\ldots Eta model \ldots'
\end{comment}
\end{quotation}
Fixed.

\begin{quotation}
\begin{comment}
\item Line 466: `\ldots Cartesian \ldots'
\end{comment}
\end{quotation}
Fixed.

\begin{quotation}
\begin{comment}
\item Table 1: instead of showing these strange numbers like 13.6, -623 and 3480 it would be helpful to say `unstable'.
\end{comment}
\end{quotation}
Done.

\begin{quotation}
\begin{comment}
\item Figure 3: Replot this figure.  The lines need to be thicker, e.g. the SLEVE line is almost invisible.
\end{comment}
\end{quotation}
Done, and made the SLEVE colour darker to improve contrast.

\begin{quotation}
\begin{comment}
\item Figure 5: The BTF $\theta$ line needs to be thicker.
\end{comment}
\end{quotation}
Figure 5 now replaced by figure 7.

\bibliographystyle{ametsoc2014}
\bibliography{references}
\end{document}

### 2015-10-28 src/tf-cutcell-comparison/tf-cutcell-comparison.tex
remove plots that aren't needed, add GW cut cell 50dz result

@@
 \bibpunct{(}{)}{;}{a}{}{,}
-\newcommand{\TODO}[1]{\textcolor{purple}{TODO: \emph{#1}}}
 
@@
 Results are compared with thermal advection tests results, showing differences in tracer density between the numeric and analytic solutions at $t = \SI{18000}{\second}$ on (c) the BTF grid, and (d) the cut cell grid.
-The results are convergent, with the exception of the cut cell grid at $\Delta z = \SI{500}{\meter}$, and errors are found in the lowest layers on the cut cell grids.
-\TODO{add GW cut cell dz=50m once the test has completed}}
+The results are convergent, except for errors found in the lowest layers on the cut cell grids.}
 	\label{fig:gw-sampleLine}

### 2015-10-29 src/tf-cutcell-comparison/tf-cutcell-comparison.tex
minor tweaks from Hilary's feedback

@@
 \end{align}
-with \(s_1\) and \(s_2\) are the scale heights of large-scale and small-scale terrain respectively.
+and \(s_1\) and \(s_2\) are the scale heights of large-scale and small-scale terrain respectively.
 The exponent \(n\) was introduced by \citet{leuenberger2010} in order to increase cell thickness in the layers nearest the ground, allowing longer timesteps.
@@
 Because the velocity field is non-divergent, the flow accelerates over mountain ridges and the tracer travels \SI{1577.4}{\meter} further compared to advection in the purely horizontal velocity field.
-Tracer height is unchanged downwind of the mountains because advection is along the terrain following coordinate surface.
+Tracer height is unchanged downwind of the mountains because advection is parallel to the surface.
 
@@
 The cubic upwind-biased advection scheme takes an approach for treating small cut cells that differs from other existing approaches by adjusting weightings to ensure that advection remains upwind-biased near small cells.
-The technique is already implemented in OpenFOAM and will be described in greater detail a future publication.
+The implementation of this technique in OpenFOAM is available at \url{https://github.com/hertzsprung/AtmosFOAM/tree/shaw-weller-2015-mwr} and will be described in greater detail a future publication.
 Combined with semi-implicit timestepping and a grid generator that preserves cell length in the direction of the flow, small cells did not impose additional timestep constraints.

### 2015-10-29 src/tf-cutcell-comparison/tf-cutcell-comparison.tex
parenthesise more eqn refs, update l2 error norms and fix a few points of grammar

@@
 Here we describe the formulation of the terrain following grids and the method of cut cell grid construction.
-The techniques presented are used to define the grids for the experiments in the subsequent section.
+The techniques presented are used to define the grids for the experiments in section~\ref{sec:results}.
 
@@
 The results of the cubic upwind-biased scheme on TF and regular grids are comparable with those for the fourth-order centred scheme from \citet{schaer2002}.
-Error is largest on the BTF grid with \(\ell_2 = \num{0.107}\) but is significantly reduced on the SLEVE grid with \(\ell_2 = \num{0.0146}\).
+Error is largest on the BTF grid with \(\ell_2 = \num{0.112}\) but is significantly reduced on the SLEVE grid with \(\ell_2 = \num{0.0146}\).
 The error is approximately halved by changing from the SLEVE grid to the cut cell grid.
@@
 The same sensitivity is found using terrain following coordinates and centred differences in space and time, as implemented by \citet{schaer2002}.
-The upwind-biased cubic scheme is less sensitive to the same change, with the $\ell_2$ error decreasing from \num{0.107} to \num{0.104}.
+The upwind-biased cubic scheme is less sensitive to the same change, with the $\ell_2$ error decreasing from \num{0.112} to \num{0.105}.
 It is likely that changes in the domain width affect the wave power spectrum of the discrete terrain profile.
@@
 Because the velocity field is non-divergent, the flow accelerates over mountain ridges and the tracer travels \SI{1577.4}{\meter} further compared to advection in the purely horizontal velocity field.
-Tracer height is unchanged downwind of the mountains because advection is parallel to the surface.
+Tracer height is unchanged downwind of the mountains because advection is parallel to the coordinate surfaces.
 
@@
 
-The test was integrated forward by 5 hours on the BTF, SLEVE and cut cell grids, and a regular grid with flat terrain.
+The test was integrated forward by 5 hours using a timestep $\Delta t = \SI{100}{\second}$ on the BTF, SLEVE and cut cell grids, and a regular grid with flat terrain.
 Maximum vertical velocities are presented in figure~\ref{fig:resting}.
@@
 Following the gravity waves test, the simulation is integrated forward by \SI{18000}{\second}, by which time the tracer initially upwind of the mountain will have cleared the mountain range.
-Hence, the analytic solution, $\phi_T$, can be found by considering the vertical displacement of the tracer profile by the terrain following velocity field.
+Hence, the analytic solution, $\phi_T$, can be found by considering the vertical displacement of the tracer profile by the terrain following velocity field:
 \begin{align}
@@
 Errors were also large on the cut cell grids in the second terrain following tracer advection test.
-This result suggests that, in the gravity waves test, that potential temperature anomalies in the cut cell grids are primarily due to advection errors.
+This result suggests that, in the gravity waves test, potential temperature errors in the cut cell grids are primarily due to advection errors.
 

### 2015-12-11 src/tf-cutcell-comparison/responses2.tex
create document to respond to secound round of comments

\documentclass{article}
\usepackage{fullpage}
\usepackage{url}
\usepackage{natbib}
\usepackage{enumitem}
\usepackage{siunitx}
\usepackage{amsmath}
\usepackage{mathtools}
\newlist{comment}{enumerate}{1}
\setlist[comment,1]{label={\arabic{commenti}.},leftmargin=*,resume=comment}

\begin{document}
\section*{Review \#1}
\begin{quotation}
This revised manuscript is improved from its original version. The Sch\"{a}r mountain wave test with various grid resolutions and the new test for theta advection are quite informative in identifying the potential source of errors in the low-level theta field in the cut-cell mountain-wave simulation. However, as indicated in my previous review, I still believe that this paper should be more focused, with more emphasis on what's new and interesting in these test cases comparing terrain-following and cut-cell coordinates, and less attention to results that are obvious or previously documented. More specific comments and suggestions are included below.

Sections 4a and 4b – Section 4a presents results for the Sch\"{a}r advection test case for various vertical coordinates and advection schemes; there is little new here that is not already covered in the \citet{schaer2002} paper. The advection tests in section 4b provide an interesting counter example to this test. I would recommend switching the emphasis in these two sections, presenting the advected tracer contours in a new Fig. 3 for the terrain-following advection, and just including results from the Sch\"{a}r case in table 1 for comparison purposes.
\end{quotation}
\begin{quotation}
 Lines 131-147 on cut-cell grid generation - It's still not clear how or why a cell vertex  is moved laterally in constructing the cut-cell mesh. Taking Fig. 2 as an example, it  seems that following the authors' directions to move all vertices lying underneath the  orography on the uniform grid up to the surface would produce four grid cells as  shown in the figure below rather than the three cells shown in Fig. 2. This four-cell structure appears to be consistent with the generation of thin cells as the authors  point out in Fig. 5c. If lateral displacement of a vertex is allowed, how is it decided  whether it should be moved and how is it decided where along the surface it should  be located (i.e. the location of point v in Fig. 2)?
\end{quotation}
\begin{quotation}
	Section 3b - Finite difference linear advection model - This advection model is solving the same flux form equation (4) as in the previous subsection. Thus, the advection operator in (6) should be in terms of $\partial (u\phi)/\partial x$.  The finite differencing in (6) can also be considered a finite-volume representation, rewriting it as 
	\begin{align*}
		\frac{\partial u \phi}{\partial x} &= \frac{1}{\Delta x} \left( u_{i+1/2} F_{i+1/2} - u_{i - 1/2} F_{i - 1/2}\right), \\
		F_{i+1/2} &= \frac{1}{12} \left( - \phi_{i+2} + 7 \phi_{i+1} + 7 \phi_i - \phi_{i-1} \right).
	\end{align*}
\end{quotation}
\begin{quotation}
 Lines 265-267 and table 1 – The authors are unnecessarily complicating their tests  by comparing results from the cut-cell grid and a regular grid. These grids should be identical for this test and merit no further comparative consideration. There is no need to simulate with both grids, show their identical behavior in table 1, and then  state that the same behavior ``is to be expected.''
\end{quotation}
\begin{quotation}
 Lines 271-279 and Table 1 – The $\ell_2$ error norms have changed by more than an order  of magnitude from those reported in the original manuscript. What changed?
\end{quotation}
\begin{quotation}
 Lines 280-285 – Reviewer 2 questioned the large change in the $\ell_2$ error norm when 
 the domain width was changed from 301 to 300 km. I share the concern expressed 
 by this reviewer and find the authors' added explanation (``It's likely that changes in 
 the domain width affect the wave power spectrum of the discrete terrain profile.'') to 
 be inadequate. If the results are sensitive to the exact location of the lateral 
 boundaries, then the numerical experiment is not well designed; the boundaries 
 should be moved far enough away from the mountain so that this sensitivity 
 disappears.  Is it possible that the authors shifted the location of the terrain by a half 
 grid interval when they changed the domain width by one grid interval? This would 
 occur if they just defined the center of the terrain profile to be in the center of the 
 domain. If that’s the case, then the results are understandable since numerically, it's a 
 different terrain shape. However, the changing behavior would have nothing to do 
 with the domain width, and thus this sensitivity test should then be removed from 
 the paper altogether. One would get the same sensitivity leaving the domain width 
 fixed and just shifting the terrain location by one-half grid interval.  
\end{quotation}
\begin{quotation}
 Lines 321-322 – When the flow is aligned with the grid there is no cross coordinate 
 flow and therefore the numerics are essentially reduced to a one-dimensional 
 advection equation. I don't see how grid distortion is even an issue. 
\end{quotation}
\begin{quotation}
 Lines 364-368 – The authors did not need to run these resting atmosphere 
 simulations to reach the conclusion stated in this paragraph. As stated in my first 
 review, this test case does not challenge the numerics on a cut-cell grid since the horizontal pressure-gradient errors will obviously be reduced in the vicinity of the 
 low level inversion where the coordinates surfaces are perfectly horizontal. A better 
 test case might be to specify an inversion layer that intersects the terrain, since on 
 the cut-cell grid, it's only just adjacent to the terrain where pressure gradient 
 calculations between cell centers are not horizontal.   
\end{quotation}
\begin{quotation}
	Section 4d – Sch\"{a}r mountain-wave tests. These tests are more interesting now that 
 the case has been simulated with a variety of grid resolutions. However, I'm 
 surprised that the analyses of these results seem to focus on behavior far 
 downstream of the mountain, as is evident in Figs. 6 and 7. I'm not aware of any 
 published results that extend beyond x = 25 km in Fig. 6. Since the amplitude of the 
 disturbances are decaying with downstream distance, small inaccuracies in behavior 
 of the absorbing layer and boundary conditions have larger impact such that the 
 relative errors in comparison to the analytic solution become large. (Note, for 
 example the erroneous extension of negative perturbation theta contours to the 
 surface at x = 25 km in Fig. 6.)  All we learn from Fig. 7b is that errors in theta near 
 the surface in the cut-cell cases, which must have been produced in the vicinity of the 
 terrain, continue to be swept downstream. I think it would be more informative to 
 focus on solution differences between the BTF and cut-cell results near the terrain as 
 is done in Fig. 8 for the terrain-following theta advection. The authors could display 
 these difference fields for the various grid resolutions in exactly the same manner as 
 in Fig. 8. The theta advection test presented by the authors is a very nice complement 
 to the gravity-wave tests. It confirms that the BTF solution is behaving properly and 
 therefore lends support to interpreting differences in the two grids as inaccuracies in 
 the cut cell treatment. I would expect that the difference fields in the Sch\"{a}r case are 
 at least qualitatively similar to those exhibited in the theta advection case, which 
 would further confirm the relevance of this advection test as well as the conclusion 
 that errors in the advection of theta through the cut cells may be responsible for the 
 anomalies appearing with this grid.
\end{quotation}
\begin{quotation}
 Lines 479-480 – As mentioned above, for the horizontal advection test, shouldn't the  results for the cut-cell and regular grids be identical?
\end{quotation}
\begin{quotation}
 Lines 481-482 – Emphasizing an error reduction of two orders of magnitude for this 
 resting atmosphere case with cut cells compared to a terrain-following coordinate is 
 not very informative. In both cases the errors appear to be quite small, and thus the 
 more relevant question is whether these differences are of any realistic significance. 
\end{quotation}

\section*{Review \#2}
\begin{quotation}
Overall, the authors have been very responsive to my comments and suggestions, and the manuscript improved considerably.
However, there are still a few minor aspects that need attention:
\begin{comment}
\item From equation (7a) it is clear that the symbol mu needs a physical unit which is 1/s. The statement in line 195
that mu is dimensionless is therefore wrong and needs to be corrected.
\end{comment}
\end{quotation}
\begin{quotation}
	\begin{comment}
	\item Line 241: $\phi$ is defined as the tracer density. Therefore, the definition of $\phi_0$ needs the physical unit \si{\kilogram\per\meter\cubed}
	\end{comment}
\end{quotation}
\begin{quotation}
	\begin{comment}
	\item Page 18, test c): The initial conditions look incomplete. Please provide information about the initial density or pressure initialization. Do you assume that the pressure at z=0 m is 1000 hPa?
	\end{comment}
\end{quotation}
\begin{quotation}
	\begin{comment}
	\item Test d) and e): As for test c), the descriptions of the initial conditions looks incomplete. Is the initial w zero in d,e?
\end{comment}
\end{quotation}
\begin{quotation}
	\begin{comment}
	\item Line 389: mu needs the physical unit 1/s.
\end{comment}
\end{quotation}
\begin{quotation}
	\begin{comment}
	\item Line 440: It would be better to say 'The functional form of the initial density field ...potential temperature field (Eq. (22)) in the gravity wave test.
	\end{comment}
\end{quotation}
\begin{quotation}
	\begin{comment}
	\item Line 445: While it is mathematically possible to initialize the tracer density with 288 (needs physical units!) at the ground, it does not make any physical sense. However, it is okay to leave it for this idealized test case, but a more physical choice is clearly more desirable.
	\end{comment}
\end{quotation}
\begin{quotation}
	\begin{comment}
	\item Line 449: Do you refer to the normalized $\ell_2$ error here?
	\end{comment}
\end{quotation}
\begin{quotation}
	\begin{comment}
	\item Check the references for some phrases that need to be capitalized, as e.g. Appalachian in line 571 or Lagrangian in line 567 
	\end{comment}
\end{quotation}

\bibliographystyle{ametsoc2014}
\bibliography{references}
\end{document}

### 2015-12-11 src/tf-cutcell-comparison/responses2.tex
minor amendments for reviewer 2, round 2

@@
 \usepackage{mathtools}
+\usepackage{xcolor}
 \newlist{comment}{enumerate}{1}
 \setlist[comment,1]{label={\arabic{commenti}.},leftmargin=*,resume=comment}
+\newcommand{\TODO}[1]{\textcolor{purple}{TODO: \emph{#1}}}
 
@@
 \end{quotation}
+Removed `dimensionless' on line \TODO{xxx} and now included correct units where $\mu$ is specified on lines \TODO{xxx} and \TODO{xxx}.
+
 \begin{quotation}
@@
 \end{quotation}
+Added physical units to $\phi_0$ and in the caption for table 1.
+
 \begin{quotation}
@@
 \end{quotation}
+Yes, amended line \TODO{xxx} to state that $p_0 = \SI{1000}{\hecto\pascal}$, which applies to all tests that use the model.
+
 \begin{quotation}
@@
 \end{quotation}
+The value of $p_0$ applies to tests (c) and (d).
+For test (d), line \TODO{xxx} now states that $w = \SI{0}{\meter\per\second}$.
+For test (e), the velocity field is defined by the streamfunction in equation (26), and $w = \SI{0}{\meter\per\second}$ only where $\mathrm{d}h/\mathrm{d}x = 0$.
+
 \begin{quotation}
@@
 \end{quotation}
+Added, many thanks.
+
 \begin{quotation}
@@
 \end{quotation}
+Instead of amending this line, we now refer to the field as a potential temperature field, not a tracer density field.  Further details are provided in response to the subsequent comment.
+
 \begin{quotation}
@@
 \end{quotation}
+The field is really the potential temperature field from the gravity waves test.
+We have modified this section so that the tracer density field is now called a potential temperature field.
+Lines \TODO{xxx twice in the abstract, in the intro and conclusion} have also been amended accordingly so that this test is no longer refered to as a tracer advection test..
+Equation (\TODO{xxx}) now has $\theta_0 = \SI{288}{\kelvin}$.
+
 \begin{quotation}
@@
 \end{quotation}
+Yes, amended the manuscript to state this.
 \begin{quotation}
@@
 \end{quotation}
+Corrected these, also `Navier-Stokes' and `Met Office Unified Model'.
 

### 2015-12-11 src/tf-cutcell-comparison/tf-cutcell-comparison.tex
minor amendments for reviewer 2, round 2

@@
 A series of standard two-dimensional tests using idealised terrain are performed: tracer advection in a prescribed horizontal velocity field, a test starting from stably stratified initial conditions, and orographically induced gravity waves described by nonhydrostatic dynamics.
-In addition, two new tracer advection tests are formulated having a velocity field that is everywhere tangential to the terrain following coordinate surfaces.
+In addition, two new advection tests are formulated having a velocity field that is everywhere tangential to the terrain following coordinate surfaces.
 These new tests present a challenge to the advection scheme on cut cell grids.
-The results of the advection tests demonstrate that tracer accuracy depends primarily upon alignment of the flow with the grid rather than the grid orthogonality.
+The results of the advection tests demonstrate that accuracy depends primarily upon alignment of the flow with the grid rather than the grid orthogonality.
 As expected, the cut cell grid maintains greater accuracy for horizontal advection and in the test of stationary, stratified flow.
@@
 In section 3 we give the governing equations, outline the model from \citet{weller-shahrokhi2014} and describe the modification which improves stability for long timesteps in the presence of steep orography.
-Section 4 analyses the results from three tracer advection tests, a test of a stably stratified atmosphere initially at rest, and orographically induced gravity waves.
+Section 4 analyses the results from three advection tests, a test of a stably stratified atmosphere initially at rest, and orographically induced gravity waves.
 Concluding remarks are made in section 5.
@@
 \end{subequations}
-where \(\rho\) is the density, \(\mathbf{u}\) is the velocity field, \(\mathbf{g}\) is the gravitational acceleration, \(c_p\) is the heat capacity at constant pressure, \(\theta = T \left(p_0/p\right)^\kappa\) is the potential temperature, \(T\) is the temperature, \(p\) is the pressure, \(p_0\) is a reference pressure, \(\Pi = \left(p / p_0 \right)^\kappa\) is the Exner function of pressure, and \(\kappa = R/c_p\) is the gas constant to heat capacity ratio.
-\(\mu\) is a dimensionless damping function used for the sponge layer in the gravity waves test in section~\ref{sec:results}\ref{sec:gw}.
+where \(\rho\) is the density, \(\mathbf{u}\) is the velocity field, \(\mathbf{g}\) is the gravitational acceleration, \(c_p\) is the heat capacity at constant pressure, \(\theta = T \left(p_0/p\right)^\kappa\) is the potential temperature, \(T\) is the temperature, \(p\) is the pressure, \(p_0 = \SI{1000}{\hecto\pascal}\) is a reference pressure, \(\Pi = \left(p / p_0 \right)^\kappa\) is the Exner function of pressure, and \(\kappa = R/c_p\) is the gas constant to heat capacity ratio.
+\(\mu\) is a damping function used for the sponge layer in the gravity waves test in section~\ref{sec:results}\ref{sec:gw}.
 
@@
 Fourth, as specified by \citet{schaer2002}, a test of orographically-induced gravity waves is performed.
-Finally, another tracer advection test is formulated that transports a stably stratified thermal profile in a terrain following velocity field.
+Finally, another advection test is formulated that transports a stably stratified thermal profile in a terrain following velocity field.
 No explicit diffusion is used in any of the tests.
@@
 \end{align}
-where $A_x = \SI{25}{\kilo\meter}$, $A_z = \SI{3}{\kilo\meter}$ are the horizontal and vertical half-widths respectively, and $\phi_0 = 1$ is the maximum density of the tracer.
+where $A_x = \SI{25}{\kilo\meter}$, $A_z = \SI{3}{\kilo\meter}$ are the horizontal and vertical half-widths respectively, and $\phi_0 = \SI{1}{\kilogram\per\meter\cubed}$ is the maximum density of the tracer.
 At $t = \SI{0}{\second}$, the tracer is centred at $(x_0, z_0) = (\SI{-50}{\kilo\meter}, \SI{9}{\kilo\meter})$ so that the tracer is upwind of the mountain and well above the maximum terrain height of \SI{3}{\kilo\meter}.
@@
 The initial thermodynamic conditions are in discrete hydrostatic balance in the vertical direction, having a reference potential temperature of $\theta(z = 0) = \SI{288}{\kelvin}$ and constant stability with Brunt-V\"ais\"al\"a frequency $N = \SI{0.01}{\per\second}$ everywhere, except for a more stable layer of $N = \SI{0.02}{\per\second}$ between $\SI{2}{\kilo\meter} \leq z \leq \SI{3}{\kilo\meter}$.
-The damping function, \(\mu\), is set to zero.
+The damping function, \(\mu\), is set to \SI{0}{\per\second}.
 Unlike \citet{klemp2011}, there is no eddy diffusion in the equation set.
@@
 
-A uniform horizontal wind $u = \SI{10}{\meter\per\second}$ is prescribed in the interior domain and at the inlet boundary.
+A uniform horizontal wind $(u, w) = (10, 0)\:\si{\meter\per\second}$ is prescribed in the interior domain and at the inlet boundary.
 No normal flow is imposed at the top and bottom boundaries and the velocity field has a zero gradient outlet boundary condition.
@@
 \end{align}
-where $\overline{\mu} = 1.2$ is the damping coefficient, $z_B = \SI{20}{\kilo\meter}$ is the bottom of the sponge layer, $H = \SI{30}{\kilo\meter}$ is the top of the domain, $x_0 = \SI{-150}{\kilo\meter}$ is the leftmost limit of the domain and $x_I = \SI{-140}{\kilo\meter}$ is the rightmost extent of the inlet sponge layer.
+where $\overline{\mu} = \SI{1.2}{\per\second}$ is the damping coefficient, $z_B = \SI{20}{\kilo\meter}$ is the bottom of the sponge layer, $H = \SI{30}{\kilo\meter}$ is the top of the domain, $x_0 = \SI{-150}{\kilo\meter}$ is the leftmost limit of the domain and $x_I = \SI{-140}{\kilo\meter}$ is the rightmost extent of the inlet sponge layer.
 The sponge layer is only active on faces whose normal is vertical so that it damps vertical momentum only.
@@
 This variation may be due to differences in the wind fields between grids, or errors in the advection of potential temperature, amongst other possible causes.
-To help establish the primary source of error, a new advection test is formulated in which the initial potential temperature field from the gravity waves test is used as a tracer density field.
-To eliminate any differences in wind fields, the tracer is advected in a fixed, terrain-following velocity field that mimics the flow in the gravity waves test.
+To help establish the primary source of error, a new advection test is formulated in which the initial potential temperature field from the gravity waves test is used.
+To eliminate any differences in wind fields, the field is advected in a fixed, terrain-following velocity field that mimics the flow in the gravity waves test.
 
@@
 
-The initial tracer density field, \(\phi\), and its boundary conditions, are the same as those of the initial potential temperature field in the gravity waves test.
-Following the gravity waves test, the simulation is integrated forward by \SI{18000}{\second}, by which time the tracer initially upwind of the mountain will have cleared the mountain range.
-Hence, the analytic solution, $\phi_T$, can be found by considering the vertical displacement of the tracer profile by the terrain following velocity field:
+The potential temperature field, \(\theta\), and its boundary conditions, are the same as those of the initial potential temperature field in the gravity waves test.
+Following the gravity waves test, the simulation is integrated forward by \SI{18000}{\second}, by which time the potential temperature initially upwind of the mountain will have cleared the mountain range.
+Hence, the analytic solution, $\theta_T$, can be found by considering the vertical displacement of the thermal profile by the terrain following velocity field:
 \begin{align}
-	\phi_T(x, z) = \phi_0 \exp \left( \frac{N^2}{g} z^\star(x, z) \right) 
+	\theta_T(x, z) = \theta_0 \exp \left( \frac{N^2}{g} z^\star(x, z) \right) 
 \end{align}
-where the tracer density at $z = 0$, $\phi_0 = 288$, and the transform, $z^\star$, is given by rearranging equation~(\ref{eqn:btf}).
+where the potential temperature at $z = 0$, $\theta_0 = \SI{288}{\kelvin}$, and the transform, $z^\star$, is given by rearranging equation~(\ref{eqn:btf}).
 
-Enlargements of the error field near the mountain are shown in figure~\ref{fig:thermalAdvection} at $\Delta z = \SI{50}{\meter}$ with contours in tracer density overlayed.
-Errors are too small to be visible on the BTF grid with a maximum error of \num{4.26e-4}.
+Enlargements of the error field near the mountain are shown in figure~\ref{fig:thermalAdvection} at $\Delta z = \SI{50}{\meter}$ with contours of potential temperature overlayed.
+Errors are too small to be visible on the BTF grid with a maximum $\ell_2$ error of \num{4.26e-4}.
 However, on the cut cell grid, the maximum error is \num{0.302}.
@@
 
-For comparison with the potential temperature anomalies in the gravity waves test, vertical profiles of tracer density error are taken at $x = \SI{50}{\kilo\meter}$.
+For comparison with the potential temperature anomalies in the gravity waves test, vertical profiles of potential temperature error are taken at $x = \SI{50}{\kilo\meter}$.
 As seen in figure~\ref{fig:gw-theta}c, errors are negligible on the BTF grids, but figure~\ref{fig:gw-theta}d reveals significant errors in the lowest layers of the cut cell grids that were advected from the mountain peaks.
@@
 
-While the magnitude and structure of tracer error on the cut cell grids differs from potential temperature anomalies in the gravity waves test, results on the BTF grids are in close agreement in both tests but not on the cut cell grids.
+While the magnitude and structure of error on the cut cell grids in this test differs from potential temperature anomalies in the gravity waves test, results on the BTF grids are in close agreement in both tests but not on the cut cell grids.
 Therefore, it is likely that anomalies on the cut cell grids in the gravity waves test are primarily due to errors in the advection of potential temperature through cut cells.
@@
 Conversely, in the first terrain following tracer advection test, errors were large on the SLEVE and cut cell grids where velocities were misaligned with the grids.
-Errors were also large on the cut cell grids in the second terrain following tracer advection test.
+Errors were also large on the cut cell grids in the second terrain following thermal advection test.
 This result suggests that, in the gravity waves test, potential temperature errors in the cut cell grids are primarily due to advection errors.
@@
 \begin{table*}
-	\caption{Minimum and maximum tracer magnitudes and \(\ell_2\) error norms, defined by equation~(\ref{eqn:l2-error}), at \(t = \SI{10000}{\second}\) in the horizontal and terrain following tracer advection tests using centred linear and cubic upwind-biased schemes.
+	\caption{Minimum and maximum tracer densities (\si{\kilogram\per\meter\cubed}) and \(\ell_2\) error norms, defined by equation~(\ref{eqn:l2-error}), at \(t = \SI{10000}{\second}\) in the horizontal and terrain following tracer advection tests using centred linear and cubic upwind-biased schemes.
 For the horizontal advection test, \(\ell_2\) error norms, minimum and maximum values are given for the fourth order scheme using the modified code from \citet{schaer2002}.}

### 2015-12-11 src/tf-cutcell-comparison/responses2.tex
amend 4th order scheme description, more explanation of pressure initialisation, temporary responses to reviewer 1, round 2

@@
 \end{quotation}
+\TODO{redo everything using slanted cell grids}
+
 \begin{quotation}
@@
 \end{quotation}
+Amended as suggested, many thanks.
+
 \begin{quotation}
@@
 \end{quotation}
+\TODO{agreed, although the comparison with the fourth order scheme in table 1 is perhaps useful?}
+
 \begin{quotation}
@@
 \end{quotation}
+In the first draft, the $\ell_2$ values had not been normalised, such that $\ell_2 = \sqrt{\sum_c \left(\phi - \phi_T\right)^2 \mathcal{V}_c}$.
+This was fixed in the second draft so that $\ell_2$ was calculated correctly according to equation (13).
 \begin{quotation}
@@
 \end{quotation}
+Yes, the change in domain width causes the terrain to be shifted by $\Delta x / 2$, and it is this which results in a change in accuracy.
+We had not intended to perform a sensitivity test, but only to clarify the specification for the \citet{schaer2002} advection test.
+\TODO{update the manuscript to put less emphasis on this}
+
 \begin{quotation}
@@
 \end{quotation}
+The advection scheme fits a multidimensional polynomial using a stencil of cell centroids, hence grid distortions will affect the resulting least squares fit.
+If the test used a dimensionally-split advection scheme with a terrain-following coordinate transform, then the test would indeed reduce to a one-dimensional problem.
+\TODO{clarify in the manuscript}
+
 \begin{quotation}
@@
 \end{quotation}
+\TODO{good idea, we should do this, but first see if anyone else does something similar (zaengl2012?).  Should we abandon the klemp2011 test entirely or just put less emphasis on its results?}
+
 \begin{quotation}
@@
 \end{quotation}
+\TODO{the gravity wave amplitude decreases but the advection errors do not.
+hence, this is a good choice for the sample line.
+we considered plotting heatmaps, but found that it was not as easy to distinguish between the gravity waves and the advection errors.}
+
 \begin{quotation}
@@
 \end{quotation}
+This was a mistake: table 1 confirmed that min, max and $\ell_2$ errors were identical on both grids.
+This line now reads, `in the horizontal advection test, accuracy was greatest on the cut cell grid'.
+
 \begin{quotation}
@@
 \end{quotation}
+\TODO{this conclusion may change when we run the new resting atmosphere test}
 
@@
 Yes, amended line \TODO{xxx} to state that $p_0 = \SI{1000}{\hecto\pascal}$, which applies to all tests that use the model.
+Line \TODO{xxx} now cites \citet{weller-shahrokhi2014} which explains how the Exner function of pressure is calculated so that it is in discrete hydrostatic balance with the prescribed potential temperature file.
 

### 2015-12-11 src/tf-cutcell-comparison/tf-cutcell-comparison.tex
amend 4th order scheme description, more explanation of pressure initialisation, temporary responses to reviewer 1, round 2

@@
 It uses terrain following coordinates and it is configured with leapfrog timestepping and either second-order centred differences, or a fourth-order centred difference scheme given by:
+\begin{subequations}
 \begin{align}
-	\frac{\partial \phi_i}{\partial x} \approx \frac{1}{\Delta x} \left( \frac{1}{12} \phi_{i-2} - \frac{2}{3} \phi_{i-1} + \frac{2}{3} \phi_{i+1} - \frac{1}{12} \phi_{i+2} \right)
+	\frac{\partial u \phi}{\partial x} &\approx \frac{1}{\Delta x} \left( u_{i+\frac{1}{2}} F_{i+\frac{1}{2}} - u_{i - \frac{1}{2}} F_{i - \frac{1}{2}}\right) \\
+	F_{i+\frac{1}{2}} &= \frac{1}{12} \left( - \phi_{i+2} + 7 \phi_{i+1} + 7 \phi_i - \phi_{i-1} \right)
 \end{align}
-and similarly for \(\partial \phi_i / \partial z\).
+\end{subequations}
+and similarly for \(\partial (w \phi) / \partial z\).
 
@@
 
-The initial thermodynamic conditions are in discrete hydrostatic balance in the vertical direction, having a reference potential temperature of $\theta(z = 0) = \SI{288}{\kelvin}$ and constant stability with Brunt-V\"ais\"al\"a frequency $N = \SI{0.01}{\per\second}$ everywhere, except for a more stable layer of $N = \SI{0.02}{\per\second}$ between $\SI{2}{\kilo\meter} \leq z \leq \SI{3}{\kilo\meter}$.
+The initial potential temperature field has $\theta(z = 0) = \SI{288}{\kelvin}$ and a constant static stability with Brunt-V\"ais\"al\"a frequency $N = \SI{0.01}{\per\second}$ everywhere, except for a more stable layer of $N = \SI{0.02}{\per\second}$ between $\SI{2}{\kilo\meter} \leq z \leq \SI{3}{\kilo\meter}$.
+The Exner function of pressure is calculated so that it is in discrete hydrostatic balance in the vertical direction \citep{weller-shahrokhi2014}.
 The damping function, \(\mu\), is set to \SI{0}{\per\second}.
@@
 Cut cell grids reduced errors in two of the five tests.
-First, in the horizontal advection test, tracer accuracy on the cut cell grid was almost as good as accuracy on a regular grid with no mountain.
+First, in the horizontal advection test, accuracy was greatest on the cut cell grid.
 Second, in the resting atmosphere test, spurious vertical velocities were two orders of magnitude smaller on the cut cell grid compared with the terrain following grids.

### 2015-12-14 src/tf-cutcell-comparison/tf-cutcell-comparison.tex
remove cut cell results pending rerun on slanted cell meshes

@@
 \bibpunct{(}{)}{;}{a}{}{,}
+
+\newcommand{\TODO}[1]{\textcolor{purple}{TODO: \emph{#1}}}
 
@@
 
+\TODO{put less emphasis on this, just clarify the correct test setup to obtain Schaer's results.}
 The centred linear scheme is sensitive to grid-scale changes in the terrain profile: when the domain width is reduced to \SI{300}{\kilo\meter}, the $\ell_2$ error on the BTF grid rises from \num{0.284} to \num{0.432}.
@@
 This test determines whether the primary source of numerical error is due to non-orthogonality or misalignment of the flow with grid layers.
+\TODO{explain how grid distortion could reduce accuracy of the advection scheme}
 
@@
 Errors are too small to be visible on the BTF grid with a maximum $\ell_2$ error of \num{4.26e-4}.
-However, on the cut cell grid, the maximum error is \num{0.302}.
+However, on the cut cell grid, the maximum error is \TODO{\num{999}}.
 Advection errors are apparent around mountainous terrain, with small cells having the largest errors.
@@
 \hline
-Horizontal        & Centred linear      & \(\ell_2\) error & 0        & 0.284   & 0.0316          & 0.0304         & 0.0304      \\
-                  &                     & min              & 0        & -0.275  & -0.0252         & -0.0251        & -0.0251     \\
-                  &                     & max              & 1        & 0.925   & 0.985           & 0.985          & 0.985       \\
+Horizontal        & Centred linear      & \(\ell_2\) error & 0        & 0.284   & 0.0316          & 999         & 0.0304      \\
+                  &                     & min              & 0        & -0.275  & -0.0252         & 999        & -0.0251     \\
+                  &                     & max              & 1        & 0.925   & 0.985           & 999          & 0.985       \\
                   & Fourth order        & \(\ell_2\) error & 0        & 0.0938  & 0.00244         & {---}          & 0.00234     \\
@@
                   &                     & max              & 1        & 1.00    & 0.984           & {---}          & 0.983       \\
-                  & Cubic upwind-biased & \(\ell_2\) error & 0        & 0.112   & 0.0146          & 0.00784        & 0.00784     \\
-                  &                     & min              & 0        & -0.0464 & -0.0106         & -0.000674      & -0.00674    \\
-                  &                     & max              & 1        & 0.922   & 0.982           & 0.983          & 0.983       \\
+                  & Cubic upwind-biased & \(\ell_2\) error & 0        & 0.112   & 0.0146          & 999        & 0.00784     \\
+                  &                     & min              & 0        & -0.0464 & -0.0106         & 999      & -0.00674    \\
+                  &                     & max              & 1        & 0.922   & 0.982           & 999          & 0.983       \\
 \hline
-Terrain following & Centred linear      & \(\ell_2\) error & 0        & 0.0341  & 0.235           & {unstable}     & {---}        \\
-	          &                     & min              & 0        & -0.0245 & -0.120          & {unstable}     & {---}        \\
-		  &                     & max              & 1        & 0.985   & 0.950           & {unstable}     & {---}        \\
-		  & Cubic upwind-biased & \(\ell_2\) error & 0        & 0.0209  & 0.162           & 0.181          & {---}        \\
-                  &                     & min              & 0        & -0.0110 & -0.0263         & -0.0284        & {---}        \\
-                  &                     & max              & 1        & 0.983   & 0.865           & 0.851          & {---}        \\
+Terrain following & Centred linear      & \(\ell_2\) error & 0        & 0.0341  & 0.235           & 999     & {---}        \\
+	          &                     & min              & 0        & -0.0245 & -0.120          & 999     & {---}        \\
+		  &                     & max              & 1        & 0.985   & 0.950           & 999     & {---}        \\
+		  & Cubic upwind-biased & \(\ell_2\) error & 0        & 0.0209  & 0.162           & 999          & {---}        \\
+                  &                     & min              & 0        & -0.0110 & -0.0263         & 999        & {---}        \\
+                  &                     & max              & 1        & 0.983   & 0.865           & 999          & {---}        \\
 \hline
@@
 \hline
-500	& 1.01 &  1.68 \\
-300	& 1.01 &  4.33 \\
-250	& 1.01 &  5.42 \\
-200	& 1.01 &  25.4 \\
-150	& 1.01 &  10.0 \\
-125	& 1.01 &  10.1 \\
-100	& 1.01 &  36.3 \\
-75	& 1.01 &  52.9 \\
-50	& 1.01 &  75.3 \\
+500	& 1.01 &  999 \\
+300	& 1.01 &  999 \\
+250	& 1.01 &  999 \\
+200	& 1.01 &  999 \\
+150	& 1.01 &  999 \\
+125	& 1.01 &  999 \\
+100	& 1.01 &  999 \\
+75	& 1.01 &  999 \\
+50	& 1.01 &  999 \\
 \hline

### 2015-12-14 src/tf-cutcell-comparison/tf-cutcell-comparison.tex
swap out figure for slanted cell figure

@@
 
+\TODO{revise these three paragraphs to describe the slanted cell mesh}
 Cut cell grids were created by starting from a uniform grid, moving all cell vertices lying underneath the orography up to the surface.
@@
 	\centering
-	\includegraphics{../fig-snappyHexMesh/fig-snappyHexMesh.pdf}
+	\includegraphics{../fig-grid-generation/fig-grid-generation.pdf}
 	%
-	\caption{Illustrative behaviour of the \texttt{snappyHexMesh} tool in cut cell grid creation.
-The terrain surface, denoted by a thick grey line, is taken from the lower boundary of the BTF grid.
-The surface is intersected with a rectangular grid shown by thin solid lines.
-Vertex $v$ is `snapped' to the surface, moving from the position marked by the open circle to that of the closed circle.
-The resultant cut cell grid is shown by thick dashed lines.
-With a typical cut cell technique, cell $c$ would be a pentagon, as shown by the dotted line, but \texttt{snappyHexMesh} produces a quadrangular cell.}
-	\label{fig:snappyHexMesh}
+	\caption{Illustration of a slanted cell grid (a) before, and (b) after construction.
+	The terrain surface, denoted by a heavy dotted line, intersects a uniform rectangular grid comprising four cells, $c_1$, $c_2$, $c_3$ and $c_4$.
+The cell vertices, marked by open circles, are moved upwards to the points at which the terrain intersects vertical cell edges, marked by filled circles.
+Cells that have no volume are removed.
+Where a cell has two vertices occupying the same point, the zero-length edge that joins those vertices is removed.
+In this illustration, cell $c_4$ is removed because it has no volume, and the zero-length edge at point $p$ is removed to create a triangular cell, $c_3$.}
+	\label{fig:grid-generation}
 \end{figure*}

### 2015-12-14 src/tf-cutcell-comparison/tf-cutcell-comparison.tex
rewrite cut cell grid generation description

@@
 
-\TODO{revise these three paragraphs to describe the slanted cell mesh}
-Cut cell grids were created by starting from a uniform grid, moving all cell vertices lying underneath the orography up to the surface.
-Then, the OpenFOAM utility \texttt{snappyHexMesh} was used to remove cells whose centres lay below the surface.
-In addition, the utility displaces boundary vertices so that they are `snapped' to the surface \citep{openfoam2015}.  An example of the resulting cut cell grid is shown in figure~\ref{fig:grid}c.
-
-There are two details of grid construction which mean that resulting cut cell grids can differ slightly from a typical cut cell grid as described by \citet{adcroft1997}.
-First, when \texttt{snappyHexMesh} moves vertices onto the terrain surface, some points are moved horizontally.
-Second, it does not create new points necessary for pentagonal cells.
-An example of these effects is shown in figure~\ref{fig:snappyHexMesh} in which vertex $v$ has been moved vertically and horizontally to avoid creating a small cell.
-The resultant grid, shown by thick dashed lines, contains triangular and quadrangular cells.
-With a typical shaving method, cell $c$ would be pentagonal, as shown by the dotted line.
-However, \texttt{snappyHexMesh} creates a quadrilateral cell.
+Cut cell grids are generated in a different way to the typical shaving technique described by \citet{adcroft1997}.
+Starting from a uniform grid, all cell vertices that lie beneath the orography are moved up to the surface.
+Where all four of a cell's vertices are moved, the cell has zero volume and so it is removed.
+Where two vertices at the same horizontal location are moved up to the surface they will occupy the same point; this results in a zero-length edge that is removed to create a triangular cell.
+Figure~\ref{fig:grid-generation} shows how a $2 \times 2$-cell, uniform grid is transformed into a cut cell grid.
+Cell $c_4$ is removed because it has zero volume, and the zero-length edge at point $p$ is removed to create a triangular cell, $c_3$.
 

### 2015-12-17 src/tf-cutcell-comparison/responses2.tex
move points down as well as up

@@
 \end{quotation}
-\TODO{redo everything using slanted cell grids}
+We have found an alternative method of generating meshes that does not require the use of the \texttt{snappyHexMesh} utility.
+This new method does not move vertices laterally and is more straightforward to describe.
+This new method creates meshes that retain the long, thin cells from the old method.
+The new method is described in section 2 and figure 2.
+All test results on cut cell grids have been updated.
 

### 2015-12-17 src/tf-cutcell-comparison/tf-cutcell-comparison.tex
move points down as well as up

@@
 Starting from a uniform grid, all cell vertices that lie beneath the orography are moved up to the surface.
-Where all four of a cell's vertices are moved, the cell has zero volume and so it is removed.
+Additionally, to avoid creating very thin cells, all vertices up to $\Delta z/5$ above the orography are moved down to the surface.
+\TODO{should I include an illustration of how this extra tweak helps?
+can I incorporate it into the existing figure?} Where all four of a cell's vertices are moved, the cell has zero volume and so it is removed.
 Where two vertices at the same horizontal location are moved up to the surface they will occupy the same point; this results in a zero-length edge that is removed to create a triangular cell.

### 2015-12-18 src/tf-cutcell-comparison/tf-cutcell-comparison.tex
TODO for updating grid generation figure

@@
 \TODO{should I include an illustration of how this extra tweak helps?
-can I incorporate it into the existing figure?} Where all four of a cell's vertices are moved, the cell has zero volume and so it is removed.
+can I incorporate it into the existing figure?
+yes, by adding a row above the existing top row, and moving the rightmost terrain point closer to the upper right vertex in $c_2$} Where all four of a cell's vertices are moved, the cell has zero volume and so it is removed.
 Where two vertices at the same horizontal location are moved up to the surface they will occupy the same point; this results in a zero-length edge that is removed to create a triangular cell.

### 2015-12-18 src/tf-cutcell-comparison/responses2.tex
remove discussion of CTCS sensitivity to terrain profile/domain width

@@
 We had not intended to perform a sensitivity test, but only to clarify the specification for the \citet{schaer2002} advection test.
-\TODO{update the manuscript to put less emphasis on this}
+
+The paragraph discussing the sensitivity of CTCS to the change in mountain profile has been removed.  Line \TODO{xxx} clarifies the test setup required to reproduce \citet{schaer2002}'s results.
 
@@
 \end{quotation}
-\TODO{good idea, we should do this, but first see if anyone else does something similar (zaengl2012?).  Should we abandon the klemp2011 test entirely or just put less emphasis on its results?}
+\TODO{good idea.
+zaengl2012 did something similar by raising the mountain to 4km, but we think that simply moving the inversion layer is preferable because the mesh remains constant.
+Should we abandon the klemp2011 test entirely or just put less emphasis on its results?}
 

### 2015-12-18 src/tf-cutcell-comparison/tf-cutcell-comparison.tex
remove discussion of CTCS sensitivity to terrain profile/domain width

@@
 
-The domain is \SI{301}{\kilo\meter} wide and \SI{25}{\kilo\meter} high, discretized onto a grid with \(\Delta x = \SI{1}{\kilo\meter}\) and \(\Delta z^\star = \SI{500}{\meter}\).
-The domain specified by \citet{schaer2002} is \SI{300}{\kilo\meter} between the outermost cell centres where tracer values are specified.
-In order to reproduce the result from \citet{schaer2002}, the domain has been extended horizontally by \(\Delta x/2\) \si{\meter} in both directions so that the distance between the outermost cell centres is still \SI{300}{\kilo\meter}.
+The domain width is \SI{301}{\kilo\meter}, taken as the horizontal distance between the inlet and outlet boundaries.
+The domain is \SI{25}{\kilo\meter} high, discretized onto a grid with \(\Delta x = \SI{1}{\kilo\meter}\) and \(\Delta z^\star = \SI{500}{\meter}\).
+Note that \citet{schaer2002} measured the domain width as \SI{300}{\kilo\meter} between the outermost cell centres where tracer values are specified.
+The two domain sizes are equivalent irrespective of the measure.
 
@@
 Tracer minima and maxima for the centred linear and fourth order schemes are lower than those presented by \citet{schaer2002} because no interpolation is used to calculate the streamfunction.
-
-\TODO{put less emphasis on this, just clarify the correct test setup to obtain Schaer's results.}
-The centred linear scheme is sensitive to grid-scale changes in the terrain profile: when the domain width is reduced to \SI{300}{\kilo\meter}, the $\ell_2$ error on the BTF grid rises from \num{0.284} to \num{0.432}.
-The same sensitivity is found using terrain following coordinates and centred differences in space and time, as implemented by \citet{schaer2002}.
-The upwind-biased cubic scheme is less sensitive to the same change, with the $\ell_2$ error decreasing from \num{0.112} to \num{0.105}.
-It is likely that changes in the domain width affect the wave power spectrum of the discrete terrain profile.
-Grid-scale oscillations persist in the centred linear scheme but are damped by the upwind-biased scheme.
 

### 2015-12-18 src/tf-cutcell-comparison/responses2.tex
improve motivation for TF tracer advection test

@@
 \end{quotation}
-The advection scheme fits a multidimensional polynomial using a stencil of cell centroids, hence grid distortions will affect the resulting least squares fit.
+The advection scheme fits a multidimensional polynomial using a stencil of cell centroids, hence grid distortions could affect the resulting least squares fit.
 If the test used a dimensionally-split advection scheme with a terrain-following coordinate transform, then the test would indeed reduce to a one-dimensional problem.
-\TODO{clarify in the manuscript}
+We have provided a more detailed motivation of this test in the introductory paragraph starting at line \TODO{xxx}.
 

### 2015-12-18 src/tf-cutcell-comparison/tf-cutcell-comparison.tex
improve motivation for TF tracer advection test

@@
 Here, we formulate a new tracer advection test in which the velocity field is everywhere tangential to the basic terrain following coordinate surfaces.
+On the BTF grid, the flow is then aligned with the grid, but the data in the multidimensional advection stencil is not uniformly distributed because the BTF grid is non-orthogonal.
+Conversely, on the cut cell grid, the flow is misaligned with the grid but, except in the lowest layer, the grid is orthogonal.
 This test determines whether the primary source of numerical error is due to non-orthogonality or misalignment of the flow with grid layers.
-\TODO{explain how grid distortion could reduce accuracy of the advection scheme}
 

### 2015-12-18 src/tf-cutcell-comparison/tf-cutcell-comparison.tex
add note for justifying x=50km sample line

@@
 A vertical profile is taken at $x = \SI{50}{\kilo\meter}$, marked by the dashed line in figure~\ref{fig:gw-theta}, with results shown for the BTF grids in figure~\ref{fig:gw-sampleLine}a and on the cut cell grids in figure~\ref{fig:gw-sampleLine}b.
+\TODO{say why x=50km is a good place to take these measurements.}
 On all grids, potential temperature differences increase with height in the lowest \SI{1200}{\meter} at $x = \SI{50}{\kilo\meter}$, in agreement with the results seen in figure~\ref{fig:gw-theta}.

### 2015-12-18 src/tf-cutcell-comparison/tf-cutcell-comparison.tex
better illustrate how to avoid very thin cells

@@
 	\caption{Illustration of a slanted cell grid (a) before, and (b) after construction.
-	The terrain surface, denoted by a heavy dotted line, intersects a uniform rectangular grid comprising four cells, $c_1$, $c_2$, $c_3$ and $c_4$.
-The cell vertices, marked by open circles, are moved upwards to the points at which the terrain intersects vertical cell edges, marked by filled circles.
+	The terrain surface, denoted by a heavy dotted line, intersects a uniform rectangular grid comprising six cells, $c_1, \ldots, c_6$.
+The cell vertices, marked by open circles, are moved to the points at which the terrain intersects vertical cell edges, marked by filled circles.
 Cells that have no volume are removed.
 Where a cell has two vertices occupying the same point, the zero-length edge that joins those vertices is removed.
-In this illustration, cell $c_4$ is removed because it has no volume, and the zero-length edge at point $p$ is removed to create a triangular cell, $c_3$.}
+In this illustration, cells $c_5$ and $c_6$ are removed because they have no volume, and the zero-length edge at point $q$ is removed to create a triangular cell, $c_4$.
+Point $p$ is moved down because it is within $\Delta z/5$ of the surface, avoiding the creation of a thin cell.}
 	\label{fig:grid-generation}

### 2016-01-04 src/tf-cutcell-comparison/tf-cutcell-comparison.tex
remove TODO for improving slanted cell figure

@@
 Additionally, to avoid creating very thin cells, all vertices up to $\Delta z/5$ above the orography are moved down to the surface.
-\TODO{should I include an illustration of how this extra tweak helps?
-can I incorporate it into the existing figure?
-yes, by adding a row above the existing top row, and moving the rightmost terrain point closer to the upper right vertex in $c_2$} Where all four of a cell's vertices are moved, the cell has zero volume and so it is removed.
+Where all four of a cell's vertices are moved, the cell has zero volume and so it is removed.
 Where two vertices at the same horizontal location are moved up to the surface they will occupy the same point; this results in a zero-length edge that is removed to create a triangular cell.

### 2016-01-04 src/tf-cutcell-comparison/tf-cutcell-comparison.tex
linear results no longer unstable, update cell area ratios for cut cell meshes

@@
 
-Using the centred linear scheme on the cut cell grid, numerical instability is generated by flow through small cells: when interpolating \(\phi\) onto faces, flux from a large cell into a small cell means that the downwind cell is given greater weighting, effectively making the scheme downwind-biased rather than upwind-biased.
-The maximum Courant number is less than one in all of the advection tests, hence it is not the source of this instability.
-
 We conclude from this test that accuracy depends upon alignment of the flow with the grid, and accuracy is not significantly reduced by grid distortions.
@@
 \hline
-500	& 1.01 &  999 \\
-300	& 1.01 &  999 \\
-250	& 1.01 &  999 \\
-200	& 1.01 &  999 \\
-150	& 1.01 &  999 \\
-125	& 1.01 &  999 \\
-100	& 1.01 &  999 \\
-75	& 1.01 &  999 \\
-50	& 1.01 &  999 \\
+500	& 1.01 &  1.68 \\
+300	& 1.01 &  7.33 \\
+250	& 1.01 &  5.44 \\
+200	& 1.01 &  5.20 \\
+150	& 1.01 &  10.0 \\
+125	& 1.01 &  10.2 \\
+100	& 1.01 &  10.0 \\
+75	& 1.01 &  10.3 \\
+50	& 1.01 &  10.4 \\
 \hline

### 2016-01-05 src/tf-cutcell-comparison/responses2.tex
Remove detailed discussion of horizontal advection
And update advection statistics table for slanted cell grids.

@@
 
-Sections 4a and 4b – Section 4a presents results for the Sch\"{a}r advection test case for various vertical coordinates and advection schemes; there is little new here that is not already covered in the \citet{schaer2002} paper.
+Sections 4a and 4b -- Section 4a presents results for the Sch\"{a}r advection test case for various vertical coordinates and advection schemes; there is little new here that is not already covered in the \citet{schaer2002} paper.
 The advection tests in section 4b provide an interesting counter example to this test.
@@
 \end{quotation}
+\TODO{figure 3 will change to show some TF advection results}.  The discussion of the horizontal advection results starting at line \TODO{xxx} has been rewritten more succinctly.
+
 \begin{quotation}
@@
 \end{quotation}
-\TODO{agreed, although the comparison with the fourth order scheme in table 1 is perhaps useful?}
+\TODO{agreed, although the comparison with the fourth order scheme in table 1 is perhaps useful?  I have not removed it yet.}
 
@@
 \end{quotation}
-\TODO{good idea.
-zaengl2012 did something similar by raising the mountain to 4km, but we think that simply moving the inversion layer is preferable because the mesh remains constant.
-Should we abandon the klemp2011 test entirely or just put less emphasis on its results?}
+\TODO{good idea.  Should we abandon the klemp2011 test entirely or just put less emphasis on its results?}
 

### 2016-01-05 src/tf-cutcell-comparison/tf-cutcell-comparison.tex
Remove detailed discussion of horizontal advection
And update advection statistics table for slanted cell grids.

@@
 
-Tracer contours at \(t = \SI{0}{\second}, \SI{5000}{\second}\) and \(\SI{10000}{\second}\) are shown in Figure~\ref{fig:advection-tracer}.
-The results are compared on the BTF grid for the centred linear scheme (\ref{fig:advection-tracer}a) and the fourth order scheme from \citet{schaer2002} (\ref{fig:advection-tracer}c), and the upwind-biased cubic scheme on the cut cell grid (\ref{fig:advection-tracer}b), and BTF grid (\ref{fig:advection-tracer}d).
-Tracer errors at \(t = \SI{10000}{\second}\) are shown for the fourth order and upwind-biased cubic schemes in figures \ref{fig:advection-tracer}e and \ref{fig:advection-tracer}f respectively.
-
-By \(t = \SI{10000}{\second}\), the tracer suffers from distortion on the BTF grid using the centred linear scheme and some artifacts remain about the mountain peak.
-The results are visually identical to those using the modified Sch\"{a}r model with centred linear differencing.
-The tracer has spread vertically due to increased numerical errors when the tracer is transported between layers.
-Distortions are reduced by using the fourth order scheme from \citet{schaer2002} (figure~\ref{fig:advection-tracer}c), but dispersion errors remain with grid-scale oscillations that travel in the opposite direction to the wind (figure~\ref{fig:advection-tracer}e).
-The results from the centred linear and fourth order schemes are slightly worse than the respective results from \citet{schaer2002} (their figure 6a and figure 8) because of the difference in the discretisation of the streamfunction, $\Psi$.
-
-Using the upwind-biased cubic scheme, tracer magnitude and shape are well-preserved on all grids, both above the mountain at \(t = \SI{5000}{\second}\) and past the mountain at \(t = \SI{10000}{\second}\).
-In this test, advection is most accurate on the cut cell grid (figure~\ref{fig:advection-tracer}b) and regular grid (not shown).
-As found by \citet{good2014}, the result is the same on both grids.
-This is to be expected since the wind is zero in the region of the ground and flow aloft is aligned with the grids.
-On the BTF grid, the tracer is less distorted by the cubic upwind-biased scheme (figure~\ref{fig:advection-tracer}d) compared to the centred linear scheme (figure~\ref{fig:advection-tracer}a) or fourth order scheme (figure~\ref{fig:advection-tracer}c).
-
 Minimum and maximum tracer values and \(\ell_2\) error norms on the BTF, SLEVE, cut cell and regular grids are summarised in table~\ref{tab:advection}, where the \(\ell_2\) error norm is defined as 
@@
 Error is largest on the BTF grid with \(\ell_2 = \num{0.112}\) but is significantly reduced on the SLEVE grid with \(\ell_2 = \num{0.0146}\).
-The error is approximately halved by changing from the SLEVE grid to the cut cell grid.
+Advection is most accurate on the cut cell and regular grids, with \(\ell_2\) approximately half of that on the SLEVE grid.
 Tracer minima and maxima for the centred linear and fourth order schemes are lower than those presented by \citet{schaer2002} because no interpolation is used to calculate the streamfunction.
@@
 
-The spatial domain, mountain profile, initial tracer profile and discretisation are the same as those in the horizontal tracer advection test.
+The spatial domain, mountain profile, initial tracer profile and discretisation are the same as those in the horizontal tracer advection test, except for the timestep \(\Delta t = \SI{20}{\second}\).
 The velocity field is defined using a streamfunction, $\Psi$, so that the discrete velocity field is non-divergent and follows the BTF coordinate surfaces given by equation~(\ref{eqn:btf}) such that
@@
 
+Tracer contours at \(t = \SI{0}{\second}, \SI{5000}{\second}\) and \(\SI{10000}{\second}\) are shown in Figure~\ref{fig:advection-tracer}.
+\TODO{discuss figure~\ref{fig:advection-tracer}}
+Tracer errors at \(t = \SI{10000}{\second}\) are shown for the \TODO{schemes} in figures \ref{fig:advection-tracer}x and \ref{fig:advection-tracer}x respectively.
+
 $\ell_2$ errors and tracer extrema for this test are compared with the horizontal advection results in table~\ref{tab:advection}.
 In the terrain following velocity field, tracer accuracy is greatest on the BTF grid.
-Using the cubic upwind-biased scheme, errors are about ten times larger on the SLEVE and cut cell grids compared to the BTF grid.
+Errors are about ten times larger on the SLEVE and cut cell grids compared to the BTF grid.
 
@@
 \hline
-Horizontal        & Centred linear      & \(\ell_2\) error & 0        & 0.284   & 0.0316          & 999         & 0.0304      \\
-                  &                     & min              & 0        & -0.275  & -0.0252         & 999        & -0.0251     \\
-                  &                     & max              & 1        & 0.925   & 0.985           & 999          & 0.985       \\
+Horizontal        & Centred linear      & \(\ell_2\) error & 0        & 0.284   & 0.0316          & 0.0304         & 0.0304      \\
+                  &                     & min              & 0        & -0.275  & -0.0252         & -0.0251        & -0.0251     \\
+                  &                     & max              & 1        & 0.925   & 0.985           & 0.985          & 0.985       \\
                   & Fourth order        & \(\ell_2\) error & 0        & 0.0938  & 0.00244         & {---}          & 0.00234     \\
@@
                   &                     & max              & 1        & 1.00    & 0.984           & {---}          & 0.983       \\
-                  & Cubic upwind-biased & \(\ell_2\) error & 0        & 0.112   & 0.0146          & 999        & 0.00784     \\
-                  &                     & min              & 0        & -0.0464 & -0.0106         & 999      & -0.00674    \\
-                  &                     & max              & 1        & 0.922   & 0.982           & 999          & 0.983       \\
+                  & Cubic upwind-biased & \(\ell_2\) error & 0        & 0.112   & 0.0146          & 0.00784        & 0.00784     \\
+                  &                     & min              & 0        & -0.0464 & -0.0106         & -0.00674       & -0.00674    \\
+                  &                     & max              & 1        & 0.922   & 0.982           & 0.983          & 0.983       \\
 \hline
-Terrain following & Centred linear      & \(\ell_2\) error & 0        & 0.0341  & 0.235           & 999     & {---}        \\
-	          &                     & min              & 0        & -0.0245 & -0.120          & 999     & {---}        \\
-		  &                     & max              & 1        & 0.985   & 0.950           & 999     & {---}        \\
-		  & Cubic upwind-biased & \(\ell_2\) error & 0        & 0.0209  & 0.162           & 999          & {---}        \\
-                  &                     & min              & 0        & -0.0110 & -0.0263         & 999        & {---}        \\
-                  &                     & max              & 1        & 0.983   & 0.865           & 999          & {---}        \\
+Terrain following & Centred linear      & \(\ell_2\) error & 0        & 0.0341  & 0.235           & 0.363          & {---}        \\
+	          &                     & min              & 0        & -0.0245 & -0.120          & -1.24          & {---}        \\
+		  &                     & max              & 1        & 0.985   & 0.950           & 1.01           & {---}        \\
+		  & Cubic upwind-biased & \(\ell_2\) error & 0        & 0.0209  & 0.162           & 0.181          & {---}        \\
+                  &                     & min              & 0        & -0.0110 & -0.0263         & -0.0284        & {---}        \\
+                  &                     & max              & 1        & 0.983   & 0.865           & 0.851          & {---}        \\
 \hline

### 2016-01-05 src/tf-cutcell-comparison/responses2.tex
justify sample line position

@@
 \end{quotation}
-\TODO{the gravity wave amplitude decreases but the advection errors do not.
-hence, this is a good choice for the sample line.
-we considered plotting heatmaps, but found that it was not as easy to distinguish between the gravity waves and the advection errors.}
+We considered plotting heatmaps but found that it was not easy to distinguish between gravity waves and numerical errors.
+The position of the sample line was chosen so that numerical errors would be revealed more clearly: far away from the mountain, the wave amplitude is small but the potential temperature errors remain constant in the lee of the mountain.
+The manuscript now explains this choice at line \TODO{xxx}.
 

### 2016-01-05 src/tf-cutcell-comparison/tf-cutcell-comparison.tex
justify sample line position

@@
 A vertical profile is taken at $x = \SI{50}{\kilo\meter}$, marked by the dashed line in figure~\ref{fig:gw-theta}, with results shown for the BTF grids in figure~\ref{fig:gw-sampleLine}a and on the cut cell grids in figure~\ref{fig:gw-sampleLine}b.
-\TODO{say why x=50km is a good place to take these measurements.}
+The position is chosen to be far away from the mountain where the gravity wave amplitude is small in order to better reveal numerical errors.
 On all grids, potential temperature differences increase with height in the lowest \SI{1200}{\meter} at $x = \SI{50}{\kilo\meter}$, in agreement with the results seen in figure~\ref{fig:gw-theta}.

### 2016-01-06 src/tf-cutcell-comparison/responses2.tex
plot and discuss TF advection results

@@
 \end{quotation}
-\TODO{figure 3 will change to show some TF advection results}.  The discussion of the horizontal advection results starting at line \TODO{xxx} has been rewritten more succinctly.
+Terrain following advection results on BTF and cut cell grids are now shown in figure 3, with the accompanying discussion starting at line \TODO{xxx}.
+The analysis of the horizontal advection tests has been rewritten more succinctly and starts at line \TODO{xxx}.
 
@@
 All test results on cut cell grids have been updated.
+\TODO{note changes that have occurred: centred linear scheme no longer unstable in TF advection; larger errors in thermal advection test?}
 

### 2016-01-06 src/tf-cutcell-comparison/tf-cutcell-comparison.tex
plot and discuss TF advection results

@@
 Tracer contours at \(t = \SI{0}{\second}, \SI{5000}{\second}\) and \(\SI{10000}{\second}\) are shown in Figure~\ref{fig:advection-tracer}.
-\TODO{discuss figure~\ref{fig:advection-tracer}}
-Tracer errors at \(t = \SI{10000}{\second}\) are shown for the \TODO{schemes} in figures \ref{fig:advection-tracer}x and \ref{fig:advection-tracer}x respectively.
+The most accurate results are found on the BTF grid using the upwind-biased cubic scheme (\ref{fig:advection-tracer}a).
+As seen in figure~\ref{fig:advection-tracer}c, errors are less than \num{0.02} in magnitude and are confined to the expected region of the tracer.
+At \(t = \SI{5000}{\second}\) the tracer is distorted by the terrain-following velocity field.
+On the BTF grid, the tracer correctly returns to its original shape having cleared the mountain by \(t = \SI{10000}{\second}\), but this is not the case with centred linear scheme on the cut cell grid (figure~\ref{fig:advection-tracer}b).
+Here, the tracer has spread vertically due to increased numerical errors when the tracer is transported between layers.
+Dispersion errors are apparent with grid-scale oscillations that travel in the opposite direction to the wind (figure~\ref{fig:advection-tracer}d) and some artifacts remain above the mountain peak.
+
+Similar accuracy is obtained for both centred linear and upwind-biased cubic schemes on the BTF grid (figures~\ref{fig:advection-tracer}c and~\ref{fig:advection-tracer}e).
+However, results are substantially improved by using the upwind-biased cubic scheme on the cut cell grid (figures~\ref{fig:advection-tracer}d and~\ref{fig:advection-tracer}f).
+Results on the SLEVE grid are comparable to those on the cut cell grid except that the artifacts above the mountain peak with the centred linear scheme on the cut cell grid are not present on the SLEVE grid (not shown).
 
@@
 %
-	\caption{Horizontally advected tracer contours at \(t = \SI{0}{\second}\), \SI{5000}{\second} and \SI{10000}{\second} using (a) centred linear scheme on the BTF grid, (b) the upwind-biased cubic scheme on the cut cell grid, (c) the fourth order scheme from \citet{schaer2002} on the BTF grid, and (d) the upwind-biased cubic scheme on the BTF grid with contour intervals every 0.1.
-Errors on the BTF grid at \(t = \SI{10000}{\second}\) are shown for (e) the fourth order scheme from \citet{schaer2002}, and (f) the upwind-biased cubic scheme, with contour intervals every 0.01.
+	\caption{Tracer contours advected in a terrain following velocity field at \(t = \SI{0}{\second}\), \SI{5000}{\second} and \SI{10000}{\second} using (a) the upwind-biased cubic scheme on the BTF grid, and (b) the centred linear scheme on the cut cell grid with contour intervals every 0.1.
+Errors at \(t = \SI{10000}{\second}\) are shown for (c) the the upwind-biased cubic scheme on the BTF grid, (d) the centred linear scheme on the cut cell grid, (e) the centred linear scheme on the BTF grid, and (f) the upwind-biased cubic scheme on the cut cell grid with contour intervals every 0.01.
 Negative contours denoted by dotted lines.
-The terrain profile is also shown immediately above the $x$ axis.
-Subfigures (c) and (e) produced using the modified version of the code from \citet{schaer2002}.}
+The terrain profile is also shown immediately above the $x$ axis.}
 	\label{fig:advection-tracer}

### 2016-01-08 src/tf-cutcell-comparison/tf-cutcell-comparison.tex
tweak panel ordering of advection contours and update discussion accordingly

@@
 
-Tracer contours at \(t = \SI{0}{\second}, \SI{5000}{\second}\) and \(\SI{10000}{\second}\) are shown in Figure~\ref{fig:advection-tracer}.
-The most accurate results are found on the BTF grid using the upwind-biased cubic scheme (\ref{fig:advection-tracer}a).
-As seen in figure~\ref{fig:advection-tracer}c, errors are less than \num{0.02} in magnitude and are confined to the expected region of the tracer.
-At \(t = \SI{5000}{\second}\) the tracer is distorted by the terrain-following velocity field.
-On the BTF grid, the tracer correctly returns to its original shape having cleared the mountain by \(t = \SI{10000}{\second}\), but this is not the case with centred linear scheme on the cut cell grid (figure~\ref{fig:advection-tracer}b).
+Tracer contours at \(t = \SI{0}{\second}, \SI{5000}{\second}\) and \(\SI{10000}{\second}\) are shown in Figure~\ref{fig:advection-tracer} using the centred linear scheme on the BTF grid and cut cell grid (\ref{fig:advection-tracer}a and~\ref{fig:advection-tracer}b respectively).
+At \(t = \SI{5000}{\second}\), the tracer is distorted by the terrain-following velocity field.
+On the BTF grid, the tracer correctly returns to its original shape having cleared the mountain by \(t = \SI{10000}{\second}\), but this is not the case with centred linear scheme on the cut cell grid.
 Here, the tracer has spread vertically due to increased numerical errors when the tracer is transported between layers.
@@
 
-Similar accuracy is obtained for both centred linear and upwind-biased cubic schemes on the BTF grid (figures~\ref{fig:advection-tracer}c and~\ref{fig:advection-tracer}e).
-However, results are substantially improved by using the upwind-biased cubic scheme on the cut cell grid (figures~\ref{fig:advection-tracer}d and~\ref{fig:advection-tracer}f).
+A small improvement is obtained on the BTF grid by using the upwind-biased cubic scheme: as seen in figure~\ref{fig:advection-tracer}e, errors are less than \num{0.02} in magnitude and errors are confined to the expected region of the tracer.
+However, results are substantially improved by using the upwind-biased cubic scheme on the cut cell grid (figure~\ref{fig:advection-tracer}f).
 Results on the SLEVE grid are comparable to those on the cut cell grid except that the artifacts above the mountain peak with the centred linear scheme on the cut cell grid are not present on the SLEVE grid (not shown).
@@
 %
-	\caption{Tracer contours advected in a terrain following velocity field at \(t = \SI{0}{\second}\), \SI{5000}{\second} and \SI{10000}{\second} using (a) the upwind-biased cubic scheme on the BTF grid, and (b) the centred linear scheme on the cut cell grid with contour intervals every 0.1.
-Errors at \(t = \SI{10000}{\second}\) are shown for (c) the the upwind-biased cubic scheme on the BTF grid, (d) the centred linear scheme on the cut cell grid, (e) the centred linear scheme on the BTF grid, and (f) the upwind-biased cubic scheme on the cut cell grid with contour intervals every 0.01.
+	\caption{Tracer contours advected in a terrain following velocity field at \(t = \SI{0}{\second}\), \SI{5000}{\second} and \SI{10000}{\second} using the centred linear scheme on (a) the BTF grid, and (b) the cut cell grid with contour intervals every 0.1.
+Errors at \(t = \SI{10000}{\second}\) are shown for (c) the the centred linear scheme on the BTF grid, (d) the centred linear scheme on the cut cell grid, (e) the upwind-biased cubic scheme on the BTF grid, and (f) the upwind-biased cubic scheme on the cut cell grid with contour intervals every 0.01.
 Negative contours denoted by dotted lines.

### 2016-01-08 src/tf-cutcell-comparison/tf-cutcell-comparison.tex
update advection stats

@@
                   &                     & max              & 1        & 1.00    & 0.984           & {---}          & 0.983       \\
-                  & Cubic upwind-biased & \(\ell_2\) error & 0        & 0.112   & 0.0146          & 0.00784        & 0.00784     \\
+                  & Cubic upwind-biased & \(\ell_2\) error & 0        & 0.112   & 0.0146          & 0.000784       & 0.00784     \\
                   &                     & min              & 0        & -0.0464 & -0.0106         & -0.00674       & -0.00674    \\
@@
 \hline
-Terrain following & Centred linear      & \(\ell_2\) error & 0        & 0.0341  & 0.235           & 0.363          & {---}        \\
-	          &                     & min              & 0        & -0.0245 & -0.120          & -1.24          & {---}        \\
-		  &                     & max              & 1        & 0.985   & 0.950           & 1.01           & {---}        \\
-		  & Cubic upwind-biased & \(\ell_2\) error & 0        & 0.0209  & 0.162           & 0.181          & {---}        \\
-                  &                     & min              & 0        & -0.0110 & -0.0263         & -0.0284        & {---}        \\
+Terrain following & Centred linear      & \(\ell_2\) error & 0        & 0.0338  & 0.235           & 0.363          & {---}        \\
+	          &                     & min              & 0        & -0.0242 & -0.120          & -1.24          & {---}        \\
+		  &                     & max              & 1        & 0.984   & 0.950           & 1.01           & {---}        \\
+		  & Cubic upwind-biased & \(\ell_2\) error & 0        & 0.0207  & 0.162           & 0.181          & {---}        \\
+                  &                     & min              & 0        & -0.0110 & -0.0263         & -0.0285        & {---}        \\
                   &                     & max              & 1        & 0.983   & 0.865           & 0.851          & {---}        \\

### 2016-01-09 src/tf-cutcell-comparison/tf-cutcell-comparison.tex
include resting atmosphere results for high/low inversion layer

@@
 
-The initial potential temperature field has $\theta(z = 0) = \SI{288}{\kelvin}$ and a constant static stability with Brunt-V\"ais\"al\"a frequency $N = \SI{0.01}{\per\second}$ everywhere, except for a more stable layer of $N = \SI{0.02}{\per\second}$ between $\SI{2}{\kilo\meter} \leq z \leq \SI{3}{\kilo\meter}$.
+Tests were performed with two different stability profiles, both having an initial potential temperature field with $\theta(z = 0) = \SI{288}{\kelvin}$ and a constant static stability with Brunt-V\"ais\"al\"a frequency $N = \SI{0.01}{\per\second}$ everywhere, except for a more stable layer of $N = \SI{0.02}{\per\second}$.
+Figure~\ref{fig:resting}a shows where this inversion layer is positioned in the two tests: the `high inversion' test follows \citet{klemp2011}, placing the layer between $\SI{2}{\kilo\meter} \leq z \leq \SI{3}{\kilo\meter}$; the `low inversion' test is designed to challenge the numerics on the cut cell grid by placing the inversion layer between $\SI{0.5}{\kilo\meter} \leq z \leq \SI{1.5}{\kilo\meter}$ so that it intersects the terrain.
+
 The Exner function of pressure is calculated so that it is in discrete hydrostatic balance in the vertical direction \citep{weller-shahrokhi2014}.
@@
 
-The test was integrated forward by 5 hours using a timestep $\Delta t = \SI{100}{\second}$ on the BTF, SLEVE and cut cell grids, and a regular grid with flat terrain.
-Maximum vertical velocities are presented in figure~\ref{fig:resting}.
-In agreement with \citet{klemp2011}, vertical velocities are larger on more distorted grids.
-However, magnitudes are smaller comparing results on the terrain following grids with those from \citet{klemp2011}.
-The results presented in figure~\ref{fig:resting}, which use a curl-free pressure gradient, have maximum spurious values of $w$ of \SI{0.33}{\meter\per\second} on the BTF grid, compared with a maximum of \(\sim \SI{7}{\meter\per\second}\) found by \citet{klemp2011} using their improved horizontal pressure gradient formulation.
-The results in figure~\ref{fig:resting} have the same maximum errors as \citet{weller-shahrokhi2014} but, due to the more stable split into implicitly and explicitly treated terms (described in the appendix), the errors decay over time due to the dissipative nature of the advection scheme.
-
+The tests were integrated forward by 5 hours using a timestep $\Delta t = \SI{100}{\second}$ on the BTF, SLEVE and cut cell grids, and a regular grid with flat terrain.
+The results presented in figure~\ref{fig:resting}b, which use a curl-free pressure gradient, have maximum spurious values of $w$ of \SI{0.37}{\meter\per\second} on the BTF grid with the high inversion layer, compared with a maximum of \(\sim \SI{7}{\meter\per\second}\) found by \citet{klemp2011} using their improved horizontal pressure gradient formulation.
 Unlike the result from \citet{klemp2011}, the SLEVE grid does not significantly reduce vertical velocities compared to the BTF grid.
-However, errors are two orders of magnitude smaller on the cut cell grid with vertical velocities of \(\sim \SI{1e-4}{\meter\per\second}\).
+For the high inversion test, errors are two orders of magnitude smaller on the cut cell grid with vertical velocities of \(\sim \SI{1e-4}{\meter\per\second}\), but this advantage is lost when the inversion layer is lowered to intersect the terrain.
 The smallest error of \(\sim \SI{1e-10}{\meter\per\second}\) is found on the regular grid.
+The results for the high inversion test in figure~\ref{fig:resting}b have maximum errors that are comparable with \citet{weller-shahrokhi2014} but, due to the more stable split into implicitly and explicitly treated terms (described in the appendix), the errors decay over time due to the dissipative nature of the advection scheme.
 
@@
 
-In summary, spurious velocities in the resting atmosphere test were similar on both terrain following grids, with lower errors compared to those from \citet{klemp2011}.
-The maximum vertical velocity was significantly decreased on the cut cell grid, so we conclude that non-orthogonality, or lack of alignment of the grid with surfaces of constant gravitational potential are a significant cause of numerical error in this test.
+In summary, spurious velocities in the resting atmosphere test were similar on terrain following and cut cell grids, with lower errors compared to those from \citet{klemp2011}.
+\TODO{we previously concluded, ``The maximum vertical velocity was significantly decreased on the cut cell grid, so we conclude that non-orthogonality, or lack of alignment of the grid with surfaces of constant gravitational potential are a significant cause of numerical error in this test.'' --- but what can we say now we have the high/low inversion tests?}
 
@@
 %
-	\caption{Maximum spurious vertical velocity, \(w\) (\si{\meter\per\second}), in the resting atmosphere test with results on BTF, SLEVE, cut cell and regular grids using the model from \citet{weller-shahrokhi2014} which includes a curl-free pressure gradient formulation.}
+	\caption{Setup and results of a stratified atmosphere initially at rest.
+Tests are performed on four grids for two different stability profiles, with panel (a) showing the placement of the inversion layer in the two profiles.
+The low inversion is positioned so that it intersects the terrain, shown immediately above the $x$ axis.
+In each test, the inversion layer has a Brunt-V\"ais\"al\"a frequency $N = \SI{0.02}{\per\second}$ and $N = \SI{0.01}{\per\second}$ elsewhere.  
+Panel (b) shows the maximum magnitude of spurious vertical velocity, \(w\) (\si{\meter\per\second}), with results on BTF, SLEVE, cut cell and regular grids using the model from \citet{weller-shahrokhi2014} which includes a curl-free pressure gradient formulation.
+Results for the high inversion test are shown with solid lines, the low inversion test with dashed lines.}
 	\label{fig:resting}

### 2016-01-09 src/tf-cutcell-comparison/responses2.tex
add responses for resting atmosphere test

@@
 \begin{quotation}
- Lines 131-147 on cut-cell grid generation - It's still not clear how or why a cell vertex  is moved laterally in constructing the cut-cell mesh.
+ Lines 131-147 on cut-cell grid generation -- It's still not clear how or why a cell vertex  is moved laterally in constructing the cut-cell mesh.
 Taking Fig.
@@
 All test results on cut cell grids have been updated.
-\TODO{note changes that have occurred: centred linear scheme no longer unstable in TF advection; larger errors in thermal advection test?}
 
 \begin{quotation}
-	Section 3b - Finite difference linear advection model - This advection model is solving the same flux form equation (4) as in the previous subsection.
+	Section 3b -- Finite difference linear advection model - This advection model is solving the same flux form equation (4) as in the previous subsection.
 Thus, the advection operator in (6) should be in terms of $\partial (u\phi)/\partial x$.
@@
 \begin{quotation}
- Lines 265-267 and table 1 – The authors are unnecessarily complicating their tests  by comparing results from the cut-cell grid and a regular grid.
+ Lines 265-267 and table 1 -- The authors are unnecessarily complicating their tests  by comparing results from the cut-cell grid and a regular grid.
 These grids should be identical for this test and merit no further comparative consideration.
@@
 \end{quotation}
-\TODO{agreed, although the comparison with the fourth order scheme in table 1 is perhaps useful?  I have not removed it yet.}
+We have removed the discussion but we believe that the inclusion of results on the regular grid is still useful because it enables a comparison with the fourth order results from \citet{schaer2002} which are not available on the cut cell grid.
 
 \begin{quotation}
- Lines 271-279 and Table 1 – The $\ell_2$ error norms have changed by more than an order  of magnitude from those reported in the original manuscript. What changed?
+ Lines 271-279 and Table 1 -- The $\ell_2$ error norms have changed by more than an order  of magnitude from those reported in the original manuscript. What changed?
 \end{quotation}
@@
 \begin{quotation}
- Lines 280-285 – Reviewer 2 questioned the large change in the $\ell_2$ error norm when 
+ Lines 280-285 -- Reviewer 2 questioned the large change in the $\ell_2$ error norm when 
  the domain width was changed from 301 to 300 km. I share the concern expressed 
@@
 \begin{quotation}
- Lines 321-322 – When the flow is aligned with the grid there is no cross coordinate 
+ Lines 321-322 -- When the flow is aligned with the grid there is no cross coordinate 
  flow and therefore the numerics are essentially reduced to a one-dimensional 
@@
 \begin{quotation}
- Lines 364-368 – The authors did not need to run these resting atmosphere 
+ Lines 364-368 -- The authors did not need to run these resting atmosphere 
  simulations to reach the conclusion stated in this paragraph. As stated in my first 
@@
 \end{quotation}
-\TODO{good idea.  Should we abandon the klemp2011 test entirely or just put less emphasis on its results?}
+We have now included test results with an inversion layer that intersects the terrain, which have similar maximum vertical velocity magnitudes on terrain following and cut cell grids.
+We have also included results for the original test to enable comparison with \citet{klemp2011}.
 
 \begin{quotation}
-	Section 4d – Sch\"{a}r mountain-wave tests. These tests are more interesting now that 
+	Section 4d -- Sch\"{a}r mountain-wave tests. These tests are more interesting now that 
  the case has been simulated with a variety of grid resolutions. However, I'm 
@@
 \begin{quotation}
- Lines 479-480 – As mentioned above, for the horizontal advection test, shouldn't the  results for the cut-cell and regular grids be identical?
+ Lines 479-480 -- As mentioned above, for the horizontal advection test, shouldn't the  results for the cut-cell and regular grids be identical?
 \end{quotation}
@@
 \begin{quotation}
- Lines 481-482 – Emphasizing an error reduction of two orders of magnitude for this 
+ Lines 481-482 -- Emphasizing an error reduction of two orders of magnitude for this 
  resting atmosphere case with cut cells compared to a terrain-following coordinate is 
@@
 \end{quotation}
-\TODO{this conclusion may change when we run the new resting atmosphere test}
+The advantage of the cut cell grid is not seen when the inversion layer is lowered to intersect the terrain.
+We have removed the emphasis on the error reduction from the concluding remarks of section 4c.
 

### 2016-01-11 src/tf-cutcell-comparison/responses2.tex
simplify discussion of resting atmosphere, only include low inversion results

@@
 \end{quotation}
-We have now included test results with an inversion layer that intersects the terrain, which have similar maximum vertical velocity magnitudes on terrain following and cut cell grids.
-We have also included results for the original test to enable comparison with \citet{klemp2011}.
+As suggested, the inversion layer has been lowered to intersects the terrain.
+Similar maximum vertical velocity magnitudes are found on terrain following and cut cell grids.
+Section 4c has been updated accordingly.
 
@@
 The advantage of the cut cell grid is not seen when the inversion layer is lowered to intersect the terrain.
-We have removed the emphasis on the error reduction from the concluding remarks of section 4c.
+Section 4c has been rewritten to discuss the new results and this conclusion has been removed.
+Also removed is the paragraph previously starting at line 488, stating that there was a trade-off between pressure gradient calculations and advection errors.
 

### 2016-01-11 src/tf-cutcell-comparison/tf-cutcell-comparison.tex
simplify discussion of resting atmosphere, only include low inversion results

--- 2016-01-08 src/tf-cutcell-comparison/tf-cutcell-comparison.tex
@@
 An idealised terrain profile is defined along with a stably stratified atmosphere at rest in hydrostatic balance.
-The analytic solution is time-invariant, but numerical errors in calculating the horizontal pressure gradient can give rise to spurious velocities which become more severe over steeper terrain \citep{klemp2011}.
+The analytic solution is time-invariant, but numerical errors in calculating the pressure gradient can give rise to spurious velocities which become more severe over steeper terrain \citep{klemp2011}.
+Cut cell grids are often suggested as a technique for reducing these spurious circulations \citep{yamazaki-satomura2010,jebens2011,good2014}.
 
@@
 
-The initial potential temperature field has $\theta(z = 0) = \SI{288}{\kelvin}$ and a constant static stability with Brunt-V\"ais\"al\"a frequency $N = \SI{0.01}{\per\second}$ everywhere, except for a more stable layer of $N = \SI{0.02}{\per\second}$ between $\SI{2}{\kilo\meter} \leq z \leq \SI{3}{\kilo\meter}$.
+The initial potential temperature field has $\theta(z = 0) = \SI{288}{\kelvin}$ and a constant static stability with Brunt-V\"ais\"al\"a frequency $N = \SI{0.01}{\per\second}$ everywhere, except for a more stable layer of $N = \SI{0.02}{\per\second}$.
+In order to challenge pressure gradient calculations on the cut cell grid, this inversion layer is placed between $\SI{0.5}{\kilo\meter} \leq z \leq \SI{1.5}{\kilo\meter}$ so that it intersects the terrain, unlike \citet{klemp2011} who position the inversion layer above the mountain.
 The Exner function of pressure is calculated so that it is in discrete hydrostatic balance in the vertical direction \citep{weller-shahrokhi2014}.
@@
 The test was integrated forward by 5 hours using a timestep $\Delta t = \SI{100}{\second}$ on the BTF, SLEVE and cut cell grids, and a regular grid with flat terrain.
-Maximum vertical velocities are presented in figure~\ref{fig:resting}.
-In agreement with \citet{klemp2011}, vertical velocities are larger on more distorted grids.
-However, magnitudes are smaller comparing results on the terrain following grids with those from \citet{klemp2011}.
-The results presented in figure~\ref{fig:resting}, which use a curl-free pressure gradient, have maximum spurious values of $w$ of \SI{0.33}{\meter\per\second} on the BTF grid, compared with a maximum of \(\sim \SI{7}{\meter\per\second}\) found by \citet{klemp2011} using their improved horizontal pressure gradient formulation.
-The results in figure~\ref{fig:resting} have the same maximum errors as \citet{weller-shahrokhi2014} but, due to the more stable split into implicitly and explicitly treated terms (described in the appendix), the errors decay over time due to the dissipative nature of the advection scheme.
-
+Maximum vertical velocities are presented in figure~\ref{fig:resting} and are similar on the BTF, SLEVE and cut cell grids.
+The the largest vertical velocity of \SI{0.37}{\meter\per\second} was found on the BTF grid, compared with a maximum of \(\sim \SI{7}{\meter\per\second}\) found by \citet{klemp2011} using their improved horizontal pressure gradient formulation.
 Unlike the result from \citet{klemp2011}, the SLEVE grid does not significantly reduce vertical velocities compared to the BTF grid.
-However, errors are two orders of magnitude smaller on the cut cell grid with vertical velocities of \(\sim \SI{1e-4}{\meter\per\second}\).
-The smallest error of \(\sim \SI{1e-10}{\meter\per\second}\) is found on the regular grid.
 
@@
 
-In summary, spurious velocities in the resting atmosphere test were similar on both terrain following grids, with lower errors compared to those from \citet{klemp2011}.
-The maximum vertical velocity was significantly decreased on the cut cell grid, so we conclude that non-orthogonality, or lack of alignment of the grid with surfaces of constant gravitational potential are a significant cause of numerical error in this test.
+The results in figure~\ref{fig:resting} have maximum errors that are comparable with \citet{weller-shahrokhi2014} but, due to the more stable split into implicitly and explicitly treated terms (described in the appendix), the errors decay over time due to the dissipative nature of the advection scheme.
+
+In summary, spurious velocities in the resting atmosphere test were similarly small on all irregular grids, with the cut cell grid offering no improvement over terrain following grids.
 
@@
 We found that the accuracy of the upwind-biased cubic advection scheme depended upon alignment of the flow with the grid rather than on grid distortions.
-
-Spurious vertical velocities were small in the resting initial state test, reaching a maximum of $\sim \SI{0.35}{\meter\per\second}$ on the BTF grid, compared to a maximum of $\sim \SI{7}{\meter\per\second}$ found by \citet{klemp2011}.
+Spurious vertical velocities in the resting atmosphere test were similar on terrain following and cut cell grids.
 In the gravity waves test, vertical velocities were in good agreement with the reference solution from \citet{melvin2010} across all grids.
 
-Cut cell grids reduced errors in two of the five tests.
-First, in the horizontal advection test, accuracy was greatest on the cut cell grid.
-Second, in the resting atmosphere test, spurious vertical velocities were two orders of magnitude smaller on the cut cell grid compared with the terrain following grids.
-
-Conversely, in the first terrain following tracer advection test, errors were large on the SLEVE and cut cell grids where velocities were misaligned with the grids.
+Cut cell grids reduced errors in the horizontal advection test.
+Conversely, in the terrain following tracer advection test, errors were large on the SLEVE and cut cell grids where velocities were misaligned with the grids.
 Errors were also large on the cut cell grids in the second terrain following thermal advection test.
 This result suggests that, in the gravity waves test, potential temperature errors in the cut cell grids are primarily due to advection errors.
-
-The tests we have presented demonstrate the trade-off between errors in calculating pressure gradient and advection terms: pressure gradient errors can be reduced by using a cut cell grid but, when flow follows the terrain, advection errors are reduced using a terrain following grid.
 
@@
 %
-	\caption{Maximum spurious vertical velocity, \(w\) (\si{\meter\per\second}), in the resting atmosphere test with results on BTF, SLEVE, cut cell and regular grids using the model from \citet{weller-shahrokhi2014} which includes a curl-free pressure gradient formulation.}
+	\caption{Maximum magnitude of spurious vertical velocity, \(w\) (\si{\meter\per\second}), in the resting atmosphere test with results on BTF, SLEVE, cut cell and regular grids using the model from \citet{weller-shahrokhi2014} which includes a curl-free pressure gradient formulation.}
 	\label{fig:resting}

### 2016-01-12 src/tf-cutcell-comparison/WellerAppendix.tex

@@
 
-The semi-implicit technique involves combining the momentum \ref{eq:momentum},
-continuity \ref{eq:cont} and $\theta$ \ref{eq:theta} equations
-and the equation of state \ref{eq:state} to form a Helmholtz equation
+The semi-implicit technique involves combining the momentum (\ref{eq:momentum}),
+continuity (\ref{eq:cont}) and $\theta$ (\ref{eq:theta}) equations
+and the equation of state (\ref{eq:state}) to form a Helmholtz equation
 to be solved implicitly, as described by \citet{weller-shahrokhi2014}.
@@
 the faces of cell $c$, and $n_{f}=1$ if $\mathbf{d}_{f}$ points
-outwards from the cell and $n_{f}=$-1 otherwise. Equation \ref{eq:GaussDiv}
+outwards from the cell and $n_{f}=$-1 otherwise. Equation (\ref{eq:GaussDiv})
 is now a sum over a sum since $(HV)_{f}$ is one element of a matrix-vector
@@
 \end{equation}
-Equation \ref{eq:contD} can now be approximated by:
+Equation (\ref{eq:contD}) can now be approximated by:
 \begin{equation}
@@
 the approach taken by \citet{weller-shahrokhi2014}. However, the
-numerical solution of equation \ref{eq:contDd} turns out to be unstable
+numerical solution of equation (\ref{eq:contDd}) turns out to be unstable
 when using a large time-step on highly non-orthogonal meshes associated

### 2016-01-12 src/tf-cutcell-comparison/fig-gravityWaves-theta.tex
discuss small cell treatment by cubicUpwindCPCFit, tweaks to figures

--- 2015-10-19 src/tf-cutcell-comparison/fig-gravityWaves-theta.tex
@@
 \documentclass{article}
-\usepackage{showframe}
-\usepackage[paperwidth=6.2in,paperheight=5.6in,margin=0in]{geometry}
+\usepackage[paperwidth=6.2in,paperheight=2.6in,margin=0in]{geometry}
 \usepackage{mathptmx}%{times}
@@
 \usepackage{color}
+\usepackage{tikz}
 
@@
 \TPMargin{3pt}
-\textblockcolour{white}
-\begin{textblock}{0.75}(1.4,0.5)
-	\normalsize
-	\centering
-	(a)
-\end{textblock}
-\begin{textblock}{0.75}(1.2,6.5)
-	\normalsize
-	\centering
-	(b)
-\end{textblock}
-\begin{textblock}{0.75}(9.05,6.5)
-	\normalsize
-	\centering
-	(c)
-\end{textblock}
-\begin{textblock}{0.75}(1.2,11.3)
-	\normalsize
-	\centering
-	(d)
-\end{textblock}
-\begin{textblock}{0.75}(9.05,11.3)
-	\normalsize
-	\centering
-	(e)
+\begin{textblock}{0.75}(11.7,0.3)
+\begin{tikzpicture}[
+  scale=0.25
+]
+\draw [dashed, ultra thick] (0,0) -- (0,18);
+\end{tikzpicture}
 \end{textblock}
 \centering
-\includegraphics[width=6in]{../gravityWaves-btf-250m-300dz/18000/thetaDiffOverview.pdf}
-\vspace*{0.1in}
-\includegraphics[width=3in]{../gravityWaves-btf-250m-300dz/18000/thetaDiffMountain.pdf}
-\includegraphics[width=3in]{../gravityWaves-cutCell-250m-300dz/18000/thetaDiffMountain.pdf}
-\includegraphics[width=3in]{../thermalAdvection-btf-250m-300dz/18000/tracerDiffMountain.pdf}
-\includegraphics[width=3in]{../thermalAdvection-cutCell-250m-300dz/18000/tracerDiffMountain.pdf}
-
-\newpage
-
-250m GW
-
-\includegraphics[width=3in]{../gravityWaves-btf-250m-250dz/18000/thetaDiffMountain.pdf}
-\includegraphics[width=3in]{../gravityWaves-cutCell-250m-250dz/18000/thetaDiffMountain.pdf}
-
-250m adv
-
-\includegraphics[width=3in]{../thermalAdvection-btf-250m-250dz/18000/tracerDiffMountain.pdf}
-\includegraphics[width=3in]{../thermalAdvection-cutCell-250m-250dz/18000/tracerDiffMountain.pdf}
-
-\newpage
-200m GW
-
-\includegraphics[width=3in]{../gravityWaves-btf-250m-200dz/18000/thetaDiffMountain.pdf}
-\includegraphics[width=3in]{../gravityWaves-cutCell-250m-200dz/18000/thetaDiffMountain.pdf}
-
-200m adv
-
-\includegraphics[width=3in]{../thermalAdvection-btf-250m-200dz/18000/tracerDiffMountain.pdf}
-\includegraphics[width=3in]{../thermalAdvection-cutCell-250m-200dz/18000/tracerDiffMountain.pdf}
-
-\newpage
-
-150m GW
-
-\includegraphics[width=3in]{../gravityWaves-btf-250m-150dz/18000/thetaDiffMountain.pdf}
-\includegraphics[width=3in]{../gravityWaves-cutCell-250m-150dz/18000/thetaDiffMountain.pdf}
-
-150m adv
-
-\includegraphics[width=3in]{../thermalAdvection-btf-250m-150dz/18000/tracerDiffMountain.pdf}
-\includegraphics[width=3in]{../thermalAdvection-cutCell-250m-150dz/18000/tracerDiffMountain.pdf}
+\includegraphics[width=6in]{../gravityWaves-btf-250m-50dz/18000/thetaDiffOverview.pdf} \\
+\includegraphics[height=5in,angle=270]{thetaDiffOverview_theta_diff.eps}
 \end{document}

### 2016-01-12 src/tf-cutcell-comparison/tf-cutcell-comparison.tex

@@
 Maximum vertical velocities are presented in figure~\ref{fig:resting} and are similar on the BTF, SLEVE and cut cell grids.
-The the largest vertical velocity of \SI{0.37}{\meter\per\second} was found on the BTF grid, compared with a maximum of \(\sim \SI{7}{\meter\per\second}\) found by \citet{klemp2011} using their improved horizontal pressure gradient formulation.
+The largest vertical velocity of \SI{0.37}{\meter\per\second} was found on the BTF grid, compared with a maximum of \(\sim \SI{7}{\meter\per\second}\) found by \citet{klemp2011} using their improved horizontal pressure gradient formulation.
 Unlike the result from \citet{klemp2011}, the SLEVE grid does not significantly reduce vertical velocities compared to the BTF grid.

### 2016-01-19 src/tf-cutcell-comparison/tf-cutcell-comparison.tex
update resting atmosphere discussion

@@
 
-The initial potential temperature field has $\theta(z = 0) = \SI{288}{\kelvin}$ and a constant static stability with Brunt-V\"ais\"al\"a frequency $N = \SI{0.01}{\per\second}$ everywhere, except for a more stable layer of $N = \SI{0.02}{\per\second}$.
-In order to challenge pressure gradient calculations on the cut cell grid, this inversion layer is placed between $\SI{0.5}{\kilo\meter} \leq z \leq \SI{1.5}{\kilo\meter}$ so that it intersects the terrain, unlike \citet{klemp2011} who position the inversion layer above the mountain.
+Tests were performed with two different stability profiles, both having an initial potential temperature field has $\theta(z = 0) = \SI{288}{\kelvin}$ and a constant static stability with Brunt-V\"ais\"al\"a frequency $N = \SI{0.01}{\per\second}$ everywhere, except for a more stable layer of $N = \SI{0.02}{\per\second}$.
+Figure~\ref{fig:resting}a shows where this inversion layer is positioned in the two tests: the `high inversion' test follows \citet{klemp2011}, placing the layer between $\SI{2}{\kilo\meter} \leq z \leq \SI{3}{\kilo\meter}$; the `low inversion' test is designed to challenge the pressure gradient calculations on the cut cell grid by placing the inversion layer between $\SI{0.5}{\kilo\meter} \leq z \leq \SI{1.5}{\kilo\meter}$ so that it intersects the terrain.
+
 The Exner function of pressure is calculated so that it is in discrete hydrostatic balance in the vertical direction \citep{weller-shahrokhi2014}.
@@
 
-The test was integrated forward by 5 hours using a timestep $\Delta t = \SI{100}{\second}$ on the BTF, SLEVE and cut cell grids, and a regular grid with flat terrain.
-Maximum vertical velocities are presented in figure~\ref{fig:resting} and are similar on the BTF, SLEVE and cut cell grids.
-The largest vertical velocity of \SI{0.37}{\meter\per\second} was found on the BTF grid, compared with a maximum of \(\sim \SI{7}{\meter\per\second}\) found by \citet{klemp2011} using their improved horizontal pressure gradient formulation.
-Unlike the result from \citet{klemp2011}, the SLEVE grid does not significantly reduce vertical velocities compared to the BTF grid.
+The test was integrated forward by 5 hours using a timestep $\Delta t = \SI{100}{\second}$ on the BTF, SLEVE and cut cell grids.
+Maximum vertical velocities are presented in figure~\ref{fig:resting}b and are similar on the BTF, SLEVE and cut cell grids.
+For the high inversion test, the largest vertical velocity of \SI{0.37}{\meter\per\second} was found on the BTF grid after \SI{400}{\second}, compared with a maximum of \(\sim \SI{7}{\meter\per\second}\) found by \citet{klemp2011} using their improved horizontal pressure gradient formulation.
+Errors are two orders of magnitude smaller on the cut cell grid with vertical velocities of \(\sim \SI{1e-4}{\meter\per\second}\), but this advantage is lost when the inversion layer is lowered to intersect the terrain.
+Unlike the result from \citet{klemp2011}, the SLEVE grid does not further reduce vertical velocities compared to the BTF grid.
+This implies that the numerics we are using are less sensitive to grid distortions.
 
@@
 
-The results in figure~\ref{fig:resting} have maximum errors that are comparable with \citet{weller-shahrokhi2014} but, due to the more stable split into implicitly and explicitly treated terms (described in the appendix), the errors decay over time due to the dissipative nature of the advection scheme.
+The results in figure~\ref{fig:resting}b have maximum errors that are comparable with \citet{weller-shahrokhi2014} but, due to the more stable split into implicitly and explicitly treated terms (described in the appendix), the errors decay over time due to the dissipative nature of the advection scheme.
 
-In summary, spurious velocities in the resting atmosphere test were similarly small on all irregular grids, with the cut cell grid offering no improvement over terrain following grids.
+In summary, we reproduce the result found by \citet{good2014} that cut cells can reduce spurious velocities over orography.
+However, in addition, we find that, with the right numerics, these errors can be very small on a BTF grid.
+We also find that, if changes in stratification intersect cut cells, spurious velocities on cut cell grids are comparable with those on TF grids.
 
@@
 %
-	\caption{Maximum magnitude of spurious vertical velocity, \(w\) (\si{\meter\per\second}), in the resting atmosphere test with results on BTF, SLEVE, cut cell and regular grids using the model from \citet{weller-shahrokhi2014} which includes a curl-free pressure gradient formulation.}
+	\caption{Setup and results of a stratified atmosphere initially at rest.
+Tests are performed on four grids for two different stability profiles, with panel (a) showing the placement of the inversion layer in the two profiles.
+The low inversion is positioned so that it intersects the terrain, shown immediately above the $x$ axis.
+In each test, the inversion layer has a Brunt-V\"ais\"al\"a frequency $N = \SI{0.02}{\per\second}$ and $N = \SI{0.01}{\per\second}$ elsewhere.  
+Panel (b) shows the maximum magnitude of spurious vertical velocity, \(w\) (\si{\meter\per\second}), with results on BTF, SLEVE, cut cell and regular grids using the model from \citet{weller-shahrokhi2014} which includes a curl-free pressure gradient formulation.
+Results for the high inversion test are shown with solid lines, the low inversion test with dashed lines.}
 	\label{fig:resting}

### 2016-01-19 src/tf-cutcell-comparison/tf-cutcell-comparison.tex
update advection and grid stats

@@
                   &                     & max              & 1        & 1.00    & 0.984           & {---}          & 0.983       \\
-                  & Cubic upwind-biased & \(\ell_2\) error & 0        & 0.112   & 0.0146          & 0.000784       & 0.00784     \\
+                  & Cubic upwind-biased & \(\ell_2\) error & 0        & 0.112   & 0.0146          & 0.00784        & 0.00784     \\
                   &                     & min              & 0        & -0.0464 & -0.0106         & -0.00674       & -0.00674    \\
@@
 \hline
-Terrain following & Centred linear      & \(\ell_2\) error & 0        & 0.0338  & 0.235           & 0.363          & {---}        \\
-	          &                     & min              & 0        & -0.0242 & -0.120          & -1.24          & {---}        \\
-		  &                     & max              & 1        & 0.984   & 0.950           & 1.01           & {---}        \\
+Terrain following & Centred linear      & \(\ell_2\) error & 0        & 0.0338  & 0.235           & 0.374          & {---}        \\
+	          &                     & min              & 0        & -0.0242 & -0.120          & -1.26          & {---}        \\
+		  &                     & max              & 1        & 0.984   & 0.950           & 1.11           & {---}        \\
 		  & Cubic upwind-biased & \(\ell_2\) error & 0        & 0.0207  & 0.162           & 0.181          & {---}        \\
-                  &                     & min              & 0        & -0.0110 & -0.0263         & -0.0285        & {---}        \\
+                  &                     & min              & 0        & -0.0109 & -0.0263         & -0.0284        & {---}        \\
                   &                     & max              & 1        & 0.983   & 0.865           & 0.851          & {---}        \\
@@
 500	& 1.01 &  1.68 \\
-300	& 1.01 &  7.33 \\
-250	& 1.01 &  5.44 \\
-200	& 1.01 &  5.20 \\
-150	& 1.01 &  10.0 \\
-125	& 1.01 &  10.2 \\
-100	& 1.01 &  10.0 \\
-75	& 1.01 &  10.3 \\
-50	& 1.01 &  10.4 \\
+300	& 1.01 &  4.11 \\
+250	& 1.01 &  3.52 \\
+200	& 1.01 &  6.04 \\
+150	& 1.01 &  6.46 \\
+125	& 1.01 &  6.12 \\
+100	& 1.01 &  6.22 \\
+75	& 1.01 &  5.98 \\
+50	& 1.01 &  6.29 \\
 \hline

### 2016-01-19 src/tf-cutcell-comparison/tf-cutcell-comparison.tex
increase tolerance on slanted cell meshes to 2dz/5

@@
 Starting from a uniform grid, all cell vertices that lie beneath the orography are moved up to the surface.
-Additionally, to avoid creating very thin cells, all vertices up to $\Delta z/5$ above the orography are moved down to the surface.
+Additionally, to avoid creating very thin cells, all vertices up to $2 \Delta z/5$ above the orography are moved down to the surface.
 Where all four of a cell's vertices are moved, the cell has zero volume and so it is removed.
@@
 In this illustration, cells $c_5$ and $c_6$ are removed because they have no volume, and the zero-length edge at point $q$ is removed to create a triangular cell, $c_4$.
-Point $p$ is moved down because it is within $\Delta z/5$ of the surface, avoiding the creation of a thin cell.}
+Point $p$ is moved down because it is within $2 \Delta z/5$ of the surface, avoiding the creation of a thin cell.}
 	\label{fig:grid-generation}

### 2016-01-21 src/tf-cutcell-comparison/tf-cutcell-comparison.tex
add TODOs, more plots

@@
 Enlargements of the error field near the mountain are shown in figure~\ref{fig:thermalAdvection} at $\Delta z = \SI{50}{\meter}$ with contours of potential temperature overlayed.
-Errors are too small to be visible on the BTF grid with a maximum $\ell_2$ error of \num{4.26e-4}.
+Errors are too small to be visible on the BTF grid with a maximum $\ell_2$ error of \TODO{\num{999}}.
 However, on the cut cell grid, the maximum error is \TODO{\num{999}}.
@@
 As seen in figure~\ref{fig:gw-theta}c, errors are negligible on the BTF grids, but figure~\ref{fig:gw-theta}d reveals significant errors in the lowest layers of the cut cell grids that were advected from the mountain peaks.
-The magnitude of error is about ten times larger than the potential temperature anomalies found on the cut cell grids in the gravity waves test seen in figure~\ref{fig:gw-theta}b.
-We offer two explanations for this difference: first, there are no restoring forces in the linear advection model and, second, the wind fields in the two tests are different.
+\TODO{the following is no longer true.
+in fact, so far, the reverse appears to be true: errors are larger in the GW tests compared to the thermal advection tests ...
+``The magnitude of error is about ten times larger than the potential temperature anomalies found on the cut cell grids in the gravity waves test seen in figure~\ref{fig:gw-theta}b.
+We offer two explanations for this difference: first, there are no restoring forces in the linear advection model and, second, the wind fields in the two tests are different.''}
 
@@
 %
-	\caption{Differences in potential temperature between the start and end of the gravity waves test on the BTF grid with $\Delta z = \SI{50}{\meter}$.
+	\caption{\TODO{replot this} Differences in potential temperature between the start and end of the gravity waves test on the BTF grid with $\Delta z = \SI{50}{\meter}$.
 The dashed line at $x = \SI{50}{\kilo\meter}$ marks the position of the vertical profile in figure~\ref{fig:gw-sampleLine}.
@@
 %
-	\caption{Vertical profiles of potential temperature differences between the start and end of the gravity waves test on (a) the BTF grid, and (b) the cut cell grid.
+	\caption{\TODO{awaiting more, higher resolution, results for this figure} Vertical profiles of potential temperature differences between the start and end of the gravity waves test on (a) the BTF grid, and (b) the cut cell grid.
 Results are compared with thermal advection tests results, showing differences in tracer density between the numeric and analytic solutions at $t = \SI{18000}{\second}$ on (c) the BTF grid, and (d) the cut cell grid.
@@
 %
-	\caption{Error in tracer density in the thermal advection test at a resolution of $\Delta z = \SI{50}{\meter}$ on (a) the BTF grid, and (b) the cut cell grid.
+	\caption{\TODO{replot these} Error in tracer density in the thermal advection test at a resolution of $\Delta z = \SI{50}{\meter}$ on (a) the BTF grid, and (b) the cut cell grid.
 Errors are negligible on the BTF grid, but on the cut cell grid errors are generated near mountainous terrain and are advected horizontally on the lee side.

### 2016-01-21 src/tf-cutcell-comparison/tf-cutcell-comparison.tex
stronger conclusions, better selection of GW mesh plots

@@
 \abstract{Terrain following coordinates are widely used in operational models but the cut cell method has been proposed as an alternative that can more accurately represent atmospheric dynamics over steep orography.
-Because the type of grid is usually chosen during model implementation, it is typically necessary to use different models in order to compare the accuracy of different grids.
-In contrast, here a single C-grid finite volume model is used to enable a like-for-like comparison of terrain following and cut cell grids.
-A series of standard two-dimensional tests using idealised terrain are performed: tracer advection in a prescribed horizontal velocity field, a test starting from stably stratified initial conditions, and orographically induced gravity waves described by nonhydrostatic dynamics.
-In addition, two new advection tests are formulated having a velocity field that is everywhere tangential to the terrain following coordinate surfaces.
-These new tests present a challenge to the advection scheme on cut cell grids.
-The results of the advection tests demonstrate that accuracy depends primarily upon alignment of the flow with the grid rather than the grid orthogonality.
-As expected, the cut cell grid maintains greater accuracy for horizontal advection and in the test of stationary, stratified flow.
+Because the type of grid is usually chosen during model implementation, it becomes necessary to use different models to compare the accuracy of different grids.
+In contrast, here a C-grid finite volume model enables a like-for-like comparison of terrain following and cut cell grids.
+A series of standard two-dimensional tests using idealised terrain are performed: tracer advection in a prescribed horizontal velocity field, a test starting from resting initial conditions, and orographically induced gravity waves described by nonhydrostatic dynamics.
+In addition, three new tests are formulated: a more challenging resting atmosphere case, and two new advection tests having a velocity field that is everywhere tangential to the terrain following coordinate surfaces.
+These new tests present a challenge on cut cell grids.
+The results of the advection tests demonstrate that accuracy depends primarily upon alignment of the flow with the grid rather than grid orthogonality.
+A resting atmosphere is well-maintained on all grids.
 In the gravity waves test, results on all grids are in good agreement with existing results from the literature, although terrain following velocity fields lead to errors on cut cell grids.
-Due to semi-implicit timestepping and an upwind-biased, explicit advection scheme, there were no timestep restrictions associated with small cut cells.}
+Due to semi-implicit timestepping and an upwind-biased, explicit advection scheme, there are no timestep restrictions associated with small cut cells.
+We do not find the significant advantages of cut cells or smoothed coordinates that other authors find.}
 
@@
 Note that \citet{schaer2002} measured the domain width as \SI{300}{\kilo\meter} between the outermost cell centres where tracer values are specified.
-The two domain sizes are equivalent irrespective of the measure.
+Both formulations create a cell centre (or mass point) rather than a cell face (or horizontal velocity point) over the top of the highest peak which is crucial for the accuracy of the centred advection schemes.
 
@@
 
-The test was executed on the BTF, SLEVE and cut cell grids, and on a regular grid with flat terrain using a centred linear scheme and the upwind-biased cubic scheme.
+The test was executed on the BTF, SLEVE and cut cell grids using a centred linear scheme and the upwind-biased cubic scheme.
 Results were also obtained on BTF and SLEVE grids with the fourth order scheme from \citet{schaer2002} using the modified version of their code.
@@
 Error is largest on the BTF grid with \(\ell_2 = \num{0.112}\) but is significantly reduced on the SLEVE grid with \(\ell_2 = \num{0.0146}\).
-Advection is most accurate on the cut cell and regular grids, with \(\ell_2\) approximately half of that on the SLEVE grid.
+Advection is most accurate on the cut cell grid, with \(\ell_2\) approximately half of that on the SLEVE grid.
 Tracer minima and maxima for the centred linear and fourth order schemes are lower than those presented by \citet{schaer2002} because no interpolation is used to calculate the streamfunction.
@@
 The results of the horizontal advection test show that numerical errors are due either to misalignment of the flow with the grid, or to grid distortions.
-Using the upwind-biased cubic scheme, distortions in the grid do not significantly distort the tracer.
+In the following section, we propose a new test in order to identify the cause of the errors.
 
@@
 With increasing resolutions up to $\Delta z = \SI{50}{\meter}$, the orography intersects more layers and becomes better resolved.
-Three of the cut cell grids are shown in figure~\ref{fig:gw-meshes} at $\Delta z = \SI{300}{\meter}$, $\SI{250}{\meter}$ and $\SI{200}{\meter}$.
-Small cells are visible on the $\SI{200}{\meter}$ grid but, on the $\SI{250}{\meter}$ grid, the grid generator has merged the small cells with those in the layer above.
+Three of the cut cell grids are shown in figure~\ref{fig:gw-meshes} at $\Delta z = \SI{300}{\meter}$, $\SI{200}{\meter}$ and $\SI{150}{\meter}$.
+Small cells are visible on the $\SI{150}{\meter}$ grid but, on the $\SI{200}{\meter}$ grid, the small cells are merged with those in the layer above.
 
@@
 \subsection{Terrain following advection of thermal profile}
-The potential temperature anomalies in the gravity waves test varied between terrain following and cut cell grids of the same resolution.
-This variation may be due to differences in the wind fields between grids, or errors in the advection of potential temperature, amongst other possible causes.
+The potential temperature anomalies in the gravity waves test do not converge with resolution when using the cut cell grids.
+This may be due to differences in the wind fields between grids, or errors in the advection of potential temperature, amongst other possible causes.
 To help establish the primary source of error, a new advection test is formulated in which the initial potential temperature field from the gravity waves test is used.
@@
 As seen in figure~\ref{fig:gw-theta}c, errors are negligible on the BTF grids, but figure~\ref{fig:gw-theta}d reveals significant errors in the lowest layers of the cut cell grids that were advected from the mountain peaks.
-\TODO{the following is no longer true.
-in fact, so far, the reverse appears to be true: errors are larger in the GW tests compared to the thermal advection tests ...
-``The magnitude of error is about ten times larger than the potential temperature anomalies found on the cut cell grids in the gravity waves test seen in figure~\ref{fig:gw-theta}b.
-We offer two explanations for this difference: first, there are no restoring forces in the linear advection model and, second, the wind fields in the two tests are different.''}
 
@@
 We found that the accuracy of the upwind-biased cubic advection scheme depended upon alignment of the flow with the grid rather than on grid distortions.
-Spurious vertical velocities in the resting atmosphere test were similar on terrain following and cut cell grids.
+Spurious vertical velocities in the resting atmosphere tests were similar on terrain following and cut cell grids.
 In the gravity waves test, vertical velocities were in good agreement with the reference solution from \citet{melvin2010} across all grids.
@@
 The implementation of this technique in OpenFOAM is available at \url{https://github.com/hertzsprung/AtmosFOAM/tree/shaw-weller-2015-mwr} and will be described in greater detail a future publication.
-Combined with semi-implicit timestepping and a grid generator that preserves cell length in the direction of the flow, small cells did not impose additional timestep constraints.
+Combined with semi-implicit timestepping and a new cut cell generation technique that preserves cell length in the direction of the flow, small cells did not impose additional timestep constraints.
+By using a suitable multidimensional advection scheme and a curl-free pressure gradient formulation, we did not find significant advantages of cut cells or smoothed coordinate systems unlike \citet{good2014,klemp2011,schaer2002}.
+In contrast, errors that do not reduce with resolution are on cut cell grids.  No significant problems were found when using basic terrain following grids.
 
@@
 	%
-	\caption{Illustration of a slanted cell grid (a) before, and (b) after construction.
+	\caption{Illustration of a cut cell grid (a) before, and (b) after construction.
 	The terrain surface, denoted by a heavy dotted line, intersects a uniform rectangular grid comprising six cells, $c_1, \ldots, c_6$.
@@
 	%
-	\caption{Cut cell grids used for the gravity waves and thermal advection tests at (a) $\Delta z = \SI{300}{\meter}$, (b) $\Delta z = \SI{250}{\meter}$, and (c) $\Delta z = \SI{200}{\meter}$.
+	\caption{Cut cell grids used for the gravity waves and thermal advection tests at (a) $\Delta z = \SI{300}{\meter}$, (b) $\Delta z = \SI{200}{\meter}$, and (c) $\Delta z = \SI{150}{\meter}$.
 The mountain peak $h_0 = \SI{250}{\meter}$.
-At $\Delta z = \SI{250}{\meter}$, the grid creation process has merged small cells with the cells in the layer above but, at $\Delta z = \SI{200}{\meter}$, small cells have been retained.
+At $\Delta z = \SI{300}{\meter}$ and $\Delta z = \SI{200}{\meter}$, the grid creation process has merged small cells with the cells in the layer above but, at $\Delta z = \SI{150}{\meter}$, small cells have been retained.
 The full two dimensional grids are \SI{300}{\kilo\meter} wide and \SI{30}{\kilo\meter} high.

### 2016-02-02 src/tf-cutcell-comparison/tf-cutcell-comparison.tex
update errors in thermalAdvection discussion

@@
 Enlargements of the error field near the mountain are shown in figure~\ref{fig:thermalAdvection} at $\Delta z = \SI{50}{\meter}$ with contours of potential temperature overlayed.
-Errors are too small to be visible on the BTF grid with a maximum $\ell_2$ error of \TODO{\num{999}}.
-However, on the cut cell grid, the maximum error is \TODO{\num{999}}.
-Advection errors are apparent around mountainous terrain, with small cells having the largest errors.
+Errors are only just visible on the BTF grid with an $\ell_2$ error of \num{1.12e-7}.
+However, on the cut cell grid, the error is about ten times larger.
+Advection errors are apparent around mountainous terrain, with small cells having some of the largest errors.
 These errors are advected horizontally along the lee slope, forming stripes.
@@
 %
-	\caption{\TODO{replot these} Error in tracer density in the thermal advection test at a resolution of $\Delta z = \SI{50}{\meter}$ on (a) the BTF grid, and (b) the cut cell grid.
+	\caption{Error in tracer density in the thermal advection test at a resolution of $\Delta z = \SI{50}{\meter}$ on (a) the BTF grid, and (b) the cut cell grid.
 Errors are negligible on the BTF grid, but on the cut cell grid errors are generated near mountainous terrain and are advected horizontally on the lee side.

### 2016-02-05 src/tf-cutcell-comparison/responses2.tex
add TODOs, more plots
strengthen conclusion
more corrections to refer to 'potential temperature' rather than 'tracer density', other small fixes and include line numbers in responses

@@
 \setlist[comment,1]{label={\arabic{commenti}.},leftmargin=*,resume=comment}
-\newcommand{\TODO}[1]{\textcolor{purple}{TODO: \emph{#1}}}
 
 \begin{document}
+In addition to the responses below, we have strengthened the abstract and conclusions, saying that we do not find any significant problems with results on BTF grids and we do not find significant advantages of cut cells or smoothed coordinates that are found by other authors.
+
 \section*{Review \#1}
@@
 \end{quotation}
-Terrain following advection results on BTF and cut cell grids are now shown in figure 3, with the accompanying discussion starting at line \TODO{xxx}.
-The analysis of the horizontal advection tests has been rewritten more succinctly and starts at line \TODO{xxx}.
+Terrain following advection results on BTF and cut cell grids are now shown in figure 3, with the accompanying discussion starting at line 282.
+The analysis of the horizontal advection tests has been rewritten more succinctly and starts at line 248.
 
@@
 All test results on cut cell grids have been updated.
+Table 3 has also been updated.
 
@@
 
-The paragraph discussing the sensitivity of CTCS to the change in mountain profile has been removed.  Line \TODO{xxx} clarifies the test setup required to reproduce \citet{schaer2002}'s results.
+The paragraph discussing the sensitivity of CTCS to the change in mountain profile has been removed.  Line 218 clarifies the test setup required to reproduce \citet{schaer2002}'s results.
 
@@
 If the test used a dimensionally-split advection scheme with a terrain-following coordinate transform, then the test would indeed reduce to a one-dimensional problem.
-We have provided a more detailed motivation of this test in the introductory paragraph starting at line \TODO{xxx}.
+We have provided a more detailed motivation of this test in the introductory paragraph starting at line 258.
 
@@
 The position of the sample line was chosen so that numerical errors would be revealed more clearly: far away from the mountain, the wave amplitude is small but the potential temperature errors remain constant in the lee of the mountain.
-The manuscript now explains this choice at line \TODO{xxx}.
+The manuscript now explains this choice at line 398.
 
@@
 \end{quotation}
-This was a mistake: table 1 confirmed that min, max and $\ell_2$ errors were identical on both grids.
-This line now reads, `in the horizontal advection test, accuracy was greatest on the cut cell grid'.
+This was a mistake: table 1 confirmed that min, max and $\ell_2$ errors were identical on both grids.  This line has been removed.
 
@@
 Section 4c has been rewritten to discuss the new results and this conclusion has been removed.
-Also removed is the paragraph previously starting at line 488, stating that there was a trade-off between pressure gradient calculations and advection errors.
+Also removed is the paragraph starting at line 488 in the previous manuscript draft, stating that there was a trade-off between pressure gradient calculations and advection errors.
 
@@
 \end{quotation}
-Removed `dimensionless' on line \TODO{xxx} and now included correct units where $\mu$ is specified on lines \TODO{xxx} and \TODO{xxx}.
+Removed `dimensionless' on line 190 and now included correct units where $\mu$ is specified on lines 325 and 370.
 
@@
 \end{quotation}
-Yes, amended line \TODO{xxx} to state that $p_0 = \SI{1000}{\hecto\pascal}$, which applies to all tests that use the model.
-Line \TODO{xxx} now cites \citet{weller-shahrokhi2014} which explains how the Exner function of pressure is calculated so that it is in discrete hydrostatic balance with the prescribed potential temperature file.
+Yes, amended line 189 to state that $p_0 = \SI{1000}{\hecto\pascal}$, which applies to all tests that use the model.
+Line 324 now cites \citet{weller-shahrokhi2014} which explains how the Exner function of pressure is calculated so that it is in discrete hydrostatic balance with the prescribed potential temperature file.
 
@@
 The value of $p_0$ applies to tests (c) and (d).
-For test (d), line \TODO{xxx} now states that $w = \SI{0}{\meter\per\second}$.
+For test (d), line 357 now states that $w = \SI{0}{\meter\per\second}$.
 For test (e), the velocity field is defined by the streamfunction in equation (26), and $w = \SI{0}{\meter\per\second}$ only where $\mathrm{d}h/\mathrm{d}x = 0$.
@@
 We have modified this section so that the tracer density field is now called a potential temperature field.
-Lines \TODO{xxx twice in the abstract, in the intro and conclusion} have also been amended accordingly so that this test is no longer refered to as a tracer advection test..
-Equation (\TODO{xxx}) now has $\theta_0 = \SI{288}{\kelvin}$.
+Lines 19, 22, 102, 459, 683 and 686--689 have also been amended accordingly so that this test is no longer refered to as a tracer advection test.
+Equation (28) now has $\theta_0 = \SI{288}{\kelvin}$.
 

### 2016-02-05 src/tf-cutcell-comparison/tf-cutcell-comparison.tex
more corrections to refer to 'potential temperature' rather than 'tracer density', other small fixes and include line numbers in responses

@@
 \bibpunct{(}{)}{;}{a}{}{,}
-
-\newcommand{\TODO}[1]{\textcolor{purple}{TODO: \emph{#1}}}
 
@@
 Section 2 presents the formulation of the terrain following and cut cell grids used in the experiments that follow.
-In section 3 we give the governing equations, outline the model from \citet{weller-shahrokhi2014} and describe the modification which improves stability for long timesteps in the presence of steep orography.
+In section 3 we give the governing equations and outline the model from \citet{weller-shahrokhi2014}.
 Section 4 analyses the results from three advection tests, a test of a stably stratified atmosphere initially at rest, and orographically induced gravity waves.
@@
 Where two vertices at the same horizontal location are moved up to the surface they will occupy the same point; this results in a zero-length edge that is removed to create a triangular cell.
-Figure~\ref{fig:grid-generation} shows how a $2 \times 2$-cell, uniform grid is transformed into a cut cell grid.
-Cell $c_4$ is removed because it has zero volume, and the zero-length edge at point $p$ is removed to create a triangular cell, $c_3$.
+Figure~\ref{fig:grid-generation} shows how a $2 \times 3$-cell, uniform grid is transformed into a cut cell grid.
+Cells $c_5$ and $c_6$ are removed because they have zero volume, and the zero-length edge at point $q$ is removed to create a triangular cell, $c_3$.
+Point $p$ is moved down because it is within $2\Delta z/5$ of the surface, avoiding the creation of a very thin cell.
 
-Some thin cells are generated but, unlike most cut cell grids, cells are typically made smaller in height but their width is unaltered.
+Some small cells are generated but, unlike most cut cell grids, cells are typically made smaller in height but their width is unaltered.
 A grid that has these thin cells can be seen in figure~\ref{fig:gw-meshes}c.
@@
 These errors are advected horizontally along the lee slope, forming stripes.
-The same error structure is present on all cut cell grids in which the terrain intersects the grid layers.
+The same error structure is present on all cut cell grids.
 
 For comparison with the potential temperature anomalies in the gravity waves test, vertical profiles of potential temperature error are taken at $x = \SI{50}{\kilo\meter}$.
-As seen in figure~\ref{fig:gw-theta}c, errors are negligible on the BTF grids, but figure~\ref{fig:gw-theta}d reveals significant errors in the lowest layers of the cut cell grids that were advected from the mountain peaks.
+As seen in figure~\ref{fig:gw-sampleLine}c, errors are negligible on the BTF grids, but figure~\ref{fig:gw-sampleLine}d reveals significant errors in the lowest layers of the cut cell grids that were advected from the mountain peaks.
 
@@
 Across all tests, a high degree of accuracy was achieved for all grids.
-Even on the highly-distorted BTF grid, which has previously been found to give poor results \citep{schaer2002,klemp2011,good2014}, errors were often small in the tests presented here.
+Even on the highly-distorted BTF grid errors were often small in the tests presented here.
 In the first two tests, tracers were advected by horizontal and terrain following velocity fields.
@@
 Conversely, in the terrain following tracer advection test, errors were large on the SLEVE and cut cell grids where velocities were misaligned with the grids.
-Errors were also large on the cut cell grids in the second terrain following thermal advection test.
+Errors were also large on the cut cell grids in the terrain following thermal advection test.
 This result suggests that, in the gravity waves test, potential temperature errors in the cut cell grids are primarily due to advection errors.
@@
 The low inversion is positioned so that it intersects the terrain, shown immediately above the $x$ axis.
-In each test, the inversion layer has a Brunt-V\"ais\"al\"a frequency $N = \SI{0.02}{\per\second}$ and $N = \SI{0.01}{\per\second}$ elsewhere.  
+In each test, the inversion layer has a Brunt-V\"ais\"al\"a frequency $N = \SI{0.02}{\per\second}$, and $N = \SI{0.01}{\per\second}$ elsewhere.  
 Panel (b) shows the maximum magnitude of spurious vertical velocity, \(w\) (\si{\meter\per\second}), with results on BTF, SLEVE, cut cell and regular grids using the model from \citet{weller-shahrokhi2014} which includes a curl-free pressure gradient formulation.
@@
 %
-	\caption{\TODO{replot this} Differences in potential temperature between the start and end of the gravity waves test on the BTF grid with $\Delta z = \SI{50}{\meter}$.
+	\caption{Differences in potential temperature between the start and end of the gravity waves test on the BTF grid with $\Delta z = \SI{50}{\meter}$.
 The dashed line at $x = \SI{50}{\kilo\meter}$ marks the position of the vertical profile in figure~\ref{fig:gw-sampleLine}.
@@
 %
-	\caption{\TODO{awaiting more, higher resolution, results for this figure} Vertical profiles of potential temperature differences between the start and end of the gravity waves test on (a) the BTF grid, and (b) the cut cell grid.
-Results are compared with thermal advection tests results, showing differences in tracer density between the numeric and analytic solutions at $t = \SI{18000}{\second}$ on (c) the BTF grid, and (d) the cut cell grid.
+	\caption{Vertical profiles of potential temperature differences between the start and end of the gravity waves test on (a) the BTF grid, and (b) the cut cell grid.
+Results are compared with thermal advection tests results, showing differences in potential temperature between the numeric and analytic solutions at $t = \SI{18000}{\second}$ on (c) the BTF grid, and (d) the cut cell grid.
 The results are convergent, except for errors found in the lowest layers on the cut cell grids.}
@@
 %
-	\caption{Error in tracer density in the thermal advection test at a resolution of $\Delta z = \SI{50}{\meter}$ on (a) the BTF grid, and (b) the cut cell grid.
+	\caption{Error in potential temperature (measured in \si{\kelvin}) in the thermal advection test at a resolution of $\Delta z = \SI{50}{\meter}$ on (a) the BTF grid, and (b) the cut cell grid.
 Errors are negligible on the BTF grid, but on the cut cell grid errors are generated near mountainous terrain and are advected horizontally on the lee side.
-Contours in tracer density are overlayed.
+Contours of the potential temperature field at $t = \SI{18000}{\second}$ are overlayed.
 Axes are in units of \si{\meter}.}
